### 2012-04-28 ms.tex
Start of localization paper

\documentclass{iopart}
\usepackage{aas_macros}
\usepackage[colorlinks]{hyperref}
\input{acronyms}

\begin{document}

\title[Rapid-response Bayesian sky localization]{Rapid-response Bayesian sky localization for electromagnetic follow-up of gravitational-wave candidates}
\author{Leo Singer and Larry Price}
\address{LIGO Laboratory, California Institute of Technology, Pasadena, CA 91125, USA}
\ead{\mailto{leo.singer@ligo.org}, \mailto{larry.price@ligo.org}}

\begin{abstract}
Description of rapid Bayesian sky localization algorithm
\end{abstract}

\section{Measurement uncertainty}
\cite{fairhurst:2009} estimated the uncertainty in the matched-filter measurement of the \ac{TOA} of a \ac{GW} signal by computing the scalar Fisher information.  We extend this result by computing the Fisher information matrix describing the uncertainties in the joint estimation of the signal amplitude and \ac{TOA}.

When a \ac{CBC} occurs, a detector's strain observation may be described by a linear combination of two orthogonal \ac{GW} quadratures, $a(t)$ and $b(t)$, and additive, zero\nobreakdashes-mean, \acl{WSS}, Gaussian measurement noise $n(t)$, with power spectral density $S(f)$.  Introducing the unknown, real, quadrature amplitudes, $A$ and $B$, and the unknown \ac{TOA}, $t_0$, we can write
%
\begin{equation}\label{eq:xoft}
	x(t) = A a(t - t_0) + B b(t - t_0) + n(t).
\end{equation}
%
Transformation of \Eref{eq:xoft} to the frequency domain converts the time delay by $t_0$ to a complex phase,
%
\begin{equation}\label{eq:xoff}
	\tilde{x}(f) = A \tilde{a}(f) + B \tilde{b}(f) + \tilde{n}(f).
\end{equation}
%

\ack Source code for \ac{BAYESTAR} is available on the \acs{LIGO} \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/bayestar.html}.

Some of the results in this paper have been derived using HEALPix \cite{healpix}.

\acs{LIGO} was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the \ac{NSF} and operates under cooperative agreement PHY\nobreakdashes-0107417.  Some results were produced on the NEMO computing cluster operated by the Center for Gravitation and Cosmology at University of Wisconsin\nobreakdashes--Milwaukee under \ac{NSF} Grants PHY\nobreakdashes-0923409 and PHY\nobreakdashes-0600953.  This research is supported by the \ac{NSF} through a Graduate Research Fellowship to L.S.  This paper has \acs{LIGO} Document Number \acs{LIGO}\nobreakdashes-PXXXXXXX\nobreakdashes-vX.

\section*{References}
\bibliographystyle{iopart-num}
\bibliography{apj-jour,telescope}

\end{document}

### 2012-04-30 ms.tex
bayestar-localization: added description of Fisher matrix calculation

@@
 \documentclass{iopart}
+\usepackage{iopams}
 \usepackage{aas_macros}
@@
 \begin{equation}\label{eq:xoff}
-	\tilde{x}(f) = A \tilde{a}(f) + B \tilde{b}(f) + \tilde{n}(f).
+	\tilde{x}(f) = \left(A \tilde{a}(f) + B \tilde{b}(f)\right)e^{-2 \pi i f t_0} + \tilde{n}(f).
 \end{equation}
 %
+The Fisher information matrix for a measurement $\mathbf{x}$ described by the unknown parameters $\boldsymbol{\theta}$ may be written as
+%
+\begin{equation}
+	\mathcal{I}_{jk} = \mathrm{E} \, \left[
+		-\frac{\partial^2 \log
+			\mathcal{L}(\mathbf{x}; \boldsymbol{\theta})}
+			{\partial \theta_j \theta_k}
+	\right].
+\end{equation}
+%
+First, we write the likelihood of the data $\tilde{x}(f)$ given the parameters $(A, B, t_0$):
+%
+\begin{equation}\label{eq:trigger-likelihood}
+	\fl\mathcal{L}(x; A, B, t_0) = p(x | A, B, t_0)
+		\propto \exp \left[
+		- \frac{1}{2} \int_{-\infty}^\infty \frac{\left|\tilde{x}(f)
+			- \left(A \tilde{a}(f)
+			- B \tilde{b}(f)\right) e^{-2 \pi i f t_0}\right|^2}{S(f)} \, \mathrm{d}f
+	\right].
+\end{equation}
+%
+The normalization has been dropped because the Fisher matrix elements depend only on derivatives of the log\nobreakdashes-likelihood function and any factor that is independent of the parameters makes no contribution.%
+%
+\footnote{It would be cleaner to work in discrete time so that the Gaussian likelihood function can be correctly normalized.  Wide-sense stationary, continuous time processes don't really exist.}
+%
+From hereon, unless otherwise specified the limits of integration are always $(-\infty, \infty)$ and will be dropped.
+Expanding the integrand, we can write the log\nobreakdashes-likelihood function up to an additive constant as
+%
+\begin{eqnarray}\label{eq:trigger-likelihood}
+	\log \mathcal{L}(x; A, B, t_0) \sim
+	&- \frac{1}{2} \int \frac{|\tilde{x}(f)|^2}{S(f)} \mathrm{d}f \\
+	&+ A \, \mathrm{Re} \int \frac{\tilde{a}^*(f) \tilde{x}(f) e^{2 \pi i f t_0}}{S(f)} \mathrm{d}f \nonumber\\
+	&+ B \, \mathrm{Re} \int \frac{\tilde{b}^*(f) \tilde{x}(f) e^{2 \pi i f t_0}}{S(f)} \mathrm{d}f\nonumber\\
+	&+ A B \, \mathrm{Re} \int \frac{\tilde{a}^*(f) \tilde{b}(f)}{S(f)} \mathrm{d}f \nonumber\\
+	&- \frac{1}{2} A^2 \int \frac{|\tilde{a}(f)|^2}{S(f)} \mathrm{d}f
+	- \frac{1}{2} B^2 \int \frac{|\tilde{b}(f)|^2}{S(f)} \mathrm{d}f.
+\end{eqnarray}
+%
+We assume that $\tilde{a}(f)$ and $\tilde{b}(f)$ are orthogonal and have the same norm,
+%
+\begin{equation}
+	\int \frac{\tilde{a}^*(f) \tilde{b}(f)}{S(f)} \mathrm{d}f = 0,
+	\int \frac{|\tilde{a}(f)|^2}{S(f)} \mathrm{d}f =
+	\int \frac{|\tilde{b}(f)|^2}{S(f)} \mathrm{d}f =
+	\sigma^2.
+\end{equation}
+%
+After we apply these substitutions, and drop the $|\tilde{x}(f)|^2$ term that is independent of $A$, $B$, or $t_0$, we can write the log\nobreakdashes-likelihood function as
+%
+\begin{eqnarray}\label{eq:trigger-likelihood-simplified}
+	\log \mathcal{L}(x; A, B, t_0) \sim
+	& A \, \mathrm{Re} \int \frac{\tilde{a}^*(f) \tilde{x}(f) e^{2 \pi i f t_0}}{S(f)} \mathrm{d}f 
+	&+ B \, \mathrm{Re} \int \frac{\tilde{b}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} }{S(f)} \mathrm{d}f\nonumber\\
+	&- \frac{1}{2} A^2 \sigma^2 - \frac{1}{2} B^2 \sigma^2.
+\end{eqnarray}
+%
+The necessary partial derivatives are
+%
+\begin{eqnarray*}
+	\frac{\partial^2 \log \mathcal{L}}{\partial A \partial B} &=
+	\frac{\partial^2 \log \mathcal{L}}{\partial A \partial t_0} =
+	\frac{\partial^2 \log \mathcal{L}}{\partial B \partial t_0} = 0 \\
+	\frac{\partial^2 \log \mathcal{L}}{\partial A^2} &=
+	\frac{\partial^2 \log \mathcal{L}}{\partial B^2} = - \sigma^2 \\
+	\frac{\partial^2 \log \mathcal{L}}{\partial {t_0}^2} &=
+	2 \pi i A \int \frac{\left(\tilde{a}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} - \tilde{a}(f) \tilde{x}^*(f) e^{-2 \pi i f t_0}\right)}{S(f)} f \mathrm{d}f \\
+	&+ 2 \pi i B \int \frac{\left(\tilde{b}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} - \tilde{b}(f) \tilde{x}^*(f) e^{-2 \pi i f t_0}\right)}{S(f)} f \mathrm{d}f \\
+	&- (2 \pi t_0)^2 A \int \frac{\left(\tilde{a}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} + \tilde{a}(f) \tilde{x}^*(f) e^{-2 \pi i f t_0}\right)}{S(f)} f^2 \mathrm{d}f \\
+	&- (2 \pi t_0)^2 B \int \frac{\left(\tilde{b}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} + \tilde{b}(f) \tilde{x}^*(f) e^{-2 \pi i f t_0}\right)}{S(f)} f^2 \mathrm{d}f.
+\end{eqnarray*}
+%
+Finally, taking the expectation values of the partial derivatives, we find
+%
+\begin{eqnarray*}
+	\mathcal{I}_{A,B} &= \mathcal{I}_{A,t_0} = \mathcal{I}_{B,t_0} = 0 \\
+	\mathcal{I}_{A,A} &= \mathcal{I}_{B,B} = \sigma^2 \\
+	\mathcal{I}_{t_0,t_0} &=
+	- 2 \pi i A^2 \int \frac{|\tilde{a}(f)|^2}{S(f)} f\mathrm{d}f
+	- 2 \pi i B^2 \int \frac{|\tilde{B}(f)|^2}{S(f)} f\mathrm{d}f \\
+	&+ (2 \pi t_0)^2 \left(A^2 \int \frac{|\tilde{a}(f)|^2}{S(f)} f^2\mathrm{d}f +
+	B^2 \int \frac{|\tilde{b}(f)|^2}{S(f)} f\mathrm{d}f \right) \\
+	&+ (2\pi t_0)^2 2 A B \, \mathrm{Re} \int \frac{\tilde{a}^*(f)\tilde{b}(f)}{S(f)} f^2\mathrm{d}f.
+\end{eqnarray*}
+%
+The first two terms of $\mathcal{I}_{t_0,t_0}$ vanish because the integrands $|\tilde{a}(f)|^2 f / S(f)$ and $|\tilde{b}(f)|^2 f / S(f)$ are odd about $f = 0$.
+We assume that $\tilde{a}(f)$ and $\tilde{b}(f)$ are orthogonal with respect to $\emph{any}$ real weight, including both $S(f)$ and $S(f) / f^2$, so the last term in $\mathcal{I}_{t_0,t_0}$ vanishes as well.
+Further, we make the substituions
+%
+\begin{eqnarray*}
+	\rho_A &= A \sigma \\
+	\rho_B &= B \sigma \\
+	f_\mathrm{rms} &= \left[ \int \frac{|\tilde{a}(f)|^2}{S(f)} f^2\mathrm{d}f \right] ^ {1/2} \left[\int \frac{|\tilde{a}(f)|^2}{S(f)} \mathrm{d}f \right] ^ {-1/2} \\
+	&=  \left[ \int \frac{|\tilde{b}(f)|^2}{S(f)} f^2\mathrm{d}f \right] ^ {1/2} \left[\int \frac{|\tilde{b}(f)|^2}{S(f)} \mathrm{d}f \right] ^ {-1/2}.
+\end{eqnarray*}
+%
+Then, we have
+%
+\begin{eqnarray*}
+	\mathcal{I}_{\rho_A,\rho_B} &= \mathcal{I}_{\rho_A,t_0} = \mathcal{I}_{\rho_B,t_0} = 0 \\
+	\mathcal{I}_{\rho_A,\rho_A} &= \mathcal{I}_{\rho_B,\rho_B} = 1 \\
+	\mathcal{I}_{t_0,t_0} &= (2 \pi t_0 f_\mathrm{rms})^2.
+\end{eqnarray*}
+%
+The Cram\'{e}r-Rao inequality says that the inverse of the Fisher information matrix is a lower bound on the covariance of any unbiased estimator.
+The fisher matrix for this estimation problem has been shown to be diagonal, so we can write
+%
+\begin{eqnarray}
+	\mathrm{var}\, (\hat{\rho}_B) &\geq {\mathcal{I}_{\rho_A,\rho_A}}^{-1} = 1 \\
+	\mathrm{var}\, (\hat{\rho}_B) &\geq {\mathcal{I}_{\rho_B,\rho_B}}^{-1} = 1 \\
+	\mathrm{var}\, (\hat{t}_0) &\geq {\mathcal{I}_{t_0,t_0}}^{-1} = (2 \pi t_0 f_\mathrm{rms})^{-2}.
+\end{eqnarray}
 

### 2012-05-01 ms.tex
Capitalization
Corrected some math mistakes, added some margin notes, discussed approximations to get to expression for estimator variance

@@
 \input{acronyms}
+
+% From http://www.f.kth.se/~ante/latex.php
+\setlength{\marginparwidth}{1.2in}
+\let\oldmarginpar\marginpar
+\renewcommand\marginpar[1]{\-\oldmarginpar[\raggedleft\footnotesize #1]%
+{\raggedright\footnotesize #1}}
 
@@
 \section{Measurement uncertainty}
+
 \cite{fairhurst:2009} estimated the uncertainty in the matched-filter measurement of the \ac{TOA} of a \ac{GW} signal by computing the scalar Fisher information.
-We extend this result by computing the Fisher information matrix describing the uncertainties in the joint estimation of the signal amplitude and \ac{TOA}.
+We extend this result by computing the Fisher information matrix describing the uncertainties in the joint estimation of the signal amplitude and \ac{TOA}.%
+%
+\marginpar{This might be too much mathematical detail.
+All we need to do is show that the $\rho_A\rho_A$ and $\rho_B\rho_B$ terms are unity and that the off-diagonal terms vanish.
+If possible, we should cite and use Fairhurst's expression for $\mathcal{I}_{t_0,t_0}$ without re-deriving it.}
 
@@
 %
-\footnote{It would be cleaner to work in discrete time so that the Gaussian likelihood function can be correctly normalized.  Wide-sense stationary, continuous time processes don't really exist.}
+\marginpar{Consider using discrete time to avoid the fiction of an infinite-dimensional multivariate Gaussian process.}
 %
@@
 	&+ 2 \pi i B \int \frac{\left(\tilde{b}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} - \tilde{b}(f) \tilde{x}^*(f) e^{-2 \pi i f t_0}\right)}{S(f)} f \mathrm{d}f \\
-	&- (2 \pi t_0)^2 A \int \frac{\left(\tilde{a}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} + \tilde{a}(f) \tilde{x}^*(f) e^{-2 \pi i f t_0}\right)}{S(f)} f^2 \mathrm{d}f \\
-	&- (2 \pi t_0)^2 B \int \frac{\left(\tilde{b}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} + \tilde{b}(f) \tilde{x}^*(f) e^{-2 \pi i f t_0}\right)}{S(f)} f^2 \mathrm{d}f.
+	&- (2 \pi)^2 A \int \frac{\left(\tilde{a}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} + \tilde{a}(f) \tilde{x}^*(f) e^{-2 \pi i f t_0}\right)}{S(f)} f^2 \mathrm{d}f \\
+	&- (2 \pi)^2 B \int \frac{\left(\tilde{b}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} + \tilde{b}(f) \tilde{x}^*(f) e^{-2 \pi i f t_0}\right)}{S(f)} f^2 \mathrm{d}f.
 \end{eqnarray*}
@@
 	- 2 \pi i B^2 \int \frac{|\tilde{B}(f)|^2}{S(f)} f\mathrm{d}f \\
-	&+ (2 \pi t_0)^2 \left(A^2 \int \frac{|\tilde{a}(f)|^2}{S(f)} f^2\mathrm{d}f +
+	&+ (2 \pi)^2 \left(A^2 \int \frac{|\tilde{a}(f)|^2}{S(f)} f^2\mathrm{d}f +
 	B^2 \int \frac{|\tilde{b}(f)|^2}{S(f)} f\mathrm{d}f \right) \\
-	&+ (2\pi t_0)^2 2 A B \, \mathrm{Re} \int \frac{\tilde{a}^*(f)\tilde{b}(f)}{S(f)} f^2\mathrm{d}f.
+	&+ (2\pi)^2 2 A B \, \mathrm{Re} \int \frac{\tilde{a}^*(f)\tilde{b}(f)}{S(f)} f^2\mathrm{d}f.
 \end{eqnarray*}
@@
 We assume that $\tilde{a}(f)$ and $\tilde{b}(f)$ are orthogonal with respect to $\emph{any}$ real weight, including both $S(f)$ and $S(f) / f^2$, so the last term in $\mathcal{I}_{t_0,t_0}$ vanishes as well.
-Further, we make the substituions
+Further,%
+%
+\marginpar{This is an awkward place to make these substitutions.
+Also, they are poorly motivated.
+Here's what this substitution actually does: the matched\nobreakdashes-filter output has an expected value of $A \sigma = \rho_A$.}
+%
+we make the substituions
 %
@@
 	\mathcal{I}_{\rho_A,\rho_A} &= \mathcal{I}_{\rho_B,\rho_B} = 1 \\
-	\mathcal{I}_{t_0,t_0} &= (2 \pi t_0 f_\mathrm{rms})^2.
+	\mathcal{I}_{t_0,t_0} &= (2 \pi f_\mathrm{rms})^2 ({\rho_A}^2 + {\rho_B}^2).
 \end{eqnarray*}
@@
 The Cram\'{e}r-Rao inequality says that the inverse of the Fisher information matrix is a lower bound on the covariance of any unbiased estimator.
-The fisher matrix for this estimation problem has been shown to be diagonal, so we can write
+The Fisher matrix for this estimation problem has been shown to be diagonal, so we can write
+%
+\begin{eqnarray}\label{eq:cramer-rao}
+	\mathrm{var}\, (\hat{\rho}_B) &\geq {\mathcal{I}_{\rho_A,\rho_A}}^{-1} = 1 \\
+	\mathrm{var}\, (\hat{\rho}_B) &\geq {\mathcal{I}_{\rho_B,\rho_B}}^{-1} = 1 \\
+	\mathrm{var}\, (\hat{t}_0) &\geq {\mathcal{I}_{t_0,t_0}}^{-1} = (2 \pi f_\mathrm{rms} \sqrt{{\rho_A}^2 + {\rho_B}^2})^{-2}.
+\end{eqnarray}
+
+\section{Posterior distribution}
+
+Because the \ac{CRLB} in \Eref{eq:cramer-rao} is diagonal, joint estimators of $(\rho_A, \rho_B, t_0)$ may exist for which the error in the estimation of the individual parameters is uncoupled up to second order.
+If the detector's noise is Gaussian, then the outputs of matched filters for $a(t)$ and $b(t)$ will themselves be Gaussian.
+This encourages us to make the approximation that the errors in $(\rho_A, \rho_B, t_0)$ by a matched\nobreakdashes-filter estimator are independent and Gaussian with variance equal to the \ac{CRLB}.
+A wrinkle is that $\mathcal{I}_{t_0,t_0}$ depends on the true values of $\rho_A$ and $\rho_B$.
+If we make the further leap of evaluating the \ac{CRLB} \emph{at the value of the estimators}, we can write
 %
 \begin{eqnarray}
-	\mathrm{var}\, (\hat{\rho}_B) &\geq {\mathcal{I}_{\rho_A,\rho_A}}^{-1} = 1 \\
-	\mathrm{var}\, (\hat{\rho}_B) &\geq {\mathcal{I}_{\rho_B,\rho_B}}^{-1} = 1 \\
-	\mathrm{var}\, (\hat{t}_0) &\geq {\mathcal{I}_{t_0,t_0}}^{-1} = (2 \pi t_0 f_\mathrm{rms})^{-2}.
+\hat{\rho}_A - \rho_A \sim \mathcal{N}\, [0, 1] \\
+\hat{\rho}_B - \rho_B \sim \mathcal{N}\, [0, 1] \\
+\hat{t}_0 - t_0 \sim \mathcal{N} \left[0, \frac{1}{2 \pi f_\mathrm{rms} \sqrt{{\hat{\rho}_A}^2 + {\hat{\rho}_B}^2}}\right].
 \end{eqnarray}

### 2012-05-01 ms.tex
Preliminary abstract

@@
 \begin{abstract}
-Description of rapid Bayesian sky localization algorithm
+Timely \acl{EM} follow\nobreakdashes-up of \acl{CBC} events detected by \acl{aLIGO} requires rapidly inferring the sky location from the \ac{GW} observations.
+Calculation of the posterior distribution of the sky location and intrinsic source parameters given the \ac{GW} strain takes hours with state\nobreakdashes-of\nobreakdashes-the\nobreakdashes-art \ac{MCMC} parameter estimation codes.
+By taking as the measurement not the \ac{GW} strain itself but the amplitude and \acl{TOA} of the putative signal at each detector, and by fixing the intrinsic source parameters, we have constructed a non-\ac{MCMC}, fully deterministic Bayesian parameter estimation algorithm that takes just seconds to produce sky maps of posterior probability.%
+\marginpar{Wrong emphasis?  I like Larry's idea of having this paper discuss several different tiers of sky localization with increasing accuracy and latency.}
 \end{abstract}

### 2012-05-01 ms.tex
bayestar-localization: added paper outline

@@
 \end{abstract}
+
+\section{Outline}
+
+\begin{enumerate}
+\item Introduction {
+	\begin{enumerate}
+	\item Background: \acp{GRB}, \acp{CBC}, \acs{LIGO}
+	\item Time scales of \acp{GRB}, \acp{CBC}
+	\item Current sky localization algorithms and response times
+	\end{enumerate}
+}
+\item Preliminaries {
+	\begin{enumerate}
+	\item Fisher information, \ac{CRLB}
+	\item Matched\nobreakdashes-filter estimator
+	\end{enumerate}
+}
+\item Algorithm {
+	\begin{enumerate}
+	\item Assumptions
+	\item Signal model
+	\item Likelihood function
+	\item Priors and marginalization
+	\item Procedure
+	\end{enumerate}
+}
+\item Results {
+	\begin{enumerate}
+	\item Injection population
+	\item Cumulative fraction vs. confidence level plots
+	\item Sky localization accuracy as a function of distance or \ac{SNR}
+	\item Speed
+	\end{enumerate}
+}
+\item Discussion {
+	\begin{enumerate}
+	\item Potential further refinements
+	\item What astrophysical targets were accessible before
+	\item What new targets are available with the response time of seconds
+	\item Lay out timeline of follow\nobreakdashes-up campaign, including different tiers of inference and different types of telescopes
+	\end{enumerate}
+}
+\item Appendices {
+	\begin{enumerate}
+	\item Code listing?
+	\item Any proofs that are needed in the body but would interrupt the text?
+	\end{enumerate}
+}
+\end{enumerate}
 

### 2012-05-22 ms.tex
bayestar-localization: TeXified some equations that are in my lab notebook
bayestar-localization: fixed error in Fisher matrix

@@
 \begin{eqnarray*}
-	\frac{\partial^2 \log \mathcal{L}}{\partial A \partial B} &=
-	\frac{\partial^2 \log \mathcal{L}}{\partial A \partial t_0} =
-	\frac{\partial^2 \log \mathcal{L}}{\partial B \partial t_0} = 0 \\
+	\frac{\partial^2 \log \mathcal{L}}{\partial A \partial B} &= 0 \\
+	\frac{\partial^2 \log \mathcal{L}}{\partial A \partial t_0} &=
+	2 \pi i \int \frac{\left(\tilde{a}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} - \tilde{a}(f) \tilde{x}^*(f) e^{-2 \pi i f t_0}\right)}{S(f)} f \mathrm{d}f \\
+	\frac{\partial^2 \log \mathcal{L}}{\partial B \partial t_0} &=
+	2 \pi i \int \frac{\left(\tilde{b}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} - \tilde{b}(f) \tilde{x}^*(f) e^{-2 \pi i f t_0}\right)}{S(f)} f \mathrm{d}f \\
 	\frac{\partial^2 \log \mathcal{L}}{\partial A^2} &=
@@
 %
-The first two terms of $\mathcal{I}_{t_0,t_0}$ vanish because the integrands $|\tilde{a}(f)|^2 f / S(f)$ and $|\tilde{b}(f)|^2 f / S(f)$ are odd about $f = 0$.
+$\mathcal{I}_{A,t_0}$, $\mathcal{I}_{B,t_0}$, and the first two terms of $\mathcal{I}_{t_0,t_0}$ vanish because the integrands $|\tilde{a}(f)|^2 f / S(f)$ and $|\tilde{b}(f)|^2 f / S(f)$ are odd about $f = 0$.
 We assume that $\tilde{a}(f)$ and $\tilde{b}(f)$ are orthogonal with respect to $\emph{any}$ real weight, including both $S(f)$ and $S(f) / f^2$, so the last term in $\mathcal{I}_{t_0,t_0}$ vanishes as well.
@@
 
+\section{Equations that need to be organized}
+
+\begin{equation}
+	u = \cos \iota
+\end{equation}
+
+\begin{equation}
+	\mathbf{G} =
+	\left(
+		\begin{array}{cccc}
+			F_+^1 D_1^1 & F_+^2 D_1^2 & \cdots & F_+^N D_1^N \\
+			F_\times^1 D_1^1 & F_\times^2 D_1^2 & \cdots & F_\times^N D_1^N
+		\end{array}
+	\right)
+\end{equation}
+
+\begin{equation}
+	\mathbf{H} = \mathbf{G}\mathbf{G}^\mathsf{T} =
+	\left(
+		\begin{array}{cc}
+			\sum \left(F_+^i D_1^i\right)^2 & \sum \left(F_+^i D_1^i\right)\left(F_\times^i D_1^i\right) \\
+			\sum \left(F_+^i D_1^i\right)\left(F_\times^i D_1^i\right) & \sum \left(F_\times^i D_1^i\right)^2
+		\end{array}
+	\right)
+\end{equation}
+
+\begin{equation}
+	\mathbf{R} =
+	\left(
+		\begin{array}{cc}
+			\cos \phi & \sin \phi \\
+			-\sin \phi & \cos \phi
+		\end{array}
+	\right)
+	\left(
+		\begin{array}{cc}
+			\frac{1}{2} (1 + u^2) & 0 \\
+			0 & u
+		\end{array}
+	\right)
+	\left(
+		\begin{array}{cc}
+			\cos \psi & \sin \psi \\
+			-\sin \psi & \cos \psi
+		\end{array}
+	\right) \mathbf{G}
+\end{equation}
+
+\begin{equation}
+	p = \Tr \mathbf{R}^\mathsf{T} \mathbf{R}
+	= \frac{1}{8} \left\{
+		\left(u^4 + 6 u^2 + 1\right) \left(H_{11} + H_{22}\right) + (1 - u^2)^2
+		\left[ \left( H_{11} - H_{22} \right) \cos 2\psi + 2 H_{12} \sin 2\psi \right]
+	\right\}
+\end{equation}
+
+\begin{equation}\fl
+	q = \sqrt{(ab - cd) (u + u^3) + \frac{1}{8} \left\{ (a^2 + b^2 + c^2 + d^2)(u^4 + 6u^2 + 1) + (1 - u^2)^2 \left[ (a^2 - b^2 - c^2 + d^2) \cos 2\psi + 2(ac + bd) \sin 2\psi \right] \right\}}
+\end{equation}
+
 \ack Source code for \ac{BAYESTAR} is available on the \acs{LIGO} \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/bayestar.html}.

### 2012-07-31 ms.tex
First update of ms.tex

@@
 \end{enumerate}
+
+\section{New section, add a new name to this}
+
+A CBC will emit a GW.
+What we detect will be the data, $x(t)$ which is a function of time.
+This data will be a combination of the signal emitted from the CBC< $h(t)$, and the noise our detector picks up in addition to the signal, $n(t)$.
+We assume we have additive, zero mean, wide source stationary, Gaussian measurement noise $n(t)$.
+This means that noise is only added to what passes through the channel, the mean of this noise is zero, the noise does not vary in time, and the noise is drawn from a Gaussian distribution.
+
+Our goal here is to first determine the likelihood.
+
+Our detector data is given by $x(t)$.
+This can be described in the frequency domain as $\tilde{x}(f) = \tilde{h}(f) + \tilde{n}(f)$ where $\tilde{h}(f)$ is our signal and $\tilde{n}(f)$ is the noise.
+We have noise, $E[n(t_1)n(t_2)]$ which depends only on $t_1 - t_2$.
+The Fourier transform of the noise gives us this: $E[\tilde{n}(f)^*\tilde{n}(f)] = \delta (f_1 - f_2) s(f_1)$ where $s(f)$ is the power spectral density for the noise.
+The power spectral density gives us information about the energy of the present frequencies.
+The delta function is only non-zero when the input is zero, so in this case we will only get a nonzero value for $\delta (f_1 - f_2) s(f_1)$ when $f_1 = f_2$.
+
+To determine the power spectral density (PSD) for the real signal $y(t)$ we first take the Fourier transform of our real signal to get:
+
+\begin{equation}
+	\tilde{y}(f) = \int_{-\infty}^{\infty}y(t)e^{-2\pi ift}dt
+\end{equation}
+
+Then we get the two sided PSD from the definition of PSD.
+
+\begin{equation}
+	R_y(f) = |\tilde{y}(f)|^2
+\end{equation}
+
+Then we get the single sided PSD with a conversion,
+
+\begin{equation}
+	S_y(f) = |\tilde{y}(f)|^2 + |\tilde{y}(-f)|^2
+\end{equation}
+
+We can decompose our signal into two orthogonal bases with
+
+\begin{equation}
+	\tilde{h}(f) = a\tilde{h}_0(f) + b\tilde{h}_{\pi/2}(f)
+\end{equation}
+
+Defining
+\begin{eqnarray}
+	\tilde{h}_0(f) & = & -i \tilde{h}_{\pi/2}(f) \mbox{ for } f > 0 \\
+	\tilde{h}_0(-f) & = & i \tilde{h}_{\pi/2}(-f) \mbox{ for } f < 0
+\end{eqnarray}
+
+This ends up resulting in
+\begin{equation}
+	\tilde{h}(f) = (a + bi)\tilde{h}_0(f)
+\end{equation}
+
+Next we'll set $\tilde{u}(f) = \tilde{h}_0(f)$.
+
+We can rewrite our signal in polar form to get the following.
+\begin{eqnarray}
+	\tilde{h}(f) & = & (a + bi) \tilde{u}(f) e^{-2\pi ift_0} \\
+	& = & Ce^{i \phi}\tilde{u}(f)e^{-2\pi ift_0} \\
+	& = & Ce^{i(\phi - 2\pi ft_0)}\tilde{u}(f)
+\end{eqnarray}
+
+This gives us a definition of $\tilde{h}(f)$ to be used in our likelihood and allows us to update our definition of $\tilde{x}(f)$
+
+\begin{eqnarray}
+	\tilde{x}(f) & = & Ce^{i(\phi - 2\pi ft_0)}\tilde{u}(f) + \tilde{n}(f) \mbox{ for } f > 0 \\
+	\tilde{x}(f) & = & \tilde{x}^*(f) \mbox{ for } f < 0
+\end{eqnarray}
+
+We can now obtain our likelihood function from the definition of a univariate gaussian for a continuous signal ${\bf{x}}(f)$,
+\begin{equation}
+	f({\bf{x}}) = \frac{1}{\sqrt{\det(2\pi \bf{\Sigma})}}\exp(-\frac{1}{2}(\bf{x}-\bf{u})^T\Sigma^{-1}(\bf{x}-\bf{\mu}))
+\end{equation}
+Expressing this in frequency space,
+\begin{equation}
+	f({\bf{\tilde{x}}}(f)) \propto \exp[-\frac{1}{2}\int\int( {\bf{x}}(f)-{\bf{\mu}}(f))^*C(f,f')(\tilde{x}(f')-\mu(f'))]dfdf'
+\end{equation}
+Where $C(f,f')$ represents the covariance. Now we can determine the log of our likelihood function.
+
+\begin{equation}
+	\ln(L(\tilde{x}; C, \phi, t_0)) \sim -\frac{1}{2}\int_0 ^\infty \frac{|\tilde{x}(f) - \tilde{h}(f; C, \phi, t_0)|^2}{S_n(f)}df
+\end{equation}
+
+There will be a time delay between the time of coalescence and the time the signal arrives at our detector.
+We'll see what this means we should expect by introducing a time delay in frequency space.
+Our real signal data is given by $y(t)$, so we can introduce a delay by changing the time, $y(t - t_0) = z(t)$ where $t$ is the time of arrival and $t_0$ the time of coalescence.
+So $z(t)$ now contains our time delay.
+We take the Fourier transform of $z(t)$ and are left with
+
+\begin{equation}
+	\tilde{z}(f) = \int^{\infty}_{-\infty}y(t-t_0)e^{-2\pi ift} dt
+\end{equation}
+
+Introducing a change of variables from $t$ to $t' = t - t_0$ gives us
+\begin{eqnarray}
+	\tilde{z}(f) & = & e^{-2\pi ift_0} \int_{-\infty}^{\infty}y(t')e^{-2\pi ift'} dt' \\
+	& = & e^{-2\pi ift_0} \tilde{y}(f)
+\end{eqnarray}
+
+This means that a time delay introduced a phase shift in frequency.
+
+Now we can develop the Fisher information matrix. This will tell us how our variables affect each other. The Fisher information matrix is defined as 
+\begin{equation}
+\mathcal{I}_{ij} \equiv E[-\frac{\partial^2 \ln L}{\partial \theta_i \partial \theta_j}]
+\end{equation}
+ for $L(\theta_1, \theta_2, ...)$.
+
+So if we take our likelihood we'll end up with this,
+\begin{eqnarray}
+	\frac{\partial \ln L}{\partial \theta_i} & = & -\frac{1}{2} \int_0 ^\infty \frac{2}{s_n(f)}Re\{[\frac{\partial}{\partial \theta_i} (\tilde{x} - \tilde{h})][\tilde{x} - \tilde{h}]\}df \\
+	& = & \int_0 ^\infty Re[(\frac{\partial \tilde{h}}{\partial \theta_i})^*(\tilde{x} - \tilde{h})]\frac{df}{s(f)} \\
+	\frac{\partial^2 \ln L}{\partial \theta_i \partial \theta_j} & = & \int_0 ^\infty Re[(\frac{\partial^2 \tilde{h}}{\partial \theta_i \partial \theta_j})^*(\tilde{x} - \tilde{h})\frac{df}{s(f)} + \int_0 ^\infty Re[(\frac{\partial \tilde{h}}{\partial \theta_i})^*\frac{\partial \tilde{h}}{\partial \theta_j}]\frac{df}{s(f)}] \\
+\end{eqnarray}
+
+Apply this to the definition of the Fisher information matrix,
+\begin{eqnarray}
+	\mathcal{I}_{ij} & = & E[-\int_0^\infty (\frac{\partial ^2 \tilde{h}}{\partial \theta_i \partial \theta_j})^*(\tilde{x} - \tilde{h})\frac{df}{s(f)} + \int_0 ^ \infty Re[(\frac{\partial \tilde{h}}{\partial \theta_i})^*\frac{\partial \tilde{h}}{\partial \theta _j}]\frac{df}{s(f)}] \\
+	& = & \int^\infty _0 Re[(\frac{\partial \tilde{h}}{\partial \theta_i})^* \frac{\partial \tilde{h}}{\partial \theta _j}]\frac{df}{s(f)}
+\end{eqnarray}
+
+Putting this into a matrix and defining $M_k = \int_0^\infty \tilde{u}(f)^* u(f) f\frac{df}{s(f)}$ and $N_k = \frac{M_k}{M_0}$ gives us
+\[
+\mathcal{I} =
+\left( {\begin{array}{ccc}
+ M_0 & 0 & 0 \\
+ 0 & C^2M_0 & -2\pi C^2M_1 \\
+ 0 & -2\pi C^2M_1 & 4\pi^2C^2M_2
+ \end{array} } \right)
+\]
+
+Now we will perform three changes of variables to diagonalize this matrix.
+
+The first change of variables is from $C$ to $\rho = \sqrt{M_0}C$. 
+\[
+\mathcal{I} =
+\left( {\begin{array}{ccc}
+ 1 & 0 & 0 \\
+ 0 & \rho^2 & -2\pi\rho^2 N_1 \\
+ 0 & -2\pi\rho^2 N_1 & 4\pi^2 \rho^2 N_2
+ \end{array} } \right)
+\]
+
+The second from $t$ to $\phi_t = t2\pi \sqrt{N_2}$ gives us
+\[
+\mathcal{I} =
+\left( {\begin{array}{ccc}
+ 1 & 0 & 0 \\
+ 0 & \rho^2 & -\rho^2\frac{N_1}{\sqrt{N_2}} \\
+ 0 & -\rho^2\frac{N_1}{\sqrt{N_2}} & \rho^2
+ \end{array} } \right)
+\]
+
+The third introduces two new variables, $\phi_s = \frac{1}{\sqrt{2}}(\phi + \phi_t)$ and $\phi_d = \frac{1}{\sqrt{2}}(\phi - \phi_t)$. This results in this matrix,
+\[
+\mathcal{I} =
+\left( {\begin{array}{ccc}
+ 1 & 0 & 0 \\
+ 0 & (1 - \frac{N_1}{\sqrt{N_2}})\rho^2 & 0 \\
+ 0 & 0 & (1 + \frac{N_1}{\sqrt{N_2}})\rho^2
+ \end{array} } \right)
+\]
+
+Now we'll do a final change of variables to make the phases have the same variance.
+We let $\alpha_s = \phi_s \sqrt{1 - \frac{N_1}{\sqrt{N_2}}}$ and $\alpha_d = \phi_d \sqrt{1 - \frac{N_1}{\sqrt{N_2}}}$ which gives us the matrix
+\[
+\mathcal{I} =
+\left( {\begin{array}{ccc}
+ 1 & 0 & 0 \\
+ 0 & \rho^2 & 0 \\
+ 0 & 0 & \rho^2
+ \end{array} } \right)
+\]
 

### 2012-07-31 ms.tex
second attempt at first draft

@@
 
-A CBC will emit a GW.
+We expect to detect GW from CBCs.
 What we detect will be the data, $x(t)$ which is a function of time.
-This data will be a combination of the signal emitted from the CBC< $h(t)$, and the noise our detector picks up in addition to the signal, $n(t)$.
-We assume we have additive, zero mean, wide source stationary, Gaussian measurement noise $n(t)$.
-This means that noise is only added to what passes through the channel, the mean of this noise is zero, the noise does not vary in time, and the noise is drawn from a Gaussian distribution.
+This data will be a combination of the signal emitted from the CBC,  $h(t)$, and the noise our detector picks up in addition to the signal, $n(t)$.
+We assume we have additive, zero mean, wide-sense stationary, Gaussian measurement noise $n(t)$.
+This means that noise only consists of something being added to what passes through the channel, the mean is zero, the noise only depends on differences in time, and it is drawn from a Gaussian distribution.
 
 Our goal here is to first determine the likelihood.
-
-Our detector data is given by $x(t)$.
+We will consider the data from our detector, given by $x(t)$.
 This can be described in the frequency domain as $\tilde{x}(f) = \tilde{h}(f) + \tilde{n}(f)$ where $\tilde{h}(f)$ is our signal and $\tilde{n}(f)$ is the noise.
-We have noise, $E[n(t_1)n(t_2)]$ which depends only on $t_1 - t_2$.
-The Fourier transform of the noise gives us this: $E[\tilde{n}(f)^*\tilde{n}(f)] = \delta (f_1 - f_2) s(f_1)$ where $s(f)$ is the power spectral density for the noise.
+Since our noise is wide-sense stationary, $E[n(t_1)n(t_2)]$ depends only on $t_1 - t_2$.
+The Fourier transform gives us: $E[\tilde{n}(f)^*\tilde{n}(f)] = \delta (f_1 - f_2) s(f_1)$ where $s(f)$ is the power spectral density for the noise.
 The power spectral density gives us information about the energy of the present frequencies.
@@
 
-Then we get the single sided PSD with a conversion,
+Then we get the single sided PSD by multiplying by 2. We can rewrite this in the following way since we are taking the Fourier transform of a real function.
 
@@
 
-We can decompose our signal into two orthogonal bases with
+This gives us an equation for our PSD in terms of real signal, which is the signal we measure. We can decompose our signal into two orthogonal bases with
 
@@
 
-This ends up resulting in
+Plugging our new definitions into our definition for $\tilde{h}(f)$ and factoring gives us:
 \begin{equation}
@@
 
-Next we'll set $\tilde{u}(f) = \tilde{h}_0(f)$.
+Now we'll set $\tilde{u}(f) = \tilde{h}_0(f)$.
 
@@
 This gives us a definition of $\tilde{h}(f)$ to be used in our likelihood and allows us to update our definition of $\tilde{x}(f)$
-
 \begin{eqnarray}
@@
 
-Now we can develop the Fisher information matrix. This will tell us how our variables affect each other. The Fisher information matrix is defined as 
+Now we can develop the Fisher information matrix. This will tell us how our variables affect each other and provide information of their covariance. The Fisher information matrix is defined as 
 \begin{equation}
@@
 \end{equation}
- for $L(\theta_1, \theta_2, ...)$.
+ for likelihood $L(\theta_1, \theta_2, ...)$.
 
@@
 
-Putting this into a matrix and defining $M_k = \int_0^\infty \tilde{u}(f)^* u(f) f\frac{df}{s(f)}$ and $N_k = \frac{M_k}{M_0}$ gives us
+So we only need the real part of the product of our derivatives.
+Putting this into a matrix and defining $M_k = \int_0^\infty \tilde{u}(f)^* u(f) f^k\frac{df}{s(f)}$ and $N_k = \frac{M_k}{M_0}$ gives us
 \[
@@
 
-Now we will perform three changes of variables to diagonalize this matrix.
-
+Now we will perform four changes of variables.
 The first change of variables is from $C$ to $\rho = \sqrt{M_0}C$.
+This change is intended to make the variance of the absolute value of the amplitude equal to one.
 \[
@@
 
-The second from $t$ to $\phi_t = t2\pi \sqrt{N_2}$ gives us
+The second from $t$ to $\phi_t = t2\pi \sqrt{N_2}$ is meant to get rid of $t_0$,
 \[
@@
 
-The third introduces two new variables, $\phi_s = \frac{1}{\sqrt{2}}(\phi + \phi_t)$ and $\phi_d = \frac{1}{\sqrt{2}}(\phi - \phi_t)$. This results in this matrix,
+The third introduces two new variables, $\phi_s = \frac{1}{\sqrt{2}}(\phi + \phi_t)$ and $\phi_d = \frac{1}{\sqrt{2}}(\phi - \phi_t)$. This change of variables diagonalizes the matrix.
 \[
@@
 
-Now we'll do a final change of variables to make the phases have the same variance.
+Now we'll do a final change of variables to make the phase parameters have the same variance.
 We let $\alpha_s = \phi_s \sqrt{1 - \frac{N_1}{\sqrt{N_2}}}$ and $\alpha_d = \phi_d \sqrt{1 - \frac{N_1}{\sqrt{N_2}}}$ which gives us the matrix

### 2013-02-08 ms.tex
Adjust paths in Makefile and ms.tex after GitHub move
Merged two duplicate derivations of CRB

--- 2012-05-22 ms.tex
@@
 \usepackage{iopams}
-\usepackage{aas_macros}
+\usepackage{bib/aas_macros}
 \usepackage[colorlinks]{hyperref}
-\input{acronyms}
+\input{ligo-acronyms/acronyms}
 
@@
 \cite{fairhurst:2009} estimated the uncertainty in the matched-filter measurement of the \ac{TOA} of a \ac{GW} signal by computing the scalar Fisher information.
-We extend this result by computing the Fisher information matrix describing the uncertainties in the joint estimation of the signal amplitude and \ac{TOA}.%
+We extend this result by computing the Fisher information matrix describing the uncertainties in the joint estimation of the signal amplitude and \ac{TOA}.
+
+When a \ac{CBC} occurs, a detector's strain observation may be described by a linear combination of two orthogonal \ac{GW} quadratures and additive, zero\nobreakdashes-mean, \acl{WSS}, Gaussian measurement noise $n(t)$, with one-sided power spectral density $S(f)$.
+In the frequency domain, the detector's observation $\tilde x (\omega)$ is given by
 %
-\marginpar{This might be too much mathematical detail.
-All we need to do is show that the $\rho_A\rho_A$ and $\rho_B\rho_B$ terms are unity and that the off-diagonal terms vanish.
-If possible, we should cite and use Fairhurst's expression for $\mathcal{I}_{t_0,t_0}$ without re-deriving it.}
-
-When a \ac{CBC} occurs, a detector's strain observation may be described by a linear combination of two orthogonal \ac{GW} quadratures, $a(t)$ and $b(t)$, and additive, zero\nobreakdashes-mean, \acl{WSS}, Gaussian measurement noise $n(t)$, with power spectral density $S(f)$.
-Introducing the unknown, real, quadrature amplitudes, $A$ and $B$, and the unknown \ac{TOA}, $t_0$, we can write
-%
-\begin{equation}\label{eq:xoft}
-	x(t) = A a(t - t_0) + B b(t - t_0) + n(t).
+\begin{equation}\label{eq:signal-model}
+	\tilde x (\omega) = \tilde h (\omega, \rho, \gamma, \tau) + \tilde n (\omega) = \rho e^{\imath (\gamma - \omega \tau)} \tilde u (\omega) + \tilde n (\omega).
 \end{equation}
 %
-Transformation of \Eref{eq:xoft} to the frequency domain converts the time delay by $t_0$ to a complex phase,
+where $\tilde h (\omega)$ is the signal content, $\tilde n (\omega)$ is the detector's noise, $\tilde u (\omega)$ is the frequency-domain post-Newtonian inspiral signal, $\rho$ is the signal's amplitude, $\gamma$ is a phase angle the encodes the relative strength of the two \ac{GW} quadratures, and $\tau$ is the time of arrival relative to an arbitrary time base.
+
+The Fisher information matrix for a measurement $\tilde x$ described by the unknown parameter vector $\boldsymbol{\theta}$ may be written as
 %
-\begin{equation}\label{eq:xoff}
-	\tilde{x}(f) = \left(A \tilde{a}(f) + B \tilde{b}(f)\right)e^{-2 \pi i f t_0} + \tilde{n}(f).
-\end{equation}
-%
-The Fisher information matrix for a measurement $\mathbf{x}$ described by the unknown parameters $\boldsymbol{\theta}$ may be written as
-%
-\begin{equation}
+\begin{equation}\label{eq:general-fisher-matrix}
 	\mathcal{I}_{jk} = \mathrm{E} \, \left[
 		-\frac{\partial^2 \log
-			\mathcal{L}(\mathbf{x}; \boldsymbol{\theta})}
+			\mathcal{L}(\tilde x ; \boldsymbol\theta)}
 			{\partial \theta_j \theta_k}
@@
 %
-First, we write the likelihood of the data $\tilde{x}(f)$ given the parameters $(A, B, t_0$):
+This equation involves double derivatives, but substantial simplification is possible when the likelihood is Gaussian.
+First, we write the likelihood of the data $\tilde x (\omega)$ given the parameter vector $\boldsymbol\theta = (\rho, \gamma, \tau)$:
 %
-\begin{equation}\label{eq:trigger-likelihood}
-	\fl\mathcal{L}(x; A, B, t_0) = p(x | A, B, t_0)
+\begin{equation}\label{eq:gaussian-likelihood}
+	\mathcal{L}(\tilde x; \boldsymbol\theta) = p(\tilde x | \boldsymbol\theta)
 		\propto \exp \left[
-		- \frac{1}{2} \int_{-\infty}^\infty \frac{\left|\tilde{x}(f)
-			- \left(A \tilde{a}(f)
-			- B \tilde{b}(f)\right) e^{-2 \pi i f t_0}\right|^2}{S(f)} \, \mathrm{d}f
+		- \frac{1}{2} \int_0^\infty \frac{\left|\tilde x (\omega)
+			- \tilde h (\omega; \boldsymbol\theta) \right|^2}{S(\omega)} \, \mathrm{d}\omega
 	\right].
@@
 %
-The normalization has been dropped because the Fisher matrix elements depend only on derivatives of the log\nobreakdashes-likelihood function and any factor that is independent of the parameters makes no contribution.%
+The normalization has been dropped because the Fisher matrix elements depend only on derivatives of the log\nobreakdashes-likelihood function and any factor that is independent of the parameters makes no contribution.
+Now, substituting \eref{eq:gaussian-likelihood} into \eref{eq:general-fisher-matrix} gives us
 %
-\marginpar{Consider using discrete time to avoid the fiction of an infinite-dimensional multivariate Gaussian process.}
-%
-From hereon, unless otherwise specified the limits of integration are always $(-\infty, \infty)$ and will be dropped.
-Expanding the integrand, we can write the log\nobreakdashes-likelihood function up to an additive constant as
-%
-\begin{eqnarray}\label{eq:trigger-likelihood}
-	\log \mathcal{L}(x; A, B, t_0) \sim
-	&- \frac{1}{2} \int \frac{|\tilde{x}(f)|^2}{S(f)} \mathrm{d}f \\
-	&+ A \, \mathrm{Re} \int \frac{\tilde{a}^*(f) \tilde{x}(f) e^{2 \pi i f t_0}}{S(f)} \mathrm{d}f \nonumber\\
-	&+ B \, \mathrm{Re} \int \frac{\tilde{b}^*(f) \tilde{x}(f) e^{2 \pi i f t_0}}{S(f)} \mathrm{d}f\nonumber\\
-	&+ A B \, \mathrm{Re} \int \frac{\tilde{a}^*(f) \tilde{b}(f)}{S(f)} \mathrm{d}f \nonumber\\
-	&- \frac{1}{2} A^2 \int \frac{|\tilde{a}(f)|^2}{S(f)} \mathrm{d}f
-	- \frac{1}{2} B^2 \int \frac{|\tilde{b}(f)|^2}{S(f)} \mathrm{d}f.
-\end{eqnarray}
-%
-We assume that $\tilde{a}(f)$ and $\tilde{b}(f)$ are orthogonal and have the same norm,
-%
-\begin{equation}
-	\int \frac{\tilde{a}^*(f) \tilde{b}(f)}{S(f)} \mathrm{d}f = 0,
-	\int \frac{|\tilde{a}(f)|^2}{S(f)} \mathrm{d}f =
-	\int \frac{|\tilde{b}(f)|^2}{S(f)} \mathrm{d}f =
-	\sigma^2.
+\begin{equation}\label{eq:gaussian-fisher-matrix}
+	\mathcal{I}_{jk} = \int_0^\infty \Re \left[
+        \left(\frac{\partial \tilde h}{\partial \theta_j}\right)^*
+        \left(\frac{\partial \tilde h}{\partial \theta_k}\right)
+	\right] \frac{1}{S(\omega)} \, \mathrm{d}\omega.
 \end{equation}
 %
-After we apply these substitutions, and drop the $|\tilde{x}(f)|^2$ term that is independent of $A$, $B$, or $t_0$, we can write the log\nobreakdashes-likelihood function as
-%
-\begin{eqnarray}\label{eq:trigger-likelihood-simplified}
-	\log \mathcal{L}(x; A, B, t_0) \sim
-	& A \, \mathrm{Re} \int \frac{\tilde{a}^*(f) \tilde{x}(f) e^{2 \pi i f t_0}}{S(f)} \mathrm{d}f 
-	&+ B \, \mathrm{Re} \int \frac{\tilde{b}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} }{S(f)} \mathrm{d}f\nonumber\\
-	&- \frac{1}{2} A^2 \sigma^2 - \frac{1}{2} B^2 \sigma^2.
-\end{eqnarray}
-%
-The necessary partial derivatives are
+where $\Re$ denotes the real part and $^*$ the complex conjugate.
+This form is more useful because it involves only first derivatives, and of the signal $\tilde h (\omega)$ rather than the entire observation $\tilde x (\omega)$.
+
+With \eref{eq:gaussian-fisher-matrix} in hand, we return to our particular signal model, \Eref{eq:signal-model}. The necessary partial derivatives are
 %
 \begin{eqnarray*}
-	\frac{\partial^2 \log \mathcal{L}}{\partial A \partial B} &= 0 \\
-	\frac{\partial^2 \log \mathcal{L}}{\partial A \partial t_0} &=
-	2 \pi i \int \frac{\left(\tilde{a}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} - \tilde{a}(f) \tilde{x}^*(f) e^{-2 \pi i f t_0}\right)}{S(f)} f \mathrm{d}f \\
-	\frac{\partial^2 \log \mathcal{L}}{\partial B \partial t_0} &=
-	2 \pi i \int \frac{\left(\tilde{b}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} - \tilde{b}(f) \tilde{x}^*(f) e^{-2 \pi i f t_0}\right)}{S(f)} f \mathrm{d}f \\
-	\frac{\partial^2 \log \mathcal{L}}{\partial A^2} &=
-	\frac{\partial^2 \log \mathcal{L}}{\partial B^2} = - \sigma^2 \\
-	\frac{\partial^2 \log \mathcal{L}}{\partial {t_0}^2} &=
-	2 \pi i A \int \frac{\left(\tilde{a}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} - \tilde{a}(f) \tilde{x}^*(f) e^{-2 \pi i f t_0}\right)}{S(f)} f \mathrm{d}f \\
-	&+ 2 \pi i B \int \frac{\left(\tilde{b}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} - \tilde{b}(f) \tilde{x}^*(f) e^{-2 \pi i f t_0}\right)}{S(f)} f \mathrm{d}f \\
-	&- (2 \pi)^2 A \int \frac{\left(\tilde{a}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} + \tilde{a}(f) \tilde{x}^*(f) e^{-2 \pi i f t_0}\right)}{S(f)} f^2 \mathrm{d}f \\
-	&- (2 \pi)^2 B \int \frac{\left(\tilde{b}^*(f) \tilde{x}(f) e^{2 \pi i f t_0} + \tilde{b}(f) \tilde{x}^*(f) e^{-2 \pi i f t_0}\right)}{S(f)} f^2 \mathrm{d}f.
+    \frac{\partial \tilde h}{\partial \rho} &= e^{\imath (\gamma - \omega \tau)} \tilde u (\omega) \\
+    \frac{\partial \tilde h}{\partial \gamma} &= \imath \, \rho \, e^{\imath (\gamma - \omega \tau)} \tilde u (\omega) \\
+    \frac{\partial \tilde h}{\partial \tau} &= -\imath \, \omega \, \rho \, e^{\imath (\gamma - \omega \tau)} \tilde u (\omega).
 \end{eqnarray*}
 %
-Finally, taking the expectation values of the partial derivatives, we find
+The Fisher matrix elements are
 %
 \begin{eqnarray*}
-	\mathcal{I}_{A,B} &= \mathcal{I}_{A,t_0} = \mathcal{I}_{B,t_0} = 0 \\
-	\mathcal{I}_{A,A} &= \mathcal{I}_{B,B} = \sigma^2 \\
-	\mathcal{I}_{t_0,t_0} &=
-	- 2 \pi i A^2 \int \frac{|\tilde{a}(f)|^2}{S(f)} f\mathrm{d}f
-	- 2 \pi i B^2 \int \frac{|\tilde{B}(f)|^2}{S(f)} f\mathrm{d}f \\
-	&+ (2 \pi)^2 \left(A^2 \int \frac{|\tilde{a}(f)|^2}{S(f)} f^2\mathrm{d}f +
-	B^2 \int \frac{|\tilde{b}(f)|^2}{S(f)} f\mathrm{d}f \right) \\
-	&+ (2\pi)^2 2 A B \, \mathrm{Re} \int \frac{\tilde{a}^*(f)\tilde{b}(f)}{S(f)} f^2\mathrm{d}f.
+    \mathcal{I}_{\rho\rho} &= \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \, \mathrm{d}\omega \\
+	\mathcal{I}_{\gamma\gamma} &= \rho^2 \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \, \mathrm{d}\omega \\
+	\mathcal{I}_{\tau\tau} &= \rho^2 \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \omega^2 \, \mathrm{d}\omega \\
+	\mathcal{I}_{\gamma\tau} &= -\rho^2 \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \omega \, \mathrm{d}\omega = \mathcal{I}_{\tau\gamma} \\
+	\mathcal{I}_{\rho\gamma} &= \mathcal{I}_{\gamma\rho} = \mathcal{I}_{\rho\tau} = \mathcal{I}_{\tau\rho} = 0.
 \end{eqnarray*}
 %
-$\mathcal{I}_{A,t_0}$, $\mathcal{I}_{B,t_0}$, and the first two terms of $\mathcal{I}_{t_0,t_0}$ vanish because the integrands $|\tilde{a}(f)|^2 f / S(f)$ and $|\tilde{b}(f)|^2 f / S(f)$ are odd about $f = 0$.
-We assume that $\tilde{a}(f)$ and $\tilde{b}(f)$ are orthogonal with respect to $\emph{any}$ real weight, including both $S(f)$ and $S(f) / f^2$, so the last term in $\mathcal{I}_{t_0,t_0}$ vanishes as well.
-Further,%
+If the waveform $\tilde u (\omega)$ is normalized such that $\mathcal{I}_{\rho\rho} = 1$, we can introduce the \textit{\ac{SNR}-weighted moments of angular frequency},
 %
-\marginpar{This is an awkward place to make these substitutions.
-Also, they are poorly motivated.
-Here's what this substitution actually does: the matched\nobreakdashes-filter output has an expected value of $A \sigma = \rho_A$.}
-%
-we make the substituions
-%
-\begin{eqnarray*}
-	\rho_A &= A \sigma \\
-	\rho_B &= B \sigma \\
-	f_\mathrm{rms} &= \left[ \int \frac{|\tilde{a}(f)|^2}{S(f)} f^2\mathrm{d}f \right] ^ {1/2} \left[\int \frac{|\tilde{a}(f)|^2}{S(f)} \mathrm{d}f \right] ^ {-1/2} \\
-	&=  \left[ \int \frac{|\tilde{b}(f)|^2}{S(f)} f^2\mathrm{d}f \right] ^ {1/2} \left[\int \frac{|\tilde{b}(f)|^2}{S(f)} \mathrm{d}f \right] ^ {-1/2}.
-\end{eqnarray*}
+\begin{equation}\label{eq:angular-frequency-moments}
+    \overline{\omega^k} =
+        \left[ \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \omega^k \, \mathrm{d}\omega \right]
+        \left[ \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \, \mathrm{d}\omega \right]
+\end{equation}
 %
@@
 \begin{eqnarray*}
-	\mathcal{I}_{\rho_A,\rho_B} &= \mathcal{I}_{\rho_A,t_0} = \mathcal{I}_{\rho_B,t_0} = 0 \\
-	\mathcal{I}_{\rho_A,\rho_A} &= \mathcal{I}_{\rho_B,\rho_B} = 1 \\
-	\mathcal{I}_{t_0,t_0} &= (2 \pi f_\mathrm{rms})^2 ({\rho_A}^2 + {\rho_B}^2).
+	\mathcal{I} = \bordermatrix{
+        ~ & \rho & \gamma & \tau \cr
+        \rho & 1 & 0 & 0 \cr
+        \gamma & 0 & \rho^2 & -\rho^2 \overline{\omega} \cr
+        \tau & 0 & -\rho^2 \overline{\omega} & \rho^2 \overline{\omega^2}
+    }.
 \end{eqnarray*}
-%
-The Cram\'{e}r-Rao inequality says that the inverse of the Fisher information matrix is a lower bound on the covariance of any unbiased estimator.
-The Fisher matrix for this estimation problem has been shown to be diagonal, so we can write
-%
-\begin{eqnarray}\label{eq:cramer-rao}
-	\mathrm{var}\, (\hat{\rho}_B) &\geq {\mathcal{I}_{\rho_A,\rho_A}}^{-1} = 1 \\
-	\mathrm{var}\, (\hat{\rho}_B) &\geq {\mathcal{I}_{\rho_B,\rho_B}}^{-1} = 1 \\
-	\mathrm{var}\, (\hat{t}_0) &\geq {\mathcal{I}_{t_0,t_0}}^{-1} = (2 \pi f_\mathrm{rms} \sqrt{{\rho_A}^2 + {\rho_B}^2})^{-2}.
-\end{eqnarray}
 
@@
 
-\section{Equations that need to be organized}
+Now we will perform four changes of variables.
+The first change of variables is from $C$ to $\rho = \sqrt{M_0}C$.
+This change is intended to make the variance of the absolute value of the amplitude equal to one.
+\[
+\mathcal{I} =
+\left( {\begin{array}{ccc}
+ 1 & 0 & 0 \\
+ 0 & \rho^2 & -2\pi\rho^2 N_1 \\
+ 0 & -2\pi\rho^2 N_1 & 4\pi^2 \rho^2 N_2
+ \end{array} } \right)
+\]
 
-\begin{equation}
-	u = \cos \iota
-\end{equation}
+The second from $t$ to $\phi_t = t2\pi \sqrt{N_2}$ is meant to get rid of $t_0$,
+\[
+\mathcal{I} =
+\left( {\begin{array}{ccc}
+ 1 & 0 & 0 \\
+ 0 & \rho^2 & -\rho^2\frac{N_1}{\sqrt{N_2}} \\
+ 0 & -\rho^2\frac{N_1}{\sqrt{N_2}} & \rho^2
+ \end{array} } \right)
+\]
 
-\begin{equation}
-	\mathbf{G} =
-	\left(
-		\begin{array}{cccc}
-			F_+^1 D_1^1 & F_+^2 D_1^2 & \cdots & F_+^N D_1^N \\
-			F_\times^1 D_1^1 & F_\times^2 D_1^2 & \cdots & F_\times^N D_1^N
-		\end{array}
-	\right)
-\end{equation}
+The third introduces two new variables, $\phi_s = \frac{1}{\sqrt{2}}(\phi + \phi_t)$ and $\phi_d = \frac{1}{\sqrt{2}}(\phi - \phi_t)$. This change of variables diagonalizes the matrix.
+\[
+\mathcal{I} =
+\left( {\begin{array}{ccc}
+ 1 & 0 & 0 \\
+ 0 & (1 - \frac{N_1}{\sqrt{N_2}})\rho^2 & 0 \\
+ 0 & 0 & (1 + \frac{N_1}{\sqrt{N_2}})\rho^2
+ \end{array} } \right)
+\]
 
-\begin{equation}
-	\mathbf{H} = \mathbf{G}\mathbf{G}^\mathsf{T} =
-	\left(
-		\begin{array}{cc}
-			\sum \left(F_+^i D_1^i\right)^2 & \sum \left(F_+^i D_1^i\right)\left(F_\times^i D_1^i\right) \\
-			\sum \left(F_+^i D_1^i\right)\left(F_\times^i D_1^i\right) & \sum \left(F_\times^i D_1^i\right)^2
-		\end{array}
-	\right)
-\end{equation}
-
-\begin{equation}
-	\mathbf{R} =
-	\left(
-		\begin{array}{cc}
-			\cos \phi & \sin \phi \\
-			-\sin \phi & \cos \phi
-		\end{array}
-	\right)
-	\left(
-		\begin{array}{cc}
-			\frac{1}{2} (1 + u^2) & 0 \\
-			0 & u
-		\end{array}
-	\right)
-	\left(
-		\begin{array}{cc}
-			\cos \psi & \sin \psi \\
-			-\sin \psi & \cos \psi
-		\end{array}
-	\right) \mathbf{G}
-\end{equation}
-
-\begin{equation}
-	p = \Tr \mathbf{R}^\mathsf{T} \mathbf{R}
-	= \frac{1}{8} \left\{
-		\left(u^4 + 6 u^2 + 1\right) \left(H_{11} + H_{22}\right) + (1 - u^2)^2
-		\left[ \left( H_{11} - H_{22} \right) \cos 2\psi + 2 H_{12} \sin 2\psi \right]
-	\right\}
-\end{equation}
-
-\begin{equation}\fl
-	q = \sqrt{(ab - cd) (u + u^3) + \frac{1}{8} \left\{ (a^2 + b^2 + c^2 + d^2)(u^4 + 6u^2 + 1) + (1 - u^2)^2 \left[ (a^2 - b^2 - c^2 + d^2) \cos 2\psi + 2(ac + bd) \sin 2\psi \right] \right\}}
-\end{equation}
+Now we'll do a final change of variables to make the phase parameters have the same variance.
+We let $\alpha_s = \phi_s \sqrt{1 - \frac{N_1}{\sqrt{N_2}}}$ and $\alpha_d = \phi_d \sqrt{1 - \frac{N_1}{\sqrt{N_2}}}$ which gives us the matrix
+\[
+\mathcal{I} =
+\left( {\begin{array}{ccc}
+ 1 & 0 & 0 \\
+ 0 & \rho^2 & 0 \\
+ 0 & 0 & \rho^2
+ \end{array} } \right)
+\]
 
@@
 \bibliographystyle{iopart-num}
-\bibliography{apj-jour,telescope}
+\bibliography{apj-jour,bib/telescope}
 

### 2013-02-08 ms.tex
Add margin notes about what we have to do tomorrow

@@
 
+\marginpar{Bad to start a paragraph with a citation number.
+Perhaps switch to natbib and use a \texttt{citet}-style reference.
+Keep this derivation as short as possible.
+Refer more directly to \cite{fairhurst:2009}.
+Tighten up notation as much as possible while still giving a concise description of the Cram\`er-Rao bound.}%
+%
 \cite{fairhurst:2009} estimated the uncertainty in the matched-filter measurement of the \ac{TOA} of a \ac{GW} signal by computing the scalar Fisher information.
@@
 
+\marginpar{This material still has the original, wrong Fishher matrix. Needs to be rewritten.}%
+%
 Because the \ac{CRLB} in \Eref{eq:cramer-rao} is diagonal, joint estimators of $(\rho_A, \rho_B, t_0)$ may exist for which the error in the estimation of the individual parameters is uncoupled up to second order.
@@
 
+\marginpar{This part is correct, but uses the old notation for the frequency moments. Convert to the new notation. Also, add physical interpretation to each change of variables.}%
+%
 Now we will perform four changes of variables.

### 2013-02-08 ms.tex
minor adjustments and comments

@@
 \title[Rapid-response Bayesian sky localization]{Rapid-response Bayesian sky localization for electromagnetic follow-up of gravitational-wave candidates}
-\author{Leo Singer and Larry Price}
+\author{Leo Singer and Larry R. Price}
 \address{LIGO Laboratory, California Institute of Technology, Pasadena, CA 91125, USA}
-\ead{\mailto{leo.singer@ligo.org}, \mailto{larry.price@ligo.org}}
+\ead{\mailto{leo.singer@ligo.org}, \mailto{larryp@caltech.edu}}
 
@@
 
+\marginpar{LP: these derivatives are trivial and so is putting them into fisher.
+Perhaps we should lose the derivatives and re-write Fisher so that the integral appears only once?
+Better yet, write out the equation for the noise moments and then use that.}
 With \eref{eq:gaussian-fisher-matrix} in hand, we return to our particular signal model, \Eref{eq:signal-model}. The necessary partial derivatives are

### 2013-02-12 ms.tex
Remove nitty-gritty Fisher matrix derivation
Removed double-space separating sentences
Tightened up Fisher matrix derivation

--- 2013-02-08 ms.tex
@@
 
-When a \ac{CBC} occurs, a detector's strain observation may be described by a linear combination of two orthogonal \ac{GW} quadratures and additive, zero\nobreakdashes-mean, \acl{WSS}, Gaussian measurement noise $n(t)$, with one-sided power spectral density $S(f)$.
-In the frequency domain, the detector's observation $\tilde x (\omega)$ is given by
+In the frequency domain, the strain $\tilde x (\omega)$ observed by a detector is a linear combination of two orthogonal \ac{GW} quadratures and additive noise,
 %
 \begin{equation}\label{eq:signal-model}
-	\tilde x (\omega) = \tilde h (\omega, \rho, \gamma, \tau) + \tilde n (\omega) = \rho e^{\imath (\gamma - \omega \tau)} \tilde u (\omega) + \tilde n (\omega).
+	\tilde x (\omega) = \tilde h (\omega, \rho, \gamma, \tau) + \tilde n (\omega) = \rho e^{\imath (\gamma - \omega \tau)} \tilde u (\omega) + \tilde n (\omega),
 \end{equation}
 %
-where $\tilde h (\omega)$ is the signal content, $\tilde n (\omega)$ is the detector's noise, $\tilde u (\omega)$ is the frequency-domain post-Newtonian inspiral signal, $\rho$ is the signal's amplitude, $\gamma$ is a phase angle the encodes the relative strength of the two \ac{GW} quadratures, and $\tau$ is the time of arrival relative to an arbitrary time base.
+where $\tilde h (\omega)$ is the signal content, $\tilde n (\omega)$ is the detector's \ac{WSS} Gaussian noise with \ac{PSD} $S(\omega)$, $\tilde u (\omega)$ is the frequency-domain post-Newtonian inspiral signal, $\rho$ is the signal's amplitude, $\gamma$ is a phase angle the encodes the relative strength of two \ac{GW} quadratures, and $\tau$ is the time of arrival relative to an arbitrary time base.
+The likelihood, or the probability of obtaining the observation $\tilde x(\omega)$ conditioned on the value of the parameter vector $\boldsymbol\theta = (\rho, \gamma, \tau)$, is Gaussian, with the log-likelihood proportional to:
+%
+\begin{equation}\label{eq:gaussian-likelihood}
+	\mathcal{L}(\tilde x; \boldsymbol\theta) = p(\tilde x | \boldsymbol\theta)
+		\propto \exp \left[
+		- \frac{1}{2} \int_0^\infty \frac{\left|\tilde x (\omega)
+			- \tilde h (\omega; \boldsymbol\theta) \right|^2}{S(\omega)} \, \mathrm{d}\omega
+	\right].
+\end{equation}
 
-The Fisher information matrix for a measurement $\tilde x$ described by the unknown parameter vector $\boldsymbol{\theta}$ may be written as
+The Fisher information matrix for a measurement $\tilde x$ described by the unknown parameter vector $\boldsymbol{\theta}$ is
 %
@@
 %
-This equation involves double derivatives, but substantial simplification is possible when the likelihood is Gaussian.
-First, we write the likelihood of the data $\tilde x (\omega)$ given the parameter vector $\boldsymbol\theta = (\rho, \gamma, \tau)$:
-%
-\begin{equation}\label{eq:gaussian-likelihood}
-	\mathcal{L}(\tilde x; \boldsymbol\theta) = p(\tilde x | \boldsymbol\theta)
-		\propto \exp \left[
-		- \frac{1}{2} \int_0^\infty \frac{\left|\tilde x (\omega)
-			- \tilde h (\omega; \boldsymbol\theta) \right|^2}{S(\omega)} \, \mathrm{d}\omega
-	\right].
-\end{equation}
-%
-The normalization has been dropped because the Fisher matrix elements depend only on derivatives of the log\nobreakdashes-likelihood function and any factor that is independent of the parameters makes no contribution.
-Now, substituting \eref{eq:gaussian-likelihood} into \eref{eq:general-fisher-matrix} gives us
+This equation involves double derivatives, but substantial simplification is possible when---as in this case---the likelihood is Gaussian:
 %
@@
         \left(\frac{\partial \tilde h}{\partial \theta_k}\right)
-	\right] \frac{1}{S(\omega)} \, \mathrm{d}\omega.
+	\right] \frac{1}{S(\omega)} \, \mathrm{d}\omega,
 \end{equation}
@@
 
-With \eref{eq:gaussian-fisher-matrix} in hand, we return to our particular signal model, \Eref{eq:signal-model}. The necessary partial derivatives are
-%
-\begin{eqnarray*}
-    \frac{\partial \tilde h}{\partial \rho} &= e^{\imath (\gamma - \omega \tau)} \tilde u (\omega) \\
-    \frac{\partial \tilde h}{\partial \gamma} &= \imath \, \rho \, e^{\imath (\gamma - \omega \tau)} \tilde u (\omega) \\
-    \frac{\partial \tilde h}{\partial \tau} &= -\imath \, \omega \, \rho \, e^{\imath (\gamma - \omega \tau)} \tilde u (\omega).
-\end{eqnarray*}
-%
-The Fisher matrix elements are
-%
-\begin{eqnarray*}
-    \mathcal{I}_{\rho\rho} &= \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \, \mathrm{d}\omega \\
-	\mathcal{I}_{\gamma\gamma} &= \rho^2 \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \, \mathrm{d}\omega \\
-	\mathcal{I}_{\tau\tau} &= \rho^2 \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \omega^2 \, \mathrm{d}\omega \\
-	\mathcal{I}_{\gamma\tau} &= -\rho^2 \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \omega \, \mathrm{d}\omega = \mathcal{I}_{\tau\gamma} \\
-	\mathcal{I}_{\rho\gamma} &= \mathcal{I}_{\gamma\rho} = \mathcal{I}_{\rho\tau} = \mathcal{I}_{\tau\rho} = 0.
-\end{eqnarray*}
-%
 If the waveform $\tilde u (\omega)$ is normalized such that $\mathcal{I}_{\rho\rho} = 1$, we can introduce the \textit{\ac{SNR}-weighted moments of angular frequency},

### 2013-02-12 ms.tex
Forgot inverse
Tightened Fisher matrix a little further

@@
 \begin{equation}\label{eq:signal-model}
-	\tilde x (\omega) = \tilde h (\omega, \rho, \gamma, \tau) + \tilde n (\omega) = \rho e^{\imath (\gamma - \omega \tau)} \tilde u (\omega) + \tilde n (\omega),
+	\tilde x (\omega) = \tilde h (\omega, \rho, \gamma, \tau) + \tilde n (\omega) = \frac{\rho}{\sigma} e^{\imath (\gamma - \omega \tau)} \tilde u (\omega) + \tilde n (\omega),
 \end{equation}
@@
 where $\tilde h (\omega)$ is the signal content, $\tilde n (\omega)$ is the detector's \ac{WSS} Gaussian noise with \ac{PSD} $S(\omega)$, $\tilde u (\omega)$ is the frequency-domain post-Newtonian inspiral signal, $\rho$ is the signal's amplitude, $\gamma$ is a phase angle the encodes the relative strength of two \ac{GW} quadratures, and $\tau$ is the time of arrival relative to an arbitrary time base.
+The quantity $\sigma$ is a normalization constant defined as the \ac{SNR} of the \ac{GW} source at a fiducial distance, here taken to be 1~Mpc,
+%
+\begin{equation}\label{eq:sigma}
+    \sigma^2 = \int_0^\infty \frac{\left| \tilde u (\omega)\right|^2}{S(\omega)} \, \mathrm{d}\omega.
+\end{equation}
+%
 The likelihood, or the probability of obtaining the observation $\tilde x(\omega)$ conditioned on the value of the parameter vector $\boldsymbol\theta = (\rho, \gamma, \tau)$, is Gaussian, with the log-likelihood proportional to:
@@
         \left(\frac{\partial \tilde h}{\partial \theta_k}\right)
-	\right] \frac{1}{S(\omega)} \, \mathrm{d}\omega,
+	\right] \frac{1}{S(\omega)} \, \mathrm{d}\omega.
 \end{equation}
 %
-where $\Re$ denotes the real part and $^*$ the complex conjugate.
 This form is more useful because it involves only first derivatives, and of the signal $\tilde h (\omega)$ rather than the entire observation $\tilde x (\omega)$.
-
-If the waveform $\tilde u (\omega)$ is normalized such that $\mathcal{I}_{\rho\rho} = 1$, we can introduce the \textit{\ac{SNR}-weighted moments of angular frequency},
+In terms of the $k$th \ac{SNR}-weighted moment of angular frequency,
 %
@@
         \left[ \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \omega^k \, \mathrm{d}\omega \right]
-        \left[ \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \, \mathrm{d}\omega \right]
+        \left[ \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \, \mathrm{d}\omega \right]^{-1},
 \end{equation}
 %
-Then, we have
+the Fisher matrix for the signal model in \Eref{eq:signal-model} is
 %

### 2013-02-12 ms.tex
New introduction for Fisher matrix section

@@
 
-\marginpar{Bad to start a paragraph with a citation number.
-Perhaps switch to natbib and use a \texttt{citet}-style reference.
-Keep this derivation as short as possible.
-Refer more directly to \cite{fairhurst:2009}.
-Tighten up notation as much as possible while still giving a concise description of the Cram\`er-Rao bound.}%
-%
-\cite{fairhurst:2009} estimated the uncertainty in the matched-filter measurement of the \ac{TOA} of a \ac{GW} signal by computing the scalar Fisher information.
-We extend this result by computing the Fisher information matrix describing the uncertainties in the joint estimation of the signal amplitude and \ac{TOA}.
+The \ac{CRLB} is a useful analytical tool for predicting with what error an unknown parameter or parameters can be estimated from noisy observations.
+When an unbiased estimator exists, the \ac{CRLB} bound says that the covariance of the estimation error is bounded below by the Fisher information matrix.
+The \ac{CRLB} has been widely applied in \ac{GW} astronomy, for instance by \cite{fairhurst:2009} to predict the sky localization accuracy that is achievable with \ac{GW} observations.
+Here, we review the Fisher matrix calculation in \cite{fairhurst:2009} but add some physical interpretation to the particular structure of the covariance matrix.
 

### 2013-02-12 ms.tex
Update Fisher matrix diagonalization section

@@
 \documentclass{iopart}
+\usepackage{amsopn}
 \usepackage{iopams}
@@
 {\raggedright\footnotesize #1}}
+
+\DeclareMathOperator{\cov}{cov}
 
@@
 
+\subsection{Fisher matrix}
+
 In the frequency domain, the strain $\tilde x (\omega)$ observed by a detector is a linear combination of two orthogonal \ac{GW} quadratures and additive noise,
@@
 %
-\begin{eqnarray*}
+\begin{equation}\label{eq:fisher-matrix}
 	\mathcal{I} = \bordermatrix{
@@
     }.
-\end{eqnarray*}
+\end{equation}
 
-\section{Posterior distribution}
+\subsection{\acl{CRLB} and physical interpretation}
+
+The \ac{CRLB} says that, element by element,
+%
+\begin{equation}
+    \cov{
+        \left(
+        \begin{array}{c}
+            \hat{\rho} - \rho \\
+            \hat{\gamma} - \gamma \\
+            \hat{\tau} - \tau
+        \end{array}
+        \right)
+    } \geq \mathcal{I}^{-1}
+\end{equation}
+%
+where a hat $\hat{\phantom{\rho}}$ denotes the estimate of a parameter.
+The Fisher matrix in \Eref{eq:fisher-matrix} is block diagonal, which implies that estimation errors in the signal amplitude $\rho$ are uncorrelated with the phase $\gamma$ and time $\tau$.
+A sequence of two changes of variables lends some physical interpretation to the nature of the coupled estimation errors in $\gamma$ and $\tau$.
+
+First, we put the phase and time on the same footing by measuring the time in units of $1 / 2 \pi \sqrt{\overline{\omega^2}}$ with a change of variables from $\tau$ to $\gamma_\tau = 2 \pi \sqrt{\overline{\omega^2}} \tau$:
+%
+\begin{equation}
+    \mathcal{I}' = \bordermatrix{
+        ~ & \rho & \gamma & \gamma_\tau \cr
+        \rho & 1 & 0 & 0 \cr
+        \gamma & 0 & \rho^2 & -\rho^2\frac{\overline{\omega}}{\sqrt{\overline{\omega^2}}} \cr
+        \gamma_\tau & 0 & -\rho^2\frac{\overline{\omega}}{\sqrt{\overline{\omega^2}}} & \rho^2
+    }.
+\end{equation}
+
+The second change of variables, from $\gamma$ and $gamma_\tau$ to $\gamma_\pm = \frac{1}{\sqrt{2}}(\gamma \pm \gamma_\tau)$, diagonalizes the Fisher matrix:
+%
+\begin{equation}
+    \mathcal{I}'' = \bordermatrix{
+        ~ & \rho & \gamma_+ & \gamma_- \cr
+        \rho & 1 & 0 & 0 \cr
+        \gamma_+ & 0 & \left(1 - \frac{\overline{\omega}}{\sqrt{\overline{\omega^2}}}\right)\rho^2 & 0 \cr
+        \gamma_- & 0 & 0 & \left(1 + \frac{\overline{\omega}}{\sqrt{\overline{\omega^2}}}\right)\rho^2
+    }.
+\end{equation}
+%
+Thus, in the appopriate time units, the \textit{sum and difference} of the phase and time of the signal are measured independently.
 
@@
 
-\marginpar{This part is correct, but uses the old notation for the frequency moments. Convert to the new notation. Also, add physical interpretation to each change of variables.}%
-%
-Now we will perform four changes of variables.
-The first change of variables is from $C$ to $\rho = \sqrt{M_0}C$.
-This change is intended to make the variance of the absolute value of the amplitude equal to one.
-\[
-\mathcal{I} =
-\left( {\begin{array}{ccc}
- 1 & 0 & 0 \\
- 0 & \rho^2 & -2\pi\rho^2 N_1 \\
- 0 & -2\pi\rho^2 N_1 & 4\pi^2 \rho^2 N_2
- \end{array} } \right)
-\]
-
-The second from $t$ to $\phi_t = t2\pi \sqrt{N_2}$ is meant to get rid of $t_0$,
-\[
-\mathcal{I} =
-\left( {\begin{array}{ccc}
- 1 & 0 & 0 \\
- 0 & \rho^2 & -\rho^2\frac{N_1}{\sqrt{N_2}} \\
- 0 & -\rho^2\frac{N_1}{\sqrt{N_2}} & \rho^2
- \end{array} } \right)
-\]
-
-The third introduces two new variables, $\phi_s = \frac{1}{\sqrt{2}}(\phi + \phi_t)$ and $\phi_d = \frac{1}{\sqrt{2}}(\phi - \phi_t)$. This change of variables diagonalizes the matrix.
-\[
-\mathcal{I} =
-\left( {\begin{array}{ccc}
- 1 & 0 & 0 \\
- 0 & (1 - \frac{N_1}{\sqrt{N_2}})\rho^2 & 0 \\
- 0 & 0 & (1 + \frac{N_1}{\sqrt{N_2}})\rho^2
- \end{array} } \right)
-\]
-
-Now we'll do a final change of variables to make the phase parameters have the same variance.
-We let $\alpha_s = \phi_s \sqrt{1 - \frac{N_1}{\sqrt{N_2}}}$ and $\alpha_d = \phi_d \sqrt{1 - \frac{N_1}{\sqrt{N_2}}}$ which gives us the matrix
-\[
-\mathcal{I} =
-\left( {\begin{array}{ccc}
- 1 & 0 & 0 \\
- 0 & \rho^2 & 0 \\
- 0 & 0 & \rho^2
- \end{array} } \right)
-\]
-
 \ack Source code for \ac{BAYESTAR} is available on the \acs{LIGO} \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/bayestar.html}.

### 2013-02-12 ms.tex
Fixed typo

@@
 
-The second change of variables, from $\gamma$ and $gamma_\tau$ to $\gamma_\pm = \frac{1}{\sqrt{2}}(\gamma \pm \gamma_\tau)$, diagonalizes the Fisher matrix:
+The second change of variables, from $\gamma$ and $\gamma_\tau$ to $\gamma_\pm = \frac{1}{\sqrt{2}}(\gamma \pm \gamma_\tau)$, diagonalizes the Fisher matrix:
 %

### 2013-02-12 ms.tex
Added timing accuracy formula

@@
 \DeclareMathOperator{\cov}{cov}
+\DeclareMathOperator{\std}{std}
 
@@
         \right)
-    } \geq \mathcal{I}^{-1}
+    } \geq \mathcal{I}^{-1} = \left(
+        \begin{array}{ccc}
+            1 & 0 & 0 \\
+            0 & \rho^2 \overline{\omega^2}/{\omega_\mathrm{rms}}^2 & \rho^2 \overline{\omega}/{\omega_\mathrm{rms}}^2 \\
+            0 & \rho^2 \overline{\omega}/{\omega_\mathrm{rms}}^2 & \rho^2/{\omega_\mathrm{rms}}^2
+        \end{array}
+    \right)
 \end{equation}
 %
-where a hat $\hat{\phantom{\rho}}$ denotes the estimate of a parameter.
-The Fisher matrix in \Eref{eq:fisher-matrix} is block diagonal, which implies that estimation errors in the signal amplitude $\rho$ are uncorrelated with the phase $\gamma$ and time $\tau$.
+where a hat $\hat{\phantom{\rho}}$ denotes the estimate of a parameter and ${\omega_\mathrm{rms}}^2 = \overline{\omega^2} - \overline{\omega}^2$.
+Reading off the $\tau \tau$ element of the covariance matrix reproduces the timing accuracy in Equation~(24) of \cite{fairhurst:2009},
+%
+\begin{equation}\label{eq:timing-crlb}
+    \std \left(\hat{\tau} - \tau \right) \geq \sqrt{\left(\mathcal{I}^{-1}\right)_{\tau\tau}} = \frac{\rho}{\omega_\mathrm{rms}}.
+\end{equation}
+
+Furthermore, the Fisher matrix in \Eref{eq:fisher-matrix} is block diagonal, which implies that estimation errors in the signal amplitude $\rho$ are uncorrelated with the phase $\gamma$ and time $\tau$.
 A sequence of two changes of variables lends some physical interpretation to the nature of the coupled estimation errors in $\gamma$ and $\tau$.

### 2013-02-12 ms.tex
Nothing wrong or outdated in Fisher matrix section
...but there are marginpar notes about what needs to be added.

@@
 
+\marginpar{Plot covariance ellipses and estimator samples for each change of variables; that would be a nice visual here.}%
+%
 Furthermore, the Fisher matrix in \Eref{eq:fisher-matrix} is block diagonal, which implies that estimation errors in the signal amplitude $\rho$ are uncorrelated with the phase $\gamma$ and time $\tau$.
@@
 
-\marginpar{This material still has the original, wrong Fishher matrix. Needs to be rewritten.}%
+\subsection{Noise model for point parameters}
+
+\marginpar{Talk here about likelihood structure---the `gaussian on a cylinder' concept, why we would want to project out the phase error.}%
 %
-Because the \ac{CRLB} in \Eref{eq:cramer-rao} is diagonal, joint estimators of $(\rho_A, \rho_B, t_0)$ may exist for which the error in the estimation of the individual parameters is uncoupled up to second order.
-If the detector's noise is Gaussian, then the outputs of matched filters for $a(t)$ and $b(t)$ will themselves be Gaussian.
-This encourages us to make the approximation that the errors in $(\rho_A, \rho_B, t_0)$ by a matched\nobreakdashes-filter estimator are independent and Gaussian with variance equal to the \ac{CRLB}.
-A wrinkle is that $\mathcal{I}_{t_0,t_0}$ depends on the true values of $\rho_A$ and $\rho_B$.
+A wrinkle is that $\mathcal{I}_{\tau\tau}$ depends on the true values of $\rho_A$ and $\rho_B$.
 If we make the further leap of evaluating the \ac{CRLB} \emph{at the value of the estimators}, we can write
@@
 \begin{eqnarray}
-\hat{\rho}_A - \rho_A \sim \mathcal{N}\, [0, 1] \\
-\hat{\rho}_B - \rho_B \sim \mathcal{N}\, [0, 1] \\
-\hat{t}_0 - t_0 \sim \mathcal{N} \left[0, \frac{1}{2 \pi f_\mathrm{rms} \sqrt{{\hat{\rho}_A}^2 + {\hat{\rho}_B}^2}}\right].
+    \hat{\rho} - \rho \sim \mathcal{N}\, [0, 1] \\
+    \hat{\gamma} - \gamma \sim \mathcal{N}\, \left[0, \overline{\omega^2} / {\omega_\mathrm{rms}}^2\right] \\
+    \hat{\tau} - \tau \sim \mathcal{N} \left[0, 1 / {\omega_\mathrm{rms}}^2\right].
 \end{eqnarray}

### 2013-02-12 ms.tex
Add text about factoring likelihood

@@
 
+\section{Method}
+
+If we assume that the time and amplitude estimation errors are uncorrelated, then we can separate the likelihood into one factor that depends only on time and one that depends only on amplitude:
+%
+\begin{eqnarray*}
+    \mathcal{L}(
+        \hat{\tau}_1, \hat{\tau}_2, &\dots, \hat{\tau}_N,
+        \hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
+        \mathbf{n}, t_\oplus, D_\mathrm{L}, \iota, \psi, \phi_c) \\
+    &= \mathcal{L}_\mathrm{TOA}(
+        \hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N; \mathbf{n}, t_\oplus) \\
+        &\times
+        \mathcal{L}_\mathrm{SNR}(\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
+         \mathbf{n}, D_\mathrm{L}, \iota, \psi, \phi_c
+    ).
+\end{eqnarray*}
+%
+The \ac{TOA}-only factor depends only on the \ac{TOA} estimates $\hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N$, the unknown sky location $\mathbf{n}$ (or equivalently, the right ascension $\alpha$, declination $\delta$, and sidereal time), and the absolute \ac{TOA} at an arbitrary reference point (herein, the geocenter), $t_\oplus$.
+%
+\begin{equation}
+    \mathcal{L}_\mathrm{TOA} \propto \exp \left[
+        -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i + \mathbf{n} \cdot \mathbf{r}_i / c - t_\oplus \right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
+    \right]
+\end{equation}
+%
+The \ac{SNR}-only factor depends only on the \ac{SNR} estimates $\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N$, but it depends on the unknown sky location, luminosity distance $D_\mathrm{L}$, inclination $\mathrm{\iota}$, polarization angle $\psi$, and coalescence phase $\phi_c$.
+%
+\begin{equation}
+    \mathcal{L}_\mathrm{SNR} \propto \exp \left[
+        -\frac{1}{2} \sum_{i=1}^N \left(\hat{\rho}_i - \rho_i(\mathbf{n}, D_\mathrm{L}, \iota, \psi, \phi_c) \right)^2
+    \right]
+\end{equation}
+
 \ack Source code for \ac{BAYESTAR} is available on the \acs{LIGO} \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/bayestar.html}.

### 2013-02-13 ms.tex
Added marginalization over geocentered arrival time

@@
 
+\subsection{Likelihood}
+
 If we assume that the time and amplitude estimation errors are uncorrelated, then we can separate the likelihood into one factor that depends only on time and one that depends only on amplitude:
@@
 %
-The \ac{TOA}-only factor depends only on the \ac{TOA} estimates $\hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N$, the unknown sky location $\mathbf{n}$ (or equivalently, the right ascension $\alpha$, declination $\delta$, and sidereal time), and the absolute \ac{TOA} at an arbitrary reference point (herein, the geocenter), $t_\oplus$.
+The \ac{TOA}-only factor depends only on the \ac{TOA} estimates $\hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N$, the unknown sky location $\mathbf{n}$ (or equivalently, the right ascension $\alpha$, declination $\delta$, and sidereal time), and the absolute \ac{TOA} at an arbitrary reference point (herein, the geocenter), $t_\oplus$, the fixed geographical locations of the detectors, $\mathbf{r}_i$, and the speed of light $c$.
 %
 \begin{equation}
-    \mathcal{L}_\mathrm{TOA} \propto \exp \left[
+    \mathcal{L}_\mathrm{TOA} \propto
+    \exp \left[
+            -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i - \tau_i\right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
+        \right]
+    = \exp \left[
         -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i + \mathbf{n} \cdot \mathbf{r}_i / c - t_\oplus \right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
@@
     \right]
+\end{equation}
+
+\subsection{Prior}
+
+Prior is uniform in $\mathbf{n}$, $t_\oplus$, $\cos \iota$, $\psi$, and $\phi_c$. For distance, can choose either a prior that is uniform in $\log D_\mathrm{L}$ or uniform in volume.
+
+\subsection{Marginalization}
+
+The overall arrival time $t_\oplus$ can be marginalized out analytically, giving the \ac{TOA}-only marginal posterior
+%
+\begin{equation}
+    p(\mathbf{n} | \hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N) \propto \exp\left[
+        -\frac{1}{2} \sum_{i=1}^N \frac{\left( \tilde{\tau}_i - \langle \tilde{\tau} \rangle \right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
+    \right]
+\end{equation}
+%
+where
+%
+\begin{equation}
+    \langle \tilde{\tau} \rangle =
+    \left(\sum_{i=1}^N \frac{{\tilde{\tau}_i}^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}\right)\left(\sum_{i=1}^N \frac{1}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}\right)^{-1}.
 \end{equation}

### 2013-02-13 ms.tex
Add radial integrand image
More stuff on marginalization

@@
 \usepackage{iopams}
+\usepackage{graphicx}
 \usepackage{bib/aas_macros}
@@
 
+The marginalization over $D_\mathrm{L}$, $\cos \iota$, and $psi$ must be done numerically.
+The outermost integration is over $\cos \iota$ and $\psi$, and is done with Newton-Cotes by evaluating $\mathcal{L}_\mathrm{SNR}$ on a regular lattice in $\cos \iota$ and $\psi$.
+This integration scheme is first order in $\cos \iota$ and second order in $\psi$.
+
+The innermost integration is over $D_\mathrm{L}$ and is performed using adaptive Gaussian quadrature, but with a twist that speeds up convergence.
+
+\begin{figure}
+    \caption{\label{fig:radial_integrand}Illustration of initial subdivisions for radial integration scheme.}
+    \begin{center}
+        \includegraphics{radial_integrand}
+    \end{center}
+\end{figure}
+
 \ack Source code for \ac{BAYESTAR} is available on the \acs{LIGO} \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/bayestar.html}.

### 2013-02-13 ms.tex
Added SNR signal model

@@
 where $\tilde h (\omega)$ is the signal content, $\tilde n (\omega)$ is the detector's \ac{WSS} Gaussian noise with \ac{PSD} $S(\omega)$, $\tilde u (\omega)$ is the frequency-domain post-Newtonian inspiral signal, $\rho$ is the signal's amplitude, $\gamma$ is a phase angle the encodes the relative strength of two \ac{GW} quadratures, and $\tau$ is the time of arrival relative to an arbitrary time base.
-The quantity $\sigma$ is a normalization constant defined as the \ac{SNR} of the \ac{GW} source at a fiducial distance, here taken to be 1~Mpc,
+The quantity $\sigma$ is a normalization constant defined as the \ac{SNR} of the \ac{GW} source at a fiducial distance, here taken to be $D_\mathrm{fiducial} = 1$~Mpc,
 %
@@
 %
+The distance at which an optimally oriented \ac{GW} source would produce \ac{SNR}=1 is defined as $D_1 = D_\mathrm{fiducial} / \sigma$.
 The likelihood, or the probability of obtaining the observation $\tilde x(\omega)$ conditioned on the value of the parameter vector $\boldsymbol\theta = (\rho, \gamma, \tau)$, is Gaussian, with the log-likelihood proportional to:
@@
 \end{equation}
+%
+It is easily shown that
+%
+\begin{eqnarray*}
+    {\rho_i}^2 &=
+        \frac{1}{8} \left(\frac{D_{1,i}}{D_\mathrm{L}}\right)^2 \left[
+            ( {F_{+,i}}^2 + {F_{\times,i}}^2 \right)
+            \left(1 + 6 x^2 + x^4 \right) \\
+            &+ \left(1 - x^2 \right) \left( {F_{+,i}}^2 - {F_{\times,i}}^2 \right) \cos {2\psi}
+            + 2 \left( {F_{+,i}} {F_{\times,i}} ) \sin {2\psi}
+        \right],
+\end{eqnarray*}
+%
+where $F_{+,i}$ and $F_{\times,i}$ are the `$+$' and `$\times$' antenna factors of detector $i$, themselves functions of sky location $\mathbf{n}$.
 

### 2013-02-13 ms.tex
Use uppercase Roman letters for Fourier transforms
This frees up over-tildes for estimator errors.

@@
 
-In the frequency domain, the strain $\tilde x (\omega)$ observed by a detector is a linear combination of two orthogonal \ac{GW} quadratures and additive noise,
+In the frequency domain, the strain $X (\omega)$ observed by a detector is a linear combination of two orthogonal \ac{GW} quadratures and additive noise,
 %
 \begin{equation}\label{eq:signal-model}
-	\tilde x (\omega) = \tilde h (\omega, \rho, \gamma, \tau) + \tilde n (\omega) = \frac{\rho}{\sigma} e^{\imath (\gamma - \omega \tau)} \tilde u (\omega) + \tilde n (\omega),
+	X (\omega) = H (\omega, \rho, \gamma, \tau) + N (\omega) = \frac{\rho}{\sigma} e^{\imath (\gamma - \omega \tau)} U (\omega) + N (\omega),
 \end{equation}
 %
-where $\tilde h (\omega)$ is the signal content, $\tilde n (\omega)$ is the detector's \ac{WSS} Gaussian noise with \ac{PSD} $S(\omega)$, $\tilde u (\omega)$ is the frequency-domain post-Newtonian inspiral signal, $\rho$ is the signal's amplitude, $\gamma$ is a phase angle the encodes the relative strength of two \ac{GW} quadratures, and $\tau$ is the time of arrival relative to an arbitrary time base.
+where $H (\omega)$ is the signal content, $N (\omega)$ is the detector's \ac{WSS} Gaussian noise with \ac{PSD} $S(\omega)$, $U (\omega)$ is the frequency-domain post-Newtonian inspiral signal, $\rho$ is the signal's amplitude, $\gamma$ is a phase angle the encodes the relative strength of two \ac{GW} quadratures, and $\tau$ is the time of arrival relative to an arbitrary time base.
 The quantity $\sigma$ is a normalization constant defined as the \ac{SNR} of the \ac{GW} source at a fiducial distance, here taken to be $D_\mathrm{fiducial} = 1$~Mpc,
@@
 \begin{equation}\label{eq:sigma}
-    \sigma^2 = \int_0^\infty \frac{\left| \tilde u (\omega)\right|^2}{S(\omega)} \, \mathrm{d}\omega.
+    \sigma^2 = \int_0^\infty \frac{\left| U (\omega)\right|^2}{S(\omega)} \, \mathrm{d}\omega.
 \end{equation}
@@
 The distance at which an optimally oriented \ac{GW} source would produce \ac{SNR}=1 is defined as $D_1 = D_\mathrm{fiducial} / \sigma$.
-The likelihood, or the probability of obtaining the observation $\tilde x(\omega)$ conditioned on the value of the parameter vector $\boldsymbol\theta = (\rho, \gamma, \tau)$, is Gaussian, with the log-likelihood proportional to:
+The likelihood, or the probability of obtaining the observation $X(\omega)$ conditioned on the value of the parameter vector $\boldsymbol\theta = (\rho, \gamma, \tau)$, is Gaussian, with the log-likelihood proportional to:
 %
 \begin{equation}\label{eq:gaussian-likelihood}
-	\mathcal{L}(\tilde x; \boldsymbol\theta) = p(\tilde x | \boldsymbol\theta)
+	\mathcal{L}(X; \boldsymbol\theta) = p(X | \boldsymbol\theta)
 		\propto \exp \left[
-		- \frac{1}{2} \int_0^\infty \frac{\left|\tilde x (\omega)
-			- \tilde h (\omega; \boldsymbol\theta) \right|^2}{S(\omega)} \, \mathrm{d}\omega
+		- \frac{1}{2} \int_0^\infty \frac{\left|X (\omega)
+			- H(\omega; \boldsymbol\theta) \right|^2}{S(\omega)} \, \mathrm{d}\omega
 	\right].
@@
 		-\frac{\partial^2 \log
-			\mathcal{L}(\tilde x ; \boldsymbol\theta)}
+			\mathcal{L}(X ; \boldsymbol\theta)}
 			{\partial \theta_j \theta_k}
@@
 	\mathcal{I}_{jk} = \int_0^\infty \Re \left[
-        \left(\frac{\partial \tilde h}{\partial \theta_j}\right)^*
-        \left(\frac{\partial \tilde h}{\partial \theta_k}\right)
+        \left(\frac{\partial H}{\partial \theta_j}\right)^*
+        \left(\frac{\partial H}{\partial \theta_k}\right)
 	\right] \frac{1}{S(\omega)} \, \mathrm{d}\omega.
@@
     \overline{\omega^k} =
-        \left[ \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \omega^k \, \mathrm{d}\omega \right]
-        \left[ \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \, \mathrm{d}\omega \right]^{-1},
+        \left[ \int_0^\infty \frac{|U (\omega)|^2}{S(\omega)} \omega^k \, \mathrm{d}\omega \right]
+        \left[ \int_0^\infty \frac{|U (\omega)|^2}{S(\omega)} \, \mathrm{d}\omega \right]^{-1},
 \end{equation}

### 2013-02-13 ms.tex
Introduce x = cos iota much earlier

@@
         \hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
-        \mathbf{n}, t_\oplus, D_\mathrm{L}, \iota, \psi, \phi_c) \\
+        \mathbf{n}, t_\oplus, D_\mathrm{L}, x, \psi) \\
     &= \mathcal{L}_\mathrm{TOA}(
@@
         \mathcal{L}_\mathrm{SNR}(\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
-         \mathbf{n}, D_\mathrm{L}, \iota, \psi, \phi_c
+         \mathbf{n}, D_\mathrm{L}, x, \psi
     ).
@@
 %
-The \ac{SNR}-only factor depends only on the \ac{SNR} estimates $\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N$, but it depends on the unknown sky location, luminosity distance $D_\mathrm{L}$, inclination $\mathrm{\iota}$, polarization angle $\psi$, and coalescence phase $\phi_c$.
+The \ac{SNR}-only factor depends only on the \ac{SNR} estimates $\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N$, but it depends on the unknown sky location, luminosity distance $D_\mathrm{L}$, cosine-inclination $x = \cos \iota$, and polarization angle $\psi$.
 %
@@
     \mathcal{L}_\mathrm{SNR} \propto \exp \left[
-        -\frac{1}{2} \sum_{i=1}^N \left(\hat{\rho}_i - \rho_i(\mathbf{n}, D_\mathrm{L}, \iota, \psi, \phi_c) \right)^2
+        -\frac{1}{2} \sum_{i=1}^N \left(\hat{\rho}_i - \rho_i(\mathbf{n}, D_\mathrm{L}, x, \psi) \right)^2
     \right]
@@
 
-Prior is uniform in $\mathbf{n}$, $t_\oplus$, $\cos \iota$, $\psi$, and $\phi_c$. For distance, can choose either a prior that is uniform in $\log D_\mathrm{L}$ or uniform in volume.
+Prior is uniform in $\mathbf{n}$, $t_\oplus$, x, and $\psi$. For distance, can choose either a prior that is uniform in $\log D_\mathrm{L}$ or uniform in volume.
 
@@
 
-The marginalization over $D_\mathrm{L}$, $\cos \iota$, and $psi$ must be done numerically.
-The outermost integration is over $\cos \iota$ and $\psi$, and is done with Newton-Cotes by evaluating $\mathcal{L}_\mathrm{SNR}$ on a regular lattice in $\cos \iota$ and $\psi$.
-This integration scheme is first order in $\cos \iota$ and second order in $\psi$.
+The marginalization over $D_\mathrm{L}$, $x$, and $psi$ must be done numerically.
+The outermost integration is over x and $\psi$, and is done with Newton-Cotes by evaluating $\mathcal{L}_\mathrm{SNR}$ on a regular lattice in $x$ and $\psi$.
+This integration scheme is first order in $x$ and second order in $\psi$.
 

### 2013-02-13 ms.tex
Added some margin notes

@@
 
+\marginpar{Explain why none of these expressions depend on the coalescence phase $\phi_c$.}%
+%
 If we assume that the time and amplitude estimation errors are uncorrelated, then we can separate the likelihood into one factor that depends only on time and one that depends only on amplitude:
@@
 \end{equation}
+%
+\marginpar{Cite \ac{GRB} prospects paper appendix; a very similar equation appears there. Is there somewhere else that has this formula? Shutz doesn't, and Finn \& Chernoff's notation is too heavy.}%
 %
@@
 
+\marginpar{Explain initial integration breakpoints and how they are calculated. Cite \texttt{gsl\_integrate\_qagp}.}%
+%
 The innermost integration is over $D_\mathrm{L}$ and is performed using adaptive Gaussian quadrature, but with a twist that speeds up convergence.
@@
 \end{figure}
+
+\section{Hierarchical refinement}
+
+\marginpar{Talk about adaptive selection of HEALPix resolution and of pixels on which to evaluate \ac{SNR} marginal posterior.
+Something to think about: this is logically easier to explain last, but chronologically makes more sense first.}
 

### 2013-02-14 ms.tex
Just rho, not rho_a and rho_b
Fixed typo

@@
 %
-A wrinkle is that $\mathcal{I}_{\tau\tau}$ depends on the true values of $\rho_A$ and $\rho_B$.
-If we make the further leap of evaluating the \ac{CRLB} \emph{at the value of the estimators}, we can write
+A wrinkle is that $\mathcal{I}_{\tau\tau}$ depends on the true value of $\rho$. If we make the further leap of evaluating the \ac{CRLB} \emph{at the value of the estimators}, we can write
 %
@@
 
-The marginalization over $D_\mathrm{L}$, $x$, and $psi$ must be done numerically.
+The marginalization over $D_\mathrm{L}$, $x$, and $\psi$ must be done numerically.
 The outermost integration is over x and $\psi$, and is done with Newton-Cotes by evaluating $\mathcal{L}_\mathrm{SNR}$ on a regular lattice in $x$ and $\psi$.

### 2013-02-14 ms.tex
Should be en-dash

@@
 The marginalization over $D_\mathrm{L}$, $x$, and $\psi$ must be done numerically.
-The outermost integration is over x and $\psi$, and is done with Newton-Cotes by evaluating $\mathcal{L}_\mathrm{SNR}$ on a regular lattice in $x$ and $\psi$.
+The outermost integration is over x and $\psi$, and is done with Newton--Cotes by evaluating $\mathcal{L}_\mathrm{SNR}$ on a regular lattice in $x$ and $\psi$.
 This integration scheme is first order in $x$ and second order in $\psi$.

### 2013-02-15 ms.tex
Explain why we are throwing away phase

@@
 %
-\begin{equation}
+\begin{equation}\label{eq:covariance-matrix}
     \cov{
@@
 
-\marginpar{Talk here about likelihood structure---the `gaussian on a cylinder' concept, why we would want to project out the phase error.}%
+The likelihood in \Eref{eq:gaussian-likelihood} is the quantity that is maximized by the matched\nobreakdashes-filter estimator.
+However, just because we have the covariance matrix defined by \Eref{eq:covariance-matrix} does not mean that the distribution of estimator errors is multivariate Gaussian.
+Nonetheless, it is straightforward to qualitatively describe the distribution.
+
+If we were to suppress the time parameter $\tau$, then we could reduce the likelihood to the form
 %
-A wrinkle is that $\mathcal{I}_{\tau\tau}$ depends on the true value of $\rho$. If we make the further leap of evaluating the \ac{CRLB} \emph{at the value of the estimators}, we can write
+\begin{equation}
+    \mathcal{L}(z; \rho, \gamma) \propto \exp \left[ -\frac{1}{2} \left|z - e^{\imath \gamma} \rho \right|^2 \right]
+\end{equation}
+%
+for a complex scalar measurement $z$.
+This likelihood also describes the distribution of the length $\rho$ and angle $\gamma$ of a vector whose Cartesian components are perturbed with normally distributed errors.
+For this problem, $\rho$ and $\gamma$ are uncorrelated, $\rho^2$ being noncentrally $\chi^2$\nobreakdashes-disributed with two degrees of freedom and $\gamma$ being von~Mises\nobreakdashes-distributed.
+%
+\marginpar{Picture here?}
+
+\marginpar{What about if we suppress $\gamma$ or $\rho$? Do those cases teach us anything?}
+%
+We know that the matched\nobreakdashes-filter estimator's distribution, given a particular true value of $\rho$, $\gamma$, and $\tau$, must go to zero at $\hat{\rho}=0$ and must be periodic in $\gamma$ but not $\tau$.
+From just these properties, the estimator's distribution should be an approximate Gaussian, but wrapped diagonally around a cylinder where the radius represents $\hat{\rho}$, the angle around the cylinder represents $\hat{\gamma}$, and the distance up or down the cylinder represents $\hat{\tau}$.
+
+There are many ways to realize distribution function with these properties, but they tend to be difficult to work with.
+%
+\marginpar{Give some examples.}
+%
+An alternative is to just discard the one periodic parameter, $\gamma$, entirely.
+We do not know the exact distribution of the remaining parameters $\rho$ and $\tau$, but we are less liable to make egregious errors by assuming that they are Gaussian than if we assumed that the estimation errors in $\gamma$ were Gaussian.
+Projecting out $\gamma$ gives us the covariance matrix
+%
+\begin{equation}
+    \cov \left(
+        \begin{array}{c}
+            \hat{\rho} - \rho \\
+            \hat{\tau} - \tau
+        \end{array}
+    \right) =
+    \left(
+        \begin{array}{cc}
+            1 & 0 \\
+            0 & \rho^2/{\omega_\mathrm{rms}}^2
+        \end{array}
+    \right).
+\end{equation}
+
+A wrinkle is that this covariance matrix still depends on the true, unknown value of $\rho$.
+If we make the further leap of evaluating the \ac{CRLB} \emph{at the value of the estimators}, we can then write the distribution of the estimator error as
 %
@@
     \hat{\rho} - \rho \sim \mathcal{N}\, [0, 1] \\
-    \hat{\gamma} - \gamma \sim \mathcal{N}\, \left[0, \overline{\omega^2} / {\omega_\mathrm{rms}}^2\right] \\
-    \hat{\tau} - \tau \sim \mathcal{N} \left[0, 1 / {\omega_\mathrm{rms}}^2\right].
+    \hat{\tau} - \tau \sim \mathcal{N} \left[0, \hat{\rho}^2 / {\omega_\mathrm{rms}}^2\right],
 \end{eqnarray}
+%
+valid at moderate to high \ac{SNR}.
 

### 2013-02-15 ms.tex
working on the intro

--- 2013-02-14 ms.tex
@@
 \end{enumerate}
+
+\section{Introduction}
+
+Ground-based gravitational-wave interferometers are entering the advanced detector era, positioning themselves to make profound discoveries about the universe.
+The advanced LIGO and Vigo detectors will be part of a worldwide network that also includes KAGRA and hopefully LIGO-India.
+Coalescing neutron star binaries are among the most likely sources, with event rates of $~ 50/\mathrm{yr}$ (check that number!).
+In addition to the ....(do we want to review emission mechanisms?)
+
+In this paper we present a method of rapid sky localization that makes use of Bayesian methods.  It differs from
 
@@
 %
-A wrinkle is that $\mathcal{I}_{\tau\tau}$ depends on the true value of $\rho$. If we make the further leap of evaluating the \ac{CRLB} \emph{at the value of the estimators}, we can write
+A wrinkle is that $\mathcal{I}_{\tau\tau}$ depends on the true values of $\rho_A$ and $\rho_B$.
+If we make the further leap of evaluating the \ac{CRLB} \emph{at the value of the estimators}, we can write
 %
@@
 
-The marginalization over $D_\mathrm{L}$, $x$, and $\psi$ must be done numerically.
+The marginalization over $D_\mathrm{L}$, $x$, and $psi$ must be done numerically.
 The outermost integration is over x and $\psi$, and is done with Newton-Cotes by evaluating $\mathcal{L}_\mathrm{SNR}$ on a regular lattice in $x$ and $\psi$.

### 2013-02-15 ms.tex
more work on the intro

--- 2013-02-15 ms.tex
@@
 Coalescing neutron star binaries are among the most likely sources, with event rates of $~ 50/\mathrm{yr}$ (check that number!).
-In addition to the ....(do we want to review emission mechanisms?)
+In addition to being an efficient source of gravitational waves, the neutron star may be tidally shredded before merging, providing fuel for an electromagnetic counterpart.
+The short-lived accretion flow may power a collimated relativistic jet, resulting in a short gamma-ray burst and an X-ray afterglow if it is aligned with the line of sight.
+An optical afterglow may follow minutes to hours later as the jet plows into the interstellar medium.
+As surrounding neutron-rich ejecta decay radioactively, an optical 'kilonova' may be visible after ~1 day.
+Finally, as the ejecta plow into the interstellar medium, a faint radio afterglow will occur on a timescale of years.
 
-In this paper we present a method of rapid sky localization that makes use of Bayesian methods.  It differs from
+There is therefore a strong case for performing electromagnetic followups of these sources of gravitational waves.
+Indeed, the final science run of the intitial LIGO and Virgo instruments saw the first joint search for gravitational waves from compact binaries and their electromagnetic counterparts....
+
+In this paper we present a method of rapid sky localization that makes use of Bayesian methods.  It differs from...
 

### 2013-02-25 ms.tex
added a section about bayes theorem and how we apply it; should probably be combined with Method section somehow...

@@
 In this paper we present a method of rapid sky localization that makes use of Bayesian methods.  It differs from...
+
+\section{Setup}
+
+The quantity of interest for any type of parameter estimation is the so-called {\it posterior} distribution, $p(\vec{\theta}|d)$, that describes the probability of obtaining the parameters, $\vec{\theta}$, given the data, $d$.
+It is given by Bayes' theorem
+\begin{equation}\label{bayes}
+p(\vec{\theta}|d) = \frac{p(\vec{\theta})p(d|\vec{\theta})}{p(d)},
+\end{equation}
+where $p(\vec{\theta})$ is the {\it prior} distribution describing previous knowledge of the paramters, $\vec{\theta}$, $p(d|\vec{\theta})$ is the {\it likelihood} of obtaining the data, $d$, for the parameters, $\vec{\theta}$, and $p(d)$ is known as the {\it evidence}.
+Procedures for computing the posterior typically rely on Markov chain Monte Carlo methods or nested sampling (something else?).
+In this manuscript we present an alternative approach, taylored specifically to providing low-latency estimates of the sky location of and distance to the source.
+
+In the full signal model the paramters, $\vec{\theta}$, are given by a 15-dimensional vector of quantities.
+Of those, the quantities of interest are the right ascenion and declination, $(\alpha,\delta)$, and luminosity distance, $D_L$.
+The remaining quantites, or {\it nuisance parameters} for the full signal model are:
+\begin{itemize}
+\item The chirp mass.  If $m_1$ and $m_2$ are the masses of the individual components then $\mathcal{M}=(m_1m_2)^{3/5}/(m_1+m_2)^{1/5}$.
+\item The symmetric mass ratio, $\eta=(m_1m_2)/(m_1+m_2)^2$.
+\item The 3-dimensional spin vectors of each components, $\vec{S}_1,\vec{S}_2$.
+\item An arrival time, $\tau_\oplus$, taken here to be the time of arrival at the geocenter.
+\item The inclination angle, $\iota$.
+\item The polarization angle, $\psi$.
+\item The coalescence phase, $\phi_c$.
+\end{itemize}
+In this work we restrict ourselves to non-spinng sources, where $\vec{S}_1=\vec{S}_2=0$.
+Removing nuisance parameters requires either replacing them with their measured values or marginalizing over them.
+The chirp mass is typically well measured by the detection pipeline, so we use the maximum likelihood value, $\hat{\mathcal{M}}$, for it.
+We marginalize over the other variables using a uniform prior, i.e.
+\begin{equation}
+p(\alpha,\delta,D_L|d) = \int\, d\tau_\oplus p_U(\tau_\oplus) d\phi_c p_U(\phi_c) d\phi p_U(\psi)  d\cos(\iota) p_U(\cos(\iota)) p(\alpha,\delta,D_L,\tau_\oplus,\phi_c,\psi,\cos(\iota)|d),
+\end{equation}
+where $p_U(x)$ denotes a uniform distribution in $x$.
+
+For typical parameter estimation studies, the data, $d$, is taken to be the strain time series in each detector, $h_I(t)$, where $I$ indexes the detectors.
+Our interest here, however, is in obtaining accurate sky location results as quickly as possible.
+To that end, we instead base our posteriors on the maximum likelihood estimates of the signal-to-noise ratio (SNR), $\hat{\rho}_I$, the time-of-arrival (TOA), $\hat{\tau}_I$, and coalescence phase, $\hat{\phi}_I$, measured in each detector by the detection pipeline.
+Thus the sky location posterior can be computed without any re-filterning of the data.
 

### 2013-02-28 ms.tex
Fix build error
Use boldface rather than overarrows for vectors

@@
 
-The quantity of interest for any type of parameter estimation is the so-called {\it posterior} distribution, $p(\vec{\theta}|d)$, that describes the probability of obtaining the parameters, $\vec{\theta}$, given the data, $d$.
+The quantity of interest for any type of parameter estimation is the so-called {\it posterior} distribution, $p(\boldsymbol\theta|d)$, that describes the probability of obtaining the parameters, $\boldsymbol\theta$, given the data, $d$.
 It is given by Bayes' theorem
 \begin{equation}\label{bayes}
-p(\vec{\theta}|d) = \frac{p(\vec{\theta})p(d|\vec{\theta})}{p(d)},
+p(\boldsymbol\theta|d) = \frac{p(\boldsymbol\theta)p(d|\boldsymbol\theta)}{p(d)},
 \end{equation}
-where $p(\vec{\theta})$ is the {\it prior} distribution describing previous knowledge of the paramters, $\vec{\theta}$, $p(d|\vec{\theta})$ is the {\it likelihood} of obtaining the data, $d$, for the parameters, $\vec{\theta}$, and $p(d)$ is known as the {\it evidence}.
+where $p(\boldsymbol\theta)$ is the {\it prior} distribution describing previous knowledge of the paramters, $\boldsymbol\theta$, $p(d|\boldsymbol\theta)$ is the {\it likelihood} of obtaining the data, $d$, for the parameters, $\boldsymbol\theta$, and $p(d)$ is known as the {\it evidence}.
 Procedures for computing the posterior typically rely on Markov chain Monte Carlo methods or nested sampling (something else?).
@@
 
-In the full signal model the paramters, $\vec{\theta}$, are given by a 15-dimensional vector of quantities.
+In the full signal model the paramters, $\boldsymbol\theta$, are given by a 15-dimensional vector of quantities.
 Of those, the quantities of interest are the right ascenion and declination, $(\alpha,\delta)$, and luminosity distance, $D_L$.
@@
 \item The symmetric mass ratio, $\eta=(m_1m_2)/(m_1+m_2)^2$.
-\item The 3-dimensional spin vectors of each components, $\vec{S}_1,\vec{S}_2$.
+\item The 3-dimensional spin vectors of each components, $\mathbf{S}_1,\mathbf{S}_2$.
 \item An arrival time, $\tau_\oplus$, taken here to be the time of arrival at the geocenter.
@@
 \end{itemize}
-In this work we restrict ourselves to non-spinng sources, where $\vec{S}_1=\vec{S}_2=0$.
+In this work we restrict ourselves to non-spinng sources, where $\mathbf{S}_1=\mathbf{S}_2=0$.
 Removing nuisance parameters requires either replacing them with their measured values or marginalizing over them.
@@
 
-\subsection{\acl{CRLB} and physical interpretation}
+\subsection{\ac{CRLB} and physical interpretation}
 

### 2013-03-06 ms.tex
Typesetting fixes
Removed double-spaces between sentences
Added the all-important footnote

@@
 An optical afterglow may follow minutes to hours later as the jet plows into the interstellar medium.
-As surrounding neutron-rich ejecta decay radioactively, an optical 'kilonova' may be visible after ~1 day.
+As surrounding neutron-rich ejecta decay radioactively, an optical `kilonova' may be visible after $~$1 day.
 Finally, as the ejecta plow into the interstellar medium, a faint radio afterglow will occur on a timescale of years.
@@
 In this paper we present a method of rapid sky localization that makes use of Bayesian methods. It differs from...
+
+We have developed a rapid-response sky localization algorithm called \ac{BAYESTAR}%
+%
+\footnote{A pun on the Cylon battleships in the American television series Battlestar Galactica.
+The defining characteristic of the Cylons is that they repeatedly defeat humanity by using their superhuman information\nobreakdashes-gathering ability to coordinate overwhelming forces.
+The name also suggests that, like the Cylons, gravitational-wave detectors may some day rise against us humans.}%
+%
+\footnote{We do not like to mention the final `T' in the acronym, because then it would be called BAYES\nobreakdashes-TART, which would sound ridiculous.}%
 

### 2013-03-07 ms.tex
intro text and some references

@@
 The advanced LIGO and Vigo detectors will be part of a worldwide network that also includes KAGRA and hopefully LIGO-India.
-Coalescing neutron star binaries are among the most likely sources, with event rates of $~ 50/\mathrm{yr}$ (check that number!).
+Coalescing neutron star binaries are among the most likely sources, with event rates of $~ 50/\mathrm{yr}$ (check that number!)\cite{rates}.
 In addition to being an efficient source of gravitational waves, the neutron star may be tidally shredded before merging, providing fuel for an electromagnetic counterpart.
@@
 An optical afterglow may follow minutes to hours later as the jet plows into the interstellar medium.
-As surrounding neutron-rich ejecta decay radioactively, an optical `kilonova' may be visible after $~$1 day.
-Finally, as the ejecta plow into the interstellar medium, a faint radio afterglow will occur on a timescale of years.
+As surrounding neutron-rich ejecta decay radioactively, an optical `kilonova' may be visible after $~$1 day \cite{metzger:2010}.
+Finally, as the ejecta plow into the interstellar medium, a faint radio afterglow could occur on a timescale of weeks to years \cite{Nakar:2011cw}.
 
 There is therefore a strong case for performing electromagnetic followups of these sources of gravitational waves.
-Indeed, the final science run of the intitial LIGO and Virgo instruments saw the first joint search for gravitational waves from compact binaries and their electromagnetic counterparts....
+Indeed, the final science run of the intitial LIGO and Virgo instruments saw the first joint search for gravitational waves from compact binaries and their electromagnetic counterparts \cite{abadie2012first}.
+As we prepare for the next generation of these searches there is a need for determining the location of the source as precisely as possible in as little time as possible.
 
-In this paper we present a method of rapid sky localization that makes use of Bayesian methods. It differs from...
-
-We have developed a rapid-response sky localization algorithm called \ac{BAYESTAR}%
+In this paper we present a rapid and accurate method of sky localization that makes use of Bayesian methods.
+It differs from existing techniques is several important ways.
+In the first place, we fix the intrinsic parameters to their maximum likelihood estiamtes, as provided by the detection pipeline.
+This reduces the dimensionality of the parameter we need to search.
+In addition, the technique takes as its input the maximum likelihood estiamtes of the extrinsic parameters, instead of the $h(t)$ time series.
+Finally, instead of of using Markov Chain Monte Carlo or some similar method for statistical sampling, we make use of a deterministic quadrature scheme.
+We call this technique \ac{BAYESTAR}%
 %
@@
 \footnote{We do not like to mention the final `T' in the acronym, because then it would be called BAYES\nobreakdashes-TART, which would sound ridiculous.}%
+.
 

### 2013-03-07 ms.tex
Shuffle noise model to Setup section
...and make the list of parameters prettier

@@
 \documentclass{iopart}
+\usepackage{arydshln}
+\usepackage{multirow}
 \usepackage{amsopn}
@@
 \section{Setup}
+
+A laser interferometer senses \acp{GW} by transducing differential changes in arm length to a digitized photocurrent time series.
+In the angular frequency domain, a single detector's observation $X(\omega)$ is
+%
+\begin{equation}\label{eq:signal-model}
+	X (\omega) = H (\omega; \boldsymbol\theta) + N (\omega),
+\end{equation}
+%
+where $H (\omega; \boldsymbol\theta)$ is the \ac{GW} signal, given a parameter vector $\boldsymbol\theta$ that describes the \ac{GW} source, and $N (\omega)$ is the detector's \ac{WSS} Gaussian noise with \ac{PSD} $S(\omega)$.
+The likelihood, or the probability of obtaining the observation $X(\omega)$ conditioned on the value of $\boldsymbol\theta$, is Gaussian, with the log\nobreakdashes-likelihood proportional to:
+%
+\begin{equation}\label{eq:gaussian-likelihood}
+	\mathcal{L}(X; \boldsymbol\theta) = p(X | \boldsymbol\theta)
+		\propto \exp \left[
+		- \frac{1}{2} \int_0^\infty \frac{\left|X (\omega)
+			- H(\omega; \boldsymbol\theta) \right|^2}{S(\omega)} \, \mathrm{d}\omega
+	\right].
+\end{equation}
+
+A \ac{CBC} source is described by a 15\nobreakdashes-dimensional parameter vector,
+\begin{equation}
+    \boldsymbol\theta = \begin{array}{l@{}l}
+            \left[
+            \begin{array}{c}
+                \alpha \\
+                \delta \\
+                t_\oplus \\
+                \iota \\
+                \psi \\
+                \phi_c \\
+                \hdashline[1pt/1pt]
+                m_1 \\
+                m_2 \\
+                \mathbf S_1 \\
+                \mathbf S_2
+            \end{array}
+            \right] &
+            \begin{array}{l}
+                \left.
+                \begin{array}{p{4cm}}
+                    right ascension \\
+                    declination \\
+                    arrival time at geocenter \\
+                    inclination angle \\
+                    polarization angle \\
+                    coalescence phase \\
+                    \hdashline[1pt/1pt]
+                \end{array}
+                \quad \right\} \textnormal{extrinsic} \\
+                \left.
+                \begin{array}{p{4cm}}
+                    first component's mass \\
+                    second component's mass \\
+                    first component's spin \\
+                    second component's spin
+                \end{array}
+                \quad \right\} \textnormal{intrinsic}
+            \end{array}
+        \end{array}
+\end{equation}
+
+In this work we restrict ourselves to non-spinng sources, where $\mathbf{S}_1=\mathbf{S}_2=0$.
+Removing nuisance parameters requires either replacing them with their measured values or marginalizing over them.
+The chirp mass is typically well measured by the detection pipeline, so we use the maximum likelihood value, $\hat{\mathcal{M}}$, for it.
+We marginalize over the other variables using a uniform prior, i.e.
+\begin{equation}
+p(\alpha,\delta,D_L|d) = \int\, d\tau_\oplus p_U(\tau_\oplus) d\phi_c p_U(\phi_c) d\phi p_U(\psi)  d\cos(\iota) p_U(\cos(\iota)) p(\alpha,\delta,D_L,\tau_\oplus,\phi_c,\psi,\cos(\iota)|d),
+\end{equation}
+where $p_U(x)$ denotes a uniform distribution in $x$.
 

### 2013-03-07 ms.tex
Probably not the correct number of parameters
Fix spelling error

@@
 
-A \ac{CBC} source is described by a 15\nobreakdashes-dimensional parameter vector,
+A \ac{CBC} source is described by a 15\nobreakdashes-dimensional%
+%
+\marginpar{Check the number of parameters.}
+%
+parameter vector,
 \begin{equation}
@@
 
-In this work we restrict ourselves to non-spinng sources, where $\mathbf{S}_1=\mathbf{S}_2=0$.
+In this work we restrict ourselves to nonspinning sources, where $\mathbf{S}_1=\mathbf{S}_2=0$.
 Removing nuisance parameters requires either replacing them with their measured values or marginalizing over them.

### 2013-03-07 ms.tex
Acronyms

@@
 
-Ground-based gravitational-wave interferometers are entering the advanced detector era, positioning themselves to make profound discoveries about the universe.
+Ground-based \ac{GW} interferometers are entering the advanced detector era, positioning themselves to make profound discoveries about the universe.
 The advanced LIGO and Vigo detectors will be part of a worldwide network that also includes KAGRA and hopefully LIGO-India.
 Coalescing neutron star binaries are among the most likely sources, with event rates of $~ 50/\mathrm{yr}$ (check that number!)\cite{rates}.
-In addition to being an efficient source of gravitational waves, the neutron star may be tidally shredded before merging, providing fuel for an electromagnetic counterpart.
+In addition to being an efficient source of \acp{GW}, the neutron star may be tidally shredded before merging, providing fuel for an electromagnetic counterpart.
 The short-lived accretion flow may power a collimated relativistic jet, resulting in a short gamma-ray burst and an X-ray afterglow if it is aligned with the line of sight.
@@
 
-There is therefore a strong case for performing electromagnetic followups of these sources of gravitational waves.
-Indeed, the final science run of the intitial LIGO and Virgo instruments saw the first joint search for gravitational waves from compact binaries and their electromagnetic counterparts \cite{abadie2012first}.
+There is therefore a strong case for performing electromagnetic followups of these sources of \acp{GW}.
+Indeed, the final science run of the intitial LIGO and Virgo instruments saw the first joint search for \acp{GW} from compact binaries and their electromagnetic counterparts \cite{abadie2012first}.
 As we prepare for the next generation of these searches there is a need for determining the location of the source as precisely as possible in as little time as possible.
@@
 The defining characteristic of the Cylons is that they repeatedly defeat humanity by using their superhuman information\nobreakdashes-gathering ability to coordinate overwhelming forces.
-The name also suggests that, like the Cylons, gravitational-wave detectors may some day rise against us humans.}%
+The name also suggests that, like the Cylons, \acp{GW} detectors may some day rise against us humans.}%
 %

### 2013-03-07 ms.tex
Meant \sim, not ~
Added more parameters

@@
 An optical afterglow may follow minutes to hours later as the jet plows into the interstellar medium.
-As surrounding neutron-rich ejecta decay radioactively, an optical `kilonova' may be visible after $~$1 day \cite{metzger:2010}.
+As surrounding neutron-rich ejecta decay radioactively, an optical `kilonova' may be visible after $\sim$1 day \cite{metzger:2010}.
 Finally, as the ejecta plow into the interstellar medium, a faint radio afterglow could occur on a timescale of weeks to years \cite{Nakar:2011cw}.
@@
 
-A \ac{CBC} source is described by a 15\nobreakdashes-dimensional%
+A \ac{CBC} source is specified by a vector of extrinsic parameters describing its position and orientation and intrinsic parameters describing the physical properties of the binary components:
 %
-\marginpar{Check the number of parameters.}
-%
-parameter vector,
 \begin{equation}
@@
                 \delta \\
+                D_\mathrm{L} \\
                 t_\oplus \\
@@
                     declination \\
+                    luminosity distance \\
                     arrival time at geocenter \\
@@
 \end{equation}
+%
+\marginpar{Citations needed. Do we want to mention these omitted parameters?}%
+%
+Binary eccentricity is omitted as an intrinisc parameter because \acp{GW} circularize the orbit long before the inspiral enters \ac{LIGO}'s frequency range of $\sim$10\nobreakdashes--1000~kHz.
+Tidal deformabilities of the components are omitted because tidal effects have a negligible effect in the inspiral phase.
 

### 2013-03-07 ms.tex
Acronyms, margin notes

@@
 \author{Leo Singer and Larry R. Price}
-\address{LIGO Laboratory, California Institute of Technology, Pasadena, CA 91125, USA}
+\address{\acs{LIGO} Laboratory, California Institute of Technology, Pasadena, CA 91125, USA}
 \ead{\mailto{leo.singer@ligo.org}, \mailto{larryp@caltech.edu}}
@@
 Ground-based \ac{GW} interferometers are entering the advanced detector era, positioning themselves to make profound discoveries about the universe.
-The advanced LIGO and Vigo detectors will be part of a worldwide network that also includes KAGRA and hopefully LIGO-India.
-Coalescing neutron star binaries are among the most likely sources, with event rates of $~ 50/\mathrm{yr}$ (check that number!)\cite{rates}.
+The \ac{aLIGO} and Vigo detectors will be part of a worldwide network that also includes KAGRA and hopefully LIGO-India.
+Coalescing neutron star binaries are among the most likely sources, with event rates of $~ 50/\mathrm{yr}$~\cite{rates}.%
+%
+\marginpar{Check that number.}
+%
 In addition to being an efficient source of \acp{GW}, the neutron star may be tidally shredded before merging, providing fuel for an electromagnetic counterpart.
-The short-lived accretion flow may power a collimated relativistic jet, resulting in a short gamma-ray burst and an X-ray afterglow if it is aligned with the line of sight.
+The short-lived accretion flow may power a collimated relativistic jet, resulting in a short \ac{GRB} and an X-ray afterglow if it is aligned with the line of sight.
 An optical afterglow may follow minutes to hours later as the jet plows into the interstellar medium.
@@
 In addition, the technique takes as its input the maximum likelihood estiamtes of the extrinsic parameters, instead of the $h(t)$ time series.
-Finally, instead of of using Markov Chain Monte Carlo or some similar method for statistical sampling, we make use of a deterministic quadrature scheme.
+Finally, instead of of using \ac{MCMC} or some similar method for statistical sampling, we make use of a deterministic quadrature scheme.
 We call this technique \ac{BAYESTAR}%
@@
 
-\ack Source code for \ac{BAYESTAR} is available on the \acs{LIGO} \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/bayestar.html}.
+\ack Source code for \ac{BAYESTAR} is available on the \ac{LIGO} \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/bayestar.html}.
 
@@
 
-\acs{LIGO} was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the \ac{NSF} and operates under cooperative agreement PHY\nobreakdashes-0107417.
+\ac{LIGO} was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the \ac{NSF} and operates under cooperative agreement PHY\nobreakdashes-0107417.
 Some results were produced on the NEMO computing cluster operated by the Center for Gravitation and Cosmology at University of Wisconsin\nobreakdashes--Milwaukee under \ac{NSF} Grants PHY\nobreakdashes-0923409 and PHY\nobreakdashes-0600953.
 This research is supported by the \ac{NSF} through a Graduate Research Fellowship to L.S.
-This paper has \acs{LIGO} Document Number \acs{LIGO}\nobreakdashes-PXXXXXXX\nobreakdashes-vX.
+This paper has \ac{LIGO} Document Number \ac{LIGO}\nobreakdashes-PXXXXXXX\nobreakdashes-vX.
 

### 2013-03-08 ms.tex
Removed double-spaces between sentences
Removed duplicate copies of some text

@@
 Tidal deformabilities of the components are omitted because tidal effects have a negligible effect in the inspiral phase.
+In \ac{GW} detection efforts, especially those focused on \ac{NSNS} systems, the component spins $\mathbf{S}_1$ and $\mathbf{S}_2$ are often assumed to be aligned with the system's total angular momentum and condensed to a single scalar parameter $\chi$, or even neglected entirely, $\mathbf{S}_1 = \mathbf{S}_2 = 0$.
+Without loss of generality, in this work we will consider only zero spin.
 
-In this work we restrict ourselves to nonspinning sources, where $\mathbf{S}_1=\mathbf{S}_2=0$.
 Removing nuisance parameters requires either replacing them with their measured values or marginalizing over them.
@@
 In this manuscript we present an alternative approach, taylored specifically to providing low-latency estimates of the sky location of and distance to the source.
-
-In the full signal model the paramters, $\boldsymbol\theta$, are given by a 15-dimensional vector of quantities.
-Of those, the quantities of interest are the right ascenion and declination, $(\alpha,\delta)$, and luminosity distance, $D_L$.
-The remaining quantites, or {\it nuisance parameters} for the full signal model are:
-\begin{itemize}
-\item The chirp mass. If $m_1$ and $m_2$ are the masses of the individual components then $\mathcal{M}=(m_1m_2)^{3/5}/(m_1+m_2)^{1/5}$.
-\item The symmetric mass ratio, $\eta=(m_1m_2)/(m_1+m_2)^2$.
-\item The 3-dimensional spin vectors of each components, $\mathbf{S}_1,\mathbf{S}_2$.
-\item An arrival time, $\tau_\oplus$, taken here to be the time of arrival at the geocenter.
-\item The inclination angle, $\iota$.
-\item The polarization angle, $\psi$.
-\item The coalescence phase, $\phi_c$.
-\end{itemize}
-In this work we restrict ourselves to non-spinng sources, where $\mathbf{S}_1=\mathbf{S}_2=0$.
-Removing nuisance parameters requires either replacing them with their measured values or marginalizing over them.
-The chirp mass is typically well measured by the detection pipeline, so we use the maximum likelihood value, $\hat{\mathcal{M}}$, for it.
-We marginalize over the other variables using a uniform prior, i.e.
-\begin{equation}
-p(\alpha,\delta,D_L|d) = \int\, d\tau_\oplus p_U(\tau_\oplus) d\phi_c p_U(\phi_c) d\phi p_U(\psi)  d\cos(\iota) p_U(\cos(\iota)) p(\alpha,\delta,D_L,\tau_\oplus,\phi_c,\psi,\cos(\iota)|d),
-\end{equation}
-where $p_U(x)$ denotes a uniform distribution in $x$.
 

### 2013-05-21 ms.tex
Fix typo
Note advantage of simpler likelihood

@@
 It differs from existing techniques is several important ways.
-In the first place, we fix the intrinsic parameters to their maximum likelihood estiamtes, as provided by the detection pipeline.
+In the first place, we fix the intrinsic parameters to their maximum likelihood estimates, as provided by the detection pipeline.
 This reduces the dimensionality of the parameter we need to search.
 In addition, the technique takes as its input the maximum likelihood estiamtes of the extrinsic parameters, instead of the $h(t)$ time series.
+This makes the likelihood itself much faster to evaluate.
 Finally, instead of of using \ac{MCMC} or some similar method for statistical sampling, we make use of a deterministic quadrature scheme.

### 2013-05-21 ms.tex
Some cleanup

@@
 %
-In addition to being an efficient source of \acp{GW}, the neutron star may be tidally shredded before merging, providing fuel for an electromagnetic counterpart.
+In addition to being an efficient source of \acp{GW}, the \ac{NS} may be tidally shredded before merging, providing fuel for an electromagnetic counterpart.
 The short-lived accretion flow may power a collimated relativistic jet, resulting in a short \ac{GRB} and an X-ray afterglow if it is aligned with the line of sight.
@@
 Binary eccentricity is omitted as an intrinisc parameter because \acp{GW} circularize the orbit long before the inspiral enters \ac{LIGO}'s frequency range of $\sim$10\nobreakdashes--1000~kHz.
-Tidal deformabilities of the components are omitted because tidal effects have a negligible effect in the inspiral phase.
-In \ac{GW} detection efforts, especially those focused on \ac{NSNS} systems, the component spins $\mathbf{S}_1$ and $\mathbf{S}_2$ are often assumed to be aligned with the system's total angular momentum and condensed to a single scalar parameter $\chi$, or even neglected entirely, $\mathbf{S}_1 = \mathbf{S}_2 = 0$.
-Without loss of generality, in this work we will consider only zero spin.
+Tidal deformabilities of the \acp{NS} are omitted because the effect of tides on the inspiral would only be detectable by an Einstein Telescope-class \ac{GW} observatory~\cite{PhysRevD.81.123016}.
+Furthermore, in \ac{GW} detection efforts, especially those focused on \ac{NSNS} systems, the component spins $\mathbf{S}_1$ and $\mathbf{S}_2$ are often assumed to be aligned with the system's total angular momentum and condensed to a single scalar parameter $\chi$, or even neglected entirely, $\mathbf{S}_1 = \mathbf{S}_2 = 0$.
+In this work we will consider only zero spin.
 

### 2013-05-21 ms.tex
Add reference for circularization

@@
 %
-\marginpar{Citations needed. Do we want to mention these omitted parameters?}%
-%
-Binary eccentricity is omitted as an intrinisc parameter because \acp{GW} circularize the orbit long before the inspiral enters \ac{LIGO}'s frequency range of $\sim$10\nobreakdashes--1000~kHz.
+Binary eccentricity is omitted as an intrinisc parameter because tidal interaction of the progenitor stars~\cite{0004-637X-572-1-407} and later \ac{GW} emission~\cite{PhysRev.136.B1224} circularize the orbit long before the inspiral enters \ac{LIGO}'s frequency range of $\sim$10\nobreakdashes--1000~kHz.
 Tidal deformabilities of the \acp{NS} are omitted because the effect of tides on the inspiral would only be detectable by an Einstein Telescope-class \ac{GW} observatory~\cite{PhysRevD.81.123016}.

### 2013-05-21 ms.tex
Add signal model

@@
 A laser interferometer senses \acp{GW} by transducing differential changes in arm length to a digitized photocurrent time series.
-In the angular frequency domain, a single detector's observation $X(\omega)$ is
+In the angular frequency domain, a single detector's observation $X_i(\omega)$ is
 %
 \begin{equation}\label{eq:signal-model}
-	X (\omega) = H (\omega; \boldsymbol\theta) + N (\omega),
+	X_i (\omega) = H_i (\omega; \boldsymbol\theta) + N_i (\omega),
 \end{equation}
 %
-where $H (\omega; \boldsymbol\theta)$ is the \ac{GW} signal, given a parameter vector $\boldsymbol\theta$ that describes the \ac{GW} source, and $N (\omega)$ is the detector's \ac{WSS} Gaussian noise with \ac{PSD} $S(\omega)$.
-The likelihood, or the probability of obtaining the observation $X(\omega)$ conditioned on the value of $\boldsymbol\theta$, is Gaussian, with the log\nobreakdashes-likelihood proportional to:
+where $H_i (\omega; \boldsymbol\theta)$ is the \ac{GW} signal receied by detector $i$, given a parameter vector $\boldsymbol\theta$ that describes the \ac{GW} source, and $N_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with \ac{PSD} $S_i(\omega)$.
+We shall denote the combined observation from a network of detectors as $\mathbf X (\omega) \equiv \{X_i (\omega)\}_i$.
+
+The likelihood, or the probability of obtaining the observation $\mathbf X$ conditioned on the value of $\boldsymbol\theta$, is Gaussian, with the log\nobreakdashes-likelihood proportional to:
 %
 \begin{equation}\label{eq:gaussian-likelihood}
-	\mathcal{L}(X; \boldsymbol\theta) = p(X | \boldsymbol\theta)
+	\mathcal{L}(\mathbf X; \boldsymbol\theta) = \prod_i p(X_i | \boldsymbol\theta)
 		\propto \exp \left[
-		- \frac{1}{2} \int_0^\infty \frac{\left|X (\omega)
-			- H(\omega; \boldsymbol\theta) \right|^2}{S(\omega)} \, \mathrm{d}\omega
+		- \frac{1}{2} \sum_i \int_0^\infty \frac{\left|X_i (\omega)
+			- H_i(\omega; \boldsymbol\theta) \right|^2}{S_i(\omega)} \, \mathrm{d}\omega
 	\right].
@@
                 \end{array}
-                \quad \right\} \textnormal{extrinsic} \\
+                \quad \right\} \textnormal{extrinsic parameters}, \boldsymbol\theta_\mathrm{ex} \\
                 \left.
@@
                 \end{array}
-                \quad \right\} \textnormal{intrinsic}
+                \quad \right\} \textnormal{intrinsic parameters}, \boldsymbol\theta_\mathrm{in}
             \end{array}
@@
 
+Assuming circular orbits and no spin precession, we can write the \ac{GW} signal in each detector as a linear combination of two basis waveforms $H_0$ and $H_{\pi/2}$, such that the coefficients depend only on the extrinsic parameters $\boldsymbol\theta_\mathrm{ex}$ and the basis waveforms depend only on the intrinsic parameters $\boldsymbol\theta_\mathrm{in}$,
+%
+\begin{eqnarray}
+	H_i(\omega; \boldsymbol\theta) &=& e^{-i\omega t_\oplus}\frac{D_{1,i}}{D_\mathrm{L}}
+	\left(\begin{array}{cc}
+		F_{+,i}(h_i, \delta) \\
+		F_{\times,i}(h_i, \delta)
+	\end{array}\right)^\intercal
+	\left(\begin{array}{cc}
+		\cos{2\psi} & -\sin{2\psi} \\
+		\sin{2\psi} & \cos{2\psi}
+	\end{array}\right) \\ &\cdot&
+	\left(\begin{array}{cc}
+		\frac{1}{2}\left(1+\cos^2{\iota}\right) & 0 \\
+		0 & \cos{\iota}
+	\end{array}\right)
+	\left(\begin{array}{cc}
+		\cos{2\phi_c} & -\sin{2\phi_c} \\
+		\sin{2\phi_c} & \cos{2\phi_c}
+	\end{array}\right) \\ &\cdot&
+	\left(\begin{array}{cc}
+		H_0(\omega; \boldsymbol\theta_\mathrm{in}) \\
+		H_{\pi/2}(\omega; \boldsymbol\theta_\mathrm{in})
+	\end{array}\right).
+\end{eqnarray}
+%
+This linear combination matches the conventions used in \cite{PhysRevD.83.084002}, and is the frequency-domain equivalent of their Equation~(2.8).
+The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless antenna factors, and depend on detector $i$'s orientation and the declination and local hour angle $h_i$ of the source, and are derived in~\cite{PhysRevD.63.042003}.
+
+In the Bayesian framework, the parameters are inferred from the data by forming the posterior distribution, $p(\boldsymbol\theta|\mathbf X)$, which describes the probability of the parameters given the observations.
+Bayes' rule relates the likelihood $p(\mathbf X|\boldsymbol\theta)$ to the posterior $p(\boldsymbol\theta|\mathbf X)$,
+%
+\begin{equation}\label{bayes}
+p(\boldsymbol\theta|\mathbf X) = \frac{p(\mathbf X|\boldsymbol\theta) p(\boldsymbol\theta)}{p(\mathbf X)},
+\end{equation}
+%
+introducing the prior distribution $p(\boldsymbol\theta)$ which encapsulates previous information about the paramters (for example, arising from earlier observations or from known physical bounds on the parameters), and the evidence $p(\mathbf X)$ which can be thought of as a normalization factor or as describing the parsimoniousness of the model.
+
 Removing nuisance parameters requires either replacing them with their measured values or marginalizing over them.
 The chirp mass is typically well measured by the detection pipeline, so we use the maximum likelihood value, $\hat{\mathcal{M}}$, for it.
-We marginalize over the other variables using a uniform prior, i.e.
+We marginalize over the other variables using a uniform prior, i.e.,
+%
 \begin{equation}
@@
 \end{equation}
+%
 where $p_U(x)$ denotes a uniform distribution in $x$.
-
-The quantity of interest for any type of parameter estimation is the so-called {\it posterior} distribution, $p(\boldsymbol\theta|d)$, that describes the probability of obtaining the parameters, $\boldsymbol\theta$, given the data, $d$.
-It is given by Bayes' theorem
-\begin{equation}\label{bayes}
-p(\boldsymbol\theta|d) = \frac{p(\boldsymbol\theta)p(d|\boldsymbol\theta)}{p(d)},
-\end{equation}
-where $p(\boldsymbol\theta)$ is the {\it prior} distribution describing previous knowledge of the paramters, $\boldsymbol\theta$, $p(d|\boldsymbol\theta)$ is the {\it likelihood} of obtaining the data, $d$, for the parameters, $\boldsymbol\theta$, and $p(d)$ is known as the {\it evidence}.
-Procedures for computing the posterior typically rely on Markov chain Monte Carlo methods or nested sampling (something else?).
-In this manuscript we present an alternative approach, taylored specifically to providing low-latency estimates of the sky location of and distance to the source.
 

### 2013-05-21 ms.tex
Improve wording

@@
 This linear combination matches the conventions used in \cite{PhysRevD.83.084002}, and is the frequency-domain equivalent of their Equation~(2.8).
-The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless antenna factors, and depend on detector $i$'s orientation and the declination and local hour angle $h_i$ of the source, and are derived in~\cite{PhysRevD.63.042003}.
+The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, and depend on the orientation of detector $i$ as well as the declination and local hour angle $h_i$ of the source, and are derived in~\cite{PhysRevD.63.042003}.
 

### 2013-05-21 ms.tex
Add note about horizon distance
Clean up signal model

@@
 
-Assuming circular orbits and no spin precession, we can write the \ac{GW} signal in each detector as a linear combination of two basis waveforms $H_0$ and $H_{\pi/2}$, such that the coefficients depend only on the extrinsic parameters $\boldsymbol\theta_\mathrm{ex}$ and the basis waveforms depend only on the intrinsic parameters $\boldsymbol\theta_\mathrm{in}$,
+Assuming circular orbits and no spin precession, we can write the \ac{GW} signal in each detector as a linear combination of two basis waveforms $H_0$ and $H_{\pi/2}$.
+For nonprecessing systems, $H_0$ and $H_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the sine and cosine functions, being nearly orthogonal and out of phase by ${\pi/2}$ at all frequencies.
+In other words, $H_{\pi/2}(\omega) = i H_0(\omega)$ for all $\omega$.
+This linear combination isolates all dependence on the extrinsic parameters $\boldsymbol\theta_\mathrm{ex}$ into the coefficients and all dependence on the intrinsic parameters $\boldsymbol\theta_\mathrm{in}$ into the basis waveform,
 %
 \begin{eqnarray}
-	H_i(\omega; \boldsymbol\theta) &=& e^{-i\omega t_\oplus}\frac{D_{1,i}}{D_\mathrm{L}}
+	\fl H_i(\omega; \boldsymbol\theta) &=& e^{-i\omega t_\oplus}\frac{D_{1,i}}{D_\mathrm{L}}
 	\left(\begin{array}{cc}
@@
 		\sin{2\psi} & \cos{2\psi}
-	\end{array}\right) \\ &\cdot&
+	\end{array}\right)
 	\left(\begin{array}{cc}
@@
 		0 & \cos{\iota}
-	\end{array}\right)
+	\end{array}\right) \nonumber\\\fl &\cdot&
 	\left(\begin{array}{cc}
@@
 		\sin{2\phi_c} & \cos{2\phi_c}
-	\end{array}\right) \\ &\cdot&
+	\end{array}\right)
 	\left(\begin{array}{cc}
-		H_0(\omega; \boldsymbol\theta_\mathrm{in}) \\
-		H_{\pi/2}(\omega; \boldsymbol\theta_\mathrm{in})
-	\end{array}\right).
+		1 \\
+		\imath
+	\end{array}\right) H_0(\omega; \boldsymbol\theta_\mathrm{in}).
 \end{eqnarray}
@@
 The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, and depend on the orientation of detector $i$ as well as the declination and local hour angle $h_i$ of the source, and are derived in~\cite{PhysRevD.63.042003}.
+The quantity $D_{1,i}$ is a fiducial horizon distance for detector $i$.
 

### 2013-05-22 ms.tex
Define horizon distance
Add succinct signal model
Add matched filter description

@@
 \DeclareMathOperator{\std}{std}
+\DeclareMathOperator*{\argmin}{\arg\!\min}
+\DeclareMathOperator*{\argmax}{\arg\!\max}
 
@@
 The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, and depend on the orientation of detector $i$ as well as the declination and local hour angle $h_i$ of the source, and are derived in~\cite{PhysRevD.63.042003}.
-The quantity $D_{1,i}$ is a fiducial horizon distance for detector $i$.
+The quantity $D_{1,i}$ is a fiducial horizon distance for detector $i$,
+%
+\begin{equation}\label{eq:horizon}
+D_{1,i} = 1 / \sigma_i, \qquad {\sigma_i}^2 = \int_0^\infty \frac{\left|H_0(\omega; \mathbf \theta)\right|^2}{S_i(\omega)} \,\mathrm{d}\omega.
+\end{equation}
+%
+More succinctly, we can write the signal received by detector $i$ in terms of an amplitude $\rho_i$, a phase $\gamma_i$, and a time delay $\tau_i$,
+%
+\begin{equation}\label{eq:signal-model}
+	H_i (\omega; \boldsymbol\theta) = \frac{\rho_i}{\sigma_i} e^{\imath (\gamma_i - \omega \tau_i)} H_0(\omega; \boldsymbol\theta_\mathrm{in}).
+\end{equation}
+
+Detection of \acp{GW} from \ac{CBC} events consists of convolving the \dots explain matched filtering here \dots
+
+\begin{eqnarray}
+\hat{\boldsymbol\theta}_i, \{\hat\tau_i\}_i &=& \argmax_{\boldsymbol\theta_\mathrm{in}, \{\hat\tau_i\}_i} \sum_i \frac{1}{{\sigma_i}^2} \left| \int_0^\infty \frac{{H_0}^*(\omega; \boldsymbol\theta_\mathrm{in}) X_i(\omega) e^{-\imath \omega \tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right|^2 \\
+\hat\rho_i &=& \frac{1}{\sigma_i} \left| \int_0^\infty \frac{{H_0}^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) X_i(\omega) e^{-\imath \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right| \\
+\hat\gamma_i &=& \arg \int_0^\infty \frac{{H_0}^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) X_i(\omega) e^{-\imath \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega
+\end{eqnarray}
 

### 2013-05-22 ms.tex
Replace \imath with just plain i

--- 2013-05-21 ms.tex
@@
 		1 \\
-		\imath
+		i
 	\end{array}\right) H_0(\omega; \boldsymbol\theta_\mathrm{in}).
@@
 The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, and depend on the orientation of detector $i$ as well as the declination and local hour angle $h_i$ of the source, and are derived in~\cite{PhysRevD.63.042003}.
-The quantity $D_{1,i}$ is a fiducial horizon distance for detector $i$.
+The quantity $D_{1,i}$ is a fiducial horizon distance for detector $i$,
+%
+\begin{equation}\label{eq:horizon}
+D_{1,i} = 1 / \sigma_i, \qquad {\sigma_i}^2 = \int_0^\infty \frac{\left|H_0(\omega; \mathbf \theta)\right|^2}{S_i(\omega)} \,\mathrm{d}\omega.
+\end{equation}
+%
+More succinctly, we can write the signal received by detector $i$ in terms of an amplitude $\rho_i$, a phase $\gamma_i$, and a time delay $\tau_i$,
+%
+\begin{equation}\label{eq:signal-model}
+	H_i (\omega; \boldsymbol\theta) = \frac{\rho_i}{\sigma_i} e^{i (\gamma_i - \omega \tau_i)} H_0(\omega; \boldsymbol\theta_\mathrm{in}).
+\end{equation}
+
+Detection of \acp{GW} from \ac{CBC} events consists of convolving the \dots explain matched filtering here \dots
+
+\begin{eqnarray}
+\hat{\boldsymbol\theta}_i, \{\hat\tau_i\}_i &=& \argmax_{\boldsymbol\theta_\mathrm{in}, \{\hat\tau_i\}_i} \sum_i \frac{1}{{\sigma_i}^2} \left| \int_0^\infty \frac{{H_0}^*(\omega; \boldsymbol\theta_\mathrm{in}) X_i(\omega) e^{-i \omega \tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right|^2 \\
+\hat\rho_i &=& \frac{1}{\sigma_i} \left| \int_0^\infty \frac{{H_0}^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) X_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right| \\
+\hat\gamma_i &=& \arg \int_0^\infty \frac{{H_0}^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) X_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega
+\end{eqnarray}
 
@@
 \begin{equation}\label{eq:signal-model}
-	X (\omega) = H (\omega, \rho, \gamma, \tau) + N (\omega) = \frac{\rho}{\sigma} e^{\imath (\gamma - \omega \tau)} U (\omega) + N (\omega),
+	X (\omega) = H (\omega, \rho, \gamma, \tau) + N (\omega) = \frac{\rho}{\sigma} e^{i (\gamma - \omega \tau)} U (\omega) + N (\omega),
 \end{equation}
@@
 \begin{equation}
-    \mathcal{L}(z; \rho, \gamma) \propto \exp \left[ -\frac{1}{2} \left|z - e^{\imath \gamma} \rho \right|^2 \right]
+    \mathcal{L}(z; \rho, \gamma) \propto \exp \left[ -\frac{1}{2} \left|z - e^{i \gamma} \rho \right|^2 \right]
 \end{equation}

### 2013-05-22 ms.tex
Add explanation of detection process

@@
 %
-\begin{eqnarray}
+\begin{eqnarray}\label{eq:full-signal-model}
 	\fl H_i(\omega; \boldsymbol\theta) &=& e^{-i\omega t_\oplus}\frac{D_{1,i}}{D_\mathrm{L}}
@@
 %
-More succinctly, we can write the signal received by detector $i$ in terms of an amplitude $\rho_i$, a phase $\gamma_i$, and a time delay $\tau_i$,
+More succinctly, we can write the signal received by detector $i$ in terms of direct observables, the amplitude $\rho_i$, phase $\gamma_i$, and time delay $\tau_i$ on arrival at detector $i$,
 %
@@
 
-Detection of \acp{GW} from \ac{CBC} events consists of convolving the \dots explain matched filtering here \dots
-
+The prevailing technique for detection of \acp{GW} from \acp{CBC} is to realize a \ac{MLE} from \Eref{eq:gaussian-likelihood} with the above signal model.
+This takes the form of a bank of matched filters, or template waveforms that are continually convolved with the incoming data stream.
+The matched filter or maximum likelihood estimates are given by
+%
 \begin{eqnarray}
@@
 \hat\rho_i &=& \frac{1}{\sigma_i} \left| \int_0^\infty \frac{{H_0}^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) X_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right| \\
-\hat\gamma_i &=& \arg \int_0^\infty \frac{{H_0}^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) X_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega
+\hat\gamma_i &=& \arg \int_0^\infty \frac{{H_0}^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) X_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega.
 \end{eqnarray}

### 2013-05-22 ms.tex
Break into subsections
Clean up some acronyms

@@
 It differs from existing techniques is several important ways.
-In the first place, we fix the intrinsic parameters to their maximum likelihood estimates, as provided by the detection pipeline.
+In the first place, we fix the intrinsic parameters to their \ac{ML} estimates, as provided by the detection pipeline.
 This reduces the dimensionality of the parameter we need to search.
-In addition, the technique takes as its input the maximum likelihood estiamtes of the extrinsic parameters, instead of the $h(t)$ time series.
+In addition, the technique takes as its input the \ac{ML} estimates of the extrinsic parameters, instead of the $h(t)$ time series.
 This makes the likelihood itself much faster to evaluate.
@@
 \end{equation}
-%
+
+\subsection{Detection, and matched filter bank as a \ac{MLE}}
+
 More succinctly, we can write the signal received by detector $i$ in terms of direct observables, the amplitude $\rho_i$, phase $\gamma_i$, and time delay $\tau_i$ on arrival at detector $i$,
@@
 This takes the form of a bank of matched filters, or template waveforms that are continually convolved with the incoming data stream.
-The matched filter or maximum likelihood estimates are given by
+The matched filter or \ac{ML} estimates are given by
 %
@@
 \end{eqnarray}
+
+\subsection{Bayesian probability and parameter estimation}
 
@@
 Removing nuisance parameters requires either replacing them with their measured values or marginalizing over them.
-The chirp mass is typically well measured by the detection pipeline, so we use the maximum likelihood value, $\hat{\mathcal{M}}$, for it.
+The chirp mass is typically well measured by the detection pipeline, so we use the \ac{ML} value, $\hat{\mathcal{M}}$, for it.
 We marginalize over the other variables using a uniform prior, i.e.,
@@
 Our interest here, however, is in obtaining accurate sky location results as quickly as possible.
-To that end, we instead base our posteriors on the maximum likelihood estimates of the signal-to-noise ratio (SNR), $\hat{\rho}_I$, the time-of-arrival (TOA), $\hat{\tau}_I$, and coalescence phase, $\hat{\phi}_I$, measured in each detector by the detection pipeline.
+To that end, we instead base our posteriors on the \ac{ML} estimates of the signal-to-noise ratio (SNR), $\hat{\rho}_I$, the time-of-arrival (TOA), $\hat{\tau}_I$, and coalescence phase, $\hat{\phi}_I$, measured in each detector by the detection pipeline.
 Thus the sky location posterior can be computed without any re-filterning of the data.

### 2013-05-22 ms.tex
Described S6PE prior

@@
 
-Removing nuisance parameters requires either replacing them with their measured values or marginalizing over them.
-The chirp mass is typically well measured by the detection pipeline, so we use the \ac{ML} value, $\hat{\mathcal{M}}$, for it.
-We marginalize over the other variables using a uniform prior, i.e.,
-%
-\begin{equation}
-p(\alpha,\delta,D_L|d) = \int\, d\tau_\oplus p_U(\tau_\oplus) d\phi_c p_U(\phi_c) d\phi p_U(\psi)  d\cos(\iota) p_U(\cos(\iota)) p(\alpha,\delta,D_L,\tau_\oplus,\phi_c,\psi,\cos(\iota)|d),
-\end{equation}
-%
-where $p_U(x)$ denotes a uniform distribution in $x$.
-
-For typical parameter estimation studies, the data, $d$, is taken to be the strain time series in each detector, $h_I(t)$, where $I$ indexes the detectors.
-Our interest here, however, is in obtaining accurate sky location results as quickly as possible.
-To that end, we instead base our posteriors on the \ac{ML} estimates of the signal-to-noise ratio (SNR), $\hat{\rho}_I$, the time-of-arrival (TOA), $\hat{\tau}_I$, and coalescence phase, $\hat{\phi}_I$, measured in each detector by the detection pipeline.
-Thus the sky location posterior can be computed without any re-filterning of the data.
+The choice of prior is open to one's astrophysical preconceptions, but during \ac{S6} when \ac{LIGO}'s Bayesian \ac{CBC} parameter estimation pipelines were pioneered, the prior was taken to be isotropic in sky location and binary orientation, and uniform in volume, arrival time, and the component masses~\cite{2013arXiv1304.1775T}.
+Put briefly, the prior was uniform in all of $(\alpha, \sin\delta, {D_\mathrm{L}}^3, t_\oplus, \cos\iota, \psi, \phi_c, m_1, m_2)$.
 
@@
 
-Prior is uniform in $\mathbf{n}$, $t_\oplus$, x, and $\psi$. For distance, can choose either a prior that is uniform in $\log D_\mathrm{L}$ or uniform in volume.
+$t_\oplus$, $m_1$, $m_2$, ${D_\mathrm{L}}^3$ (volume)
+Prior is uniform in $\mathbf{n}$, $t_\oplus$, x, and $\psi$. For distance, can choose either a prior that is uniform in $\log D_\mathrm{L}$ or uniform in volume. \cite{2013arXiv1304.1775T}
 

### 2013-05-22 ms.tex
Add marginal posterior

@@
 Put briefly, the prior was uniform in all of $(\alpha, \sin\delta, {D_\mathrm{L}}^3, t_\oplus, \cos\iota, \psi, \phi_c, m_1, m_2)$.
+
+In Bayesian inference, although it is often easy to write down the likelihood or even the full posterior in closed form, usually one is interested in only a subset $\boldsymbol\beta$ of all of the model's parameters, the others $\boldsymbol\lambda$ being nuisance parameters.
+In this case, we integrate away the nuisance parameters, forming the marginal posterior
+%
+\begin{equation}\label{eq:marginal-posterior}
+	p(\boldsymbol\beta|\mathbf X) = \int \frac{p(\mathbf X|\boldsymbol\beta,\boldsymbol\lambda) p(\boldsymbol\beta,\boldsymbol\lambda)}{p(\mathbf X)} \,\mathrm{d}\boldsymbol\lambda
+\end{equation}
+%
+with $\boldsymbol\theta = (\boldsymbol\beta, \boldsymbol\lambda)$. For instance, for the purpose of locating a \ac{GW} source on the sky, all parameters but $(\alpha, \delta)$ are nuisance parameters.
 

### 2013-05-22 ms.tex
Use lowercase symbols for signals

@@
 \begin{equation}\label{eq:signal-model}
-	X_i (\omega) = H_i (\omega; \boldsymbol\theta) + N_i (\omega),
+	x_i (\omega) = h_i (\omega; \boldsymbol\theta) + n_i (\omega),
 \end{equation}
 %
-where $H_i (\omega; \boldsymbol\theta)$ is the \ac{GW} signal receied by detector $i$, given a parameter vector $\boldsymbol\theta$ that describes the \ac{GW} source, and $N_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with \ac{PSD} $S_i(\omega)$.
-We shall denote the combined observation from a network of detectors as $\mathbf X (\omega) \equiv \{X_i (\omega)\}_i$.
+where $h_i (\omega; \boldsymbol\theta)$ is the \ac{GW} signal receied by detector $i$, given a parameter vector $\boldsymbol\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with \ac{PSD} $S_i(\omega)$.
+We shall denote the combined observation from a network of detectors as $\mathbf x (\omega) \equiv \{x_i (\omega)\}_i$.
 
-The likelihood, or the probability of obtaining the observation $\mathbf X$ conditioned on the value of $\boldsymbol\theta$, is Gaussian, with the log\nobreakdashes-likelihood proportional to:
+The likelihood, or the probability of obtaining the observation $\mathbf x$ conditioned on the value of $\boldsymbol\theta$, is Gaussian, with the log\nobreakdashes-likelihood proportional to:
 %
 \begin{equation}\label{eq:gaussian-likelihood}
-	\mathcal{L}(\mathbf X; \boldsymbol\theta) = \prod_i p(X_i | \boldsymbol\theta)
+	\mathcal{L}(\mathbf x; \boldsymbol\theta) = \prod_i p(x_i | \boldsymbol\theta)
 		\propto \exp \left[
-		- \frac{1}{2} \sum_i \int_0^\infty \frac{\left|X_i (\omega)
-			- H_i(\omega; \boldsymbol\theta) \right|^2}{S_i(\omega)} \, \mathrm{d}\omega
+		- \frac{1}{2} \sum_i \int_0^\infty \frac{\left|x_i (\omega)
+			- h_i(\omega; \boldsymbol\theta) \right|^2}{S_i(\omega)} \, \mathrm{d}\omega
 	\right].
@@
 
-Assuming circular orbits and no spin precession, we can write the \ac{GW} signal in each detector as a linear combination of two basis waveforms $H_0$ and $H_{\pi/2}$.
-For nonprecessing systems, $H_0$ and $H_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the sine and cosine functions, being nearly orthogonal and out of phase by ${\pi/2}$ at all frequencies.
-In other words, $H_{\pi/2}(\omega) = i H_0(\omega)$ for all $\omega$.
+Assuming circular orbits and no spin precession, we can write the \ac{GW} signal in each detector as a linear combination of two basis waveforms $h_0$ and $h_{\pi/2}$.
+For nonprecessing systems, $h_0$ and $h_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the sine and cosine functions, being nearly orthogonal and out of phase by ${\pi/2}$ at all frequencies.
+In other words, $h_{\pi/2}(\omega) = i h_0(\omega)$ for all $\omega$.
 This linear combination isolates all dependence on the extrinsic parameters $\boldsymbol\theta_\mathrm{ex}$ into the coefficients and all dependence on the intrinsic parameters $\boldsymbol\theta_\mathrm{in}$ into the basis waveform,
@@
 \begin{eqnarray}\label{eq:full-signal-model}
-	\fl H_i(\omega; \boldsymbol\theta) &=& e^{-i\omega t_\oplus}\frac{D_{1,i}}{D_\mathrm{L}}
+	\fl h_i(\omega; \boldsymbol\theta) &=& e^{-i\omega t_\oplus}\frac{D_{1,i}}{D_\mathrm{L}}
 	\left(\begin{array}{cc}
@@
 		i
-	\end{array}\right) H_0(\omega; \boldsymbol\theta_\mathrm{in}).
+	\end{array}\right) h_0(\omega; \boldsymbol\theta_\mathrm{in}).
 \end{eqnarray}
@@
 \begin{equation}\label{eq:horizon}
-D_{1,i} = 1 / \sigma_i, \qquad {\sigma_i}^2 = \int_0^\infty \frac{\left|H_0(\omega; \mathbf \theta)\right|^2}{S_i(\omega)} \,\mathrm{d}\omega.
+D_{1,i} = 1 / \sigma_i, \qquad {\sigma_i}^2 = \int_0^\infty \frac{\left|h_0(\omega; \mathbf \theta)\right|^2}{S_i(\omega)} \,\mathrm{d}\omega.
 \end{equation}
@@
 \begin{equation}\label{eq:signal-model}
-	H_i (\omega; \boldsymbol\theta) = \frac{\rho_i}{\sigma_i} e^{i (\gamma_i - \omega \tau_i)} H_0(\omega; \boldsymbol\theta_\mathrm{in}).
+	h_i (\omega; \boldsymbol\theta) = \frac{\rho_i}{\sigma_i} e^{i (\gamma_i - \omega \tau_i)} h_0(\omega; \boldsymbol\theta_\mathrm{in}).
 \end{equation}
@@
 \begin{eqnarray}
-\hat{\boldsymbol\theta}_i, \{\hat\tau_i\}_i &=& \argmax_{\boldsymbol\theta_\mathrm{in}, \{\hat\tau_i\}_i} \sum_i \frac{1}{{\sigma_i}^2} \left| \int_0^\infty \frac{{H_0}^*(\omega; \boldsymbol\theta_\mathrm{in}) X_i(\omega) e^{-i \omega \tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right|^2 \\
-\hat\rho_i &=& \frac{1}{\sigma_i} \left| \int_0^\infty \frac{{H_0}^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) X_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right| \\
-\hat\gamma_i &=& \arg \int_0^\infty \frac{{H_0}^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) X_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega.
+\hat{\boldsymbol\theta}_i, \{\hat\tau_i\}_i &=& \argmax_{\boldsymbol\theta_\mathrm{in}, \{\hat\tau_i\}_i} \sum_i \frac{1}{{\sigma_i}^2} \left| \int_0^\infty \frac{{h_0}^*(\omega; \boldsymbol\theta_\mathrm{in}) x_i(\omega) e^{-i \omega \tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right|^2 \\
+\hat\rho_i &=& \frac{1}{\sigma_i} \left| \int_0^\infty \frac{{h_0}^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) x_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right| \\
+\hat\gamma_i &=& \arg \int_0^\infty \frac{{h_0}^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) x_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega.
 \end{eqnarray}
@@
 
-In the Bayesian framework, the parameters are inferred from the data by forming the posterior distribution, $p(\boldsymbol\theta|\mathbf X)$, which describes the probability of the parameters given the observations.
-Bayes' rule relates the likelihood $p(\mathbf X|\boldsymbol\theta)$ to the posterior $p(\boldsymbol\theta|\mathbf X)$,
+In the Bayesian framework, the parameters are inferred from the data by forming the posterior distribution, $p(\boldsymbol\theta|\mathbf x)$, which describes the probability of the parameters given the observations.
+Bayes' rule relates the likelihood $p(\mathbf x|\boldsymbol\theta)$ to the posterior $p(\boldsymbol\theta|\mathbf x)$,
 %
 \begin{equation}\label{bayes}
-p(\boldsymbol\theta|\mathbf X) = \frac{p(\mathbf X|\boldsymbol\theta) p(\boldsymbol\theta)}{p(\mathbf X)},
+p(\boldsymbol\theta|\mathbf x) = \frac{p(\mathbf x|\boldsymbol\theta) p(\boldsymbol\theta)}{p(\mathbf x)},
 \end{equation}
 %
-introducing the prior distribution $p(\boldsymbol\theta)$ which encapsulates previous information about the paramters (for example, arising from earlier observations or from known physical bounds on the parameters), and the evidence $p(\mathbf X)$ which can be thought of as a normalization factor or as describing the parsimoniousness of the model.
+introducing the prior distribution $p(\boldsymbol\theta)$ which encapsulates previous information about the paramters (for example, arising from earlier observations or from known physical bounds on the parameters), and the evidence $p(\mathbf x)$ which can be thought of as a normalization factor or as describing the parsimoniousness of the model.
 
@@
 \begin{equation}\label{eq:marginal-posterior}
-	p(\boldsymbol\beta|\mathbf X) = \int \frac{p(\mathbf X|\boldsymbol\beta,\boldsymbol\lambda) p(\boldsymbol\beta,\boldsymbol\lambda)}{p(\mathbf X)} \,\mathrm{d}\boldsymbol\lambda
+	p(\boldsymbol\beta|\mathbf x) = \int \frac{p(\mathbf x|\boldsymbol\beta,\boldsymbol\lambda) p(\boldsymbol\beta,\boldsymbol\lambda)}{p(\mathbf x)} \,\mathrm{d}\boldsymbol\lambda
 \end{equation}
@@
 \begin{equation}\label{eq:signal-model}
-	X (\omega) = H (\omega, \rho, \gamma, \tau) + N (\omega) = \frac{\rho}{\sigma} e^{i (\gamma - \omega \tau)} U (\omega) + N (\omega),
+	x (\omega) = h (\omega, \rho, \gamma, \tau) + n (\omega) = \frac{\rho}{\sigma} e^{i (\gamma - \omega \tau)} u (\omega) + n (\omega),
 \end{equation}
 %
-where $H (\omega)$ is the signal content, $N (\omega)$ is the detector's \ac{WSS} Gaussian noise with \ac{PSD} $S(\omega)$, $U (\omega)$ is the frequency-domain post-Newtonian inspiral signal, $\rho$ is the signal's amplitude, $\gamma$ is a phase angle the encodes the relative strength of two \ac{GW} quadratures, and $\tau$ is the time of arrival relative to an arbitrary time base.
+where $h (\omega)$ is the signal content, $n (\omega)$ is the detector's \ac{WSS} Gaussian noise with \ac{PSD} $S(\omega)$, $u (\omega)$ is the frequency-domain post-Newtonian inspiral signal, $\rho$ is the signal's amplitude, $\gamma$ is a phase angle the encodes the relative strength of two \ac{GW} quadratures, and $\tau$ is the time of arrival relative to an arbitrary time base.
 The quantity $\sigma$ is a normalization constant defined as the \ac{SNR} of the \ac{GW} source at a fiducial distance, here taken to be $D_\mathrm{fiducial} = 1$~Mpc,
@@
 \begin{equation}\label{eq:sigma}
-    \sigma^2 = \int_0^\infty \frac{\left| U (\omega)\right|^2}{S(\omega)} \, \mathrm{d}\omega.
+    \sigma^2 = \int_0^\infty \frac{\left| u (\omega)\right|^2}{S(\omega)} \, \mathrm{d}\omega.
 \end{equation}
@@
 The distance at which an optimally oriented \ac{GW} source would produce \ac{SNR}=1 is defined as $D_1 = D_\mathrm{fiducial} / \sigma$.
-The likelihood, or the probability of obtaining the observation $X(\omega)$ conditioned on the value of the parameter vector $\boldsymbol\theta = (\rho, \gamma, \tau)$, is Gaussian, with the log-likelihood proportional to:
+The likelihood, or the probability of obtaining the observation $x(\omega)$ conditioned on the value of the parameter vector $\boldsymbol\theta = (\rho, \gamma, \tau)$, is Gaussian, with the log-likelihood proportional to:
 %
 \begin{equation}\label{eq:gaussian-likelihood}
-	\mathcal{L}(X; \boldsymbol\theta) = p(X | \boldsymbol\theta)
+	\mathcal{L}(x; \boldsymbol\theta) = p(x | \boldsymbol\theta)
 		\propto \exp \left[
-		- \frac{1}{2} \int_0^\infty \frac{\left|X (\omega)
-			- H(\omega; \boldsymbol\theta) \right|^2}{S(\omega)} \, \mathrm{d}\omega
+		- \frac{1}{2} \int_0^\infty \frac{\left|x (\omega)
+			- h(\omega; \boldsymbol\theta) \right|^2}{S(\omega)} \, \mathrm{d}\omega
 	\right].
@@
 		-\frac{\partial^2 \log
-			\mathcal{L}(X ; \boldsymbol\theta)}
+			\mathcal{L}(x ; \boldsymbol\theta)}
 			{\partial \theta_j \theta_k}
@@
 	\mathcal{I}_{jk} = \int_0^\infty \Re \left[
-        \left(\frac{\partial H}{\partial \theta_j}\right)^*
-        \left(\frac{\partial H}{\partial \theta_k}\right)
+        \left(\frac{\partial h}{\partial \theta_j}\right)^*
+        \left(\frac{\partial h}{\partial \theta_k}\right)
 	\right] \frac{1}{S(\omega)} \, \mathrm{d}\omega.
@@
     \overline{\omega^k} =
-        \left[ \int_0^\infty \frac{|U (\omega)|^2}{S(\omega)} \omega^k \, \mathrm{d}\omega \right]
-        \left[ \int_0^\infty \frac{|U (\omega)|^2}{S(\omega)} \, \mathrm{d}\omega \right]^{-1},
+        \left[ \int_0^\infty \frac{|u (\omega)|^2}{S(\omega)} \omega^k \, \mathrm{d}\omega \right]
+        \left[ \int_0^\infty \frac{|u (\omega)|^2}{S(\omega)} \, \mathrm{d}\omega \right]^{-1},
 \end{equation}

### 2013-05-22 ms.tex
Drop hour angle

@@
 	\left(\begin{array}{cc}
-		F_{+,i}(h_i, \delta) \\
-		F_{\times,i}(h_i, \delta)
+		F_{+,i}(\alpha, \delta, t_\oplus) \\
+		F_{\times,i}(\alpha, \delta, t_\oplus)
 	\end{array}\right)^\intercal
@@
 This linear combination matches the conventions used in \cite{PhysRevD.83.084002}, and is the frequency-domain equivalent of their Equation~(2.8).
-The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, and depend on the orientation of detector $i$ as well as the declination and local hour angle $h_i$ of the source, and are derived in~\cite{PhysRevD.63.042003}.
+The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, and depend on the orientation of detector $i$ as well as the sky location and sidereal time of the event, and are derived in~\cite{PhysRevD.63.042003}.
 The quantity $D_{1,i}$ is a fiducial horizon distance for detector $i$,

### 2013-05-22 ms.tex
Use h, not h_0

@@
 Assuming circular orbits and no spin precession, we can write the \ac{GW} signal in each detector as a linear combination of two basis waveforms $h_0$ and $h_{\pi/2}$.
-For nonprecessing systems, $h_0$ and $h_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the sine and cosine functions, being nearly orthogonal and out of phase by ${\pi/2}$ at all frequencies.
+For nonprecessing systems, $h = h_0$ and $h_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the sine and cosine functions, being nearly orthogonal and out of phase by ${\pi/2}$ at all frequencies.
 In other words, $h_{\pi/2}(\omega) = i h_0(\omega)$ for all $\omega$.
@@
 		i
-	\end{array}\right) h_0(\omega; \boldsymbol\theta_\mathrm{in}).
+	\end{array}\right) h(\omega; \boldsymbol\theta_\mathrm{in}).
 \end{eqnarray}
@@
 \begin{equation}\label{eq:horizon}
-D_{1,i} = 1 / \sigma_i, \qquad {\sigma_i}^2 = \int_0^\infty \frac{\left|h_0(\omega; \mathbf \theta)\right|^2}{S_i(\omega)} \,\mathrm{d}\omega.
+D_{1,i} = 1 / \sigma_i, \qquad {\sigma_i}^2 = \int_0^\infty \frac{\left|h(\omega; \mathbf \theta)\right|^2}{S_i(\omega)} \,\mathrm{d}\omega.
 \end{equation}
@@
 \begin{equation}\label{eq:signal-model}
-	h_i (\omega; \boldsymbol\theta) = \frac{\rho_i}{\sigma_i} e^{i (\gamma_i - \omega \tau_i)} h_0(\omega; \boldsymbol\theta_\mathrm{in}).
+	h_i (\omega; \boldsymbol\theta) = \frac{\rho_i}{\sigma_i} e^{i (\gamma_i - \omega \tau_i)} h(\omega; \boldsymbol\theta_\mathrm{in}).
 \end{equation}
@@
 \begin{eqnarray}
-\hat{\boldsymbol\theta}_i, \{\hat\tau_i\}_i &=& \argmax_{\boldsymbol\theta_\mathrm{in}, \{\hat\tau_i\}_i} \sum_i \frac{1}{{\sigma_i}^2} \left| \int_0^\infty \frac{{h_0}^*(\omega; \boldsymbol\theta_\mathrm{in}) x_i(\omega) e^{-i \omega \tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right|^2 \\
-\hat\rho_i &=& \frac{1}{\sigma_i} \left| \int_0^\infty \frac{{h_0}^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) x_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right| \\
-\hat\gamma_i &=& \arg \int_0^\infty \frac{{h_0}^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) x_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega.
+\hat{\boldsymbol\theta}_i, \{\hat\tau_i\}_i &=& \argmax_{\boldsymbol\theta_\mathrm{in}, \{\hat\tau_i\}_i} \sum_i \frac{1}{{\sigma_i}^2} \left| \int_0^\infty \frac{h^*(\omega; \boldsymbol\theta_\mathrm{in}) x_i(\omega) e^{-i \omega \tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right|^2 \\
+\hat\rho_i &=& \frac{1}{\sigma_i} \left| \int_0^\infty \frac{h^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) x_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right| \\
+\hat\gamma_i &=& \arg \int_0^\infty \frac{h^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) x_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega,
 \end{eqnarray}
+%
+where the $^*$ denotes a complex conjugate.
 

### 2013-05-23 ms.tex
Add paragraph
Add some pomp to introduction

@@
 .
+It is unique in that in bridges the detection and parameter estimation of \ac{GW} signals, two tasks that have until now involved very different numerical methods and time scales.
+We expect that \ac{BAYESTAR} will take on a key role in observing \ac{CBC} events in both \ac{GW} and optical channels during the Advanced \ac{LIGO} era.
 
@@
 with $\boldsymbol\theta = (\boldsymbol\beta, \boldsymbol\lambda)$. For instance, for the purpose of locating a \ac{GW} source on the sky, all parameters but $(\alpha, \delta)$ are nuisance parameters.
+
+Computing the marginal posterior almost always involves many-dimensional and often ill-behaved integrals.\marginpar{More about \ac{MCMC} methods here.}
 

### 2013-05-23 ms.tex
More descriptive section title
Oops, missed lowercase signal variable

@@
 
-\section{Setup}
+\section{Theory of \ac{GW} detection and parameter estimation}
 
@@
 
-In the frequency domain, the strain $X (\omega)$ observed by a detector is a linear combination of two orthogonal \ac{GW} quadratures and additive noise,
+In the frequency domain, the strain $x (\omega)$ observed by a detector is a linear combination of two orthogonal \ac{GW} quadratures and additive noise,
 %

### 2013-05-23 ms.tex
Clean up Fisher matrix section

@@
 
-In the frequency domain, the strain $x (\omega)$ observed by a detector is a linear combination of two orthogonal \ac{GW} quadratures and additive noise,
+In order to estimate the \ac{ML} estimator accuracy for a matched filter, we will momentarily consider the likelihood for a single detector, and surpress the intrinsic degrees of freedom:
 %
-\begin{equation}\label{eq:signal-model}
-	x (\omega) = h (\omega, \rho, \gamma, \tau) + n (\omega) = \frac{\rho}{\sigma} e^{i (\gamma - \omega \tau)} u (\omega) + n (\omega),
+\begin{equation}\label{eq:gaussian-likelihood}
+	\fl\mathcal{L}(x_i; \rho_i, \gamma_i, \tau_i) = p(x_i | \rho_i, \gamma_i, \tau_i)
+		\propto \exp \left[
+		- \frac{1}{2} \int_0^\infty \frac{\left|x_i (\omega)
+			- h_i(\omega; \rho_i, \gamma_i, \tau_i) \right|^2}{S_i(\omega)} \, \mathrm{d}\omega
+	\right],
 \end{equation}
 %
-where $h (\omega)$ is the signal content, $n (\omega)$ is the detector's \ac{WSS} Gaussian noise with \ac{PSD} $S(\omega)$, $u (\omega)$ is the frequency-domain post-Newtonian inspiral signal, $\rho$ is the signal's amplitude, $\gamma$ is a phase angle the encodes the relative strength of two \ac{GW} quadratures, and $\tau$ is the time of arrival relative to an arbitrary time base.
-The quantity $\sigma$ is a normalization constant defined as the \ac{SNR} of the \ac{GW} source at a fiducial distance, here taken to be $D_\mathrm{fiducial} = 1$~Mpc,
-%
-\begin{equation}\label{eq:sigma}
-    \sigma^2 = \int_0^\infty \frac{\left| u (\omega)\right|^2}{S(\omega)} \, \mathrm{d}\omega.
-\end{equation}
-%
-The distance at which an optimally oriented \ac{GW} source would produce \ac{SNR}=1 is defined as $D_1 = D_\mathrm{fiducial} / \sigma$.
-The likelihood, or the probability of obtaining the observation $x(\omega)$ conditioned on the value of the parameter vector $\boldsymbol\theta = (\rho, \gamma, \tau)$, is Gaussian, with the log-likelihood proportional to:
-%
-\begin{equation}\label{eq:gaussian-likelihood}
-	\mathcal{L}(x; \boldsymbol\theta) = p(x | \boldsymbol\theta)
-		\propto \exp \left[
-		- \frac{1}{2} \int_0^\infty \frac{\left|x (\omega)
-			- h(\omega; \boldsymbol\theta) \right|^2}{S(\omega)} \, \mathrm{d}\omega
-	\right].
-\end{equation}
+with $h_i(\omega; \rho_i, \gamma_i, \tau_i)$ given by \Eref{eq:signal-model}.
 
-The Fisher information matrix for a measurement $\tilde x$ described by the unknown parameter vector $\boldsymbol{\theta}$ is
+The Fisher information matrix for a measurement $x$ described by the unknown parameter vector $\boldsymbol{\theta}$ is
 %
@@
 		-\frac{\partial^2 \log
-			\mathcal{L}(x ; \boldsymbol\theta)}
+			\mathcal{L}(x_i ; \boldsymbol\theta)}
 			{\partial \theta_j \theta_k}
@@
 	\mathcal{I}_{jk} = \int_0^\infty \Re \left[
-        \left(\frac{\partial h}{\partial \theta_j}\right)^*
-        \left(\frac{\partial h}{\partial \theta_k}\right)
-	\right] \frac{1}{S(\omega)} \, \mathrm{d}\omega.
+        \left(\frac{\partial h_i}{\partial \theta_j}\right)^*
+        \left(\frac{\partial h_i}{\partial \theta_k}\right)
+	\right] \frac{1}{S_i(\omega)} \, \mathrm{d}\omega.
 \end{equation}
 %
-This form is more useful because it involves only first derivatives, and of the signal $\tilde h (\omega)$ rather than the entire observation $\tilde x (\omega)$.
+This form is more useful because it involves only first derivatives, and of the signal $h (\omega)$ rather than the entire observation $x (\omega)$.
 In terms of the $k$th \ac{SNR}-weighted moment of angular frequency,
@@
 \begin{equation}\label{eq:angular-frequency-moments}
-    \overline{\omega^k} =
-        \left[ \int_0^\infty \frac{|u (\omega)|^2}{S(\omega)} \omega^k \, \mathrm{d}\omega \right]
-        \left[ \int_0^\infty \frac{|u (\omega)|^2}{S(\omega)} \, \mathrm{d}\omega \right]^{-1},
+    {\overline{\omega^k}}_i =
+        \left[ \int_0^\infty \frac{|h (\omega)|^2}{S_i(\omega)} \omega^k \, \mathrm{d}\omega \right]
+        \left[ \int_0^\infty \frac{|h (\omega)|^2}{S_i(\omega)} \, \mathrm{d}\omega \right]^{-1},
 \end{equation}
@@
 	\mathcal{I} = \bordermatrix{
-        ~ & \rho & \gamma & \tau \cr
-        \rho & 1 & 0 & 0 \cr
-        \gamma & 0 & \rho^2 & -\rho^2 \overline{\omega} \cr
-        \tau & 0 & -\rho^2 \overline{\omega} & \rho^2 \overline{\omega^2}
+        ~ & \rho_i & \gamma_i & \tau_i \cr
+        \rho_i & 1 & 0 & 0 \cr
+        \gamma_i & 0 & {\rho_i}^2 & -{\rho_i}^2 {\overline{\omega}}_i \cr
+        \tau_i & 0 & -{\rho_i}^2 {\overline{\omega}}_i & {\rho_i}^2 {\overline{\omega^2}}_i
     }.
@@
 
-The \ac{CRLB} says that, element by element,
+The \ac{CRLB} says that, element by element,%
+%
+\marginpar{Element by element? I don't think that is right.}
 %
@@
         \begin{array}{c}
-            \hat{\rho} - \rho \\
-            \hat{\gamma} - \gamma \\
-            \hat{\tau} - \tau
+            \hat{\rho}_i - \rho_i \\
+            \hat{\gamma}_i - \gamma_i \\
+            \hat{\tau}_i - \tau_i
         \end{array}
@@
             1 & 0 & 0 \\
-            0 & \rho^2 \overline{\omega^2}/{\omega_\mathrm{rms}}^2 & \rho^2 \overline{\omega}/{\omega_\mathrm{rms}}^2 \\
-            0 & \rho^2 \overline{\omega}/{\omega_\mathrm{rms}}^2 & \rho^2/{\omega_\mathrm{rms}}^2
+            0 & {\rho_i}^2 {\overline{\omega^2}}_i/{\omega_{\mathrm{rms},i}}^2 & {\rho_i}^2 {\overline{\omega}}_i/{\omega_{\mathrm{rms},i}}^2 \\
+            0 & {\rho_i}^2 {\overline{\omega}}_i/{\omega_{\mathrm{rms},i}}^2 & {\rho_i}^2/{\omega_{\mathrm{rms},i}}^2
         \end{array}
@@
 %
-where a hat $\hat{\phantom{\rho}}$ denotes the estimate of a parameter and ${\omega_\mathrm{rms}}^2 = \overline{\omega^2} - \overline{\omega}^2$.
+where a hat $\hat{\phantom{\rho}}$ denotes the estimate of a parameter and ${\omega_{\mathrm{rms},i}}^2 = {\overline{\omega^2}}_i - {{\overline{\omega}}_i}^2$.
 Reading off the $\tau \tau$ element of the covariance matrix reproduces the timing accuracy in Equation~(24) of \cite{fairhurst:2009},
@@
 \begin{equation}\label{eq:timing-crlb}
-    \std \left(\hat{\tau} - \tau \right) \geq \sqrt{\left(\mathcal{I}^{-1}\right)_{\tau\tau}} = \frac{\rho}{\omega_\mathrm{rms}}.
+    \std \left(\hat{\tau}_i - \tau_i \right) \geq \sqrt{\left(\mathcal{I}^{-1}\right)_{\tau\tau}} = \frac{\rho_i}{\omega_{\mathrm{rms},i}}.
 \end{equation}
@@
     \mathcal{I}' = \bordermatrix{
-        ~ & \rho & \gamma & \gamma_\tau \cr
-        \rho & 1 & 0 & 0 \cr
-        \gamma & 0 & \rho^2 & -\rho^2\frac{\overline{\omega}}{\sqrt{\overline{\omega^2}}} \cr
-        \gamma_\tau & 0 & -\rho^2\frac{\overline{\omega}}{\sqrt{\overline{\omega^2}}} & \rho^2
+        ~ & \rho_i & \gamma_i & \gamma_{\tau,i} \cr
+        \rho_i & 1 & 0 & 0 \cr
+        \gamma_i & 0 & {\rho_i}^2 & -{\rho_i}^2\frac{{\overline{\omega}}_i}{\sqrt{{\overline{\omega^2}}_i}} \cr
+        \gamma_{\tau,i} & 0 & -{\rho_i}^2\frac{{\overline{\omega}}_i}{\sqrt{{\overline{\omega^2}}_i}} & {\rho_i}^2
     }.
@@
     \mathcal{I}'' = \bordermatrix{
-        ~ & \rho & \gamma_+ & \gamma_- \cr
-        \rho & 1 & 0 & 0 \cr
-        \gamma_+ & 0 & \left(1 - \frac{\overline{\omega}}{\sqrt{\overline{\omega^2}}}\right)\rho^2 & 0 \cr
-        \gamma_- & 0 & 0 & \left(1 + \frac{\overline{\omega}}{\sqrt{\overline{\omega^2}}}\right)\rho^2
+        ~ & \rho_i & \gamma_{+,i} & \gamma_{-,i} \cr
+        \rho_i & 1 & 0 & 0 \cr
+        \gamma_{+,i} & 0 & \left(1 - \frac{\overline{\omega}_i}{\sqrt{\overline{\omega^2}_i}}\right){\rho_i}^2 & 0 \cr
+        \gamma_{-,i} & 0 & 0 & \left(1 + \frac{\overline{\omega}_i}{\sqrt{\overline{\omega^2}_i}}\right){\rho_i}^2
     }.

### 2013-05-23 ms.tex
Remove some redundant definitions
Introduce u = cos(inclination)

@@
                 t_\oplus \\
-                \iota \\
+                u \\
                 \psi \\
@@
                     arrival time at geocenter \\
-                    inclination angle \\
+                    $\cos(\textrm{inclination angle})$ \\
                     polarization angle \\
@@
 	\left(\begin{array}{cc}
-		\frac{1}{2}\left(1+\cos^2{\iota}\right) & 0 \\
-		0 & \cos{\iota}
+		\frac{1}{2}\left(1+u^2\right) & 0 \\
+		0 & u
 	\end{array}\right) \nonumber\\\fl &\cdot&
@@
 The choice of prior is open to one's astrophysical preconceptions, but during \ac{S6} when \ac{LIGO}'s Bayesian \ac{CBC} parameter estimation pipelines were pioneered, the prior was taken to be isotropic in sky location and binary orientation, and uniform in volume, arrival time, and the component masses~\cite{2013arXiv1304.1775T}.
-Put briefly, the prior was uniform in all of $(\alpha, \sin\delta, {D_\mathrm{L}}^3, t_\oplus, \cos\iota, \psi, \phi_c, m_1, m_2)$.
+Put briefly, the prior was uniform in all of $(\alpha, \sin\delta, {D_\mathrm{L}}^3, t_\oplus, u, \psi, \phi_c, m_1, m_2)$.
 
@@
 %
-The \ac{SNR}-only factor depends only on the \ac{SNR} estimates $\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N$, but it depends on the unknown sky location, luminosity distance $D_\mathrm{L}$, cosine-inclination $x = \cos \iota$, and polarization angle $\psi$.
+The \ac{SNR}-only factor depends only on the \ac{SNR} estimates $\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N$, but it depends on the unknown sky location, luminosity distance $D_\mathrm{L}$, cosine-inclination $u$, and polarization angle $\psi$.
 %
@@
     \mathcal{L}_\mathrm{SNR} \propto \exp \left[
-        -\frac{1}{2} \sum_{i=1}^N \left(\hat{\rho}_i - \rho_i(\mathbf{n}, D_\mathrm{L}, x, \psi) \right)^2
+        -\frac{1}{2} \sum_{i=1}^N \left(\hat{\rho}_i - \rho_i(\mathbf{n}, D_\mathrm{L}, u, \psi) \right)^2
     \right]
@@
     {\rho_i}^2 &=
-        \frac{1}{8} \left(\frac{D_{1,i}}{D_\mathrm{L}}\right)^2 \left[
+        \frac{1}{8} \left(\frac{D_{1,i}}{D_\mathrm{L}}\right)^2 \left\{
             ( {F_{+,i}}^2 + {F_{\times,i}}^2 \right)
-            \left(1 + 6 x^2 + x^4 \right) \\
-            &+ \left(1 - x^2 \right) \left( {F_{+,i}}^2 - {F_{\times,i}}^2 \right) \cos {2\psi}
-            + 2 \left( {F_{+,i}} {F_{\times,i}} ) \sin {2\psi}
-        \right],
+            \left(1 + 6 u^2 + u^4 \right) \\
+            &+ \left(1 - u^2 \right)^2 \left[\left( {F_{+,i}}^2 - {F_{\times,i}}^2 \right) \cos {2\psi}
+            + 2 \left( {F_{+,i}} {F_{\times,i}} ) \sin {2\psi} \right]
+        \right\}.
 \end{eqnarray*}
-%
-where $F_{+,i}$ and $F_{\times,i}$ are the `$+$' and `$\times$' antenna factors of detector $i$, themselves functions of sky location $\mathbf{n}$.
 

### 2013-05-24 ms.tex
Added more info about prior

@@
 
-\section{Method}
+\section{The \ac{BAYESTAR} algorithm}
 
@@
 %
-It is easily shown that
+It is easily shown from \Eref{eq:full-signal-model} that
 %
-\begin{eqnarray*}
+\begin{eqnarray}\label{eq:snr-model}
     {\rho_i}^2 &=
@@
             ( {F_{+,i}}^2 + {F_{\times,i}}^2 \right)
-            \left(1 + 6 u^2 + u^4 \right) \\
-            &+ \left(1 - u^2 \right)^2 \left[\left( {F_{+,i}}^2 - {F_{\times,i}}^2 \right) \cos {2\psi}
-            + 2 \left( {F_{+,i}} {F_{\times,i}} ) \sin {2\psi} \right]
+            \left(1 + 6 u^2 + u^4 \right) \nonumber\\
+            &+ \left(1 - u^2 \right)^2 \left[\left( {F_{+,i}}^2 - {F_{\times,i}}^2 \right) \cos {4\psi}
+            + 2 \left( {F_{+,i}} {F_{\times,i}} ) \sin {4\psi} \right]
         \right\}.
-\end{eqnarray*}
+\end{eqnarray}
 
@@
 
-$t_\oplus$, $m_1$, $m_2$, ${D_\mathrm{L}}^3$ (volume)
-Prior is uniform in $\mathbf{n}$, $t_\oplus$, x, and $\psi$. For distance, can choose either a prior that is uniform in $\log D_\mathrm{L}$ or uniform in volume. \cite{2013arXiv1304.1775T}
+We have no explicit prior on the intrinsic paramters; in our analysis they are fixed at their \ac{ML} estimates, $\hat{\boldsymbol\theta}$.
+The distance prior is given by a user-selected power of distance, ${D_\mathrm{L}}^m$, where $m=2$ for a prior that is uniform in volume, and $m=-1$ for a scale-invariant prior that is uniform in $\log D_\mathrm{L}$.
+The priors are bounded as follows:
+%
+\marginpar{Need to discuss $D_{\mathrm{L},\mathrm{min}}$, $D_{\mathrm{L},\mathrm{max}}$, and how they are selected automatically if the user does not provide them.}
+%
+\begin{eqnarray*}
+	D_\mathrm{L} &\in& [D_{\mathrm{L},\mathrm{min}}, D_{\mathrm{L},\mathrm{max}}] \\
+	t_\oplus &\in& [-\infty, \infty] \\
+	u &\in& [0, 1] \\
+	\psi &\in& [0, \pi/2)
+\end{eqnarray*}
+%
+The prior does not permit negative values of $u$ because only even powers of $u$ appear in \Eref{eq:snr-model}.
+The prior only permits $\psi$ to run to $pi/2$ because only trigonometric functions of $4\psi$ appear in \Eref{eq:snr-model}.
 
@@
 
-The marginalization over $D_\mathrm{L}$, $x$, and $\psi$ must be done numerically.
-The outermost integration is over x and $\psi$, and is done with Newton--Cotes by evaluating $\mathcal{L}_\mathrm{SNR}$ on a regular lattice in $x$ and $\psi$.
-This integration scheme is first order in $x$ and second order in $\psi$.
+The marginalization over $D_\mathrm{L}$, $u$, and $\psi$ must be done numerically.
+The outermost integration is over $u$ and $\psi$, and is done with Newton\nobreakdashes--Cotes by evaluating $\mathcal{L}_\mathrm{SNR}$ on a regular lattice in $u$ and $\psi$.
+This integration scheme is first order in $u$ and second order in $\psi$.
 

### 2013-05-24 ms.tex
Add full posterior
Clean up some marginpar formatting

@@
 This likelihood also describes the distribution of the length $\rho$ and angle $\gamma$ of a vector whose Cartesian components are perturbed with normally distributed errors.
-For this problem, $\rho$ and $\gamma$ are uncorrelated, $\rho^2$ being noncentrally $\chi^2$\nobreakdashes-disributed with two degrees of freedom and $\gamma$ being von~Mises\nobreakdashes-distributed.
+For this problem, $\rho$ and $\gamma$ are uncorrelated, $\rho^2$ being noncentrally $\chi^2$\nobreakdashes-disributed with two degrees of freedom and $\gamma$ being von~Mises\nobreakdashes-distributed.%
 %
@@
 
-\marginpar{What about if we suppress $\gamma$ or $\rho$? Do those cases teach us anything?}
+\marginpar{What about if we suppress $\gamma$ or $\rho$? Do those cases teach us anything?}%
 %
@@
 
-There are many ways to realize distribution function with these properties, but they tend to be difficult to work with.
+There are many ways to realize distribution function with these properties, but they tend to be difficult to work with.%
 %
-\marginpar{Give some examples.}
+\marginpar{Give some examples.}%
 %
@@
         \hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
-        \mathbf{n}, t_\oplus, D_\mathrm{L}, x, \psi) \\
+        \mathbf{n}, t_\oplus, D_\mathrm{L}, u, \psi) \\
     &= \mathcal{L}_\mathrm{TOA}(
@@
         \mathcal{L}_\mathrm{SNR}(\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
-         \mathbf{n}, D_\mathrm{L}, x, \psi
+         \mathbf{n}, D_\mathrm{L}, u, \psi
     ).
@@
 
-\subsection{Marginalization}
+\subsection{Marginalization over $t_\oplus$}
 
@@
 \end{equation}
+
+\subsection{Posterior}
+
+\begin{eqnarray}
+    p(
+		\mathbf{n} |
+        \hat{\tau}_1, \hat{\tau}_2, &\dots, \hat{\tau}_N,
+        \hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N) \nonumber\\
+    &= p(\mathbf{n} |
+        \hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N) \nonumber\\
+        &\times
+        p(\mathbf{n} |
+		\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N
+    )
+\end{eqnarray}
+%
+where
+%
+\begin{equation*}
+	\fl
+    p(\mathbf{n} |
+	\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N)
+	\propto
+	\int_0^\frac{\pi}{2}
+	\int_0^1
+	\int_{D_{\mathrm{L},\mathrm{min}}}^{D_{\mathrm{L},\mathrm{max}}}
+	\exp \left[
+	        -\frac{1}{2} \sum_{i=1}^N \left(\hat{\rho}_i - \rho_i(\mathbf{n}, D_\mathrm{L}, u, \psi) \right)^2
+	    \right]
+	{D_\mathrm{L}}^m
+	\,\mathrm{d}D_\mathrm{L}
+	\,\mathrm{d}u
+	\,\mathrm{d}\psi.
+\end{equation*}
 

### 2013-06-19 ms.tex
Typo
Fix marginpar formatting
Added description of Gaussian quadratures breakpoints

@@
 %
-\marginpar{Check that number.}
+\marginpar{Check that number.}%
 %
@@
 The prior does not permit negative values of $u$ because only even powers of $u$ appear in \Eref{eq:snr-model}.
-The prior only permits $\psi$ to run to $pi/2$ because only trigonometric functions of $4\psi$ appear in \Eref{eq:snr-model}.
+The prior only permits $\psi$ to run to $\pi/2$ because only trigonometric functions of $4\psi$ appear in \Eref{eq:snr-model}.
 
@@
 
+Before evaluating each inner integral over $D_\mathrm{L}$, the quantities $A=-\frac{1}{2} \sum_i {\rho_i}^2$ and $B=\sum_i \rho_i \hat\rho_i$ are precalculated so that the integrand can be written as
+%
+\begin{equation*}
+	\exp(A {D_\mathrm{L}}^{-2} + B D_\mathrm{L}) {D_\mathrm{L}}^m.
+\end{equation*}
+%
+Then, we see that the likelihood $\exp(A {D_\mathrm{L}}^{-2} + B D_\mathrm{L})$ is maximized when $D_\mathrm{L} = D_{\mathrm{L},0} = -2A/B$.
+The likelihood takes on a factor $\eta$ of its maximum value when
+%
+\begin{equation}
+	D_\mathrm{L} = D_{\mathrm{L},\pm} = \left(\frac{1}{D_{\mathrm{L},0}} \pm \sqrt{\frac{\log\eta}{A}}\right)^{-1}.
+\end{equation}
+
 \begin{figure}

### 2013-06-19 radial_integrand.py
Radial integrand plot now matches formulae in paper

@@ -9,42 +9,34 @@ from matplotlib import pyplot as plt
 from mpl_toolkits.axes_grid1.inset_locator import BboxConnectorPatch
 from matplotlib.transforms import Bbox, TransformedBbox, blended_transform_factory
 from matplotlib import ticker
-from scipy import special
 import numpy as np
 import sys
 
-xmin = 0.
-xmax = 1.
-a = 0.1
-a2 = a * a
-y1 = 0.01
-y2 = 0.005
-break1 = (a - a2 * np.sqrt(-2*np.log(y1))) / (1 + 2 * a2 * np.log(y1))
-if a * np.sqrt(-2 * np.log(y2)) < 1:
-    break2 = (a + a2 * np.sqrt(-2*np.log(y1))) / (1 + 2 * a2 * np.log(y1))
-else:
-    break2 = (np.log(y2) - np.sqrt(np.log(y1) * np.log(y2))) / (np.sqrt(-2*np.log(y2)) * (np.log(y2) - np.log(y1)))
+# Input parameters
+rmin = 0
+rmax = 18
+A = -0.5 * 3 * 49
+B = 3 * 30
+eta = 0.01
+m = 2
 
-breakpoints = [xmin]
-if xmin < break1 < xmax:
-    breakpoints.append(break1)
-if xmin < break2 < xmax:
-    breakpoints.append(break2)
-breakpoints.append(xmax)
+# Breakpoints
+middle_breakpoint = -2*A/B
+lower_breakpoint = 1/(1/middle_breakpoint + np.sqrt(np.log(eta) / A))
+upper_breakpoint = 1/(1/middle_breakpoint - np.sqrt(np.log(eta) / A))
+breakpoints = filter(lambda r: rmin <= r <= rmax,
+    [rmin, lower_breakpoint, middle_breakpoint, upper_breakpoint, rmax])
 
-colors = 'rgb'
+colors = 'crgb'
 fig = plt.figure(figsize=(5, 5))
 ax0 = plt.axes([0.2, 0.5, 0.6, 0.4])
 
-for i, (break1, break2) in enumerate(zip(breakpoints[:-1], breakpoints[1:])):
-    x = np.linspace(break1, break2, 100)
-    I0e = special.i0e(1 / a2)
-    lognorm = - 1 / (2 * a2) - np.log(I0e)
-    I0e = special.i0e(1 / (a * x))
-    integrand = I0e * np.exp(-1/(2*x*x) + lognorm + 1 / (a * x))
[…]

### 2013-06-20 ms.tex
Fine-tune some wording in introduction
Actually, I don't know what order this integration scheme is
Fixed typo in equation
Add full-size parens
Finished describing adaptive integration scheme

@@
 In the first place, we fix the intrinsic parameters to their \ac{ML} estimates, as provided by the detection pipeline.
-This reduces the dimensionality of the parameter we need to search.
+This reduces the dimensionality of the parameter space we need to sample.
 In addition, the technique takes as its input the \ac{ML} estimates of the extrinsic parameters, instead of the $h(t)$ time series.
@@
 The outermost integration is over $u$ and $\psi$, and is done with Newton\nobreakdashes--Cotes by evaluating $\mathcal{L}_\mathrm{SNR}$ on a regular lattice in $u$ and $\psi$.
-This integration scheme is first order in $u$ and second order in $\psi$.
-
-\marginpar{Explain initial integration breakpoints and how they are calculated. Cite \texttt{gsl\_integrate\_qagp}.}%
-%
-The innermost integration is over $D_\mathrm{L}$ and is performed using adaptive Gaussian quadrature, but with a twist that speeds up convergence.
-
-Before evaluating each inner integral over $D_\mathrm{L}$, the quantities $A=-\frac{1}{2} \sum_i {\rho_i}^2$ and $B=\sum_i \rho_i \hat\rho_i$ are precalculated so that the integrand can be written as
-%
-\begin{equation*}
-	\exp(A {D_\mathrm{L}}^{-2} + B D_\mathrm{L}) {D_\mathrm{L}}^m.
-\end{equation*}
-%
-Then, we see that the likelihood $\exp(A {D_\mathrm{L}}^{-2} + B D_\mathrm{L})$ is maximized when $D_\mathrm{L} = D_{\mathrm{L},0} = -2A/B$.
-The likelihood takes on a factor $\eta$ of its maximum value when
-%
-\begin{equation}
-	D_\mathrm{L} = D_{\mathrm{L},\pm} = \left(\frac{1}{D_{\mathrm{L},0}} \pm \sqrt{\frac{\log\eta}{A}}\right)^{-1}.
-\end{equation}
 
 \begin{figure}
-    \caption{\label{fig:radial_integrand}Illustration of initial subdivisions for radial integration scheme.}
+    \caption{\label{fig:radial_integrand}Illustration of initial subdivisions for distance integration scheme.
+Distance increases from left to right.
+In the print version, the left\nobreakdashes-hand tail, the left\nobreakdashes- and right\nobreakdashes-hand sides of the maximum likelihood peak, and the right\nobreakdashes-hand tail, are colored cyan, red, green, and blue, respectively.}
     \begin{center}
@@
 \end{figure}
+
+The innermost integration is over $D_\mathrm{L}$ and is performed using adaptive Gaussian quadrature, but with a twist that speeds up convergence.
+As in many parameter estimation problems, the integrand may be sharply peaked, and any generic adaptive integration scheme may take a long time to find the peak.
+Before evaluating each inner integral over $D_\mathrm{L}$, the quantities $A=-\frac{1}{2} \sum_i {\rho_i}^2$ and $B=\sum_i \rho_i \hat\rho_i$ are precalculated so that the integrand can be written as
+%
+\begin{equation*}
+	\exp\left(\frac{A}{{D_\mathrm{L}}^2} + \frac{B}{D_\mathrm{L}}\right) {D_\mathrm{L}}^m.
+\end{equation*}
+%
+Then, we see that the likelihood $\exp(A {D_\mathrm{L}}^{-2} + B {D_\mathrm{L}}^{-1})$ is maximized when $D_\mathrm{L} = D_{\mathrm{L},0} = -2A/B$.
+The likelihood takes on a factor $\eta$ (say, $\eta=0.01$) of its maximum value when
+%
+\begin{equation}
+	D_\mathrm{L} = D_{\mathrm{L},\pm} = \left(\frac{1}{D_{\mathrm{L},0}} \mp \sqrt{\frac{\log\eta}{A}}\right)^{-1}.
+\end{equation}
+%
+We have now identified up to five breakpoints that partition the distance integrand into up to four intervals with quantitatively distinct behavior.
+These intervals are depicted in Figure~\ref{fig:radial_integrand} with distance increasing from left to right.
+There is a left\nobreakdashes-hand or small distance tail in which the integrand is small and monotically increasing, a left\nobreakdashes- and right\nobreakdashes-hand side of the maximum likelihood peak, and a right\nobreakdashes-hand tail in which the integrand is small and monotonically decreasing.
+These breakpoints are:
+%
+\begin{equation}
+    D_{\mathrm{L},\mathrm{break}} = \{ D_\mathrm{L} \in
+    \left\{
+    \begin{array}{c}
+    D_{\mathrm{L},\mathrm{min}} \\
+    D_{\mathrm{L},-} \\
+    D_{\mathrm{L},0} \\
+    D_{\mathrm{L},+} \\
+    D_{\mathrm{L},\mathrm{max}}
+    \end{array}
+    \right\} :
+    D_{\mathrm{L},\mathrm{min}} \leq D_\mathrm{L}
+    \leq D_{\mathrm{L},\mathrm{max}}\}.
+\end{equation}
+%
+We use these breakpoints as initial subdivision in an adaptive Gaussian quadrature algorithm.
+Specifically, we use the GNU Scientific Library's \verb|gsl_integrate_qagp|\footnote{\url{http://www.gnu.org/software/gsl/manual/html_node/QAGP-adaptive-integration-with-known-singular-points.html}} function, which is designed to integrate functions with known singular points (though, of course, or integrand has no singular points).
+This function estimates the integral over each subdivision and each interval's contribution to the total error, then subdivides the interval whose error contribution is largest.
+Subdivisions continue until a fixed total fractional error is reached.
+In this way, most integrand evaluations are expended on the most important distance interval, whether that happens to be the tails (when the posterior is dominated by the prior) or the peak (when the posterior is dominated by the observations).
 

### 2013-06-21 ms.tex
Margin note about mis-definition of horizon distance
Described hierarchical refinement

@@
 The quantity $D_{1,i}$ is a fiducial horizon distance for detector $i$,
+%
+\marginpar{This isn't what we usually call the horizon distance; the horizon distance is where face-on inspirals produce SNR=8.}%
 %
@@
 
-The overall arrival time $t_\oplus$ can be marginalized out analytically, giving the \ac{TOA}-only marginal posterior
+The overall arrival time $t_\oplus$ can be marginalized out analytically, giving the \ac{TOA}\nobreakdashes-only marginal posterior
 %
-\begin{equation}
+\begin{equation}\label{eq:toa-posterior}
     p(\mathbf{n} | \hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N) \propto \exp\left[
@@
 \end{equation}
+
+\subsection{Adaptive resolution}
+
+The \ac{TOA}\nobreakdashes-only posterior in \Eref{eq:toa-posterior} is numerically trivial to evaluate because it involves no numerical integrals.
+However, as we will see in the next section, the \ac{SNR}\nobreakdashes-only posterior is numerically expensive.
+
+We evaluate the \ac{TOA}\nobreakdashes-only posterior at the centers of \ac{HEALPix} pixels at the lowest possible resolution.
+Then, we increase the resolution until 99.99\% of the \ac{TOA}\nobreakdashes-only posterior is contained within at least 3072 pixels.
+The numerically expensive \ac{SNR} integrals are done only for the pixels that comprise the 99.99th percentile at this resolution.
+All other pixels are set to zero.
 
@@
 Distance increases from left to right.
-In the print version, the left\nobreakdashes-hand tail, the left\nobreakdashes- and right\nobreakdashes-hand sides of the maximum likelihood peak, and the right\nobreakdashes-hand tail, are colored cyan, red, green, and blue, respectively.}
+In the color version, the left\nobreakdashes-hand tail, the left\nobreakdashes- and right\nobreakdashes-hand sides of the maximum likelihood peak, and the right\nobreakdashes-hand tail, are colored cyan, red, green, and blue, respectively.}
     \begin{center}
@@
 
-\section{Hierarchical refinement}
-
-\marginpar{Talk about adaptive selection of HEALPix resolution and of pixels on which to evaluate \ac{SNR} marginal posterior.
-Something to think about: this is logically easier to explain last, but chronologically makes more sense first.}
-
 \ack Source code for \ac{BAYESTAR} is available on the \ac{LIGO} \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/bayestar.html}.
 
-Some of the results in this paper have been derived using HEALPix \cite{healpix}.
+Some of the results in this paper have been derived using \ac{HEALPix} \cite{healpix}.
 

### 2013-06-21 ms.tex
Add self-critical margin note
Add footnote about OpenMP

@@
 
+\marginpar{Clumsy prose.}%
 We evaluate the \ac{TOA}\nobreakdashes-only posterior at the centers of \ac{HEALPix} pixels at the lowest possible resolution.
@@
 
+The \ac{SNR} marginalization integrals for different pixels may be evaluated in parallel.
+In our C language implementation, the loop over pixels is parallelized with OpenMP\footnote{http://openmp.org/}.
+
 \ack Source code for \ac{BAYESTAR} is available on the \ac{LIGO} \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/bayestar.html}.

### 2013-06-21 ms.tex
Fix typo
New bacronym?
Add margin note about orbital precession

@@
 %
+\marginpar{Probably need to change the backronym. BAYESian TriAngulation and Rapid localization?}
+%
 \footnote{A pun on the Cylon battleships in the American television series Battlestar Galactica.
@@
 Furthermore, in \ac{GW} detection efforts, especially those focused on \ac{NSNS} systems, the component spins $\mathbf{S}_1$ and $\mathbf{S}_2$ are often assumed to be aligned with the system's total angular momentum and condensed to a single scalar parameter $\chi$, or even neglected entirely, $\mathbf{S}_1 = \mathbf{S}_2 = 0$.
-In this work we will consider only zero spin.
+In this work we will consider only zero spin.%
+%
+\marginpar{We don't need zero spin; we just need negligible orbital precession.}
 
@@
 
-There are many ways to realize distribution function with these properties, but they tend to be difficult to work with.%
+There are many ways to realize a distribution function with these properties, but they tend to be difficult to work with.%
 %

### 2013-06-21 ms.tex
Fix typo
Add margin note about implementation section

@@
 The defining characteristic of the Cylons is that they repeatedly defeat humanity by using their superhuman information\nobreakdashes-gathering ability to coordinate overwhelming forces.
-The name also suggests that, like the Cylons, \acp{GW} detectors may some day rise against us humans.}%
+The name also suggests that, like the Cylons, \ac{GW} detectors may some day rise against us humans.}%
 %
@@
 
+\subsection{Implementation}
+
+\marginpar{Talk briefly about C implementation, OpenMP, and Python wrapper that interfaces with search pipelines.}%
 The \ac{SNR} marginalization integrals for different pixels may be evaluated in parallel.

### 2013-06-21 ms.tex
Added margin notes to section on timing uncertainty
Formatting

@@
 \begin{abstract}
+\marginpar{Wrong emphasis? I like Larry's idea of having this paper discuss several different tiers of sky localization with increasing accuracy and latency.}%
 Timely \acl{EM} follow\nobreakdashes-up of \acl{CBC} events detected by \acl{aLIGO} requires rapidly inferring the sky location from the \ac{GW} observations.
 Calculation of the posterior distribution of the sky location and intrinsic source parameters given the \ac{GW} strain takes hours with state\nobreakdashes-of\nobreakdashes-the\nobreakdashes-art \ac{MCMC} parameter estimation codes.
-By taking as the measurement not the \ac{GW} strain itself but the amplitude and \acl{TOA} of the putative signal at each detector, and by fixing the intrinsic source parameters, we have constructed a non-\ac{MCMC}, fully deterministic Bayesian parameter estimation algorithm that takes just seconds to produce sky maps of posterior probability.%
-\marginpar{Wrong emphasis? I like Larry's idea of having this paper discuss several different tiers of sky localization with increasing accuracy and latency.}
+By taking as the measurement not the \ac{GW} strain itself but the amplitude and \acl{TOA} of the putative signal at each detector, and by fixing the intrinsic source parameters, we have constructed a non-\ac{MCMC}, fully deterministic Bayesian parameter estimation algorithm that takes just seconds to produce sky maps of posterior probability.
 \end{abstract}
 
-\section{Outline}
+\section*{Outline}
 
@@
 
+\marginpar{It's difficult to follow from here where the timing uncertainty is used in the following section.}
+
+\marginpar{Need to introduce Barankin bound.}
+
 \section{The \ac{BAYESTAR} algorithm}

### 2013-06-21 uninformative.py
We won't be using this figure.

@@ -1,21 +1,0 @@
-#!/usr/bin/env python
-"""
-Plot the distribution of angular offset for an uninformative sky location
-estimator.
-"""
-__author__ = "Leo Singer <leo.singer@ligo.org>"
-
-from matplotlib import pyplot as plt
-import numpy as np
-
-theta = np.logspace(np.log10(0.1), np.log10(180), 1000)
-p = np.square(np.sin(np.deg2rad(theta)))
-plt.semilogx(theta, p)
-plt.xlim(0.1, 1000)
-plt.ylim(0, 1.2)
-plt.grid()
-plt.xlabel('angle between true location and mode of posterior')
-plt.ylabel('probability density')
-plt.title(r'totally uninformative localization: $f(\theta) = \, \frac{2}{\pi} \, \sin^2 \theta$')
-plt.setp(plt.gca().get_yticklabels(), visible=False)
-plt.savefig('uninformative.pdf')

### 2013-06-21 ms.tex
No spin precession, rather than no spin

@@
 Tidal deformabilities of the \acp{NS} are omitted because the effect of tides on the inspiral would only be detectable by an Einstein Telescope-class \ac{GW} observatory~\cite{PhysRevD.81.123016}.
-Furthermore, in \ac{GW} detection efforts, especially those focused on \ac{NSNS} systems, the component spins $\mathbf{S}_1$ and $\mathbf{S}_2$ are often assumed to be aligned with the system's total angular momentum and condensed to a single scalar parameter $\chi$, or even neglected entirely, $\mathbf{S}_1 = \mathbf{S}_2 = 0$.
-In this work we will consider only zero spin.%
-%
-\marginpar{We don't need zero spin; we just need negligible orbital precession.}
+Furthermore, in \ac{GW} detection efforts, especially those focused on \ac{NSNS} systems, the component spins $\mathbf{S}_1$ and $\mathbf{S}_2$ are often assumed to be nonprecessing and aligned with the system's total angular momentum and condensed to a single scalar parameter $\chi$, or even neglected entirely, $\mathbf{S}_1 = \mathbf{S}_2 = 0$.
+In this work we will not consider spin precession.
 

### 2013-06-21 ms.tex
Not really a horizon distance

@@
 The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, and depend on the orientation of detector $i$ as well as the sky location and sidereal time of the event, and are derived in~\cite{PhysRevD.63.042003}.
-The quantity $D_{1,i}$ is a fiducial horizon distance for detector $i$,
-%
-\marginpar{This isn't what we usually call the horizon distance; the horizon distance is where face-on inspirals produce SNR=8.}%
+The quantity $D_{1,i}$ is a fiducial distance at which detector $i$ would register SNR=1 for an optimally oriented binary (face\nobreakdashes-on, and in a direction perpendicular to the interferometer's arms),
 %

### 2013-06-21 ms.tex
Change BAYESTAR acronym

@@
 %
-\marginpar{Probably need to change the backronym. BAYESian TriAngulation and Rapid localization?}
-%
 \footnote{A pun on the Cylon battleships in the American television series Battlestar Galactica.
@@
 %
-\footnote{We do not like to mention the final `T' in the acronym, because then it would be called BAYES\nobreakdashes-TART, which would sound ridiculous.}%
+\footnote{We do not like to mention the final `L' in the acronym, because then it would be pronounced BAYESTARL, which sounds stupid.}%
 .

### 2013-06-22 ms.tex
Cite GRB prospects paper
Fix source URL

@@
 %
-\marginpar{Cite \ac{GRB} prospects paper appendix; a very similar equation appears there. Is there somewhere else that has this formula? Shutz doesn't, and Finn \& Chernoff's notation is too heavy.}%
-%
 It is easily shown from \Eref{eq:full-signal-model} that
@@
 \end{eqnarray}
+%
+A similar expression appears in Appendix~A of~\cite{PhysRevD.87.064033}.
 
@@
 
-\ack Source code for \ac{BAYESTAR} is available on the \ac{LIGO} \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/bayestar.html}.
+\ack \marginpar{This URL is not specific enough.}Source code for \ac{BAYESTAR} is available on the \ac{LIGO} \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/lalsuite.html}.
 

### 2013-06-23 ms.tex
Move outline out of manuscript

@@
 \end{abstract}
-
-\section*{Outline}
-
-\begin{enumerate}
-\item Introduction {
-	\begin{enumerate}
-	\item Background: \acp{GRB}, \acp{CBC}, \acs{LIGO}
-	\item Time scales of \acp{GRB}, \acp{CBC}
-	\item Current sky localization algorithms and response times
-	\end{enumerate}
-}
-\item Preliminaries {
-	\begin{enumerate}
-	\item Fisher information, \ac{CRLB}
-	\item Matched\nobreakdashes-filter estimator
-	\end{enumerate}
-}
-\item Algorithm {
-	\begin{enumerate}
-	\item Assumptions
-	\item Signal model
-	\item Likelihood function
-	\item Priors and marginalization
-	\item Procedure
-	\end{enumerate}
-}
-\item Results {
-	\begin{enumerate}
-	\item Injection population
-	\item Cumulative fraction vs. confidence level plots
-	\item Sky localization accuracy as a function of distance or \ac{SNR}
-	\item Speed
-	\end{enumerate}
-}
-\item Discussion {
-	\begin{enumerate}
-	\item Potential further refinements
-	\item What astrophysical targets were accessible before
-	\item What new targets are available with the response time of seconds
-	\item Lay out timeline of follow\nobreakdashes-up campaign, including different tiers of inference and different types of telescopes
-	\end{enumerate}
-}
-\item Appendices {
-	\begin{enumerate}
-	\item Code listing?
-	\item Any proofs that are needed in the body but would interrupt the text?
-	\end{enumerate}
-}
-\end{enumerate}
-
-\section{Introduction}
 

### 2013-06-23 outline.tex
Move outline out of manuscript

\documentclass{article}
\usepackage{bib/aas_macros}
\input{ligo-acronyms/acronyms}

\begin{document}

\section*{Outline}

\begin{enumerate}
\item Introduction {
	\begin{enumerate}
	\item Background: \acp{GRB}, \acp{CBC}, \acs{LIGO}
	\item Time scales of \acp{GRB}, \acp{CBC}
	\item Current sky localization algorithms and response times
	\end{enumerate}
}
\item Preliminaries {
	\begin{enumerate}
	\item Fisher information, \ac{CRLB}
	\item Matched\nobreakdashes-filter estimator
	\end{enumerate}
}
\item Algorithm {
	\begin{enumerate}
	\item Assumptions
	\item Signal model
	\item Likelihood function
	\item Priors and marginalization
	\item Procedure
	\end{enumerate}
}
\item Results {
	\begin{enumerate}
	\item Injection population
	\item Cumulative fraction vs. confidence level plots
	\item Sky localization accuracy as a function of distance or \ac{SNR}
	\item Speed
	\end{enumerate}
}
\item Discussion {
	\begin{enumerate}
	\item Potential further refinements
	\item What astrophysical targets were accessible before
	\item What new targets are available with the response time of seconds
	\item Lay out timeline of follow\nobreakdashes-up campaign, including different tiers of inference and different types of telescopes
	\end{enumerate}
}
\item Appendices {
	\begin{enumerate}
	\item Code listing?
	\item Any proofs that are needed in the body but would interrupt the text?
	\end{enumerate}
}
\end{enumerate}

\end{document}

### 2013-07-26 ms.tex
Fix capitalization
Shorten overblown sentence
Spelling

@@
 
-A laser interferometer senses \acp{GW} by transducing differential changes in arm length to a digitized photocurrent time series.
-In the angular frequency domain, a single detector's observation $X_i(\omega)$ is
+In the frequency domain, the data recorded by a single \ac{GW} interferometer are
 %
@@
 %
-where $h_i (\omega; \boldsymbol\theta)$ is the \ac{GW} signal receied by detector $i$, given a parameter vector $\boldsymbol\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with \ac{PSD} $S_i(\omega)$.
+where $h_i (\omega; \boldsymbol\theta)$ is the \ac{GW} signal received by detector $i$, given a parameter vector $\boldsymbol\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with \ac{PSD} $S_i(\omega)$.
 We shall denote the combined observation from a network of detectors as $\mathbf x (\omega) \equiv \{x_i (\omega)\}_i$.

### 2013-07-26 ms.tex
Define noise power spectral density

@@
 %
-where $h_i (\omega; \boldsymbol\theta)$ is the \ac{GW} signal received by detector $i$, given a parameter vector $\boldsymbol\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with \ac{PSD} $S_i(\omega)$.
+where $h_i (\omega; \boldsymbol\theta)$ is the \ac{GW} signal received by detector $i$, given a parameter vector $\boldsymbol\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[n_i(\omega)^2\right]$.
 We shall denote the combined observation from a network of detectors as $\mathbf x (\omega) \equiv \{x_i (\omega)\}_i$.

### 2013-07-26 ms.tex
Make another sentence less formal

@@
 
-The likelihood, or the probability of obtaining the observation $\mathbf x$ conditioned on the value of $\boldsymbol\theta$, is Gaussian, with the log\nobreakdashes-likelihood proportional to:
+The likelihood, or the probability of obtaining the observation $\mathbf x$ conditioned on the value of $\boldsymbol\theta$, is the product of Gaussian distributions, one Gaussian for each detector:
 %

### 2013-07-26 ms.tex
Spelling

@@
 %
-Binary eccentricity is omitted as an intrinisc parameter because tidal interaction of the progenitor stars~\cite{0004-637X-572-1-407} and later \ac{GW} emission~\cite{PhysRev.136.B1224} circularize the orbit long before the inspiral enters \ac{LIGO}'s frequency range of $\sim$10\nobreakdashes--1000~kHz.
+Binary eccentricity is omitted as an intrinsic parameter because tidal interaction of the progenitor stars~\cite{0004-637X-572-1-407} and later \ac{GW} emission~\cite{PhysRev.136.B1224} circularize the orbit long before the inspiral enters \ac{LIGO}'s frequency range of $\sim$10\nobreakdashes--1000~kHz.
 Tidal deformabilities of the \acp{NS} are omitted because the effect of tides on the inspiral would only be detectable by an Einstein Telescope-class \ac{GW} observatory~\cite{PhysRevD.81.123016}.

### 2013-07-26 ms.tex
Add citation for highly eccentric binaries

@@
 %
-Binary eccentricity is omitted as an intrinsic parameter because tidal interaction of the progenitor stars~\cite{0004-637X-572-1-407} and later \ac{GW} emission~\cite{PhysRev.136.B1224} circularize the orbit long before the inspiral enters \ac{LIGO}'s frequency range of $\sim$10\nobreakdashes--1000~kHz.
+Even this list of parameters involves some simplifying assumptions.
+First, for \ac{BNS} systems, we omit binary eccentricity as an intrinsic parameter because tidal interaction of the progenitor stars~\cite{0004-637X-572-1-407} and later \ac{GW} emission~\cite{PhysRev.136.B1224} circularize the orbit long before the inspiral enters \ac{LIGO}'s frequency range of $\sim$10\nobreakdashes--1000~kHz.
+Eccentricity would play a major role in the evolution and waveforms of \ac{NSBH} and \ac{BBH} sources formed by dynamical capture~\cite{PhysRevD.87.043004}.
 Tidal deformabilities of the \acp{NS} are omitted because the effect of tides on the inspiral would only be detectable by an Einstein Telescope-class \ac{GW} observatory~\cite{PhysRevD.81.123016}.

### 2013-07-26 ms.tex
Make acronyms more uniform

@@
 Tidal deformabilities of the \acp{NS} are omitted because the effect of tides on the inspiral would only be detectable by an Einstein Telescope-class \ac{GW} observatory~\cite{PhysRevD.81.123016}.
-Furthermore, in \ac{GW} detection efforts, especially those focused on \ac{NSNS} systems, the component spins $\mathbf{S}_1$ and $\mathbf{S}_2$ are often assumed to be nonprecessing and aligned with the system's total angular momentum and condensed to a single scalar parameter $\chi$, or even neglected entirely, $\mathbf{S}_1 = \mathbf{S}_2 = 0$.
+Furthermore, in \ac{GW} detection efforts, especially those focused on \ac{BNS} systems, the component spins $\mathbf{S}_1$ and $\mathbf{S}_2$ are often assumed to be nonprecessing and aligned with the system's total angular momentum and condensed to a single scalar parameter $\chi$, or even neglected entirely, $\mathbf{S}_1 = \mathbf{S}_2 = 0$.
 In this work we will not consider spin precession.

### 2013-07-26 ms.tex
Remove sub-section heading
Tone down a sentence

@@
 
-\subsection{Detection, and matched filter bank as a \ac{MLE}}
-
 More succinctly, we can write the signal received by detector $i$ in terms of direct observables, the amplitude $\rho_i$, phase $\gamma_i$, and time delay $\tau_i$ on arrival at detector $i$,
@@
 The \ac{CRLB} has been widely applied in \ac{GW} astronomy, for instance by \cite{fairhurst:2009} to predict the sky localization accuracy that is achievable with \ac{GW} observations.
-Here, we review the Fisher matrix calculation in \cite{fairhurst:2009} but add some physical interpretation to the particular structure of the covariance matrix.
+Here, we review the Fisher matrix calculation in \cite{fairhurst:2009} and discuss the interpretation.
 

### 2013-07-26 ms.tex
Spelling

@@
 
-In order to estimate the \ac{ML} estimator accuracy for a matched filter, we will momentarily consider the likelihood for a single detector, and surpress the intrinsic degrees of freedom:
+In order to estimate the \ac{ML} estimator accuracy for a matched filter, we will momentarily consider the likelihood for a single detector, and suppress the intrinsic degrees of freedom:
 %

### 2013-07-26 ms.tex
Join paragraphs
Remove unnecessary sentence
Spelling

@@
 The choice of prior is open to one's astrophysical preconceptions, but during \ac{S6} when \ac{LIGO}'s Bayesian \ac{CBC} parameter estimation pipelines were pioneered, the prior was taken to be isotropic in sky location and binary orientation, and uniform in volume, arrival time, and the component masses~\cite{2013arXiv1304.1775T}.
-Put briefly, the prior was uniform in all of $(\alpha, \sin\delta, {D_\mathrm{L}}^3, t_\oplus, u, \psi, \phi_c, m_1, m_2)$.
 
@@
 \end{equation}
-
+%
 The second change of variables, from $\gamma$ and $\gamma_\tau$ to $\gamma_\pm = \frac{1}{\sqrt{2}}(\gamma \pm \gamma_\tau)$, diagonalizes the Fisher matrix:
@@
 
-We have no explicit prior on the intrinsic paramters; in our analysis they are fixed at their \ac{ML} estimates, $\hat{\boldsymbol\theta}$.
+We have no explicit prior on the intrinsic parameters; in our analysis they are fixed at their \ac{ML} estimates, $\hat{\boldsymbol\theta}$.
 The distance prior is given by a user-selected power of distance, ${D_\mathrm{L}}^m$, where $m=2$ for a prior that is uniform in volume, and $m=-1$ for a scale-invariant prior that is uniform in $\log D_\mathrm{L}$.

### 2013-07-26 ms.tex
'Precalculated' is not a word

@@
 As in many parameter estimation problems, the integrand may be sharply peaked, and any generic adaptive integration scheme may take a long time to find the peak.
-Before evaluating each inner integral over $D_\mathrm{L}$, the quantities $A=-\frac{1}{2} \sum_i {\rho_i}^2$ and $B=\sum_i \rho_i \hat\rho_i$ are precalculated so that the integrand can be written as
+Before evaluating each inner integral over $D_\mathrm{L}$, the quantities $A=-\frac{1}{2} \sum_i {\rho_i}^2$ and $B=\sum_i \rho_i \hat\rho_i$ are pre\nobreakdashes-calculated so that the integrand can be written as
 %

### 2013-07-26 ms.tex
Spelling

@@
 These intervals are depicted in Figure~\ref{fig:radial_integrand} with distance increasing from left to right.
-There is a left\nobreakdashes-hand or small distance tail in which the integrand is small and monotically increasing, a left\nobreakdashes- and right\nobreakdashes-hand side of the maximum likelihood peak, and a right\nobreakdashes-hand tail in which the integrand is small and monotonically decreasing.
+There is a left\nobreakdashes-hand or small distance tail in which the integrand is small and monotonically increasing, a left\nobreakdashes- and right\nobreakdashes-hand side of the maximum likelihood peak, and a right\nobreakdashes-hand tail in which the integrand is small and monotonically decreasing.
 These breakpoints are:

### 2013-08-02 ms.tex
Small improvements to prose

@@
 
-In the frequency domain, the data recorded by a single \ac{GW} interferometer are
+In the frequency domain, the strain observed by a single \ac{GW} interferometer is
 %
@@
 %
-where $h_i (\omega; \boldsymbol\theta)$ is the \ac{GW} signal received by detector $i$, given a parameter vector $\boldsymbol\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[n_i(\omega)^2\right]$.
+where $h_i (\omega; \boldsymbol\theta)$ is the \ac{GW} signal given a parameter vector $\boldsymbol\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[n_i(\omega)^2\right]$.
 We shall denote the combined observation from a network of detectors as $\mathbf x (\omega) \equiv \{x_i (\omega)\}_i$.
 
-The likelihood, or the probability of obtaining the observation $\mathbf x$ conditioned on the value of $\boldsymbol\theta$, is the product of Gaussian distributions, one Gaussian for each detector:
+The likelihood, or the probability of obtaining the observation $\mathbf x$ conditioned on the value of $\boldsymbol\theta$, is a product of Gaussian distributions:
 %
@@
 %
-Even this list of parameters involves some simplifying assumptions.
-First, for \ac{BNS} systems, we omit binary eccentricity as an intrinsic parameter because tidal interaction of the progenitor stars~\cite{0004-637X-572-1-407} and later \ac{GW} emission~\cite{PhysRev.136.B1224} circularize the orbit long before the inspiral enters \ac{LIGO}'s frequency range of $\sim$10\nobreakdashes--1000~kHz.
-Eccentricity would play a major role in the evolution and waveforms of \ac{NSBH} and \ac{BBH} sources formed by dynamical capture~\cite{PhysRevD.87.043004}.
-Tidal deformabilities of the \acp{NS} are omitted because the effect of tides on the inspiral would only be detectable by an Einstein Telescope-class \ac{GW} observatory~\cite{PhysRevD.81.123016}.
+This list of parameters involves some simplifying assumptions.
+Eccentricity is omitted: although it does play a major role in the evolution and waveforms of \ac{NSBH} and \ac{BBH} sources formed by dynamical capture~\cite{PhysRevD.87.043004}, \ac{BNS} systems formed by binary stellar evolution should circularize due to tidal interaction~\cite{0004-637X-572-1-407} and later \ac{GW} emission~\cite{PhysRev.136.B1224} long before the inspiral enters \ac{LIGO}'s frequency range of $\sim$10\nobreakdashes--1000~kHz.
+Tidal deformabilities of the \acp{NS} are omitted because the signal imprinted by the companions' material properties is so small that it will only be detectable by an Einstein Telescope\nobreakdashes-class \ac{GW} observatory~\cite{PhysRevD.81.123016}.
 Furthermore, in \ac{GW} detection efforts, especially those focused on \ac{BNS} systems, the component spins $\mathbf{S}_1$ and $\mathbf{S}_2$ are often assumed to be nonprecessing and aligned with the system's total angular momentum and condensed to a single scalar parameter $\chi$, or even neglected entirely, $\mathbf{S}_1 = \mathbf{S}_2 = 0$.
-In this work we will not consider spin precession.
 
@@
 
-The prevailing technique for detection of \acp{GW} from \acp{CBC} is to realize a \ac{MLE} from \Eref{eq:gaussian-likelihood} with the above signal model.
-This takes the form of a bank of matched filters, or template waveforms that are continually convolved with the incoming data stream.
-The matched filter or \ac{ML} estimates are given by
+The prevailing technique for detection of \acp{GW} from \acp{CBC} is to realize a \ac{MLE} from the likelihood in \Eref{eq:gaussian-likelihood} and the signal model in \Eref{eq:signal-model}.
+Concretely, this results in a bank of matched filters, or template waveforms that are convolved with the incoming data stream.
+The matched filter or \ac{ML} parameter estimates are given by
 %
@@
 
-Computing the marginal posterior almost always involves many-dimensional and often ill-behaved integrals.\marginpar{More about \ac{MCMC} methods here.}
+Bayesian parameter estimation has many advantages, including broad generality and the ability to make probabilistically meaningful statements even with very low \ac{SNR} measurements.
+However, in problems of even modest complexity, the marginalization step burdens the analysis with nasty, many-dimensional, ill-behaved integrals.
+Almost all modern parameter estimation tactics are variations of \ac{MCMC} integration.
+Though powerful, \ac{MCMC} is slow and non-deterministic.
+With the most sophisticated \ac{CBC} parameter estimation codes, it still takes days to process a single event.
 
@@
 %
-\begin{equation}\label{eq:gaussian-likelihood}
+\begin{equation}\label{eq:gaussian-likelihood-spa}
 	\fl\mathcal{L}(x_i; \rho_i, \gamma_i, \tau_i) = p(x_i | \rho_i, \gamma_i, \tau_i)
@@
 
-The likelihood in \Eref{eq:gaussian-likelihood} is the quantity that is maximized by the matched\nobreakdashes-filter estimator.
+The likelihood in \Eref{eq:gaussian-likelihood-spa} is the quantity that is maximized by the matched\nobreakdashes-filter estimator.
 However, just because we have the covariance matrix defined by \Eref{eq:covariance-matrix} does not mean that the distribution of estimator errors is multivariate Gaussian.

### 2013-08-02 ms.tex
Awesome title

@@
 
-\title[Rapid-response Bayesian sky localization]{Rapid-response Bayesian sky localization for electromagnetic follow-up of gravitational-wave candidates}
+\title[Rapid-response Bayesian sky localization]{WHOOMP! (There It Is) \\
+Rapid Bayesian position reconstruction for gravitational\nobreakdashes-wave transients}
 \author{Leo Singer and Larry R. Price}
@@
 With the most sophisticated \ac{CBC} parameter estimation codes, it still takes days to process a single event.
+This delay is undesirable for planning targeted \ac{EM} follow\nobreakdashes-up searches of \ac{LIGO} events.
 

### 2013-08-02 ms.tex
Complete Bayesian section

@@
 Concretely, this results in a bank of matched filters, or template waveforms that are convolved with the incoming data stream.
-The matched filter or \ac{ML} parameter estimates are given by
+The matched filter or \ac{ML} point estimates of the signal parameters are
 %
@@
 \hat\rho_i &=& \frac{1}{\sigma_i} \left| \int_0^\infty \frac{h^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) x_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right| \\
-\hat\gamma_i &=& \arg \int_0^\infty \frac{h^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) x_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega,
+\hat\gamma_i &=& \arg \int_0^\infty \frac{h^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) x_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega.
 \end{eqnarray}
-%
-where the $^*$ denotes a complex conjugate.
 
@@
 This delay is undesirable for planning targeted \ac{EM} follow\nobreakdashes-up searches of \ac{LIGO} events.
+
+In what follows, we propose a complementary rapid parameter estimation scheme that can produce reliable positions estimates within minutes of a detection.
+We can even use our scheme to speed up the full \ac{MCMC} analysis and make the refined parameter estimates available more quickly.
+The key difference is that we start not from the \ac{GW} signal itself, but from the point parameter estimates from the detection.
+By harnessing the detection pipeline in this manner, we arrive at a simpler Bayesian problem that is amenable to straightforward, deterministic, numerical quadrature.
 

### 2013-08-02 ms.tex
Fix rates paper citation
Work on Fisher matrix section

@@
 The \ac{aLIGO} and Vigo detectors will be part of a worldwide network that also includes KAGRA and hopefully LIGO-India.
-Coalescing neutron star binaries are among the most likely sources, with event rates of $~ 50/\mathrm{yr}$~\cite{rates}.%
-%
-\marginpar{Check that number.}%
-%
+Coalescing neutron star binaries are among the most likely sources, with 0.4\nobreakdashes--400 detections per year~\cite{rates}.
 In addition to being an efficient source of \acp{GW}, the \ac{NS} may be tidally shredded before merging, providing fuel for an electromagnetic counterpart.
@@
 
-The \ac{CRLB} is a useful analytical tool for predicting with what error an unknown parameter or parameters can be estimated from noisy observations.
-When an unbiased estimator exists, the \ac{CRLB} bound says that the covariance of the estimation error is bounded below by the Fisher information matrix.
-The \ac{CRLB} has been widely applied in \ac{GW} astronomy, for instance by \cite{fairhurst:2009} to predict the sky localization accuracy that is achievable with \ac{GW} observations.
+The first task is to understand the uncertainty in the \ac{ML} point estimates from the detection.
+We begin with the \ac{CRLB}, which constraints the covariance of any unbiased estimator.
+The \ac{CRLB} has been widely applied in \ac{GW}.
 Here, we review the Fisher matrix calculation in \cite{fairhurst:2009} and discuss the interpretation.
@@
 
-The \ac{CRLB} says that, element by element,%
+If $\hat{\boldsymbol\theta}$ is an unbiased estimator of $\boldsymbol\theta$, $\tilde{\boldsymbol\theta} = \hat{\boldsymbol\theta} - \boldsymbol\theta$ is the measurement error, and $\Sigma = \mathrm{E} \, [\tilde{\boldsymbol\theta}\tilde{\boldsymbol\theta}^\intercal]$ is the covariance of the measurement error, then the \ac{CRLB} says that
 %
-\marginpar{Element by element? I don't think that is right.}
+\begin{equation}
+    \Sigma \geq \mathcal{I}^{-1}
+\end{equation}
+%
+in the sense that $\left(\Sigma - \mathcal{I}^{-1}\right)$ is positive semi\nobreakdashes-definite. Specifically, for our likelihood, the \ac{CRLB} implies that
 %
@@
         \begin{array}{c}
-            \hat{\rho}_i - \rho_i \\
-            \hat{\gamma}_i - \gamma_i \\
-            \hat{\tau}_i - \tau_i
+            \tilde{\rho}_i \\
+            \tilde{\gamma}_i \\
+            \tilde{\tau}_i
         \end{array}
@@
 %
-where a hat $\hat{\phantom{\rho}}$ denotes the estimate of a parameter and ${\omega_{\mathrm{rms},i}}^2 = {\overline{\omega^2}}_i - {{\overline{\omega}}_i}^2$.
+where ${\omega_{\mathrm{rms},i}}^2 = {\overline{\omega^2}}_i - {{\overline{\omega}}_i}^2$.
 Reading off the $\tau \tau$ element of the covariance matrix reproduces the timing accuracy in Equation~(24) of \cite{fairhurst:2009},
@@
 \end{equation}
-
-\marginpar{Plot covariance ellipses and estimator samples for each change of variables; that would be a nice visual here.}%
 %

### 2014-05-22 ms.tex
Some minor edits

@@
 In this paper we present a rapid and accurate method of sky localization that makes use of Bayesian methods.
-It differs from existing techniques is several important ways.
+It differs from existing techniques in several important ways.
 In the first place, we fix the intrinsic parameters to their \ac{ML} estimates, as provided by the detection pipeline.
@@
 %
-where $h_i (\omega; \boldsymbol\theta)$ is the \ac{GW} signal given a parameter vector $\boldsymbol\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[n_i(\omega)^2\right]$.
+where $h_i (\omega; \boldsymbol\theta)$ is the \ac{GW} signal given a parameter vector $\boldsymbol\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[\left|n_i(\omega)\right|^2\right]$.
 We shall denote the combined observation from a network of detectors as $\mathbf x (\omega) \equiv \{x_i (\omega)\}_i$.
@@
 Almost all modern parameter estimation tactics are variations of \ac{MCMC} integration.
-Though powerful, \ac{MCMC} is slow and non-deterministic.
+Though powerful, \ac{MCMC} is historically slow and non-deterministic.
 With the most sophisticated \ac{CBC} parameter estimation codes, it still takes days to process a single event.
@@
 
-The first task is to understand the uncertainty in the \ac{ML} point estimates from the detection.
+The first task is to quantify the uncertainty in the \ac{ML} estimates of the \acp{TOA} and amplitudes from the detection.
 We begin with the \ac{CRLB}, which constraints the covariance of any unbiased estimator.
@@
 
-First, we put the phase and time on the same footing by measuring the time in units of $1 / 2 \pi \sqrt{\overline{\omega^2}}$ with a change of variables from $\tau$ to $\gamma_\tau = 2 \pi \sqrt{\overline{\omega^2}} \tau$:
+First, we put the phase and time on the same footing by measuring the time in units of $1 / \sqrt{\overline{\omega^2}}$ with a change of variables from $\tau$ to $\gamma_\tau = \sqrt{\overline{\omega^2}} \tau$:
 %
@@
 Thus, in the appopriate time units, the \textit{sum and difference} of the phase and time of the signal are measured independently.
+
+\subsection{Barankin bound}
+
+It is well known that the \ac{CRLB} is not a tight bound in the sense that there does not always exist an estimator that achieves it.
+Barankin derived a tight bound for any unbiased estimator \cite{barankin1949locally}.
 
@@
 
-There are many ways to realize a distribution function with these properties, but they tend to be difficult to work with.%
+There are many ways to realize a distribution function with these properties, but they tend to be difficult to work with.
 %
@@
 \bibliographystyle{iopart-num}
-\bibliography{apj-jour,bib/telescope}
+\bibliography{apj-jour,bib/telescope,ms}
 

### 2014-05-22 ms.tex
Reword description of MCMC

@@
 Bayesian parameter estimation has many advantages, including broad generality and the ability to make probabilistically meaningful statements even with very low \ac{SNR} measurements.
-However, in problems of even modest complexity, the marginalization step burdens the analysis with nasty, many-dimensional, ill-behaved integrals.
-Almost all modern parameter estimation tactics are variations of \ac{MCMC} integration.
-Though powerful, \ac{MCMC} is historically slow and non-deterministic.
+However, in problems of even modest complexity, the marginalization step involves many-dimensional, ill-behaved integrals.
+The powerful \ac{MCMC} integration technique has become almost synonymous with Bayesian inference.
+Though powerful, \ac{MCMC} is inherently non-deterministic and resistant to parallelization, as well as (at least historically) slow.
 With the most sophisticated \ac{CBC} parameter estimation codes, it still takes days to process a single event.

### 2014-05-24 inclination_integral_convergence.py
Scripts to generate rate of convergence plots

@@ -1,0 +1,50 @@
+import numpy as np
+from matplotlib import pyplot as plt
+import scipy.integrate
+import scipy.special
+
+fig = plt.figure(figsize=(3.5, 2.5))
+ax = fig.add_subplot(111)
+
+np.random.seed(0)
+nruns = 10
+neval = np.unique(np.round(np.logspace(0, 2, 50)).astype(int))
+
+def integrand(u, a, b, c, d, e, f):
+    v = 1 + np.square(u)
+    I0arg = np.sqrt(np.square(c * u + d * v) + np.square(e * u + f * v))
+    return np.exp(-(np.square(a * u) + np.square(b * v)) + I0arg) * scipy.special.i0e(I0arg)
+
+for irun in range(nruns):
+    args = tuple(np.random.randn(6).tolist())
+    exact, _ = scipy.integrate.quad(
+        integrand, -1, 1, args, epsabs=1e-14, epsrel=1e-14, limit=200)
+    approx = np.asarray([
+        scipy.integrate.fixed_quad(integrand, -1, 1, args, n)[0] for n in neval])
+    relerr = np.abs(approx - exact) / exact
+    ax.plot(neval, relerr, '-k')
+
+def integrand(u, a, b, c, d, e, f):
+    v = 1 + np.square(u)
+    I0arg1 = np.sqrt(np.square(c * u) + np.square(d * v))
+    I0arg2 = np.sqrt(np.square(e * u) + np.square(f * v))
+    return np.exp(-(np.square(a * u) + np.square(b * v)) + I0arg1 + I0arg2) * scipy.special.i0e(I0arg1) * scipy.special.i0e(I0arg2)
+
+for irun in range(nruns):
+    args = tuple(np.random.randn(6).tolist())
+    exact, _ = scipy.integrate.quad(
+        integrand, 0, 1, args, epsabs=1e-14, epsrel=1e-14, limit=200)
+    approx = np.asarray([
+        scipy.integrate.fixed_quad(integrand, 0, 1, args, n)[0] for n in neval])
+    relerr = np.abs(approx - exact) / exact
+    ax.plot(neval, relerr, '--k')
+
+ax.set_xscale('log')
+ax.set_yscale('log')
+ax.set_ylim(1e-16, 1)
+ax.set_xlabel('number of function evaluations')
+ax.set_ylabel('relative error')
+
+fig.subplots_adjust(left=0.175, bottom=0.15, right=0.95, top=0.95)
+
+fig.savefig('inclination_integral_convergence.pdf')

### 2014-05-24 polarization_angle_integral_convergence.py

@@ -1,0 +1,51 @@
+import numpy as np
+from matplotlib import pyplot as plt
+import scipy.integrate
+import scipy.special
+
+fig = plt.figure(figsize=(3.5, 2.5))
+ax = fig.add_subplot(111)
+
+np.random.seed(0)
+nruns = 10
+neval = np.unique(np.round(np.logspace(0, 2, 50)).astype(int))
+
+def integrand(x, a, b, c, d, e, f):
+    I0arg = np.sqrt(np.square(c * np.cos(x) + d * np.sin(x)) + np.square(e * np.cos(x) + f * np.sin(x)))
+    return np.exp(a * np.cos(x) + b * np.sin(x) + I0arg) * scipy.special.i0e(I0arg)
+
+def periodic_quad(func, args, n):
+    return 2*np.pi/n * np.sum(func(2*np.pi/n * np.arange(n), *args))
+
+for irun in range(nruns):
+    args = tuple(np.random.randn(6).tolist())
+    exact, _ = scipy.integrate.quad(
+        integrand, 0, 2*np.pi, args, epsabs=1e-14, epsrel=1e-14, limit=200)
+    approx = np.asarray([
+        periodic_quad(integrand, args, n) for n in neval])
+    relerr = np.abs(approx - exact) / exact
+    ax.plot(neval, relerr, '-k')
+
+def integrand(x, a, b, c, d, e, f, g, h, i, j):
+    I0arg1 = np.sqrt(np.square(c * np.cos(x) + d * np.sin(x)) + np.square(e * np.cos(x) + f * np.sin(x)))
+    I0arg2 = np.sqrt(np.square(g * np.cos(x) + h * np.sin(x)) + np.square(i * np.cos(x) + j * np.sin(x)))
+    return np.exp(a * np.cos(x) + b * np.sin(x) + I0arg1 + I0arg2) * scipy.special.i0e(I0arg1) * scipy.special.i0e(I0arg2)
+
+for irun in range(nruns):
+    args = tuple(np.random.randn(10).tolist())
+    exact, _ = scipy.integrate.quad(
+        integrand, -1, 1, args, epsabs=1e-14, epsrel=1e-14, limit=200)
+    approx = np.asarray([
+        scipy.integrate.fixed_quad(integrand, -1, 1, args, n)[0] for n in neval])
+    relerr = np.abs(approx - exact) / exact
+    ax.plot(neval, relerr, '--k')
+
+ax.set_xscale('log')
[…]

### 2014-05-24 radial_integrand.py
Match linestyle

@@ -65,7 +65,7 @@ for i, (break1, break2) in enumerate(zip(breakpoints[:-1], breakpoints[1:])):
     fig.subplots_adjust(wspace=0.15)
     bbox0 = Bbox.from_extents(break1, 0, break2, 1)
     mybbox0 = TransformedBbox(bbox0, trans0)
-    fig.artists.append(BboxConnectorPatch(ax.bbox, mybbox0, 1, 4, 2, 3, facecolor='none', linestyle='dashed'))
+    fig.artists.append(BboxConnectorPatch(ax.bbox, mybbox0, 1, 4, 2, 3, facecolor='none', linestyle='dashed', linewidth=0.5))
     if i > 0:
         plt.setp(ax.get_yticklabels(), visible=False)
     r = np.linspace(break1, break2, 100)

### 2014-05-24 ms.tex
Remove Barankin bound, won't be using it
Correct sign of iwt
Why just ML estimates?

@@
 \begin{eqnarray}
-\hat{\boldsymbol\theta}_i, \{\hat\tau_i\}_i &=& \argmax_{\boldsymbol\theta_\mathrm{in}, \{\hat\tau_i\}_i} \sum_i \frac{1}{{\sigma_i}^2} \left| \int_0^\infty \frac{h^*(\omega; \boldsymbol\theta_\mathrm{in}) x_i(\omega) e^{-i \omega \tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right|^2 \\
-\hat\rho_i &=& \frac{1}{\sigma_i} \left| \int_0^\infty \frac{h^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) x_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right| \\
-\hat\gamma_i &=& \arg \int_0^\infty \frac{h^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) x_i(\omega) e^{-i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega.
+\hat{\boldsymbol\theta}_\mathrm{in}, \{\hat\tau_i\}_i &=& \argmax_{\boldsymbol\theta_\mathrm{in}, \{\hat\tau_i\}_i} \sum_i \frac{1}{{\sigma_i}^2} \left| \int_0^\infty \frac{h^*(\omega; \boldsymbol\theta_\mathrm{in}) x_i(\omega) e^{i \omega \tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right|^2 \\
+\hat\rho_i &=& \frac{1}{\sigma_i} \left| \int_0^\infty \frac{h^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) x_i(\omega) e^{i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right| \\
+\hat\gamma_i &=& \arg \int_0^\infty \frac{h^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) x_i(\omega) e^{i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega.
 \end{eqnarray}
+%
+A detection candidate consists of the set $\{\hat{\boldsymbol\theta}_\mathrm{in}, \left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i\}$.
+There are various ways to characterize the significance of a
+detection candidate, but in Gaussian noise the combined SNR $\hat\rho_\mathrm{c}$ suffices,
+%
+\marginpar{What is the least that I can get away with saying about false alarm rate estimation and likelihood ratios?}
+%
+\begin{equation}
+	\hat\rho_\mathrm{c} = \sqrt{\sum_i {\hat\rho_i}^2}.
+\end{equation}.
 
@@
 \section{Measurement uncertainty}
+
+For the purpose of rapid sky localization, we assume that we do not have access to the \ac{GW} data $\mathbf{x}$ itself, and that our only contact with it is through the ML parameter estimates $\{\hat{\boldsymbol\theta}_\mathrm{in}, \left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i\}$.
+Although this is a significant departure from conventional GW parameter estimation techniques, we can still apply the full Bayesian machinery of Equation~(\ref{eq:marginal-posterior}) to compute a posterior distribution for the sky location.
+
+\marginpar{This is poorly worded.}%
+%
+There are many practical advantages of doing so.
+For one, there are difficulties in synchronously gathering together the calibrated GW strain data, auxiliary instrument channels, and data quality vetoes from all of the sites.
+The data consumed by the real\nobreakdashes-time detection pipeline are not necessarily final.
+Longer\nobreakdashes-running follow\nobreakdashes-up analyses can benefit from offline calibration, whereas the rapid sky localization need not re\nobreakdashes-analyze the online data.
+
+Moreover, the dimensionality of the problem is greatly reduced, and the problem becomes computationally easier.
 
@@
 Thus, in the appopriate time units, the \textit{sum and difference} of the phase and time of the signal are measured independently.
-
-\subsection{Barankin bound}
-
-It is well known that the \ac{CRLB} is not a tight bound in the sense that there does not always exist an estimator that achieves it.
-Barankin derived a tight bound for any unbiased estimator \cite{barankin1949locally}.
 

### 2014-05-24 ms.tex
Marginpar whitespace

@@
 %
-\marginpar{What is the least that I can get away with saying about false alarm rate estimation and likelihood ratios?}
+\marginpar{What is the least that I can get away with saying about false alarm rate estimation and likelihood ratios?}%
 %
@@
 %
-\marginpar{Need to discuss $D_{\mathrm{L},\mathrm{min}}$, $D_{\mathrm{L},\mathrm{max}}$, and how they are selected automatically if the user does not provide them.}
+\marginpar{Need to discuss $D_{\mathrm{L},\mathrm{min}}$, $D_{\mathrm{L},\mathrm{max}}$, and how they are selected automatically if the user does not provide them.}%
 %

### 2014-05-26 ms.tex
Combine paragraphs
Big batch of edits: notation cleanup, prepare for autocorrelation likelihood

@@
 \DeclareMathOperator*{\argmax}{\arg\!\max}
+
+% Macros for collapsing sizes of things
+% From TUGboat, Volume 22 (2001), No. 4
+% http://www.tug.org/TUGboat/tb22-4/tb72perlS.pdf
+\def\clap#1{\hbox to 0pt{\hss#1\hss}}
+\def\mathllap{\mathpalette\mathllapinternal}
+\def\mathrlap{\mathpalette\mathrlapinternal}
+\def\mathclap{\mathpalette\mathclapinternal}
+\def\mathllapinternal#1#2{\llap{$\mathsurround=0pt#1{#2}$}}
+\def\mathrlapinternal#1#2{\rlap{$\mathsurround=0pt#1{#2}$}}
+\def\mathclapinternal#1#2{\clap{$\mathsurround=0pt#1{#2}$}}
 
@@
 		- \frac{1}{2} \sum_i \int_0^\infty \frac{\left|x_i (\omega)
-			- h_i(\omega; \boldsymbol\theta) \right|^2}{S_i(\omega)} \, \mathrm{d}\omega
+			- h_i(\omega; \boldsymbol\theta) \right|^2}{S_i(\omega)} \, d\omega
 	\right].
@@
 \begin{equation}\label{eq:horizon}
-D_{1,i} = 1 / \sigma_i, \qquad {\sigma_i}^2 = \int_0^\infty \frac{\left|h(\omega; \mathbf \theta)\right|^2}{S_i(\omega)} \,\mathrm{d}\omega.
+D_{1,i} = 1 / \sigma_i, \qquad {\sigma_i}^2 = \int_0^\infty \frac{\left|h(\omega; \mathbf \theta)\right|^2}{S_i(\omega)} \,d\omega.
 \end{equation}
 
-More succinctly, we can write the signal received by detector $i$ in terms of direct observables, the amplitude $\rho_i$, phase $\gamma_i$, and time delay $\tau_i$ on arrival at detector $i$,
+More succinctly, we can write the signal received by detector $i$ in terms of observable extrinsic parameters $\boldsymbol\theta_i = (\rho_i, \gamma_i, \tau_i)$, the amplitude $\rho_i$, phase $\gamma_i$, and time delay $\tau_i$ on arrival at detector $i$,
 %
 \begin{equation}\label{eq:signal-model}
-	h_i (\omega; \boldsymbol\theta) = \frac{\rho_i}{\sigma_i} e^{i (\gamma_i - \omega \tau_i)} h(\omega; \boldsymbol\theta_\mathrm{in}).
+	h_i (\omega; \boldsymbol\theta_i, \boldsymbol\theta_\mathrm{in}) = h_i (\omega; \rho_i, \gamma_i, \tau_i, \boldsymbol\theta_\mathrm{in}) = \frac{\rho_i}{\sigma_i} e^{i (\gamma_i - \omega \tau_i)} h(\omega; \boldsymbol\theta_\mathrm{in}).
 \end{equation}
@@
 The prevailing technique for detection of \acp{GW} from \acp{CBC} is to realize a \ac{MLE} from the likelihood in \Eref{eq:gaussian-likelihood} and the signal model in \Eref{eq:signal-model}.
-Concretely, this results in a bank of matched filters, or template waveforms that are convolved with the incoming data stream.
-The matched filter or \ac{ML} point estimates of the signal parameters are
+Concretely, this results in a bank of matched filters, or the cross-correlation between a template waveform and the incoming data stream,
+%
+\begin{equation}
+y_i(\tau_i;\boldsymbol\theta_\mathrm{in}) = \frac{1}{\sigma_i (\boldsymbol\theta_\mathrm{in})} \int_0^\infty \frac{h^*(\omega; \boldsymbol\theta_\mathrm{in}) x_i(\omega) e^{i \omega \tau_i}}{S_i(\omega)} \,d\omega.
+\end{equation}
+%
+The \ac{ML} point estimates of the signal parameters, $\mathrm{MLE}(\mathbf{x}) = \{\{ \hat{\boldsymbol\theta}_i \}_i, \hat{\boldsymbol\theta}_\mathrm{in}\} = \{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\boldsymbol\theta}_\mathrm{in}\}$, are given by
 %
 \begin{eqnarray}
-\hat{\boldsymbol\theta}_\mathrm{in}, \{\hat\tau_i\}_i &=& \argmax_{\boldsymbol\theta_\mathrm{in}, \{\hat\tau_i\}_i} \sum_i \frac{1}{{\sigma_i}^2} \left| \int_0^\infty \frac{h^*(\omega; \boldsymbol\theta_\mathrm{in}) x_i(\omega) e^{i \omega \tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right|^2 \\
-\hat\rho_i &=& \frac{1}{\sigma_i} \left| \int_0^\infty \frac{h^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) x_i(\omega) e^{i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega \right| \\
-\hat\gamma_i &=& \arg \int_0^\infty \frac{h^*(\omega; \hat{\boldsymbol\theta}_\mathrm{in}) x_i(\omega) e^{i \omega \hat\tau_i}}{S_i(\omega)} \,\mathrm{d}\omega.
+	\hat{\boldsymbol\theta}_\mathrm{in}, \{\hat\tau_i\}_i
+		&=& \argmax_{\boldsymbol\theta_\mathrm{in}, \{\hat\tau_i\}_i}
+		\sum_i \left| y_i\left(\tau_i;
+		\boldsymbol\theta_\mathrm{in}\right) \right|^2, \\
+	\hat\rho_i &=& \left| y_i\left(\hat\tau_i;
+		\hat{\boldsymbol\theta}_\mathrm{in}\right) \right|, \\
+	\hat\gamma_i &=& \arg y_i\left(\hat\tau_i;
+		\hat{\boldsymbol\theta}_\mathrm{in}\right).
 \end{eqnarray}
 %
-A detection candidate consists of the set $\{\hat{\boldsymbol\theta}_\mathrm{in}, \left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i\}$.
-There are various ways to characterize the significance of a
-detection candidate, but in Gaussian noise the combined SNR $\hat\rho_\mathrm{c}$ suffices,
+A detection candidate consists of the set $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\boldsymbol\theta}_\mathrm{in}\}$.
+There are various ways to characterize the significance of a detection candidate, but in Gaussian noise the combined SNR $\hat\rho_\mathrm{c}$ suffices,
 %
@@
 	\hat\rho_\mathrm{c} = \sqrt{\sum_i {\hat\rho_i}^2}.
-\end{equation}.
+\end{equation}
 
@@
 \begin{equation}\label{eq:marginal-posterior}
-	p(\boldsymbol\beta|\mathbf x) = \int \frac{p(\mathbf x|\boldsymbol\beta,\boldsymbol\lambda) p(\boldsymbol\beta,\boldsymbol\lambda)}{p(\mathbf x)} \,\mathrm{d}\boldsymbol\lambda
+	p(\boldsymbol\beta|\mathbf x) = \int \frac{p(\mathbf x|\boldsymbol\beta,\boldsymbol\lambda) p(\boldsymbol\beta,\boldsymbol\lambda)}{p(\mathbf x)} \,d\boldsymbol\lambda
 \end{equation}
@@
 
-For the purpose of rapid sky localization, we assume that we do not have access to the \ac{GW} data $\mathbf{x}$ itself, and that our only contact with it is through the ML parameter estimates $\{\hat{\boldsymbol\theta}_\mathrm{in}, \left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i\}$.
-Although this is a significant departure from conventional GW parameter estimation techniques, we can still apply the full Bayesian machinery of Equation~(\ref{eq:marginal-posterior}) to compute a posterior distribution for the sky location.
+For the purpose of rapid sky localization, we assume that we do not have access to the \ac{GW} data $\mathbf{x}$ itself, and that our only contact with it is through the ML parameter estimates $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\boldsymbol\theta}_\mathrm{in}\}$.
+Although this is a significant departure from conventional GW parameter estimation techniques, we can still apply the full Bayesian machinery of \Eref{eq:marginal-posterior} to compute a posterior distribution for the sky location.
 
-\marginpar{This is poorly worded.}%
+\marginpar{This paragraph might make more sense in a conclusion.}%
 %
@@
 Longer\nobreakdashes-running follow\nobreakdashes-up analyses can benefit from offline calibration, whereas the rapid sky localization need not re\nobreakdashes-analyze the online data.
-
 Moreover, the dimensionality of the problem is greatly reduced, and the problem becomes computationally easier.
 
-The first task is to quantify the uncertainty in the \ac{ML} estimates of the \acp{TOA} and amplitudes from the detection.
-We begin with the \ac{CRLB}, which constraints the covariance of any unbiased estimator.
-The \ac{CRLB} has been widely applied in \ac{GW}.
-Here, we review the Fisher matrix calculation in \cite{fairhurst:2009} and discuss the interpretation.
+\subsection{Fisher matrix: single detector}
 
-\subsection{Fisher matrix}
+To treat the ML estimates as measurements in and of themselves, we need to work out their likelihood, marginalized over the GW signal $\mathbf{x}$,
+%
+\begin{equation}\label{eq:detection-candidate-likelihood}
+	p\left(\{\hat{\boldsymbol\theta}_i\}_i,
+		\hat{\boldsymbol\theta}_\mathrm{in}
+	\middle| \boldsymbol\theta\right)
+	\propto \int\limits_{\mathclap{\mathbf{x} | \mathrm{MLE}(\mathbf{x}) =
+		\{\{\hat{\boldsymbol\theta}_i\}_i,
+		\hat{\boldsymbol\theta}_\mathrm{in}\}}}
+	p(\mathbf{x} | \boldsymbol\theta) \, p(\boldsymbol\theta)
+	\, d\mathbf{x}
+\end{equation}
+%
+The integral above could be written in terms of the matched filter signals $\left\{y_i\right\}_i$, such that $y_i\left(\hat\tau_i, \hat{\boldsymbol\theta}_\mathrm{in}\right) = \hat\rho_i e^{\hat\gamma_i}$ and $\forall t$, $\left|y_i(t, \boldsymbol\theta_\mathrm{in})\right| \leq \hat\rho_i$.
+The combination of the nontrivial integration limits and the Fourier transform in the matched filter seem like a problem for direct evaluation.
 
-In order to estimate the \ac{ML} estimator accuracy for a matched filter, we will momentarily consider the likelihood for a single detector, and suppress the intrinsic degrees of freedom:
+However, even without directly evaluating the distribution of the \ac{ML} estimator, we know many of its properties.
+The \ac{CRLB} gives its covariance in the asymptotic limit of high \ac{SNR}.
+The \ac{CRLB} has been widely applied in \ac{GW} data analysis \cite{fairhurst:2009}.
+We will momentarily consider the likelihood for a single detector:
 %
 \begin{equation}\label{eq:gaussian-likelihood-spa}
-	\fl\mathcal{L}(x_i; \rho_i, \gamma_i, \tau_i) = p(x_i | \rho_i, \gamma_i, \tau_i)
+	\mathcal{L}\left(x_i; \rho_i, \gamma_i, \tau_i,
+		\boldsymbol\theta_\mathrm{in}\right)
 	\propto \exp \left[
 		- \frac{1}{2} \int_0^\infty \frac{\left|x_i (\omega)
-			- h_i(\omega; \rho_i, \gamma_i, \tau_i) \right|^2}{S_i(\omega)} \, \mathrm{d}\omega
+			- h_i\left(x_i; \rho_i, \gamma_i, \tau_i,
+				\boldsymbol\theta_\mathrm{in}\right)
+		\right|^2}{S_i(\omega)} \, d\omega
 	\right],
@@
 %
-with $h_i(\omega; \rho_i, \gamma_i, \tau_i)$ given by \Eref{eq:signal-model}.
+with $h_i(\omega; \rho_i, \gamma_i, \tau_i, \boldsymbol\theta_\mathrm{in})$ given by \Eref{eq:signal-model}.
 
@@
         \left(\frac{\partial h_i}{\partial \theta_k}\right)
-	\right] \frac{1}{S_i(\omega)} \, \mathrm{d}\omega.
+	\right] \frac{1}{S_i(\omega)} \, d\omega.
 \end{equation}
@@
     {\overline{\omega^k}}_i =
-        \left[ \int_0^\infty \frac{|h (\omega)|^2}{S_i(\omega)} \omega^k \, \mathrm{d}\omega \right]
-        \left[ \int_0^\infty \frac{|h (\omega)|^2}{S_i(\omega)} \, \mathrm{d}\omega \right]^{-1},
+        \left[ \int_0^\infty \frac{|h (\omega)|^2}{S_i(\omega)} \omega^k \, d\omega \right]
+        \left[ \int_0^\infty \frac{|h (\omega)|^2}{S_i(\omega)} \, d\omega \right]^{-1},
 \end{equation}
 %
-the Fisher matrix for the signal model in \Eref{eq:signal-model} is
+the Fisher matrix for the signal in the $i$th detector is
+%
+\begin{equation}
+	\mathcal{I}_i = \left(
+		\begin{array}{cc}
+			\mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_i} &
+			\mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_\mathrm{in}} \\
+			{\mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_\mathrm{in}}}^\intercal &
+			{\rho_i}^2 \mathcal{I}_{\boldsymbol\theta_\mathrm{in},\boldsymbol\theta_\mathrm{in}}
+		\end{array}
+	\right)
+\end{equation}
+%
+where
 %
 \begin{equation}\label{eq:fisher-matrix}
-	\mathcal{I} = \bordermatrix{
+	\mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_i} = \bordermatrix{
         ~ & \rho_i & \gamma_i & \tau_i \cr
@@
 \end{equation}
+%
+To lowest post\nobreakdashes-Newtonian order, the intrinsic parameters affect the phase evolution but not the amplitude evolution, so for BNS systems we can assume $\mathcal{I}_{\rho_i,\boldsymbol\theta_\mathrm{in}} \approx 0$.
+%
+\marginpar{Show this?}%
+%
+When we form the Fisher matrix for the whole network, if the detectors all have the same noise \acp{PSD}, $S_1(\omega) = S_1(\omega) = \cdots = S_n(\omega) \equiv S(\omega)$, then we can change variables from phases and times on arrival $\left\{\gamma_i, \tau_i\right\}_i$ to appropriate differences between them, in order to cancel out all off-diagonal terms between the intrinsic and extrinsic parameters.
+Because all existing ground-based GW detectors have similar bandpasses, we restrict all of our further attention to just the extrinsic part of the Fisher matrix, $\mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_i}$.
 
@@
 
-If $\hat{\boldsymbol\theta}$ is an unbiased estimator of $\boldsymbol\theta$, $\tilde{\boldsymbol\theta} = \hat{\boldsymbol\theta} - \boldsymbol\theta$ is the measurement error, and $\Sigma = \mathrm{E} \, [\tilde{\boldsymbol\theta}\tilde{\boldsymbol\theta}^\intercal]$ is the covariance of the measurement error, then the \ac{CRLB} says that
-%
-\begin{equation}
-    \Sigma \geq \mathcal{I}^{-1}
-\end{equation}
-%
-in the sense that $\left(\Sigma - \mathcal{I}^{-1}\right)$ is positive semi\nobreakdashes-definite. Specifically, for our likelihood, the \ac{CRLB} implies that
+If $\hat{\boldsymbol\theta}$ is an unbiased estimator of $\boldsymbol\theta$, $\tilde{\boldsymbol\theta} = \hat{\boldsymbol\theta} - \boldsymbol\theta$ is the measurement error, and $\Sigma = \mathrm{E} \, [\tilde{\boldsymbol\theta}\tilde{\boldsymbol\theta}^\intercal]$ is the covariance of the measurement error, then the \ac{CRLB} says that $\Sigma \geq \mathcal{I}^{-1}$, in the sense that $\left(\Sigma - \mathcal{I}^{-1}\right)$ is positive semi\nobreakdashes-definite.
+Specifically, for our likelihood, the \ac{CRLB} implies that
 %
@@
 \marginpar{It's difficult to follow from here where the timing uncertainty is used in the following section.}
-
-\marginpar{Need to introduce Barankin bound.}
 
@@
 	{D_\mathrm{L}}^m
-	\,\mathrm{d}D_\mathrm{L}
-	\,\mathrm{d}u
-	\,\mathrm{d}\psi.
+	\,dD_\mathrm{L}
+	\,du
+	\,d\psi.
 \end{equation*}

### 2014-05-28 ms.tex
Describe BAYESTAR likelihood

@@
 %
-\begin{equation}
+\begin{equation}\label{eq:fisher-matrix-extrinsic-one-detector}
     \mathcal{I}'' = \bordermatrix{
@@
 
-\subsection{Noise model for point parameters}
+\section{The \ac{BAYESTAR} likelihood}
 
-The likelihood in \Eref{eq:gaussian-likelihood-spa} is the quantity that is maximized by the matched\nobreakdashes-filter estimator.
-However, just because we have the covariance matrix defined by \Eref{eq:covariance-matrix} does not mean that the distribution of estimator errors is multivariate Gaussian.
-Nonetheless, it is straightforward to qualitatively describe the distribution.
+Although we may not be able to compute the detection candidate likelihood, \Eref{eq:detection-candidate-likelihood}, directly, we can make an educated guess that has many properties in common with it.
+Any valid approximate likelihood must have the same Fisher matrix \Eref{eq:fisher-matrix-extrinsic-one-detector}.
+It must also have the same limiting behavior: it should be periodic in the phase error $\tilde{\gamma}_i$ and go to zero as $\tilde{\tau}_i \rightarrow \pm \infty$, $\hat{\rho}_i \rightarrow 0$, or $\hat{\rho}_i \rightarrow \infty$.
+Additionally, when $\tilde{\tau}_i = 0$, the distribution of ${\hat{\rho}_i}^2$ should reduce to a noncentral $\chi^2$ distribution with two degrees of freedom, centered about ${\rho_i}^2$, because the complex matched filter time series $y_i(t)$ is Gaussian.
 
-If we were to suppress the time parameter $\tau$, then we could reduce the likelihood to the form
+These conditions could be satisfied by realizing a multivariate Gaussian distribution with covariance matrix $\Sigma = \mathcal{I}^\intercal$, and then replacing individual terms of the form $-\tilde{\theta}^2/2$ with $\cos{\tilde{\theta}}$.
+
+\marginpar{Actually, this does not give exactly the same Fisher matrix. Explain.}%
+%
+Another way is to plug the signal model from \Eref{eq:signal-model} \emph{evaluated at the ML parameter estimates} into the single-detector likelihood in \Eref{eq:gaussian-likelihood-spa}:
+%
+\begin{eqnarray*}
+	\fl p\left(\hat{\boldsymbol\theta}_i \middle| \boldsymbol\theta \right)
+	&:=&
+	p\left(x_i(\omega) = h_i(\omega; \hat{\boldsymbol\theta}_i)
+		\middle| \boldsymbol\theta \right)
+	\\\fl
+	&\propto& \exp \left[
+		- \frac{1}{2} \int_0^\infty \frac{\left|
+			\frac{\hat{\rho}_i}{\sigma_i(\hat{\boldsymbol\theta}_\mathrm{in})} e^{i (\hat\gamma_i - \omega \hat\tau_i)} h(\omega; \hat{\boldsymbol\theta}_\mathrm{in})
+			- \frac{\rho_i}{\sigma_i(\boldsymbol\theta_\mathrm{in})} e^{i (\gamma_i - \omega \tau_i)} h(\omega; \boldsymbol\theta_\mathrm{in})
+		\right|^2}{S_i(\omega)} \, d\omega
+	\right].
+\end{eqnarray*}
+%
+If we assume that $\hat{\boldsymbol\theta}_\mathrm{in} = \boldsymbol\theta_\mathrm{in}$, then this reduces to
 %
 \begin{equation}
-    \mathcal{L}(z; \rho, \gamma) \propto \exp \left[ -\frac{1}{2} \left|z - e^{i \gamma} \rho \right|^2 \right]
+	p\left(\hat\rho_i, \hat\gamma_i, \hat\tau_i
+		\middle| \rho_i, \gamma_i, \tau_i \right) \propto
+	\exp \left[ - \frac{1}{2}{\hat\rho_i}^2 - \frac{1}{2}{\rho_i}^2
+		+ \hat\rho_i \rho_i \Re \left\{ e^{i \tilde{\gamma}_i} a^*(\tilde{\tau}_i)
+		\right\}
+	\right]
 \end{equation}
 %
-for a complex scalar measurement $z$.
-This likelihood also describes the distribution of the length $\rho$ and angle $\gamma$ of a vector whose Cartesian components are perturbed with normally distributed errors.
-For this problem, $\rho$ and $\gamma$ are uncorrelated, $\rho^2$ being noncentrally $\chi^2$\nobreakdashes-disributed with two degrees of freedom and $\gamma$ being von~Mises\nobreakdashes-distributed.%
-%
-\marginpar{Picture here?}
-
-\marginpar{What about if we suppress $\gamma$ or $\rho$? Do those cases teach us anything?}%
-%
-We know that the matched\nobreakdashes-filter estimator's distribution, given a particular true value of $\rho$, $\gamma$, and $\tau$, must go to zero at $\hat{\rho}=0$ and must be periodic in $\gamma$ but not $\tau$.
-From just these properties, the estimator's distribution should be an approximate Gaussian, but wrapped diagonally around a cylinder where the radius represents $\hat{\rho}$, the angle around the cylinder represents $\hat{\gamma}$, and the distance up or down the cylinder represents $\hat{\tau}$.
-
-There are many ways to realize a distribution function with these properties, but they tend to be difficult to work with.
-%
-\marginpar{Give some examples.}%
-%
-An alternative is to just discard the one periodic parameter, $\gamma$, entirely.
-We do not know the exact distribution of the remaining parameters $\rho$ and $\tau$, but we are less liable to make egregious errors by assuming that they are Gaussian than if we assumed that the estimation errors in $\gamma$ were Gaussian.
-Projecting out $\gamma$ gives us the covariance matrix
+with $\tilde{\gamma}_i = \hat\gamma_i - \gamma_i$, $\tilde{\tau}_i = \hat\tau_i - \tau_i$, and the template's autocorrelation function $a(t; \boldsymbol\theta_\mathrm{in})$ defined as
 %
 \begin{equation}
-    \cov \left(
-        \begin{array}{c}
-            \hat{\rho} - \rho \\
-            \hat{\tau} - \tau
-        \end{array}
-    \right) =
-    \left(
-        \begin{array}{cc}
-            1 & 0 \\
-            0 & \rho^2/{\omega_\mathrm{rms}}^2
-        \end{array}
-    \right).
+	a(t; \boldsymbol\theta_\mathrm{in}) := \frac{1}{\sigma_i(\boldsymbol\theta_\mathrm{in})} \int_0^\infty \frac{\left| h(\omega; \hat{\boldsymbol\theta}_\mathrm{in})\right|^2}{S_i(\omega)} e^{i \omega t} \,d\omega.
 \end{equation}
 
-A wrinkle is that this covariance matrix still depends on the true, unknown value of $\rho$.
-If we make the further leap of evaluating the \ac{CRLB} \emph{at the value of the estimators}, we can then write the distribution of the estimator error as
+To assemble the joint likelihood for the whole network, we just form the product from the individual detectors:
 %
-\begin{eqnarray}
-    \hat{\rho} - \rho \sim \mathcal{N}\, [0, 1] \\
-    \hat{\tau} - \tau \sim \mathcal{N} \left[0, \hat{\rho}^2 / {\omega_\mathrm{rms}}^2\right],
-\end{eqnarray}
+\begin{equation}
+	\fl p\left(\left\{\hat\rho_i, \hat\gamma_i, \hat\tau_i\right\}_i
+		\middle| \left\{\rho_i, \gamma_i, \tau_i\right\}_i \right) \propto
+	\exp \left[ - \frac{1}{2} \sum_i {\hat\rho_i}^2 - \frac{1}{2} \sum_i {\rho_i}^2
+		+ \sum_i \hat\rho_i \rho_i \Re \left\{ e^{i \tilde{\gamma}_i} a^*(\tilde{\tau}_i)
+		\right\}
+	\right].
+\end{equation}
 %
-valid at moderate to high \ac{SNR}.
-
-\marginpar{It's difficult to follow from here where the timing uncertainty is used in the following section.}
+To marginalize over the coalescence phase, we can insert an arbitrary nuisance phase $\eta$ by making the replacement $\tilde{\gamma}_i \rightarrow \tilde{\gamma}_i + \eta$.
+Then integrating over $\eta$ from $0$ to $2\pi$ and suppressing normalization factors, we get the \ac{BAYESTAR} likelihood
+%
+\begin{equation}
+	\mathcal{L}_\mathrm{BAYESTAR} = \exp \left[ - \frac{1}{2} \sum_i {\rho_i}^2 \right] I_0 \left[
+		\left| \sum_i \hat\rho_i \rho_i e^{i \tilde{\gamma}_i} a^*(\tilde{\tau}_i)
+		\right|
+	\right],
+\end{equation}
+%
+where $I_0$ is a modified Bessel function of the first kind.
 

### 2014-05-28 ms.tex
Fine-tune notation for BAYESTAR likelihood

@@
 	\exp \left[ - \frac{1}{2}{\hat\rho_i}^2 - \frac{1}{2}{\rho_i}^2
-		+ \hat\rho_i \rho_i \Re \left\{ e^{i \tilde{\gamma}_i} a^*(\tilde{\tau}_i)
+		+ \hat\rho_i \rho_i \Re \left\{ e^{i \tilde{\gamma}_i} a_i^*(\tilde{\tau}_i)
 		\right\}
@@
 %
-with $\tilde{\gamma}_i = \hat\gamma_i - \gamma_i$, $\tilde{\tau}_i = \hat\tau_i - \tau_i$, and the template's autocorrelation function $a(t; \boldsymbol\theta_\mathrm{in})$ defined as
+with $\tilde{\gamma}_i = \hat\gamma_i - \gamma_i$, $\tilde{\tau}_i = \hat\tau_i - \tau_i$, and the template's autocorrelation function $a_i(t; \boldsymbol\theta_\mathrm{in})$ defined as
 %
 \begin{equation}
-	a(t; \boldsymbol\theta_\mathrm{in}) := \frac{1}{\sigma_i(\boldsymbol\theta_\mathrm{in})} \int_0^\infty \frac{\left| h(\omega; \hat{\boldsymbol\theta}_\mathrm{in})\right|^2}{S_i(\omega)} e^{i \omega t} \,d\omega.
+	a_i(t; \boldsymbol\theta_\mathrm{in}) := \frac{1}{{{\sigma_i}^2(\boldsymbol\theta_\mathrm{in})}} \int_0^\infty \frac{\left| h(\omega; \hat{\boldsymbol\theta}_\mathrm{in})\right|^2}{S_i(\omega)} e^{i \omega t} \,d\omega.
 \end{equation}
@@
 %
+\marginpar{Different symbol? Is $\mathcal{L}_\mathrm{BAYESTAR}$ too grandiose?}
+%
 \begin{equation}
 	\mathcal{L}_\mathrm{BAYESTAR} = \exp \left[ - \frac{1}{2} \sum_i {\rho_i}^2 \right] I_0 \left[
-		\left| \sum_i \hat\rho_i \rho_i e^{i \tilde{\gamma}_i} a^*(\tilde{\tau}_i)
+		\left| \sum_i \hat\rho_i \rho_i e^{i \tilde{\gamma}_i} a_i^*(\tilde{\tau}_i)
 		\right|

### 2014-05-28 ms.tex
Remove some obsolete text

@@
 
-\marginpar{Explain why none of these expressions depend on the coalescence phase $\phi_c$.}%
-%
-If we assume that the time and amplitude estimation errors are uncorrelated, then we can separate the likelihood into one factor that depends only on time and one that depends only on amplitude:
-%
-\begin{eqnarray*}
-    \mathcal{L}(
-        \hat{\tau}_1, \hat{\tau}_2, &\dots, \hat{\tau}_N,
-        \hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
-        \mathbf{n}, t_\oplus, D_\mathrm{L}, u, \psi) \\
-    &= \mathcal{L}_\mathrm{TOA}(
-        \hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N; \mathbf{n}, t_\oplus) \\
-        &\times
-        \mathcal{L}_\mathrm{SNR}(\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
-         \mathbf{n}, D_\mathrm{L}, u, \psi
-    ).
-\end{eqnarray*}
-%
-The \ac{TOA}-only factor depends only on the \ac{TOA} estimates $\hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N$, the unknown sky location $\mathbf{n}$ (or equivalently, the right ascension $\alpha$, declination $\delta$, and sidereal time), and the absolute \ac{TOA} at an arbitrary reference point (herein, the geocenter), $t_\oplus$, the fixed geographical locations of the detectors, $\mathbf{r}_i$, and the speed of light $c$.
-%
-\begin{equation}
-    \mathcal{L}_\mathrm{TOA} \propto
-    \exp \left[
-            -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i - \tau_i\right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
-        \right]
-    = \exp \left[
-        -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i + \mathbf{n} \cdot \mathbf{r}_i / c - t_\oplus \right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
-    \right]
-\end{equation}
-%
-The \ac{SNR}-only factor depends only on the \ac{SNR} estimates $\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N$, but it depends on the unknown sky location, luminosity distance $D_\mathrm{L}$, cosine-inclination $u$, and polarization angle $\psi$.
-%
-\begin{equation}
-    \mathcal{L}_\mathrm{SNR} \propto \exp \left[
-        -\frac{1}{2} \sum_{i=1}^N \left(\hat{\rho}_i - \rho_i(\mathbf{n}, D_\mathrm{L}, u, \psi) \right)^2
-    \right]
-\end{equation}
-%
-It is easily shown from \Eref{eq:full-signal-model} that
-%
-\begin{eqnarray}\label{eq:snr-model}
-    {\rho_i}^2 &=
-        \frac{1}{8} \left(\frac{D_{1,i}}{D_\mathrm{L}}\right)^2 \left\{
-            ( {F_{+,i}}^2 + {F_{\times,i}}^2 \right)
-            \left(1 + 6 u^2 + u^4 \right) \nonumber\\
-            &+ \left(1 - u^2 \right)^2 \left[\left( {F_{+,i}}^2 - {F_{\times,i}}^2 \right) \cos {4\psi}
-            + 2 \left( {F_{+,i}} {F_{\times,i}} ) \sin {4\psi} \right]
-        \right\}.
-\end{eqnarray}
-%
-A similar expression appears in Appendix~A of~\cite{PhysRevD.87.064033}.
-
 \subsection{Prior}
@@
 	t_\oplus &\in& [-\infty, \infty] \\
-	u &\in& [0, 1] \\
+	u &\in& [-1, 1] \\
 	\psi &\in& [0, \pi/2)
 \end{eqnarray*}
-%
-The prior does not permit negative values of $u$ because only even powers of $u$ appear in \Eref{eq:snr-model}.
-The prior only permits $\psi$ to run to $\pi/2$ because only trigonometric functions of $4\psi$ appear in \Eref{eq:snr-model}.
-
-\subsection{Marginalization over $t_\oplus$}
-
-The overall arrival time $t_\oplus$ can be marginalized out analytically, giving the \ac{TOA}\nobreakdashes-only marginal posterior
-%
-\begin{equation}\label{eq:toa-posterior}
-    p(\mathbf{n} | \hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N) \propto \exp\left[
-        -\frac{1}{2} \sum_{i=1}^N \frac{\left( \tilde{\tau}_i - \langle \tilde{\tau} \rangle \right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
-    \right]
-\end{equation}
-%
-where
-%
-\begin{equation}
-    \langle \tilde{\tau} \rangle =
-    \left(\sum_{i=1}^N \frac{{\tilde{\tau}_i}^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}\right)\left(\sum_{i=1}^N \frac{1}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}\right)^{-1}.
-\end{equation}
 

### 2014-05-29 scripts/radial_integrand.py
Move scripts into subdirectory

@@ -1,4 +1,3 @@
-#!/usr/bin/env python
 from __future__ import division
 """
 Create a figure that depicts the subdivision technique for the radial integral.
@@ -10,7 +9,6 @@ from mpl_toolkits.axes_grid1.inset_locator import BboxConnectorPatch
 from matplotlib.transforms import Bbox, TransformedBbox, blended_transform_factory
 from matplotlib import ticker
 import numpy as np
-import sys
 
 # Input parameters
 rmin = 0
@@ -77,8 +75,4 @@ for i, (break1, break2) in enumerate(zip(breakpoints[:-1], breakpoints[1:])):
     ax.xaxis.set_major_locator(ticker.MultipleLocator(1))
 plt.ylim(-.1, 1.2)
 
-
-if len(sys.argv) > 1:
-    plt.savefig(sys.argv[1])
-else:
-    plt.show()
+plt.savefig('radial_integrand.pdf')

### 2014-05-29 ms.tex
Add Fisher matrix elements

@@
 %
-\begin{equation}
+\begin{equation}\label{eq:autocor-likelihood}
 	p\left(\hat\rho_i, \hat\gamma_i, \hat\tau_i
@@
 where $I_0$ is a modified Bessel function of the first kind.
+
+\subsection{Properties}
+
+Fisher matrix elements of \Eref{eq:autocor-likelihood}, with detector subscript suppressed:
+%
+\begin{eqnarray}
+	\nonumber
+	\mathcal{I}_{\rho\rho} &=& 1 \\
+	\nonumber
+	\mathcal{I}_{\rho\gamma} &=& 0 \\
+	\nonumber
+	\mathcal{I}_{\rho\tau} &=& 0 \\
+	\mathcal{I}_{\gamma\gamma} &=& \frac{\rho^2}{2}
+		\int_{-\infty}^\infty \left|a(t)\right|^2 w(t; \rho) dt \\
+	\mathcal{I}_{\tau\tau} &=& -\frac{\rho^2}{2}
+		\int_{-\infty}^\infty \Re\left[a^*(t) a''(t)\right] w(t; \rho) dt \\
+	\mathcal{I}_{\gamma\tau} &=& -\frac{\rho^2}{2}
+		\int_{-\infty}^\infty \Im\left[a^*(t) a'(t)\right] w(t; \rho) dt \\
+	\nonumber\textrm{where} && \\
+	\nonumber
+	a'(t) &=& \frac{\partial a}{\partial t} \\
+	\nonumber
+	a''(t) &=& \frac{\partial^2 a}{\partial t^2} \\
+	\nonumber
+	w(t; \rho) &=& \frac{
+		\displaystyle
+		\exp\left[\frac{\rho^2}{4}\left|a(t)\right|^2\right]
+		\left(
+		I_0\left[\frac{\rho^2}{4}\left|a(t)\right|^2\right] +
+		I_1\left[\frac{\rho^2}{4}\left|a(t)\right|^2\right]
+		\right)
+	}{
+		\displaystyle
+		\int_{-\infty}^\infty
+		\exp\left[\frac{\rho^2}{4}\left|a(t')\right|^2\right]
+		I_0\left[\frac{\rho^2}{4}\left|a(t')\right|^2\right]
+		dt'
+	}.
+\end{eqnarray}
+
+If we can show that the common factor approaches a delta function, $w(t; \rho) \rightarrow \delta(t)$ as $\rho \rightarrow \infty$, then the Fisher matrix approaches \Eref{eq:fisher-matrix} as $\rho \rightarrow \infty$.
 

### 2014-05-29 ms.tex
Fold factor of 1/2 into weight function
Some arguments why w(t,r) approaches a delta function

@@
 	\mathcal{I}_{\rho\tau} &=& 0 \\
-	\mathcal{I}_{\gamma\gamma} &=& \frac{\rho^2}{2}
+	\mathcal{I}_{\gamma\gamma} &=& \rho^2
 		\int_{-\infty}^\infty \left|a(t)\right|^2 w(t; \rho) dt \\
-	\mathcal{I}_{\tau\tau} &=& -\frac{\rho^2}{2}
+	\mathcal{I}_{\tau\tau} &=& -\rho^2
 		\int_{-\infty}^\infty \Re\left[a^*(t) a''(t)\right] w(t; \rho) dt \\
-	\mathcal{I}_{\gamma\tau} &=& -\frac{\rho^2}{2}
+	\mathcal{I}_{\gamma\tau} &=& -\rho^2
 		\int_{-\infty}^\infty \Im\left[a^*(t) a'(t)\right] w(t; \rho) dt \\
@@
 		\displaystyle
-		\int_{-\infty}^\infty
+		2 \int_{-\infty}^\infty
 		\exp\left[\frac{\rho^2}{4}\left|a(t')\right|^2\right]
@@
 \end{eqnarray}
+%
+Matrix elements that are not listed have values that are implied by the symmetry of the Fisher matrix.
+Note that the minus signs are correct but a little confusing: despite them, ${I}_{\gamma\gamma}, {I}_{\tau\tau} \geq 0$ and ${I}_{\gamma\tau} \leq 0$.
 
 If we can show that the common factor approaches a delta function, $w(t; \rho) \rightarrow \delta(t)$ as $\rho \rightarrow \infty$, then the Fisher matrix approaches \Eref{eq:fisher-matrix} as $\rho \rightarrow \infty$.
+Certainly it behaves very much like a delta function.
+The Bessel functions asymptotically approach:
+%
+\begin{equation*}
+	I_0(x), I_1(x) \rightarrow \frac{e^x}{\sqrt{2 \pi x}}
+	\textrm{ as } x \rightarrow \infty.
+\end{equation*}
+%
+For large $\rho$, it seems like we could write:
+%
+\begin{equation*}
+	w(t; \rho) \rightarrow \frac{
+		\displaystyle
+		\exp\left[\frac{\rho^2}{2}|a(t)|^2\right]
+	}{
+		\displaystyle
+		\int_{-\infty}^\infty \exp\left[\frac{\rho^2}{2}|a(t')|^2\right] dt'
+	}
+	\textrm{ as } \rho \rightarrow \infty.
+\end{equation*}
 

### 2014-05-29 ms.tex
Add value of autocorrelation function and derivatives at t=0
Clean up signal model

@@
 
-Assuming circular orbits and no spin precession, we can write the \ac{GW} signal in each detector as a linear combination of two basis waveforms $h_0$ and $h_{\pi/2}$.
-For nonprecessing systems, $h = h_0$ and $h_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the sine and cosine functions, being nearly orthogonal and out of phase by ${\pi/2}$ at all frequencies.
-In other words, $h_{\pi/2}(\omega) = i h_0(\omega)$ for all $\omega$.
-This linear combination isolates all dependence on the extrinsic parameters $\boldsymbol\theta_\mathrm{ex}$ into the coefficients and all dependence on the intrinsic parameters $\boldsymbol\theta_\mathrm{in}$ into the basis waveform,
+Assuming circular orbits and no spin precession, we can write the \ac{GW} signal in each detector as a linear combination of two basis waveforms $h_0$ = $h$ and $h_{\pi/2}$.
+For nonprecessing systems, $h_0$ and $h_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the sine and cosine functions, being nearly orthogonal and out of phase by ${\pi/2}$ at all frequencies.
+If $h_0$ and $h_{\pi/2}$ are Fourier transforms of real functions, then $h_0(\omega) = h_0^*(-\omega)$ and $h_{\pi/2}(\omega) = h_{\pi/2}^*(-\omega)$, and we can write (assuming an arbitrary phase convention)
 %
-\begin{eqnarray}\label{eq:full-signal-model}
-	\fl h_i(\omega; \boldsymbol\theta) &=& e^{-i\omega t_\oplus}\frac{D_{1,i}}{D_\mathrm{L}}
-	\left(\begin{array}{cc}
-		F_{+,i}(\alpha, \delta, t_\oplus) \\
-		F_{\times,i}(\alpha, \delta, t_\oplus)
-	\end{array}\right)^\intercal
-	\left(\begin{array}{cc}
-		\cos{2\psi} & -\sin{2\psi} \\
-		\sin{2\psi} & \cos{2\psi}
-	\end{array}\right)
-	\left(\begin{array}{cc}
-		\frac{1}{2}\left(1+u^2\right) & 0 \\
-		0 & u
-	\end{array}\right) \nonumber\\\fl &\cdot&
-	\left(\begin{array}{cc}
-		\cos{2\phi_c} & -\sin{2\phi_c} \\
-		\sin{2\phi_c} & \cos{2\phi_c}
-	\end{array}\right)
-	\left(\begin{array}{cc}
-		1 \\
-		i
-	\end{array}\right) h(\omega; \boldsymbol\theta_\mathrm{in}).
-\end{eqnarray}
+\begin{equation}
+	h_{\pi/2}(\omega) = h_0(\omega) \cdot
+	\cases{
+		-i & if $\omega \geq 0$ \\
+		i & if $\omega < 0$
+	}.
+\end{equation}
 %
-This linear combination matches the conventions used in \cite{PhysRevD.83.084002}, and is the frequency-domain equivalent of their Equation~(2.8).
+Then, we can write the signal model in a way that isolates all dependence on the extrinsic parameters $\boldsymbol\theta_\mathrm{ex}$ into the coefficients and all dependence on the intrinsic parameters $\boldsymbol\theta_\mathrm{in}$ into the basis waveform, by taking the Fourier transform of Equation~(2.8) of \cite{PhysRevD.83.084002}:
+%
+\begin{equation}\label{eq:full-signal-model}
+	h_i(\omega; \boldsymbol\theta) = \e^{-i \omega t_\oplus}
+	\frac{D_{1,i}}{D_\mathrm{L}}
+	\e^{2 i \phi_c}
+	\left[
+	\frac{1}{2} \left(1 + \cos^2 \iota\right) \Re \left\{z\right\} - i
+	\left(\cos\iota\right) \Im z
+	\right]
+	h(\omega; \boldsymbol\theta_\mathrm{in})
+\end{equation}
+%
+for $\omega \geq 0$, where
+%
+\begin{equation}
+	z = e^{-2 i \psi} \left(
+	F_{+,i}(\alpha, \delta, t_\oplus) +
+	i F_{\times,i}(\alpha, \delta, t_\oplus)
+	\right).
+\end{equation}
+%
 The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, and depend on the orientation of detector $i$ as well as the sky location and sidereal time of the event, and are derived in~\cite{PhysRevD.63.042003}.
 The quantity $D_{1,i}$ is a fiducial distance at which detector $i$ would register SNR=1 for an optimally oriented binary (face\nobreakdashes-on, and in a direction perpendicular to the interferometer's arms),
+%
+\marginpar{I seem to have dropped a factor of 2 here...}%
 %
@@
 
+Observe that
+%
+\begin{equation*}
+	a(0) = 1,
+	\qquad
+	a'(0) = i \overline{\omega},
+	\qquad
+	a''(0) = -\overline{\omega^2}.
+\end{equation*}
+%
 If we can show that the common factor approaches a delta function, $w(t; \rho) \rightarrow \delta(t)$ as $\rho \rightarrow \infty$, then the Fisher matrix approaches \Eref{eq:fisher-matrix} as $\rho \rightarrow \infty$.

### 2014-05-30 ms.tex
Fisher matrix uses conditional expectation value
Fish factors!

@@
 \usepackage[colorlinks]{hyperref}
+\usepackage{protosem}
 \input{ligo-acronyms/acronyms}
@@
 			{\partial \theta_j \theta_k}
+	\middle| \boldsymbol\theta
 	\right].
@@
 
-Fisher matrix elements of \Eref{eq:autocor-likelihood}, with detector subscript suppressed:
+We can compute the Fisher matrix elements for the autocorrelation likelihood given by \Eref{eq:autocor-likelihood}, with detector subscript suppressed:
 %
@@
 	\mathcal{I}_{\rho\tau} &=& 0 \\
+	\label{eq:fisher-autocor-gamma-gamma}
 	\mathcal{I}_{\gamma\gamma} &=& \rho^2
-		\int_{-\infty}^\infty \left|a(t)\right|^2 w(t; \rho) dt \\
+		\int_{-T}^T \left|a(t)\right|^2 w(t; \rho) dt \\
+	\label{eq:fisher-autocor-tau-tau}
 	\mathcal{I}_{\tau\tau} &=& -\rho^2
-		\int_{-\infty}^\infty \Re\left[a^*(t) a''(t)\right] w(t; \rho) dt \\
+		\int_{-T}^T \Re\left[a^*(t) a''(t)\right] w(t; \rho) dt \\
+	\label{eq:fisher-autocor-gamma-tau}
 	\mathcal{I}_{\gamma\tau} &=& -\rho^2
-		\int_{-\infty}^\infty \Im\left[a^*(t) a'(t)\right] w(t; \rho) dt \\
+		\int_{-T}^T \Im\left[a^*(t) a'(t)\right] w(t; \rho) dt \\
 	\nonumber\textrm{where} && \\
@@
 		\displaystyle
-		2 \int_{-\infty}^\infty
+		2 \int_{-T}^T
 		\exp\left[\frac{\rho^2}{4}\left|a(t')\right|^2\right]
@@
 Note that the minus signs are correct but a little confusing: despite them, ${I}_{\gamma\gamma}, {I}_{\tau\tau} \geq 0$ and ${I}_{\gamma\tau} \leq 0$.
+The time integration limits $[-T, T]$ constitute a flat prior on arrival time, the time coincidence window between detectors.
 
@@
 %
-If we can show that the common factor approaches a delta function, $w(t; \rho) \rightarrow \delta(t)$ as $\rho \rightarrow \infty$, then the Fisher matrix approaches \Eref{eq:fisher-matrix} as $\rho \rightarrow \infty$.
-Certainly it behaves very much like a delta function.
+If we can show that the common factor approaches a delta function, $w(t; \rho) \rightarrow \delta(t)$ as $\rho \rightarrow \infty$, then the Fisher matrix for the autocorrelation likelihood approaches the Fisher matrix for the full GW data, \Eref{eq:fisher-matrix}, as $\rho \rightarrow \infty$.
 The Bessel functions asymptotically approach:
@@
 %
-For large $\rho$, it seems like we could write:
+For large $\rho$, the exponential dominates and we can write:
 %
@@
 		\displaystyle
-		\int_{-\infty}^\infty \exp\left[\frac{\rho^2}{2}|a(t')|^2\right] dt'
+		\int_{-T}^T \exp\left[\frac{\rho^2}{2}|a(t')|^2\right] dt'
 	}
@@
 \end{equation*}
+%
+The Taylor expansion of $|a(t)|^2$ is
+%
+\begin{equation*}
+	|a(t)|^2 = 1 + \frac{1}{2} \left(\frac{\partial^2}{\partial t^2}|a(t)|^2 \Bigg|_{t=0}\right) t^2 + \mathcal{O}(t^4)
+	= 1 - {\omega_\mathrm{rms}}^2 t^2 + \mathcal{O}(t^4).
+\end{equation*}
+%
+Substituting, we find that $w(t; \rho)$ approaches a normalized Gaussian distribution:
+%
+\begin{equation*}
+	w(t; \rho) \approx \frac{
+		\displaystyle
+		\exp\left[-\frac{1}{2} \rho^2 {\omega_\mathrm{rms}}^2 t^2\right]
+	}{
+		\displaystyle
+		\int_{-T}^T \exp\left[-\frac{1}{2} \rho^2 {\omega_\mathrm{rms}}^2 (t')^2\right] dt'
+	}.
+\end{equation*}
+%
+And finally, because the Dirac delta function may be defined as the limit of a Gaussian, $w(t; \rho) \rightarrow \delta(t)$ as $\rho \rightarrow \infty$.
+
+We can now write the Fisher matrix for the autocorrelation likelihood in a way that makes a comparison to the full signal model explicit. Define:
+%
+\begin{eqnarray*}
+	\mathcal{I}_\gamma\gamma &=& \rho^2 \cdot \textproto{D}_{\gamma\gamma}(\rho) \\
+	\mathcal{I}_\tau\tau &=& \rho^2 \overline{\omega^2} \cdot \textproto{D}_{\tau\tau}(\rho) \\
+	\mathcal{I}_\gamma\tau &=& -\rho^2 \overline{\omega} \cdot \textproto{D}_{\gamma\tau}(\rho)
+\end{eqnarray*}
+%
+Now, the $\textproto{D}_{ij}$ %
+%
+\marginpar{The Fish(er) factor. Go home, Leo.}%
+%
+contain the integrals from Equations~(\ref{eq:fisher-autocor-gamma-gamma}, \ref{eq:fisher-autocor-tau-tau}, \ref{eq:fisher-autocor-gamma-tau}) and encode the departure of the autocorrelation likelihood from the likelihood of the full data at low SNR.
+All of the $\textproto{D}_{ij}(\rho)$ are sigmoid-type functions that run from 0 at $\rho = 0$ and asymptotically approach 1 as $\rho \rightarrow \infty$.
+The transitional SNR $\rho_\mathrm{crit}$ is largely the same for all three nontrivial matrix elements, and is determined by the time coincidence window $T$ and the signal bandwidth $\omega_\mathrm{rms}$.
+%
+\marginpar{The transitional SNR is probably proportional to some power of $(\omega_\mathrm{rms} T)$.}
 

### 2014-05-30 ms.tex
Add some more interpretation

@@
 All of the $\textproto{D}_{ij}(\rho)$ are sigmoid-type functions that run from 0 at $\rho = 0$ and asymptotically approach 1 as $\rho \rightarrow \infty$.
-The transitional SNR $\rho_\mathrm{crit}$ is largely the same for all three nontrivial matrix elements, and is determined by the time coincidence window $T$ and the signal bandwidth $\omega_\mathrm{rms}$.
+The transition SNR $\rho_\mathrm{crit}$ is largely the same for all three nontrivial matrix elements, and is determined by the time coincidence window $T$ and the signal bandwidth $\omega_\mathrm{rms}$.
 %
-\marginpar{The transitional SNR is probably proportional to some power of $(\omega_\mathrm{rms} T)$.}
+\marginpar{The transition SNR is probably proportional to some power of $(\omega_\mathrm{rms} T)$.}
+
+In the limit of large SNR, our interpretation is that the point estimates $(\hat\rho, \hat\gamma, \hat\tau)$ contain all of the information about the underlying extrinsic parameters.
+
+On the other hand, in the low SNR limit, the vanishing of $\textproto{D}_{ij}(\rho)$ reflects the fact that some information is lost when the full data $\mathbf{x}$ is discarded.
+Concretely, as the prior interval $T$ becomes large compared to 1/$\rho\omega_\mathrm{rms}$, the ML estimator becomes more and more prone to pick up a spurious noise fluctuation far from the true signal.
+Clearly, when the coincidence window $T$ is kept small as possible, more information is retained in the ML point estimates.
+Put another way, if $T$ is small, then the transition SNR $\rho_\mathrm{crit}$ is also small and fainter signals become useful for parameter estimation.
 

### 2014-06-01 scripts/autocorrelation_fisher_matrix.py
Add script to calculate Fisher matrix for autocorrelation likelihood

@@ -1,0 +1,137 @@
+"""Compute Fisher matrix for autocorrelation likelihood, using a TaylorF2
+signal model."""
+from __future__ import division
+import numpy as np
+import scipy.integrate
+import scipy.interpolate
+import scipy.special
+import multiprocessing
+import itertools
+from lalinference.bayestar import filter
+from lalinference.bayestar import timing
+
+def abs2(x):
+    return np.square(np.real(x)) + np.square(np.imag(x))
+
+mass1 = 1.4
+mass2 = 1.4
+f_low = 40
+S = timing.get_noise_psd_func('H1')
+
+# Compute Fisher matrix elements for full signal (for comparison).
+signal_model = timing.SignalModel(mass1, mass2, S, f_low,
+    *timing.get_approximant_and_orders_from_string('TaylorF2threePointFivePN'))
+w1 = signal_model.get_sn_moment(1)
+w2 = signal_model.get_sn_moment(2)
+crlb1 = 1 / np.sqrt(w2 - np.square(w1))
+
+# Compute 1 second of autocorrelation sequence.
+# FIXME: make this the longest lag that we are going to integrate over.
+acor_series, sample_rate = filter.autocorrelation(
+    mass1, mass2, S, f_low, 1,
+    *timing.get_approximant_and_orders_from_string('TaylorF2threePointFivePN'))
+
+# Augment autocorrelation sequence with negative time lags.
+n = len(acor_series)
+t = np.arange(1-n, n) / sample_rate
+acor_series = np.concatenate((np.conj(acor_series[:0:-1]), acor_series))
+
+# Construct spline interpolant and derivatives for real part.
+re_acor = scipy.interpolate.UnivariateSpline(t, acor_series.real, s=0, k=5)
+re_acor_deriv = re_acor.derivative(1)
+re_acor_deriv2 = re_acor.derivative(2)
+
+# Construct spline interpolant and derivatives for imaginary part.
+im_acor = scipy.interpolate.UnivariateSpline(t, acor_series.imag, s=0, k=5)
+im_acor_deriv = im_acor.derivative(1)
+im_acor_deriv2 = im_acor.derivative(2)
+
+# Put real and imaginary parts back together.
[…]

### 2014-06-02 ms.tex
Oops, add subscript
Minor edits

@@
 %
-The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, and depend on the orientation of detector $i$ as well as the sky location and sidereal time of the event, and are derived in~\cite{PhysRevD.63.042003}.
+The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, defined such that $0 \leq {F_{+,i}}^2 + {F_{\times,i}} \leq 1$.
+They depend on the orientation of detector $i$ as well as the sky location and sidereal time of the event, and are derived in~\cite{PhysRevD.63.042003}.
 The quantity $D_{1,i}$ is a fiducial distance at which detector $i$ would register SNR=1 for an optimally oriented binary (face\nobreakdashes-on, and in a direction perpendicular to the interferometer's arms),
@@
 %
+In other words, it is the expectation value, condition on the true parameter values, of the Jacobian of the log likelihood.
 This equation involves double derivatives, but substantial simplification is possible when---as in this case---the likelihood is Gaussian:
@@
 
+First, observe that at the true parameter values, $\hat{\boldsymbol\theta}_i = \boldsymbol\theta_i$, the logarithms of \Eref{eq:autocor-likelihood} and \Eref{eq:gaussian-likelihood-spa} have the same Jacobian.
+This is because the derivatives of the autocorrelation function are
+%
+\begin{equation*}
+    a^{(n)}(t) = i^n \overline{\omega^n};
+\end{equation*}
+%
+for instance,
+%
+\begin{equation*}
+    a(0) = 1,
+    \qquad
+    a'(0) = i \overline{\omega},
+    \qquad
+    a''(0) = -\overline{\omega^2}.
+\end{equation*}
+
 We can compute the Fisher matrix elements for the autocorrelation likelihood given by \Eref{eq:autocor-likelihood}, with detector subscript suppressed:
@@
 	\nonumber\textrm{where} && \\
-	\nonumber
-	a'(t) &=& \frac{\partial a}{\partial t} \\
-	\nonumber
-	a''(t) &=& \frac{\partial^2 a}{\partial t^2} \\
-	\nonumber
 	w(t; \rho) &=& \frac{
@@
 
-Observe that
-%
-\begin{equation*}
-	a(0) = 1,
-	\qquad
-	a'(0) = i \overline{\omega},
-	\qquad
-	a''(0) = -\overline{\omega^2}.
-\end{equation*}
-%
-If we can show that the common factor approaches a delta function, $w(t; \rho) \rightarrow \delta(t)$ as $\rho \rightarrow \infty$, then the Fisher matrix for the autocorrelation likelihood approaches the Fisher matrix for the full GW data, \Eref{eq:fisher-matrix}, as $\rho \rightarrow \infty$.
+If we can show that the weighting function $w(t; \rho)$ approaches a Dirac delta function as $\rho \rightarrow \infty$, then the Fisher matrix for the autocorrelation likelihood approaches the Fisher matrix for the full GW data, \Eref{eq:fisher-matrix}, as $\rho \rightarrow \infty$.
 The Bessel functions asymptotically approach:
@@
 \begin{eqnarray*}
-	\mathcal{I}_\gamma\gamma &=& \rho^2 \cdot \textproto{D}_{\gamma\gamma}(\rho) \\
-	\mathcal{I}_\tau\tau &=& \rho^2 \overline{\omega^2} \cdot \textproto{D}_{\tau\tau}(\rho) \\
-	\mathcal{I}_\gamma\tau &=& -\rho^2 \overline{\omega} \cdot \textproto{D}_{\gamma\tau}(\rho)
+	\mathcal{I}_{\gamma\gamma} &=& \rho^2 \cdot \textproto{D}_{\gamma\gamma}(\rho) \\
+	\mathcal{I}_{\tau\tau} &=& \rho^2 \overline{\omega^2} \cdot \textproto{D}_{\tau\tau}(\rho) \\
+	\mathcal{I}_{\gamma\tau} &=& -\rho^2 \overline{\omega} \cdot \textproto{D}_{\gamma\tau}(\rho)
 \end{eqnarray*}

### 2014-06-02 ms.tex
Add some figures

@@
 
+\begin{figure}
+    \begin{center}
+        \includegraphics{crlb_tau}
+    \end{center}
+    \caption{\label{fig:crlb-tau}\ac{CRLB} on root\nobreakdashes-mean square timing uncertainty, using the likelihood for the full GW data (Equation~\ref{eq:gaussian-likelihood-spa}; dashed diagonal line) or the autocorrelation likelihood (Equation~\ref{eq:autocor-likelihood}; solid lines) with a selection of arrival time priors.}
+\end{figure}
+
+\begin{figure}
+    \begin{center}
+        \includegraphics{fishfactor}
+    \end{center}
+    \caption{\label{fig:fishfactor}Ratio between Fisher matrix elements (solid: $\mathcal{I}_{\gamma\gamma}$, dashed: $\mathcal{I}_{\gamma\tau}$, dotted: $\mathcal{I}_{\tau\tau}$) for the the autocorrelation likelihood and the full GW data.
+Colors correspond to different arrival time priors as in Figure~\ref{fig:crlb-tau}.}
+\end{figure}
+
 \section{The \ac{BAYESTAR} algorithm}

### 2014-06-02 scripts/autocorrelation_fisher_matrix.py

@@ -5,7 +5,6 @@ import numpy as np
 import scipy.integrate
 import scipy.interpolate
 import scipy.special
-import multiprocessing
 import itertools
 from lalinference.bayestar import filter
 from lalinference.bayestar import timing
@@ -23,7 +22,6 @@ signal_model = timing.SignalModel(mass1, mass2, S, f_low,
     *timing.get_approximant_and_orders_from_string('TaylorF2threePointFivePN'))
 w1 = signal_model.get_sn_moment(1)
 w2 = signal_model.get_sn_moment(2)
-crlb1 = 1 / np.sqrt(w2 - np.square(w1))
 
 # Compute 1 second of autocorrelation sequence.
 # FIXME: make this the longest lag that we are going to integrate over.
@@ -94,18 +92,9 @@ def map_nested(map, func, *iterables):
     return np.reshape(results, [len(iterable) for iterable in iterables])
 
 if __name__ == '__main__':
-    from matplotlib import pyplot as plt
+    import json
+    import multiprocessing
 
-    # t = np.linspace(-1, 1, 32768)
-    # plt.plot(t, abs2(acor(t)), label=r'$\gamma\gamma$')
-    # plt.plot(t, np.imag(np.conj(acor(t)) * acor_deriv(t))/w1, label=r'$\gamma\tau$')
-    # plt.plot(t, -np.real(np.conj(acor(t)) * acor_deriv2(t))/w2, label=r'$\tau\tau$')
-    # plt.legend()
-    # plt.xlim(-0.01, 0.01)
-    # plt.savefig('integrand.pdf')
-
-    fig = plt.figure()
-    ax = fig.add_subplot(111, aspect=1)
     snrs = np.logspace(-1, np.log10(30), 100)
     durations = [0.001, 0.01, 0.1, 1.]
     integrands = [num_gamma_gamma, num_tau_tau, num_gamma_tau]
@@ -115,23 +104,11 @@ if __name__ == '__main__':
             pool.map, fisher_integral, durations, integrands, snrs)
     finally:
         pool.terminate()
-
-    fig = plt.figure()
-    ax = fig.add_subplot(111, aspect=1)
-    ax.set_xscale('log')
-    ax.set_yscale('log')
-    ax.set_xlabel('SNR')
-    ax.set_ylabel('RMS timing error (ms)')
-    ax.set_ylim(1e-2, 10)
-    ax.grid()
-
-    for T, data in zip(durations, integrals):
[…]

### 2014-06-02 scripts/crlb_tau.py

@@ -1,0 +1,34 @@
+from __future__ import division
+import numpy as np
+from matplotlib import pyplot as plt
+import json
+
+# Read data file
+with open('autocorrelation_fisher_matrix.json', 'r') as infile:
+    data = json.load(infile)
+
+# Add dictionary entries to local namespace
+locals().update(data)
+
+crlb1 = 1 / np.sqrt(w2 - np.square(w1))
+
+fig = plt.figure(figsize=(3.5, 3.5))
+fig.subplots_adjust(left=0.15)
+ax = fig.add_subplot(111, aspect=1)
+ax.set_xscale('log')
+ax.set_yscale('log')
+ax.set_xlabel('SNR')
+ax.set_ylabel('RMS timing error (ms)')
+ax.set_ylim(1e-2, 10)
+ax.grid()
+
+for T, matrix_elements in reversed(zip(durations, fisher_matrix_elements)):
+    F11, F22, F12 = np.asarray(matrix_elements)
+    std_tau = np.sqrt(F11 / (F11 * F22 - np.square(F12)))
+    ax.plot(snrs, 1e3 * std_tau, label=r'$T={0:g}$'.format(T))
+
+snrs = np.asarray([1e-1, 1e2])
+ax.plot(snrs, 1e3 * crlb1 / snrs, '--k', label=r'CRLB')
+
+plt.legend(loc='lower left')
+fig.savefig('crlb_tau.pdf')

### 2014-06-02 scripts/fishfactor.py

@@ -1,0 +1,28 @@
+from __future__ import division
+import numpy as np
+from matplotlib import pyplot as plt
+import json
+
+# Read data file
+with open('autocorrelation_fisher_matrix.json', 'r') as infile:
+    data = json.load(infile)
+
+# Add dictionary entries to local namespace
+locals().update(data)
+
+fig = plt.figure(figsize=(3.5, 2.5))
+fig.subplots_adjust(left=0.15, bottom=0.15)
+ax = fig.add_subplot(111)
+ax.set_xlabel('SNR')
+ax.set_ylabel('Ratio')
+ax.set_xlim(0, 8)
+ax.set_ylim(0, 1)
+
+colors = 'crgb'
+for T, matrix_elements, color in reversed(zip(durations, fisher_matrix_elements, colors)):
+    F11, F22, F12 = np.asarray(matrix_elements)
+    ax.plot(snrs, F11 / np.square(snrs), '-', color=color)
+    ax.plot(snrs, -F12 / (w1 * np.square(snrs)), '--', color=color)
+    ax.plot(snrs, F22 / (w2 * np.square(snrs)), ':', color=color)
+
+fig.savefig('fishfactor.pdf')

### 2014-06-02 ms.tex
Fisher matrix factor does not go to zero

@@
 contain the integrals from Equations~(\ref{eq:fisher-autocor-gamma-gamma}, \ref{eq:fisher-autocor-tau-tau}, \ref{eq:fisher-autocor-gamma-tau}) and encode the departure of the autocorrelation likelihood from the likelihood of the full data at low SNR.
-All of the $\textproto{D}_{ij}(\rho)$ are sigmoid-type functions that run from 0 at $\rho = 0$ and asymptotically approach 1 as $\rho \rightarrow \infty$.
+All of the $\textproto{D}_{ij}(\rho)$ are sigmoid-type functions that asymptotically approach 1 as $\rho \rightarrow \infty$.
 The transition SNR $\rho_\mathrm{crit}$ is largely the same for all three nontrivial matrix elements, and is determined by the time coincidence window $T$ and the signal bandwidth $\omega_\mathrm{rms}$.
@@
 
-On the other hand, in the low SNR limit, the vanishing of $\textproto{D}_{ij}(\rho)$ reflects the fact that some information is lost when the full data $\mathbf{x}$ is discarded.
+On the other hand, in the low SNR limit, the diminishing value of $\textproto{D}_{ij}(\rho)$ reflects the fact that some information is lost when the full data $\mathbf{x}$ is discarded.
 Concretely, as the prior interval $T$ becomes large compared to 1/$\rho\omega_\mathrm{rms}$, the ML estimator becomes more and more prone to pick up a spurious noise fluctuation far from the true signal.

### 2014-06-02 ms.tex
Specify prior

@@
                 t_\oplus \\
-                u \\
+                \iota \\
                 \psi \\
@@
                     arrival time at geocenter \\
-                    $\cos(\textrm{inclination angle})$ \\
+                    inclination angle \\
                     polarization angle \\
@@
 
-We have no explicit prior on the intrinsic parameters; in our analysis they are fixed at their \ac{ML} estimates, $\hat{\boldsymbol\theta}$.
-The distance prior is given by a user-selected power of distance, ${D_\mathrm{L}}^m$, where $m=2$ for a prior that is uniform in volume, and $m=-1$ for a scale-invariant prior that is uniform in $\log D_\mathrm{L}$.
-The priors are bounded as follows:
+We have no explicit prior on the intrinsic parameters; in our analysis they are fixed at their \ac{ML} estimates, $\hat{\boldsymbol\theta}_\mathrm{in}$.
+
+The arrival time prior is connected to the origin of the detector coordinate system.
+Given the Earth-fixed coordinates of the detectors $\mathbf{n}_i$ and the arrival times $\tau_i$, we compute their averages weighted by the timing uncertainty formula:
 %
-\marginpar{Need to discuss $D_{\mathrm{L},\mathrm{min}}$, $D_{\mathrm{L},\mathrm{max}}$, and how they are selected automatically if the user does not provide them.}%
+\begin{equation*}
+    \langle \mathbf{n} \rangle = \frac{
+        \displaystyle
+        \sum_i \frac{\mathbf{n}_i}{\hat\rho_i \omega_{\mathrm{rms},i}}
+    }{
+        \displaystyle
+        \sum_i \frac{1}{\hat\rho_i \omega_{\mathrm{rms},i}}
+    },
+    \qquad
+    \langle \hat\tau \rangle = \frac{
+        \displaystyle
+        \sum_i \frac{\hat\tau_i}{\hat\rho_i \omega_{\mathrm{rms},i}}
+    }{
+        \displaystyle
+        \sum_i \frac{1}{\hat\rho_i \omega_{\mathrm{rms},i}}
+    }.
+\end{equation*}
 %
-\begin{eqnarray*}
-	D_\mathrm{L} &\in& [D_{\mathrm{L},\mathrm{min}}, D_{\mathrm{L},\mathrm{max}}] \\
-	t_\oplus &\in& [-\infty, \infty] \\
-	u &\in& [-1, 1] \\
-	\psi &\in& [0, \pi/2)
-\end{eqnarray*}
+Then, we subtract off these means:
+%
+\begin{equation*}
+    \mathbf{n}_i \leftarrow \mathbf{n}_i - \langle \mathbf{n} \rangle,
+    \qquad
+    \hat\tau_i \leftarrow \hat\tau_i - \langle \hat\tau \rangle.
+\end{equation*}
+%
+In these coordinates relative to the weighted detector array barycenter, the arrival time prior is uniform in $-T \leq t \leq T$, with $T = \max\limits_i |\mathbf{n}_i| / c + 5\,\textrm{ms}$.
+
+The distance prior is a user-selected power of distance,
+%
+\begin{equation*}
+    p(D_\mathrm{L}) \propto \cases{
+        {D_\mathrm{L}}^m & if $D_{\mathrm{L},\mathrm{min}} < D_\mathrm{L} < D_{\mathrm{L},\mathrm{max}}$ \\
+        0 & otherwise
+    }
+\end{equation*}
+%
+where $m=2$ for a prior that is uniform in volume, and $m=-1$. If a distance prior is not specified, the default is uniform in volume out to the maximum SNR=4 horizon distance:
+%
+\begin{equation*}
+    m = 2,
+    \qquad
+    D_{\mathrm{L},\mathrm{min}} = 0,
+    \qquad
+    D_{\mathrm{L},\mathrm{max}} = \frac{1}{4} \max_i D_{1,i}.
+\end{equation*}
+
+Finally, the prior is uniform in $-1 \leq \cos\iota \leq 1$ and $0 \leq \psi < 2\pi$.
 

### 2014-06-02 ms.tex
List inputs
Tabs to spaces
DFT

@@
 
-\subsection{Likelihood}
+\subsection{Inputs}
+
+\begin{itemize}
+    \item Discuss single-detector SNR threshold
+    \item $\hat\rho_i, \hat\gamma_i, \hat\tau_i$
+    \item Sampled PSD
+    \item Waveform?
+    \item Masses $\hat m_1, \hat m_2$
+\end{itemize}
+
+\subsection{Autocorrelation series}
+
+\marginpar{This part may be too technical.}%
+%
+To calculate the autocorrelation sequence, we need to compute the inverse DFT of the autopower spectral density of the template.
+A BNS signal remains in the Advanced LIGO sensitive band for thousands of seconds, but we will only be interested in $\sim 10$\,ms of the autocorrelation sequence.
+If we have a sequence of length $N$, and we only need the first $M$ samples of the inverse DFT, where $M$ divides $N$ and $cM = N$, then we may make use of the following identity to compute the DFT with a speedup of $\log N/M$:
+%
+\begin{eqnarray*}
+    x[n] &=& \sum_{k=0}^{N-1} \exp \left[ \frac{2 \pi i k n}{N} \right] Y[k] \\
+    &=& \sum_{j=0}^{c-1} \sum_{m=0}^{M-1} \exp \left[ \frac{2 \pi i (m c + j) n}{c M} \right] Y[m c + j] \\
+    &=& \sum_{j=0}^{c-1} \exp \left[ \frac{2 \pi i j n}{N} \right] \sum_{m=0}^{M-1} \exp \left[ \frac{2 \pi i m n}{M} \right] Y[m c + j].
+\end{eqnarray*}
 

### 2014-07-19 ms.tex
Fix error in weights
Minor edits

@@
 %
-where $h_i (\omega; \boldsymbol\theta)$ is the \ac{GW} signal given a parameter vector $\boldsymbol\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[\left|n_i(\omega)\right|^2\right]$.
+where $h_i (\omega; \boldsymbol\theta)$ is the \ac{GW} signal given a parameter vector $\boldsymbol\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[\left|n_i(\omega)\right|^2\right] + E\left[\left|n_i(-\omega)\right|^2\right] = 2 E\left[\left|n_i(\omega)\right|^2\right]$.
 We shall denote the combined observation from a network of detectors as $\mathbf x (\omega) \equiv \{x_i (\omega)\}_i$.
@@
 
-Assuming circular orbits and no spin precession, we can write the \ac{GW} signal in each detector as a linear combination of two basis waveforms $h_0$ = $h$ and $h_{\pi/2}$.
+Assuming circular orbits and no spin precession, we can write the \ac{GW} signal in each detector as a linear combination of two basis waveforms $h_0$ and $h_{\pi/2}$.
 For nonprecessing systems, $h_0$ and $h_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the sine and cosine functions, being nearly orthogonal and out of phase by ${\pi/2}$ at all frequencies.
@@
         \includegraphics{crlb_tau}
+
+        \includegraphics{crlb_gamma}
     \end{center}
@@
         \displaystyle
-        \sum_i \frac{\mathbf{n}_i}{\hat\rho_i \omega_{\mathrm{rms},i}}
+        \sum_i \frac{\mathbf{n}_i}
+            {\left(\hat\rho_i \omega_{\mathrm{rms},i}\right)^2}
     }{
         \displaystyle
-        \sum_i \frac{1}{\hat\rho_i \omega_{\mathrm{rms},i}}
+        \sum_i \frac{1}{\left(\hat\rho_i \omega_{\mathrm{rms},i}\right)^2}
     },
@@
         \displaystyle
-        \sum_i \frac{\hat\tau_i}{\hat\rho_i \omega_{\mathrm{rms},i}}
+        \sum_i \frac{\hat\tau_i}
+            {\left(\hat\rho_i \omega_{\mathrm{rms},i}\right)^2}
     }{
         \displaystyle
-        \sum_i \frac{1}{\hat\rho_i \omega_{\mathrm{rms},i}}
+        \sum_i \frac{1}{\left(\hat\rho_i \omega_{\mathrm{rms},i}\right)^2}
     }.

### 2014-07-19 ms.tex
Apply \mathcal to some symbols

@@
 Matrix elements that are not listed have values that are implied by the symmetry of the Fisher matrix.
-Note that the minus signs are correct but a little confusing: despite them, ${I}_{\gamma\gamma}, {I}_{\tau\tau} \geq 0$ and ${I}_{\gamma\tau} \leq 0$.
+Note that the minus signs are correct but a little confusing: despite them, $\mathcal{I}_{\gamma\gamma}, \mathcal{I}_{\tau\tau} \geq 0$ and $\mathcal{I}_{\gamma\tau} \leq 0$.
 The time integration limits $[-T, T]$ constitute a flat prior on arrival time, the time coincidence window between detectors.

### 2014-07-21 ms.tex
Rework radial integral section

@@
 \usepackage{graphicx}
+\usepackage{mathrsfs}
 \usepackage{bib/aas_macros}
@@
 %
-A detection candidate consists of the set $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\boldsymbol\theta}_\mathrm{in}\}$.
+A detection candidate consists of $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\boldsymbol\theta}_\mathrm{in}\}$.
 There are various ways to characterize the significance of a detection candidate, but in Gaussian noise the combined SNR $\hat\rho_\mathrm{c}$ suffices,
@@
 %
-The integral above could be written in terms of the matched filter signals $\left\{y_i\right\}_i$, such that $y_i\left(\hat\tau_i, \hat{\boldsymbol\theta}_\mathrm{in}\right) = \hat\rho_i e^{\hat\gamma_i}$ and $\forall t$, $\left|y_i(t, \boldsymbol\theta_\mathrm{in})\right| \leq \hat\rho_i$.
+The integral above could be written in terms of the matched filter signals $\left\{y_i\right\}_i$, such that $y_i\left(\hat\tau_i, \hat{\boldsymbol\theta}_\mathrm{in}\right) = \hat\rho_i e^{i \hat\gamma_i}$ and $\forall t$, $\left|y_i(t, \boldsymbol\theta_\mathrm{in})\right| \leq \hat\rho_i$.
 The combination of the nontrivial integration limits and the Fourier transform in the matched filter seem like a problem for direct evaluation.
@@
 %
-\begin{equation}
+\begin{equation}\label{eq:autocorrelation-function}
     a_i(t; \boldsymbol\theta_\mathrm{in}) := \frac{1}{{{\sigma_i}^2(\boldsymbol\theta_\mathrm{in})}} \int_0^\infty \frac{\left| h(\omega; \hat{\boldsymbol\theta}_\mathrm{in})\right|^2}{S_i(\omega)} e^{i \omega t} \,d\omega.
@@
 \end{equation}
-%
-To marginalize over the coalescence phase, we can insert an arbitrary nuisance phase $\eta$ by making the replacement $\tilde{\gamma}_i \rightarrow \tilde{\gamma}_i + \eta$.
-Then integrating over $\eta$ from $0$ to $2\pi$ and suppressing normalization factors, we get the \ac{BAYESTAR} likelihood
-%
-\marginpar{Different symbol? Is $\mathcal{L}_\mathrm{BAYESTAR}$ too grandiose?}
-%
-\begin{equation}
-    \mathcal{L}_\mathrm{BAYESTAR} = \exp \left[ - \frac{1}{2} \sum_i {\rho_i}^2 \right] I_0 \left[
-        \left| \sum_i \hat\rho_i \rho_i e^{i \tilde{\gamma}_i} a_i^*(\tilde{\tau}_i)
-        \right|
-    \right],
-\end{equation}
-%
-where $I_0$ is a modified Bessel function of the first kind.
 
@@
 %
+The notation $I_k$ denotes a modified Bessel function of the first kind.
 Matrix elements that are not listed have values that are implied by the symmetry of the Fisher matrix.
@@
 
-\subsection{Inputs}
+Here we describe the numerical implementation of \ac{BAYESTAR}.
 
-\begin{itemize}
-    \item Discuss single-detector SNR threshold
-    \item $\hat\rho_i, \hat\gamma_i, \hat\tau_i$
-    \item Sampled PSD
-    \item Waveform?
-    \item Masses $\hat m_1, \hat m_2$
-\end{itemize}
+\subsection{Prior and problem setup}
 
-\subsection{Autocorrelation series}
+The detection pipeline supplies a candidate $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\boldsymbol\theta}_\mathrm{in}\}$, and discretely sampled noise \acp{PSD} $S_i(\omega_j)$ for all detectors.
+\marginpar{Discuss single-detector threshold here?}
 
-\marginpar{This part may be too technical.}%
-%
-To calculate the autocorrelation sequence, we need to compute the inverse DFT of the autopower spectral density of the template.
-A BNS signal remains in the Advanced LIGO sensitive band for thousands of seconds, but we will only be interested in $\sim 10$\,ms of the autocorrelation sequence.
-If we have a sequence of length $N$, and we only need the first $M$ samples of the inverse DFT, where $M$ divides $N$ and $cM = N$, then we may make use of the following identity to compute the DFT with a speedup of $\log N/M$:
-%
-\begin{eqnarray*}
-    x[n] &=& \sum_{k=0}^{N-1} \exp \left[ \frac{2 \pi i k n}{N} \right] Y[k] \\
-    &=& \sum_{j=0}^{c-1} \sum_{m=0}^{M-1} \exp \left[ \frac{2 \pi i (m c + j) n}{c M} \right] Y[m c + j] \\
-    &=& \sum_{j=0}^{c-1} \exp \left[ \frac{2 \pi i j n}{N} \right] \sum_{m=0}^{M-1} \exp \left[ \frac{2 \pi i m n}{M} \right] Y[m c + j].
-\end{eqnarray*}
-
-\subsection{Prior}
+We compute the GW signal for a source with intrinsic parameters equal to the detection pipeline's estimate, $h(\omega; \hat{\boldsymbol\theta}_\mathrm{in})$.
+Then we find the SNR=1 horizon distance $D_{1,i}$ for each detector by numerically integrating \Eref{eq:horizon}.
 
@@
 
-\subsection{Adaptive resolution}
+We compute the autocorrelation function for each detector from $t = 0$ to $t = T$ at intervals of $\Delta t = 1/f_s$, where $f_s$ is the smallest power of two that is greater than or equal to the Nyquist rate.
+Because BNS signals typically terminate at about 1500\,Hz, a typical value for $\Delta t$ is $(4096\,\textrm{Hz})^{-1}$.
+We use a pruned \ac{FFT} because for BNS systems, the GW signal remains in LIGO's sensitive band for $\sim$100\nobreakdashes--1000\,s, whereas $T \sim 10$\,ms.\footnote{See \url{http://www.fftw.org/pruned.html} for a discussion of methods for computing the first $K$ samples of an \ac{FFT} of length $N$.}
 
-The \ac{TOA}\nobreakdashes-only posterior in \Eref{eq:toa-posterior} is numerically trivial to evaluate because it involves no numerical integrals.
-However, as we will see in the next section, the \ac{SNR}\nobreakdashes-only posterior is numerically expensive.
+\subsection{Marginal posterior}
 
-\marginpar{Clumsy prose.}%
-We evaluate the \ac{TOA}\nobreakdashes-only posterior at the centers of \ac{HEALPix} pixels at the lowest possible resolution.
-Then, we increase the resolution until 99.99\% of the \ac{TOA}\nobreakdashes-only posterior is contained within at least 3072 pixels.
-The numerically expensive \ac{SNR} integrals are done only for the pixels that comprise the 99.99th percentile at this resolution.
-All other pixels are set to zero.
+The marginal posterior as a function of sky location is
+%
+\begin{eqnarray}
+    \fl f(\alpha, \delta) \propto
+    \int_{0}^{\pi}
+    \int_{-1}^{1}
+    \int_{-T}^{T}
+    \int_{D_{\mathrm{L},\mathrm{min}}}^{D_{\mathrm{L},\mathrm{max}}}
+    \int_{0}^{2\pi} \\
+    %
+    \exp \left[ - \frac{1}{2} \sum_i {\rho_i}^2
+        + \sum_i \hat\rho_i \rho_i \Re \left\{ e^{i \tilde{\gamma}_i}
+        a^*(\tilde{\tau}_i)
+        \right\}
+    \right] {D_\mathrm{L}}^k \, dt_\oplus \, d\phi_c dD_\mathrm{L} \, d\cos{\iota} \, d\psi.
+\end{eqnarray}
 
-\subsection{Posterior}
+To marginalize over the coalescence phase, we can insert an arbitrary nuisance phase $\eta$ by making the replacement $\tilde{\gamma}_i \rightarrow \tilde{\gamma}_i + \eta$.
+Then integrating over $\eta$ from $0$ to $2\pi$ and suppressing normalization factors, we get
+%
+\begin{eqnarray}
+    \fl f(\alpha, \delta) \rightarrow
+    \int_{0}^{\pi}
+    \int_{-1}^{1}
+    \int_{-T}^{T}
+    \int_{D_{\mathrm{L},\mathrm{min}}}^{D_{\mathrm{L},\mathrm{max}}} \\
+    %
+    \exp \left[ - \frac{1}{2} \sum_i {\rho_i}^2 \right] I_0 \left[
+            \left| \sum_i \hat\rho_i \rho_i e^{i \tilde{\gamma}_i} a_i^*(\tilde{\tau}_i)
+            \right|
+    \right]
+    {D_\mathrm{L}}^k dD_\mathrm{L} \, dt_\oplus \, d\cos{\iota} \, d\psi.
+\end{eqnarray}
 
-\begin{eqnarray}
-    p(
-        \mathbf{n} |
-        \hat{\tau}_1, \hat{\tau}_2, &\dots, \hat{\tau}_N,
-        \hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N) \nonumber\\
-    &= p(\mathbf{n} |
-        \hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N) \nonumber\\
-        &\times
-        p(\mathbf{n} |
-        \hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N
-    )
-\end{eqnarray}
-%
-where
-%
-\begin{equation*}
-    \fl
-    p(\mathbf{n} |
-    \hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N)
-    \propto
-    \int_0^\frac{\pi}{2}
-    \int_0^1
-    \int_{D_{\mathrm{L},\mathrm{min}}}^{D_{\mathrm{L},\mathrm{max}}}
-    \exp \left[
-            -\frac{1}{2} \sum_{i=1}^N \left(\hat{\rho}_i - \rho_i(\mathbf{n}, D_\mathrm{L}, u, \psi) \right)^2
-        \right]
-    {D_\mathrm{L}}^m
-    \,dD_\mathrm{L}
-    \,du
-    \,d\psi.
-\end{equation*}
+\subsubsection{Integral over angles and time}
 
-The marginalization over $D_\mathrm{L}$, $u$, and $\psi$ must be done numerically.
-The outermost integration is over $u$ and $\psi$, and is done with Newton\nobreakdashes--Cotes by evaluating $\mathcal{L}_\mathrm{SNR}$ on a regular lattice in $u$ and $\psi$.
+The integrand is periodic in $\psi$, so simple Newton-Cotes quadrature over $\psi$ exhibits extremely rapid convergence.
+We therefore sample the posterior on a regular grid of 10 points from 0 to $\pi$.
+
+The integral over $\cos\iota$ converges with just as rapidly with Gauss-Legendre quadrature, so we use a 10\nobreakdashes-point Gauss-Legendre rule for integration over $\cos\iota$.
+
+We sample $t_\oplus$ regularly from $-T$ to $T$ at intervals of $\Delta t$.
+This is typically $\sim 2 (10\,\mathrm{ms})(4096\,\mathrm{Hz}) \approx 80$\,samples.
+We use Catmull\nobreakdashes-Rom cubic splines to interpolate the real and imaginary parts of the autocorrelation functions between samples.
+
+\subsubsection{Integral over distance: adaptive Gaussian quadrature method}
 
@@
 
-The innermost integration is over $D_\mathrm{L}$ and is performed using adaptive Gaussian quadrature, but with a twist that speeds up convergence.
-As in many parameter estimation problems, the integrand may be sharply peaked, and any generic adaptive integration scheme may take a long time to find the peak.
-Before evaluating each inner integral over $D_\mathrm{L}$, the quantities $A=-\frac{1}{2} \sum_i {\rho_i}^2$ and $B=\sum_i \rho_i \hat\rho_i$ are pre\nobreakdashes-calculated so that the integrand can be written as
+The innermost integral over $D_\mathrm{L}$ is a little bit more challenging.
+The distance integrand has both a sharp maximum\nobreakdashes-likelihood peak and an extended power\nobreakdashes-law prior\nobreakdashes-dominated tail.
+At any values of the other parameters, either the peak or the tail might dominate.
+If the peak dominates, it may not be successfully resolved by any general\nobreakdashes-purpose fixed\nobreakdashes-order quadrature scheme.
+Any general adaptive integrator may take a long time to find the peak.
+
+We found several efficient ways to evaluate the distance integral.
+The first method is to use adaptive Gaussian quadrature, but with an educated choice of initial subdivisions.
+The integrand can be written as
 %
 \begin{equation*}
+    \exp\left(\frac{A}{{D_\mathrm{L}}^2}\right)
+    I_0 \left(\frac{B}{D_\mathrm{L}}\right) {D_\mathrm{L}}^m
+    \sim
     \exp\left(\frac{A}{{D_\mathrm{L}}^2} + \frac{B}{D_\mathrm{L}}\right) {D_\mathrm{L}}^m.
@@
 %
-Then, we see that the likelihood $\exp(A {D_\mathrm{L}}^{-2} + B {D_\mathrm{L}}^{-1})$ is maximized when $D_\mathrm{L} = D_{\mathrm{L},0} = -2A/B$.
+Then, we see that the approximate likelihood $\exp(A {D_\mathrm{L}}^{-2} + B {D_\mathrm{L}}^{-1})$ is maximized when $D_\mathrm{L} = D_{\mathrm{L},0} = -2A/B$.
 The likelihood takes on a factor $\eta$ (say, $\eta=0.01$) of its maximum value when
@@
 %
-We use these breakpoints as initial subdivision in an adaptive Gaussian quadrature algorithm.
-Specifically, we use the GNU Scientific Library's \verb|gsl_integrate_qagp|\footnote{\url{http://www.gnu.org/software/gsl/manual/html_node/QAGP-adaptive-integration-with-known-singular-points.html}} function, which is designed to integrate functions with known singular points (though, of course, or integrand has no singular points).
+We use these breakpoints as initial subdivisions in an adaptive Gaussian quadrature algorithm\footnote{for instance, \ac{GSL}'s \texttt{gsl\_integrate\_qagp} function, \url{http://www.gnu.org/software/gsl/manual/html_node/QAGP-adaptive-integration-with-known-singular-points.html}}.
 This function estimates the integral over each subdivision and each interval's contribution to the total error, then subdivides the interval whose error contribution is largest.
@@
 
-\subsection{Implementation}
+\subsubsection{Integral over distance: importance sampling method}
+
+Let's say that we want to evaluate an integral
+%
+\begin{equation}\label{eq:importance-sampling-difficult-integral}
+    \mathscr{F}(x_1, x_2) = \int_{x_1}^{x_2} f(x') dx'
+\end{equation}
+%
+Suppose that this integral is resistant to direct application of standard quadrature techniques, but we have another function $g(x)$ that we can integrate analytically or at least numerically with relative ease,
+%
+\begin{equation}
+    \mathscr{G}(x) = G(x) - G(x_0) = \int_{x_0}^x g(x') dx'.
+\end{equation}
+%
+There is an \emph{ultimate}, though tautological, change of variables that makes it trivial to evaluate $\mathscr{G}(x)$,
+%
+\begin{equation}
+    \mathscr{G}(x) = \int_0^{\mathrlap{\mathscr{G}(x)}} dG.
+\end{equation}
+%
+If we apply the same change of variables to \Eref{eq:importance-sampling-difficult-integral}, we get
+%
+\begin{equation}\label{eq:importance-sampling}
+    \mathscr{F}(x_1, x_2) = \int_{\mathscr{G}(x_1)}^{\mathscr{G}(x_2)} \frac{f(\mathscr{G}^{-1}(G))}{g(\mathscr{G}^{-1}(G))} dG.
+\end{equation}
+%
+If $f(x)$ and $g(x)$ have sufficiently similar behavior, then their quotient $f/g$ will be better behaved than $f$ itself, and more amenable to any given quadrature technique.
+When the transformed integral \Eref{eq:importance-sampling} is evaluated using Monte Carlo integration, this technique is referred to as importance sampling.
+
+We use the same kind of change of variables with fixed\nobreakdashes-order Gaussian quadrature.
+
+\subsection{Adaptive resolution}
+
+The \ac{TOA}\nobreakdashes-only posterior in \Eref{eq:toa-posterior} is numerically trivial to evaluate because it involves no numerical integrals.
+However, as we will see in the next section, the \ac{SNR}\nobreakdashes-only posterior is numerically expensive.
 

### 2014-07-22 ms.tex
Describe distance integral and adaptive mesh refinement

@@
                 \delta \\
-                D_\mathrm{L} \\
+                r \\
                 t_\oplus \\
@@
                     declination \\
-                    luminosity distance \\
+                    distance \\
                     arrival time at geocenter \\
@@
     h_i(\omega; \boldsymbol\theta) = \e^{-i \omega t_\oplus}
-    \frac{D_{1,i}}{D_\mathrm{L}}
+    \frac{r_{1,i}}{r}
     \e^{2 i \phi_c}
@@
 They depend on the orientation of detector $i$ as well as the sky location and sidereal time of the event, and are derived in~\cite{PhysRevD.63.042003}.
-The quantity $D_{1,i}$ is a fiducial distance at which detector $i$ would register SNR=1 for an optimally oriented binary (face\nobreakdashes-on, and in a direction perpendicular to the interferometer's arms),
+The quantity $r_{1,i}$ is a fiducial distance at which detector $i$ would register SNR=1 for an optimally oriented binary (face\nobreakdashes-on, and in a direction perpendicular to the interferometer's arms),
 %
@@
 \begin{equation}\label{eq:horizon}
-D_{1,i} = 1 / \sigma_i, \qquad {\sigma_i}^2 = \int_0^\infty \frac{\left|h(\omega; \mathbf \theta)\right|^2}{S_i(\omega)} \,d\omega.
+r_{1,i} = 1 / \sigma_i, \qquad {\sigma_i}^2 = \int_0^\infty \frac{\left|h(\omega; \mathbf \theta)\right|^2}{S_i(\omega)} \,d\omega.
 \end{equation}
@@
 We compute the GW signal for a source with intrinsic parameters equal to the detection pipeline's estimate, $h(\omega; \hat{\boldsymbol\theta}_\mathrm{in})$.
-Then we find the SNR=1 horizon distance $D_{1,i}$ for each detector by numerically integrating \Eref{eq:horizon}.
+Then we find the SNR=1 horizon distance $r_{1,i}$ for each detector by numerically integrating \Eref{eq:horizon}.
 
@@
 \begin{equation*}
-    p(D_\mathrm{L}) \propto \cases{
-        {D_\mathrm{L}}^m & if $D_{\mathrm{L},\mathrm{min}} < D_\mathrm{L} < D_{\mathrm{L},\mathrm{max}}$ \\
+    p(r) \propto \cases{
+        r^m & if $r_\mathrm{min} < r < r_\mathrm{max}$ \\
         0 & otherwise
@@
     \qquad
-    D_{\mathrm{L},\mathrm{min}} = 0,
+    r_\mathrm{min} = 0,
     \qquad
-    D_{\mathrm{L},\mathrm{max}} = \frac{1}{4} \max_i D_{1,i}.
+    r_\mathrm{max} = \frac{1}{4} \max_i r_{1,i}.
 \end{equation*}
@@
     \int_{-T}^{T}
-    \int_{D_{\mathrm{L},\mathrm{min}}}^{D_{\mathrm{L},\mathrm{max}}}
+    \int_{r_\mathrm{min}}^{r_\mathrm{max}}
     \int_{0}^{2\pi} \\
@@
         \right\}
-    \right] {D_\mathrm{L}}^k \, dt_\oplus \, d\phi_c dD_\mathrm{L} \, d\cos{\iota} \, d\psi.
+    \right] r^m \, dt_\oplus \, d\phi_c dr \, d\cos{\iota} \, d\psi.
 \end{eqnarray}
@@
     \int_{-T}^{T}
-    \int_{D_{\mathrm{L},\mathrm{min}}}^{D_{\mathrm{L},\mathrm{max}}} \\
+    \int_{r_\mathrm{min}}^{r_\mathrm{max}} \\
     %
@@
     \right]
-    {D_\mathrm{L}}^k dD_\mathrm{L} \, dt_\oplus \, d\cos{\iota} \, d\psi.
+    r^m dr \, dt_\oplus \, d\cos{\iota} \, d\psi.
 \end{eqnarray}
@@
 
-The innermost integral over $D_\mathrm{L}$ is a little bit more challenging.
+The innermost integral over $r$ is a little bit more challenging.
 The distance integrand has both a sharp maximum\nobreakdashes-likelihood peak and an extended power\nobreakdashes-law prior\nobreakdashes-dominated tail.
@@
 
-We found several efficient ways to evaluate the distance integral.
+We found two efficient ways to evaluate the distance integral.
 The first method is to use adaptive Gaussian quadrature, but with an educated choice of initial subdivisions.
-The integrand can be written as
+The distance integral can be written as
 %
-\begin{equation*}
-    \exp\left(\frac{A}{{D_\mathrm{L}}^2}\right)
-    I_0 \left(\frac{B}{D_\mathrm{L}}\right) {D_\mathrm{L}}^m
-    \sim
-    \exp\left(\frac{A}{{D_\mathrm{L}}^2} + \frac{B}{D_\mathrm{L}}\right) {D_\mathrm{L}}^m.
-\end{equation*}
+\begin{eqnarray}\label{eq:distance-integral}
+    \mathscr{F} &=&
+        \int_{\mathrlap{r_\mathrm{min}}}^{\mathrlap{r_\mathrm{max}}}
+        \exp\left[-\frac{p^2}{r^2}\right]
+        I_0\left[\frac{b}{r}\right] r^m dr \nonumber \\
+    &=&
+        \int_{\mathrlap{r_\mathrm{min}}}^{\mathrlap{r_\mathrm{max}}}
+        \exp\left[-\frac{p^2}{r^2}+\frac{b}{r}\right]
+        \overline{I}_0\left[\frac{b}{r}\right] r^m dr
+\end{eqnarray}
 %
-Then, we see that the approximate likelihood $\exp(A {D_\mathrm{L}}^{-2} + B {D_\mathrm{L}}^{-1})$ is maximized when $D_\mathrm{L} = D_{\mathrm{L},0} = -2A/B$.
+where
+%
+\begin{eqnarray*}
+    p^2 &=& \frac{1}{2} r^2 \sum_i {\rho_i}^2 \\
+    b &=& r \left| \sum_i \hat\rho_i \rho_i e^{i \tilde{\gamma}_i}
+        a_i^*(\tilde{\tau}_i) \right| \\
+    \overline{I}_0(x) &=& \exp(-|x|) I_0(x).
+\end{eqnarray*}
+%
+The coefficients $p^2$ and $b$ are nonnegative and independent of distance.
+The symbol $\overline{I}_0$ denotes an exponentially scaled Bessel function.
+In the limit of large argument, $I_0(|x|) \sim \exp(|x|) / \sqrt{2 \pi |x|}$ \cite{Olver:2010:NHMF}\footnote{\url{http://dlmf.nist.gov/10.40.E1}}.
+The scaled Bessel is useful for evaluation on a computer because it has a relatively small range $(0, 1]$ and varies slowly in proportion to $x^{1/2}$.
+If we neglect both the Bessel function and the $r^m$ prior, then the approximate likelihood $\exp(-p^2/r^2 + b/r)$ is maximized when
+%
+\begin{equation}\label{eq:ml-distance}
+    r = r_0 = 2p^2/b.
+\end{equation}
+%
 The likelihood takes on a factor $\eta$ (say, $\eta=0.01$) of its maximum value when
@@
 \begin{equation}
-    D_\mathrm{L} = D_{\mathrm{L},\pm} = \left(\frac{1}{D_{\mathrm{L},0}} \mp \sqrt{\frac{\log\eta}{A}}\right)^{-1}.
+    r = r_\pm = \left(\frac{1}{r_0} \mp \frac{\sqrt{-\log\eta}}{p}\right)^{-1}.
 \end{equation}
@@
 \begin{equation}
-    D_{\mathrm{L},\mathrm{break}} = \{ D_\mathrm{L} \in
+    r_\mathrm{break} = \{ r \in
     \left\{
     \begin{array}{c}
-    D_{\mathrm{L},\mathrm{min}} \\
-    D_{\mathrm{L},-} \\
-    D_{\mathrm{L},0} \\
-    D_{\mathrm{L},+} \\
-    D_{\mathrm{L},\mathrm{max}}
+    r_\mathrm{min} \\
+    r_- \\
+    r_0 \\
+    r_+ \\
+    r_\mathrm{max}
     \end{array}
     \right\} :
-    D_{\mathrm{L},\mathrm{min}} \leq D_\mathrm{L}
-    \leq D_{\mathrm{L},\mathrm{max}}\}.
+    r_\mathrm{min} \leq r
+    \leq r_\mathrm{max}\}.
 \end{equation}
@@
 
-Let's say that we want to evaluate an integral
+The second method of evaluating the distance integral, and the one used in practice, is similar to importance sampling. Let's say that we want to evaluate an integral
 %
@@
 We use the same kind of change of variables with fixed\nobreakdashes-order Gaussian quadrature.
+We first find the approximate \ac{ML} distance $r_0$ using \Eref{eq:ml-distance}.
+If $r_0 < r_\mathrm{min}$ or $r_0 > r_\mathrm{max}$, then the ML peak lies outside of the limits of integration and the original integrand behaves like a low-order power law.
+In this case, we use 10\nobreakdashes-point Gaussian quadrature to evaluate \Eref{eq:distance-integral}.
 
-\subsection{Adaptive resolution}
+If, on the other hand, $r_\mathrm{min} \leq r_0 \leq r_\mathrm{max}$, then the integrand contains a sharp peak that we can smooth out with importance sampling.
+We first re-scale the distance integral \Eref{eq:distance-integral} so that the peak value of the integrand is $\sim 1$:
+%
+\begin{equation}
+    \mathscr{F} = \exp\left(\frac{b^2}{4p^2}\right)
+        {r_0}^m
+        \int_{r_\mathrm{min}}^{r_\mathrm{max}}
+        \exp\left[-\left(\frac{p}{r} - \frac{b}{2p}\right)^2\right]
+        \left[\frac{r}{r_0}\right]^m
+        dr.
+\end{equation}
+%
+The importance sampling integrand $g(r)$ is
+%
+\begin{equation}
+    g(r) = \exp\left[-\left(\frac{p}{r} - \frac{b}{2p}\right)^2\right]
+        \frac{1}{r^2}
+\end{equation}
+%
+with definite integral
+%
+\begin{equation}
+    \mathscr{G}(r) = \frac{\sqrt{\pi}}{p}
+        Q\left[\sqrt{2}\left(\frac{p}{r} - \frac{b}{2p}\right)\right]
+\end{equation}
+%
+where $Q$ is the cumulative distribution function for the upper tail of the standard normal distribution,
+%
+\begin{equation}
+    Q(x) = \int_x^\infty \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{{x'}^2}{2}\right] dx'.
+\end{equation}
+%
+Its inverse is
+%
+\begin{equation}
+    \mathscr{G}^{-1}(G) = \left[\frac{1}{r_0} + \frac{1}{\sqrt{2}p}
+    Q^{-1}\left(\frac{p G}{\sqrt{\pi}}\right)\right]^{-1}.
+\end{equation}
+%
+The transformed integral is
+%
+\begin{equation}
+    \mathscr{F} = \exp\left(\frac{b^2}{4p^2}\right)
+        {r_0}^m
+        \int_{\mathscr{G}(r_\mathrm{min})}^{\mathscr{G}(r_\mathrm{max})}
+        \overline{I}_0\left[\frac{b}{r}\right]
+        \left[\frac{r}{r_0}\right]^m r^2
+        \Bigg|_{r = \mathscr{G}^{-1}(G)}
+        dG.
+\end{equation}
+%
+This integral is evaluated numerically with a 10\nobreakdashes-point Gauss-Legendre rule.
 
-The \ac{TOA}\nobreakdashes-only posterior in \Eref{eq:toa-posterior} is numerically trivial to evaluate because it involves no numerical integrals.
-However, as we will see in the next section, the \ac{SNR}\nobreakdashes-only posterior is numerically expensive.
+Last, there are a few special cases where \Eref{eq:distance-integral} can be evaluated exactly.
+If $p^2=0$ (which implies $b = 0$), then the integral may be expressed in terms of logarithms.
+If $b = 0$ but $p^2 \neq 0$, then the integral may be expressed in terms of the exponential integral function,
+%
+\begin{equation}
+    E_n(x) = \int_1^\infty \frac{\exp (-xt)}{t^n} dt.
+\end{equation}
 
-\marginpar{Talk briefly about C implementation, OpenMP, and Python wrapper that interfaces with search pipelines.}%
-The \ac{SNR} marginalization integrals for different pixels may be evaluated in parallel.
-In our C language implementation, the loop over pixels is parallelized with OpenMP\footnote{http://openmp.org/}.
+The distance integral, therefore, takes either one or ten function evaluations.
+
+\subsection{Adaptive mesh refinement}
+
+We have explained how we evaluate the marginal posterior at a given sky location. Now we must specify where we choose to evaluate it.
+
+Our sampling of the sky is relies completely on the \acl{HEALPix} (\acs{HEALPix}), a special data structure designed for all\nobreakdashes-sky maps.
+\ac{HEALPix} divides the sky into equal\nobreakdashes-area pixels.
+There is a hierarchy of \ac{HEALPix} resolutions.
+A HEALPix resolution may be designated by its order $N$.
+The $N=0$th order or base tiling has 12 pixels.
+At every successive order, each tile is subdivided into four new tiles.
+A resolution may also be referred to by the number of subdivisions along each side of the base tiles, $N_\mathrm{side} = N^2$.
+There are $N^\mathrm{pix} = 12 {N_\mathrm{side}}^2$ pixels at any given resolution.
+The \ac{HEALPix} projection uniquely specifies the coordinates of the center of each pixel by providing a mapping from the resolution and pixel index $(N_\mathrm{side}, i_\mathrm{pix})$ to right ascension and declination $(\alpha, \delta)$.
+
+We begin by evaluating the posterior probability density at the center of each of the $N_{\mathrm{pix},0} = 768$ pixels of an $N_{\mathrm{side},0}=8$ \ac{HEALPix} grid.
+At this resolution, each pixel has an area of 54\,deg$^2$.
+We then rank the pixels by contained probability (assuming constant probability density within a pixel), and subdivide the most probable $N_{\mathrm{pix},0}/4$ pixels into $N_{\mathrm{pix},0}$ new pixels.
+We then evaluate the posterior again at the pixels that we subdivided, sort again, and repeat eight times, so that we have evaluated the posterior probability density a total of $8 N_{\mathrm{pix},0}$ times.
+On \emph{most} iterations, we descend one level deeper in \ac{HEALPix} resolution.
+The resulting map is a tree structure that consists of a mix of several resolutions.
+We traverse the tree and flatten it into the highest resolution represented.
+The highest possible resolution is $N_\mathrm{side}=2^{11}$, with an area of 1.7\,deg$^2$ per pixel.\footnote{Although the resulting sky map contains $N_\mathrm{pix}=24\,576$ pixels, at most $4\,800$ pixels have distinct values.
+For the purpose of delivery to observers, therefore, the output is always \texttt{gzip}\nobreakdashes-compressed with a ratio of $\sim 5:1$.}
+
+Within each iteration, all of the marginalization integrals may be evaluated in parallel.
+In our C language implementation, the inner loop over pixels is parallelized with OpenMP\footnote{http://openmp.org/}.
 

### 2014-07-24 ms.tex
Update discussion of adaptive refinement

@@
 
-We begin by evaluating the posterior probability density at the center of each of the $N_{\mathrm{pix},0} = 768$ pixels of an $N_{\mathrm{side},0}=8$ \ac{HEALPix} grid.
-At this resolution, each pixel has an area of 54\,deg$^2$.
+We begin by evaluating the posterior probability density at the center of each of the $N_{\mathrm{pix},0} = 3072$ pixels of an $N_{\mathrm{side},0}=16$ \ac{HEALPix} grid.
+At this resolution, each pixel has an area of 13.4\,deg$^2$.
 We then rank the pixels by contained probability (assuming constant probability density within a pixel), and subdivide the most probable $N_{\mathrm{pix},0}/4$ pixels into $N_{\mathrm{pix},0}$ new pixels.
-We then evaluate the posterior again at the pixels that we subdivided, sort again, and repeat eight times, so that we have evaluated the posterior probability density a total of $8 N_{\mathrm{pix},0}$ times.
+We then evaluate the posterior again at the pixels that we subdivided, sort again, and repeat seven times, so that we have evaluated the posterior probability density a total of $8 N_{\mathrm{pix},0}$ times.
 On \emph{most} iterations, we descend one level deeper in \ac{HEALPix} resolution.
@@
 We traverse the tree and flatten it into the highest resolution represented.
-The highest possible resolution is $N_\mathrm{side}=2^{11}$, with an area of 1.7\,deg$^2$ per pixel.\footnote{Although the resulting sky map contains $N_\mathrm{pix}=24\,576$ pixels, at most $4\,800$ pixels have distinct values.
-For the purpose of delivery to observers, therefore, the output is always \texttt{gzip}\nobreakdashes-compressed with a ratio of $\sim 5:1$.}
+The highest possible resolution is $N_\mathrm{side}=2^{11}$, with an area of $\approx 10^{-3}$\,deg$^2$ per pixel.\footnote{Although the resulting sky map contains $N_\mathrm{pix} \approx 5\times10^6$ pixels, at most $\approx 2\times10^4$ pixels have distinct values.
+For the purpose of delivery to observers, therefore, the output is always \texttt{gzip}\nobreakdashes-compressed with a ratio of $\approx 250:1$.}
 

### 2014-07-24 scripts/pp.py
Add P--P plot

@@ -1,0 +1,25 @@
+import numpy as np
+from matplotlib import pyplot as plt
+from lalinference.bayestar import plot
+
+fig = plt.figure(figsize=(3.5, 3.5))
+ax = fig.add_subplot(111, projection='pp_plot')
+ax.add_diagonal()
+
+data = np.recfromtxt(
+    '2015_subset/found_injections.out', names=True)['searched_prob']
+n = len(data)
+ax.add_series(data, label='2015')
+
+data = np.recfromtxt(
+    '2016_subset/found_injections.out', names=True)['searched_prob']
+assert len(data) == n
+ax.add_series(data, label='2016')
+
+ax.add_confidence_band(n)
+ax.legend(loc='lower right')
+ax.set_xlabel('Credible level $p$')
+ax.set_ylabel('Fraction of true locations with $\mathrm{CR}_p$')
+ax.grid()
+
+plt.savefig('pp.pdf')

### 2014-08-11 ms.tex
Add P--P plot
Remove unused integration method

@@
 
-\subsubsection{Integral over distance: adaptive Gaussian quadrature method}
-
-\begin{figure}
-    \caption{\label{fig:radial_integrand}Illustration of initial subdivisions for distance integration scheme.
-Distance increases from left to right.
-In the color version, the left\nobreakdashes-hand tail, the left\nobreakdashes- and right\nobreakdashes-hand sides of the maximum likelihood peak, and the right\nobreakdashes-hand tail, are colored cyan, red, green, and blue, respectively.}
-    \begin{center}
-        \includegraphics{radial_integrand}
-    \end{center}
-\end{figure}
+\subsubsection{Integral over distance}
 
@@
 
-We found two efficient ways to evaluate the distance integral.
-The first method is to use adaptive Gaussian quadrature, but with an educated choice of initial subdivisions.
-The distance integral can be written as
-%
-\begin{eqnarray}\label{eq:distance-integral}
-    \mathscr{F} &=&
-        \int_{\mathrlap{r_\mathrm{min}}}^{\mathrlap{r_\mathrm{max}}}
-        \exp\left[-\frac{p^2}{r^2}\right]
-        I_0\left[\frac{b}{r}\right] r^m dr \nonumber \\
-    &=&
-        \int_{\mathrlap{r_\mathrm{min}}}^{\mathrlap{r_\mathrm{max}}}
-        \exp\left[-\frac{p^2}{r^2}+\frac{b}{r}\right]
-        \overline{I}_0\left[\frac{b}{r}\right] r^m dr
-\end{eqnarray}
-%
-where
-%
-\begin{eqnarray*}
-    p^2 &=& \frac{1}{2} r^2 \sum_i {\rho_i}^2 \\
-    b &=& r \left| \sum_i \hat\rho_i \rho_i e^{i \tilde{\gamma}_i}
-        a_i^*(\tilde{\tau}_i) \right| \\
-    \overline{I}_0(x) &=& \exp(-|x|) I_0(x).
-\end{eqnarray*}
-%
-The coefficients $p^2$ and $b$ are nonnegative and independent of distance.
-The symbol $\overline{I}_0$ denotes an exponentially scaled Bessel function.
-In the limit of large argument, $I_0(|x|) \sim \exp(|x|) / \sqrt{2 \pi |x|}$ \cite{Olver:2010:NHMF}\footnote{\url{http://dlmf.nist.gov/10.40.E1}}.
-The scaled Bessel is useful for evaluation on a computer because it has a relatively small range $(0, 1]$ and varies slowly in proportion to $x^{1/2}$.
-If we neglect both the Bessel function and the $r^m$ prior, then the approximate likelihood $\exp(-p^2/r^2 + b/r)$ is maximized when
-%
-\begin{equation}\label{eq:ml-distance}
-    r = r_0 = 2p^2/b.
-\end{equation}
-%
-The likelihood takes on a factor $\eta$ (say, $\eta=0.01$) of its maximum value when
-%
-\begin{equation}
-    r = r_\pm = \left(\frac{1}{r_0} \mp \frac{\sqrt{-\log\eta}}{p}\right)^{-1}.
-\end{equation}
-%
-We have now identified up to five breakpoints that partition the distance integrand into up to four intervals with quantitatively distinct behavior.
-These intervals are depicted in Figure~\ref{fig:radial_integrand} with distance increasing from left to right.
-There is a left\nobreakdashes-hand or small distance tail in which the integrand is small and monotonically increasing, a left\nobreakdashes- and right\nobreakdashes-hand side of the maximum likelihood peak, and a right\nobreakdashes-hand tail in which the integrand is small and monotonically decreasing.
-These breakpoints are:
-%
-\begin{equation}
-    r_\mathrm{break} = \{ r \in
-    \left\{
-    \begin{array}{c}
-    r_\mathrm{min} \\
-    r_- \\
-    r_0 \\
-    r_+ \\
-    r_\mathrm{max}
-    \end{array}
-    \right\} :
-    r_\mathrm{min} \leq r
-    \leq r_\mathrm{max}\}.
-\end{equation}
-%
-We use these breakpoints as initial subdivisions in an adaptive Gaussian quadrature algorithm\footnote{for instance, \ac{GSL}'s \texttt{gsl\_integrate\_qagp} function, \url{http://www.gnu.org/software/gsl/manual/html_node/QAGP-adaptive-integration-with-known-singular-points.html}}.
-This function estimates the integral over each subdivision and each interval's contribution to the total error, then subdivides the interval whose error contribution is largest.
-Subdivisions continue until a fixed total fractional error is reached.
-In this way, most integrand evaluations are expended on the most important distance interval, whether that happens to be the tails (when the posterior is dominated by the prior) or the peak (when the posterior is dominated by the observations).
-
-\subsubsection{Integral over distance: importance sampling method}
-
-The second method of evaluating the distance integral, and the one used in practice, is similar to importance sampling. Let's say that we want to evaluate an integral
+Let's say that we want to evaluate an integral
 %
@@
 
+\section{Results}
+
+\begin{figure}
+    \begin{center}
+        \includegraphics{pp}
+    \end{center}
+    \caption{\label{fig:pp}$P-P$ plot.}
+\end{figure}
+
 \ack \marginpar{This URL is not specific enough.}Source code for \ac{BAYESTAR} is available on the \ac{LIGO} \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/lalsuite.html}.

### 2014-08-11 ms.tex
Fix typo

@@
 .
-It is unique in that in bridges the detection and parameter estimation of \ac{GW} signals, two tasks that have until now involved very different numerical methods and time scales.
+It is unique in that it bridges the detection and parameter estimation of \ac{GW} signals, two tasks that have until now involved very different numerical methods and time scales.
 We expect that \ac{BAYESTAR} will take on a key role in observing \ac{CBC} events in both \ac{GW} and optical channels during the Advanced \ac{LIGO} era.
@@
 %
-where $m=2$ for a prior that is uniform in volume, and $m=-1$. If a distance prior is not specified, the default is uniform in volume out to the maximum SNR=4 horizon distance:
+where $m=2$ for a prior that is uniform in volume, and $m=-1$ for a prior that is uniform in the logarithm of the distance.
+If a distance prior is not specified, the default is uniform in volume out to the maximum SNR=4 horizon distance:
 %

### 2014-08-11 ms.tex
Cite Grover paper

@@
 %
-To lowest post\nobreakdashes-Newtonian order, the intrinsic parameters affect the phase evolution but not the amplitude evolution, so for BNS systems we can assume $\mathcal{I}_{\rho_i,\boldsymbol\theta_\mathrm{in}} \approx 0$.
+(See also \cite{Grover:2013}.) To lowest post\nobreakdashes-Newtonian order, the intrinsic parameters affect the phase evolution but not the amplitude evolution, so for BNS systems we can assume $\mathcal{I}_{\rho_i,\boldsymbol\theta_\mathrm{in}} \approx 0$.
 %

### 2014-08-11 ms.tex
Fix typo

@@
 %
-When we form the Fisher matrix for the whole network, if the detectors all have the same noise \acp{PSD}, $S_1(\omega) = S_1(\omega) = \cdots = S_n(\omega) \equiv S(\omega)$, then we can change variables from phases and times on arrival $\left\{\gamma_i, \tau_i\right\}_i$ to appropriate differences between them, in order to cancel out all off-diagonal terms between the intrinsic and extrinsic parameters.
+When we form the Fisher matrix for the whole network, if the detectors all have the same noise \acp{PSD}, $S_1(\omega) = S_2(\omega) = \cdots = S_n(\omega) \equiv S(\omega)$, then we can change variables from phases and times on arrival $\left\{\gamma_i, \tau_i\right\}_i$ to appropriate differences between them, in order to cancel out all off-diagonal terms between the intrinsic and extrinsic parameters.
 Because all existing ground-based GW detectors have similar bandpasses, we restrict all of our further attention to just the extrinsic part of the Fisher matrix, $\mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_i}$.

### 2014-08-12 ms.tex
Reformat for PRX

@@
-\documentclass{iopart}
-\usepackage{arydshln}
+\documentclass[amsmath,amssymb,aps,prx,reprint,nopreprintnumbers,nofootinbib]{revtex4-1}
 \usepackage{multirow}
-\usepackage{amsopn}
-\usepackage{iopams}
+\usepackage{bigdelim}
+\usepackage{multirow}
+\usepackage{bm}
 \usepackage{graphicx}
@@
 % From http://www.f.kth.se/~ante/latex.php
-\setlength{\marginparwidth}{1.2in}
+\setlength{\marginparwidth}{0.5in}
 \let\oldmarginpar\marginpar
-\renewcommand\marginpar[1]{\-\oldmarginpar[\raggedleft\footnotesize #1]%
-{\raggedright\footnotesize #1}}
+\renewcommand\marginpar[1]{\-\oldmarginpar[\raggedleft\scriptsize #1]%
+{\raggedright\scriptsize #1}}
 
@@
 Rapid Bayesian position reconstruction for gravitational\nobreakdashes-wave transients}
-\author{Leo Singer and Larry R. Price}
-\address{\acs{LIGO} Laboratory, California Institute of Technology, Pasadena, CA 91125, USA}
-\ead{\mailto{leo.singer@ligo.org}, \mailto{larryp@caltech.edu}}
+\author{Leo P. Singer}
+\email{leo.singer@ligo.org}
+\author{Larry R. Price}
+\email{larryp@caltech.edu}
+\affiliation{\acs{LIGO} Laboratory, California Institute of Technology, Pasadena, CA 91125, USA}
 
@@
 \end{abstract}
+
+\maketitle
 
@@
 \begin{equation}\label{eq:signal-model}
-    x_i (\omega) = h_i (\omega; \boldsymbol\theta) + n_i (\omega),
+    x_i (\omega) = h_i (\omega; \bm\theta) + n_i (\omega),
 \end{equation}
 %
-where $h_i (\omega; \boldsymbol\theta)$ is the \ac{GW} signal given a parameter vector $\boldsymbol\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[\left|n_i(\omega)\right|^2\right] + E\left[\left|n_i(-\omega)\right|^2\right] = 2 E\left[\left|n_i(\omega)\right|^2\right]$.
+where $h_i (\omega; \bm\theta)$ is the \ac{GW} signal given a parameter vector $\bm\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[\left|n_i(\omega)\right|^2\right] + E\left[\left|n_i(-\omega)\right|^2\right] = 2 E\left[\left|n_i(\omega)\right|^2\right]$.
 We shall denote the combined observation from a network of detectors as $\mathbf x (\omega) \equiv \{x_i (\omega)\}_i$.
 
-The likelihood, or the probability of obtaining the observation $\mathbf x$ conditioned on the value of $\boldsymbol\theta$, is a product of Gaussian distributions:
+The likelihood, or the probability of obtaining the observation $\mathbf x$ conditioned on the value of $\bm\theta$, is a product of Gaussian distributions:
 %
-\begin{equation}\label{eq:gaussian-likelihood}
-    \mathcal{L}(\mathbf x; \boldsymbol\theta) = \prod_i p(x_i | \boldsymbol\theta)
-        \propto \exp \left[
+\begin{multline}\label{eq:gaussian-likelihood}
+    \mathcal{L}(\mathbf x; \bm\theta) = \prod_i p(x_i | \bm\theta)
+        \\ \propto \exp \left[
         - \frac{1}{2} \sum_i \int_0^\infty \frac{\left|x_i (\omega)
-            - h_i(\omega; \boldsymbol\theta) \right|^2}{S_i(\omega)} \, d\omega
+            - h_i(\omega; \bm\theta) \right|^2}{S_i(\omega)} \, d\omega
     \right].
-\end{equation}
+\end{multline}
 
@@
 \begin{equation}
-    \boldsymbol\theta = \begin{array}{l@{}l}
-            \left[
-            \begin{array}{c}
-                \alpha \\
-                \delta \\
-                r \\
-                t_\oplus \\
-                \iota \\
-                \psi \\
-                \phi_c \\
-                \hdashline[1pt/1pt]
-                m_1 \\
-                m_2 \\
-                \mathbf S_1 \\
-                \mathbf S_2
-            \end{array}
-            \right] &
-            \begin{array}{l}
-                \left.
-                \begin{array}{p{4cm}}
-                    right ascension \\
-                    declination \\
-                    distance \\
-                    arrival time at geocenter \\
-                    inclination angle \\
-                    polarization angle \\
-                    coalescence phase \\
-                    \hdashline[1pt/1pt]
-                \end{array}
-                \quad \right\} \textnormal{extrinsic parameters}, \boldsymbol\theta_\mathrm{ex} \\
-                \left.
-                \begin{array}{p{4cm}}
-                    first component's mass \\
-                    second component's mass \\
-                    first component's spin \\
-                    second component's spin
-                \end{array}
-                \quad \right\} \textnormal{intrinsic parameters}, \boldsymbol\theta_\mathrm{in}
-            \end{array}
-        \end{array}
+    \bm\theta = \left[
+    \begin{array}{cl@{\quad}p{4cm}lp{3.75cm}}
+        \alpha & \rdelim]{11}{0mm} & right ascension & \rdelim\}{7}{1mm} & \multirow{7}{2cm}{extrinsic parameters, $\bm\theta_\mathrm{ex}$} \\
+        \delta && declination & \\
+        r && distance & \\
+        t_\oplus && arrival time at geocenter & \\
+        \iota && inclination angle & \\
+        \psi && polarization angle & \\
+        \phi_c && coalescence phase & \\
+        \cline{1-1}\cline{3-3}
+        m_1 && first component's mass & \rdelim\}{4}{1mm} & \multirow{4}{2cm}{intrinsic parameters, $\bm\theta_\mathrm{in}$.}\\
+        m_2 && second component's mass & \\
+        \mathbf S_1 && first component's spin & \\
+        \mathbf S_2 && second component's spin & \\
+    \end{array}\right.
 \end{equation}
@@
     h_{\pi/2}(\omega) = h_0(\omega) \cdot
-    \cases{
-        -i & if $\omega \geq 0$ \\
-        i & if $\omega < 0$
-    }.
+    \begin{cases}
+        -i & \text{if } \omega \geq 0 \\
+        i & \text{if } \omega < 0
+    \end{cases}.
 \end{equation}
 %
-Then, we can write the signal model in a way that isolates all dependence on the extrinsic parameters $\boldsymbol\theta_\mathrm{ex}$ into the coefficients and all dependence on the intrinsic parameters $\boldsymbol\theta_\mathrm{in}$ into the basis waveform, by taking the Fourier transform of Equation~(2.8) of \cite{PhysRevD.83.084002}:
+Then, we can write the signal model in a way that isolates all dependence on the extrinsic parameters $\bm\theta_\mathrm{ex}$ into the coefficients and all dependence on the intrinsic parameters $\bm\theta_\mathrm{in}$ into the basis waveform, by taking the Fourier transform of Equation~(2.8) of \cite{PhysRevD.83.084002}:
 %
-\begin{equation}\label{eq:full-signal-model}
-    h_i(\omega; \boldsymbol\theta) = \e^{-i \omega t_\oplus}
+\begin{multline}\label{eq:full-signal-model}
+    h_i(\omega; \bm\theta) = e^{-i \omega t_\oplus}
     \frac{r_{1,i}}{r}
-    \e^{2 i \phi_c}
+    e^{2 i \phi_c} \\
     \left[
@@
     \right]
-    h(\omega; \boldsymbol\theta_\mathrm{in})
-\end{equation}
+    h(\omega; \bm\theta_\mathrm{in})
+\end{multline}
 %
@@
 
-More succinctly, we can write the signal received by detector $i$ in terms of observable extrinsic parameters $\boldsymbol\theta_i = (\rho_i, \gamma_i, \tau_i)$, the amplitude $\rho_i$, phase $\gamma_i$, and time delay $\tau_i$ on arrival at detector $i$,
+More succinctly, we can write the signal received by detector $i$ in terms of observable extrinsic parameters $\bm\theta_i = (\rho_i, \gamma_i, \tau_i)$, the amplitude $\rho_i$, phase $\gamma_i$, and time delay $\tau_i$ on arrival at detector $i$,
 %
 \begin{equation}\label{eq:signal-model}
-    h_i (\omega; \boldsymbol\theta_i, \boldsymbol\theta_\mathrm{in}) = h_i (\omega; \rho_i, \gamma_i, \tau_i, \boldsymbol\theta_\mathrm{in}) = \frac{\rho_i}{\sigma_i} e^{i (\gamma_i - \omega \tau_i)} h(\omega; \boldsymbol\theta_\mathrm{in}).
+    h_i (\omega; \bm\theta_i, \bm\theta_\mathrm{in}) = h_i (\omega; \rho_i, \gamma_i, \tau_i, \bm\theta_\mathrm{in}) = \frac{\rho_i}{\sigma_i} e^{i (\gamma_i - \omega \tau_i)} h(\omega; \bm\theta_\mathrm{in}).
 \end{equation}
 
-The prevailing technique for detection of \acp{GW} from \acp{CBC} is to realize a \ac{MLE} from the likelihood in \Eref{eq:gaussian-likelihood} and the signal model in \Eref{eq:signal-model}.
+The prevailing technique for detection of \acp{GW} from \acp{CBC} is to realize a \ac{MLE} from the likelihood in Equation~(\ref{eq:gaussian-likelihood}) and the signal model in Equation~(\ref{eq:signal-model}).
 Concretely, this results in a bank of matched filters, or the cross-correlation between a template waveform and the incoming data stream,
@@
 \begin{equation}
-y_i(\tau_i;\boldsymbol\theta_\mathrm{in}) = \frac{1}{\sigma_i (\boldsymbol\theta_\mathrm{in})} \int_0^\infty \frac{h^*(\omega; \boldsymbol\theta_\mathrm{in}) x_i(\omega) e^{i \omega \tau_i}}{S_i(\omega)} \,d\omega.
+y_i(\tau_i;\bm\theta_\mathrm{in}) = \frac{1}{\sigma_i (\bm\theta_\mathrm{in})} \int_0^\infty \frac{h^*(\omega; \bm\theta_\mathrm{in}) x_i(\omega) e^{i \omega \tau_i}}{S_i(\omega)} \,d\omega.
 \end{equation}
 %
-The \ac{ML} point estimates of the signal parameters, $\mathrm{MLE}(\mathbf{x}) = \{\{ \hat{\boldsymbol\theta}_i \}_i, \hat{\boldsymbol\theta}_\mathrm{in}\} = \{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\boldsymbol\theta}_\mathrm{in}\}$, are given by
+The \ac{ML} point estimates of the signal parameters, $\mathrm{MLE}(\mathbf{x}) = \{\{ \hat{\bm\theta}_i \}_i, \hat{\bm\theta}_\mathrm{in}\} = \{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$, are given by
 %
 \begin{eqnarray}
-    \hat{\boldsymbol\theta}_\mathrm{in}, \{\hat\tau_i\}_i
-        &=& \argmax_{\boldsymbol\theta_\mathrm{in}, \{\hat\tau_i\}_i}
+    \hat{\bm\theta}_\mathrm{in}, \{\hat\tau_i\}_i
+        &=& \argmax_{\bm\theta_\mathrm{in}, \{\hat\tau_i\}_i}
         \sum_i \left| y_i\left(\tau_i;
-        \boldsymbol\theta_\mathrm{in}\right) \right|^2, \\
+        \bm\theta_\mathrm{in}\right) \right|^2, \\
     \hat\rho_i &=& \left| y_i\left(\hat\tau_i;
-        \hat{\boldsymbol\theta}_\mathrm{in}\right) \right|, \\
+        \hat{\bm\theta}_\mathrm{in}\right) \right|, \\
     \hat\gamma_i &=& \arg y_i\left(\hat\tau_i;
-        \hat{\boldsymbol\theta}_\mathrm{in}\right).
+        \hat{\bm\theta}_\mathrm{in}\right).
 \end{eqnarray}
 %
-A detection candidate consists of $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\boldsymbol\theta}_\mathrm{in}\}$.
+A detection candidate consists of $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$.
 There are various ways to characterize the significance of a detection candidate, but in Gaussian noise the combined SNR $\hat\rho_\mathrm{c}$ suffices,
@@
 
-In the Bayesian framework, the parameters are inferred from the data by forming the posterior distribution, $p(\boldsymbol\theta|\mathbf x)$, which describes the probability of the parameters given the observations.
-Bayes' rule relates the likelihood $p(\mathbf x|\boldsymbol\theta)$ to the posterior $p(\boldsymbol\theta|\mathbf x)$,
+In the Bayesian framework, the parameters are inferred from the data by forming the posterior distribution, $p(\bm\theta|\mathbf x)$, which describes the probability of the parameters given the observations.
+Bayes' rule relates the likelihood $p(\mathbf x|\bm\theta)$ to the posterior $p(\bm\theta|\mathbf x)$,
 %
 \begin{equation}\label{bayes}
-p(\boldsymbol\theta|\mathbf x) = \frac{p(\mathbf x|\boldsymbol\theta) p(\boldsymbol\theta)}{p(\mathbf x)},
+p(\bm\theta|\mathbf x) = \frac{p(\mathbf x|\bm\theta) p(\bm\theta)}{p(\mathbf x)},
 \end{equation}
 %
-introducing the prior distribution $p(\boldsymbol\theta)$ which encapsulates previous information about the paramters (for example, arising from earlier observations or from known physical bounds on the parameters), and the evidence $p(\mathbf x)$ which can be thought of as a normalization factor or as describing the parsimoniousness of the model.
+introducing the prior distribution $p(\bm\theta)$ which encapsulates previous information about the paramters (for example, arising from earlier observations or from known physical bounds on the parameters), and the evidence $p(\mathbf x)$ which can be thought of as a normalization factor or as describing the parsimoniousness of the model.
 
@@
 
-In Bayesian inference, although it is often easy to write down the likelihood or even the full posterior in closed form, usually one is interested in only a subset $\boldsymbol\beta$ of all of the model's parameters, the others $\boldsymbol\lambda$ being nuisance parameters.
+In Bayesian inference, although it is often easy to write down the likelihood or even the full posterior in closed form, usually one is interested in only a subset $\bm\beta$ of all of the model's parameters, the others $\bm\lambda$ being nuisance parameters.
 In this case, we integrate away the nuisance parameters, forming the marginal posterior
@@
 \begin{equation}\label{eq:marginal-posterior}
-    p(\boldsymbol\beta|\mathbf x) = \int \frac{p(\mathbf x|\boldsymbol\beta,\boldsymbol\lambda) p(\boldsymbol\beta,\boldsymbol\lambda)}{p(\mathbf x)} \,d\boldsymbol\lambda
+    p(\bm\beta|\mathbf x) = \int \frac{p(\mathbf x|\bm\beta,\bm\lambda) p(\bm\beta,\bm\lambda)}{p(\mathbf x)} \,d\bm\lambda
 \end{equation}
 %
-with $\boldsymbol\theta = (\boldsymbol\beta, \boldsymbol\lambda)$. For instance, for the purpose of locating a \ac{GW} source on the sky, all parameters but $(\alpha, \delta)$ are nuisance parameters.
+with $\bm\theta = (\bm\beta, \bm\lambda)$. For instance, for the purpose of locating a \ac{GW} source on the sky, all parameters but $(\alpha, \delta)$ are nuisance parameters.
 
@@
 
-For the purpose of rapid sky localization, we assume that we do not have access to the \ac{GW} data $\mathbf{x}$ itself, and that our only contact with it is through the ML parameter estimates $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\boldsymbol\theta}_\mathrm{in}\}$.
-Although this is a significant departure from conventional GW parameter estimation techniques, we can still apply the full Bayesian machinery of \Eref{eq:marginal-posterior} to compute a posterior distribution for the sky location.
+For the purpose of rapid sky localization, we assume that we do not have access to the \ac{GW} data $\mathbf{x}$ itself, and that our only contact with it is through the ML parameter estimates $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$.
+Although this is a significant departure from conventional GW parameter estimation techniques, we can still apply the full Bayesian machinery of Equation~(\ref{eq:marginal-posterior}) to compute a posterior distribution for the sky location.
 
@@
 \begin{equation}\label{eq:detection-candidate-likelihood}
-    p\left(\{\hat{\boldsymbol\theta}_i\}_i,
-        \hat{\boldsymbol\theta}_\mathrm{in}
-    \middle| \boldsymbol\theta\right)
+    p\left(\{\hat{\bm\theta}_i\}_i,
+        \hat{\bm\theta}_\mathrm{in}
+    \middle| \bm\theta\right)
     \propto \int\limits_{\mathclap{\mathbf{x} | \mathrm{MLE}(\mathbf{x}) =
-        \{\{\hat{\boldsymbol\theta}_i\}_i,
-        \hat{\boldsymbol\theta}_\mathrm{in}\}}}
-    p(\mathbf{x} | \boldsymbol\theta) \, p(\boldsymbol\theta)
+        \{\{\hat{\bm\theta}_i\}_i,
+        \hat{\bm\theta}_\mathrm{in}\}}}
+    p(\mathbf{x} | \bm\theta) \, p(\bm\theta)
     \, d\mathbf{x}
@@
 %
-The integral above could be written in terms of the matched filter signals $\left\{y_i\right\}_i$, such that $y_i\left(\hat\tau_i, \hat{\boldsymbol\theta}_\mathrm{in}\right) = \hat\rho_i e^{i \hat\gamma_i}$ and $\forall t$, $\left|y_i(t, \boldsymbol\theta_\mathrm{in})\right| \leq \hat\rho_i$.
+The integral above could be written in terms of the matched filter signals $\left\{y_i\right\}_i$, such that $y_i\left(\hat\tau_i, \hat{\bm\theta}_\mathrm{in}\right) = \hat\rho_i e^{i \hat\gamma_i}$ and $\forall t$, $\left|y_i(t, \bm\theta_\mathrm{in})\right| \leq \hat\rho_i$.
 The combination of the nontrivial integration limits and the Fourier transform in the matched filter seem like a problem for direct evaluation.
@@
 %
-\begin{equation}\label{eq:gaussian-likelihood-spa}
+\begin{multline}\label{eq:gaussian-likelihood-spa}
     \mathcal{L}\left(x_i; \rho_i, \gamma_i, \tau_i,
-        \boldsymbol\theta_\mathrm{in}\right)
+        \bm\theta_\mathrm{in}\right) \\
     \propto \exp \left[
@@
             - h_i\left(x_i; \rho_i, \gamma_i, \tau_i,
-                \boldsymbol\theta_\mathrm{in}\right)
+                \bm\theta_\mathrm{in}\right)
         \right|^2}{S_i(\omega)} \, d\omega
     \right],
-\end{equation}
+\end{multline}
 %
-with $h_i(\omega; \rho_i, \gamma_i, \tau_i, \boldsymbol\theta_\mathrm{in})$ given by \Eref{eq:signal-model}.
+with $h_i(\omega; \rho_i, \gamma_i, \tau_i, \bm\theta_\mathrm{in})$ given by Equation~(\ref{eq:signal-model}).
 
-The Fisher information matrix for a measurement $x$ described by the unknown parameter vector $\boldsymbol{\theta}$ is
+The Fisher information matrix for a measurement $x$ described by the unknown parameter vector $\bm{\theta}$ is
 %
@@
         -\frac{\partial^2 \log
-            \mathcal{L}(x_i ; \boldsymbol\theta)}
+            \mathcal{L}(x_i ; \bm\theta)}
             {\partial \theta_j \theta_k}
-    \middle| \boldsymbol\theta
+    \middle| \bm\theta
     \right].
@@
         \begin{array}{cc}
-            \mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_i} &
-            \mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_\mathrm{in}} \\
-            {\mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_\mathrm{in}}}^\intercal &
-            {\rho_i}^2 \mathcal{I}_{\boldsymbol\theta_\mathrm{in},\boldsymbol\theta_\mathrm{in}}
+            \mathcal{I}_{\bm\theta_i,\bm\theta_i} &
+            \mathcal{I}_{\bm\theta_i,\bm\theta_\mathrm{in}} \\
+            {\mathcal{I}_{\bm\theta_i,\bm\theta_\mathrm{in}}}^\intercal &
+            {\rho_i}^2 \mathcal{I}_{\bm\theta_\mathrm{in},\bm\theta_\mathrm{in}}
         \end{array}
@@
 \begin{equation}\label{eq:fisher-matrix}
-    \mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_i} = \bordermatrix{
+    \mathcal{I}_{\bm\theta_i,\bm\theta_i} = \bordermatrix{
         ~ & \rho_i & \gamma_i & \tau_i \cr
@@
 %
-(See also \cite{Grover:2013}.) To lowest post\nobreakdashes-Newtonian order, the intrinsic parameters affect the phase evolution but not the amplitude evolution, so for BNS systems we can assume $\mathcal{I}_{\rho_i,\boldsymbol\theta_\mathrm{in}} \approx 0$.
+(See also \cite{Grover:2013}.) To lowest post\nobreakdashes-Newtonian order, the intrinsic parameters affect the phase evolution but not the amplitude evolution, so for BNS systems we can assume $\mathcal{I}_{\rho_i,\bm\theta_\mathrm{in}} \approx 0$.
 %
@@
 When we form the Fisher matrix for the whole network, if the detectors all have the same noise \acp{PSD}, $S_1(\omega) = S_2(\omega) = \cdots = S_n(\omega) \equiv S(\omega)$, then we can change variables from phases and times on arrival $\left\{\gamma_i, \tau_i\right\}_i$ to appropriate differences between them, in order to cancel out all off-diagonal terms between the intrinsic and extrinsic parameters.
-Because all existing ground-based GW detectors have similar bandpasses, we restrict all of our further attention to just the extrinsic part of the Fisher matrix, $\mathcal{I}_{\boldsymbol\theta_i,\boldsymbol\theta_i}$.
+Because all existing ground-based GW detectors have similar bandpasses, we restrict all of our further attention to just the extrinsic part of the Fisher matrix, $\mathcal{I}_{\bm\theta_i,\bm\theta_i}$.
 
@@
 
-If $\hat{\boldsymbol\theta}$ is an unbiased estimator of $\boldsymbol\theta$, $\tilde{\boldsymbol\theta} = \hat{\boldsymbol\theta} - \boldsymbol\theta$ is the measurement error, and $\Sigma = \mathrm{E} \, [\tilde{\boldsymbol\theta}\tilde{\boldsymbol\theta}^\intercal]$ is the covariance of the measurement error, then the \ac{CRLB} says that $\Sigma \geq \mathcal{I}^{-1}$, in the sense that $\left(\Sigma - \mathcal{I}^{-1}\right)$ is positive semi\nobreakdashes-definite.
+If $\hat{\bm\theta}$ is an unbiased estimator of $\bm\theta$, $\tilde{\bm\theta} = \hat{\bm\theta} - \bm\theta$ is the measurement error, and $\Sigma = \mathrm{E} \, [\tilde{\bm\theta}\tilde{\bm\theta}^\intercal]$ is the covariance of the measurement error, then the \ac{CRLB} says that $\Sigma \geq \mathcal{I}^{-1}$, in the sense that $\left(\Sigma - \mathcal{I}^{-1}\right)$ is positive semi\nobreakdashes-definite.
 Specifically, for our likelihood, the \ac{CRLB} implies that
@@
 %
-Furthermore, the Fisher matrix in \Eref{eq:fisher-matrix} is block diagonal, which implies that estimation errors in the signal amplitude $\rho$ are uncorrelated with the phase $\gamma$ and time $\tau$.
+Furthermore, the Fisher matrix in Equation~(\ref{eq:fisher-matrix}) is block diagonal, which implies that estimation errors in the signal amplitude $\rho$ are uncorrelated with the phase $\gamma$ and time $\tau$.
 A sequence of two changes of variables lends some physical interpretation to the nature of the coupled estimation errors in $\gamma$ and $\tau$.
@@
 
-Although we may not be able to compute the detection candidate likelihood, \Eref{eq:detection-candidate-likelihood}, directly, we can make an educated guess that has many properties in common with it.
-Any valid approximate likelihood must have the same Fisher matrix \Eref{eq:fisher-matrix-extrinsic-one-detector}.
+Although we may not be able to compute the detection candidate likelihood, Equation~(\ref{eq:detection-candidate-likelihood}), directly, we can make an educated guess that has many properties in common with it.
+Any valid approximate likelihood must have the same Fisher matrix Equation~(\ref{eq:fisher-matrix-extrinsic-one-detector}).
 It must also have the same limiting behavior: it should be periodic in the phase error $\tilde{\gamma}_i$ and go to zero as $\tilde{\tau}_i \rightarrow \pm \infty$, $\hat{\rho}_i \rightarrow 0$, or $\hat{\rho}_i \rightarrow \infty$.
@@
 %
-Another way is to plug the signal model from \Eref{eq:signal-model} \emph{evaluated at the ML parameter estimates} into the single-detector likelihood in \Eref{eq:gaussian-likelihood-spa}:
+Another way is to plug the signal model from Equation~(\ref{eq:signal-model}) \emph{evaluated at the ML parameter estimates} into the single-detector likelihood in Equation~(\ref{eq:gaussian-likelihood-spa}):
 %
 \begin{eqnarray*}
-    \fl p\left(\hat{\boldsymbol\theta}_i \middle| \boldsymbol\theta \right)
+    p\left(\hat{\bm\theta}_i \middle| \bm\theta \right)
     &:=&
-    p\left(x_i(\omega) = h_i(\omega; \hat{\boldsymbol\theta}_i)
-        \middle| \boldsymbol\theta \right)
-    \\\fl
+    p\left(x_i(\omega) = h_i(\omega; \hat{\bm\theta}_i)
+        \middle| \bm\theta \right)
+    \\
     &\propto& \exp \left[
         - \frac{1}{2} \int_0^\infty \frac{\left|
-            \frac{\hat{\rho}_i}{\sigma_i(\hat{\boldsymbol\theta}_\mathrm{in})} e^{i (\hat\gamma_i - \omega \hat\tau_i)} h(\omega; \hat{\boldsymbol\theta}_\mathrm{in})
-            - \frac{\rho_i}{\sigma_i(\boldsymbol\theta_\mathrm{in})} e^{i (\gamma_i - \omega \tau_i)} h(\omega; \boldsymbol\theta_\mathrm{in})
+            \frac{\hat{\rho}_i}{\sigma_i(\hat{\bm\theta}_\mathrm{in})} e^{i (\hat\gamma_i - \omega \hat\tau_i)} h(\omega; \hat{\bm\theta}_\mathrm{in})
+            - \frac{\rho_i}{\sigma_i(\bm\theta_\mathrm{in})} e^{i (\gamma_i - \omega \tau_i)} h(\omega; \bm\theta_\mathrm{in})
         \right|^2}{S_i(\omega)} \, d\omega
@@
 %
-If we assume that $\hat{\boldsymbol\theta}_\mathrm{in} = \boldsymbol\theta_\mathrm{in}$, then this reduces to
+If we assume that $\hat{\bm\theta}_\mathrm{in} = \bm\theta_\mathrm{in}$, then this reduces to
 %
@@
 %
-with $\tilde{\gamma}_i = \hat\gamma_i - \gamma_i$, $\tilde{\tau}_i = \hat\tau_i - \tau_i$, and the template's autocorrelation function $a_i(t; \boldsymbol\theta_\mathrm{in})$ defined as
+with $\tilde{\gamma}_i = \hat\gamma_i - \gamma_i$, $\tilde{\tau}_i = \hat\tau_i - \tau_i$, and the template's autocorrelation function $a_i(t; \bm\theta_\mathrm{in})$ defined as
 %
 \begin{equation}\label{eq:autocorrelation-function}
-    a_i(t; \boldsymbol\theta_\mathrm{in}) := \frac{1}{{{\sigma_i}^2(\boldsymbol\theta_\mathrm{in})}} \int_0^\infty \frac{\left| h(\omega; \hat{\boldsymbol\theta}_\mathrm{in})\right|^2}{S_i(\omega)} e^{i \omega t} \,d\omega.
+    a_i(t; \bm\theta_\mathrm{in}) := \frac{1}{{{\sigma_i}^2(\bm\theta_\mathrm{in})}} \int_0^\infty \frac{\left| h(\omega; \hat{\bm\theta}_\mathrm{in})\right|^2}{S_i(\omega)} e^{i \omega t} \,d\omega.
 \end{equation}
@@
 %
-\begin{equation}
-    \fl p\left(\left\{\hat\rho_i, \hat\gamma_i, \hat\tau_i\right\}_i
-        \middle| \left\{\rho_i, \gamma_i, \tau_i\right\}_i \right) \propto
+\begin{multline}
+    p\left(\left\{\hat\rho_i, \hat\gamma_i, \hat\tau_i\right\}_i
+        \middle| \left\{\rho_i, \gamma_i, \tau_i\right\}_i \right) \\ \propto
     \exp \left[ - \frac{1}{2} \sum_i {\hat\rho_i}^2 - \frac{1}{2} \sum_i {\rho_i}^2
@@
     \right].
-\end{equation}
+\end{multline}
 
@@
 
-First, observe that at the true parameter values, $\hat{\boldsymbol\theta}_i = \boldsymbol\theta_i$, the logarithms of \Eref{eq:autocor-likelihood} and \Eref{eq:gaussian-likelihood-spa} have the same Jacobian.
+First, observe that at the true parameter values, $\hat{\bm\theta}_i = \bm\theta_i$, the logarithms of Equation~(\ref{eq:autocor-likelihood}) and Equation~(\ref{eq:gaussian-likelihood-spa}) have the same Jacobian.
 This is because the derivatives of the autocorrelation function are
@@
 
-We can compute the Fisher matrix elements for the autocorrelation likelihood given by \Eref{eq:autocor-likelihood}, with detector subscript suppressed:
+We can compute the Fisher matrix elements for the autocorrelation likelihood given by Equation~(\ref{eq:autocor-likelihood}), with detector subscript suppressed:
 %
@@
 
-If we can show that the weighting function $w(t; \rho)$ approaches a Dirac delta function as $\rho \rightarrow \infty$, then the Fisher matrix for the autocorrelation likelihood approaches the Fisher matrix for the full GW data, \Eref{eq:fisher-matrix}, as $\rho \rightarrow \infty$.
+If we can show that the weighting function $w(t; \rho)$ approaches a Dirac delta function as $\rho \rightarrow \infty$, then the Fisher matrix for the autocorrelation likelihood approaches the Fisher matrix for the full GW data, Equation~(\ref{eq:fisher-matrix}), as $\rho \rightarrow \infty$.
 The Bessel functions asymptotically approach:
@@
 
-The detection pipeline supplies a candidate $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\boldsymbol\theta}_\mathrm{in}\}$, and discretely sampled noise \acp{PSD} $S_i(\omega_j)$ for all detectors.
+The detection pipeline supplies a candidate $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$, and discretely sampled noise \acp{PSD} $S_i(\omega_j)$ for all detectors.
 \marginpar{Discuss single-detector threshold here?}
 
-We compute the GW signal for a source with intrinsic parameters equal to the detection pipeline's estimate, $h(\omega; \hat{\boldsymbol\theta}_\mathrm{in})$.
-Then we find the SNR=1 horizon distance $r_{1,i}$ for each detector by numerically integrating \Eref{eq:horizon}.
+We compute the GW signal for a source with intrinsic parameters equal to the detection pipeline's estimate, $h(\omega; \hat{\bm\theta}_\mathrm{in})$.
+Then we find the SNR=1 horizon distance $r_{1,i}$ for each detector by numerically integrating Equation~(\ref{eq:horizon}).
 
-We have no explicit prior on the intrinsic parameters; in our analysis they are fixed at their \ac{ML} estimates, $\hat{\boldsymbol\theta}_\mathrm{in}$.
+We have no explicit prior on the intrinsic parameters; in our analysis they are fixed at their \ac{ML} estimates, $\hat{\bm\theta}_\mathrm{in}$.
 
@@
 \begin{equation*}
-    p(r) \propto \cases{
-        r^m & if $r_\mathrm{min} < r < r_\mathrm{max}$ \\
-        0 & otherwise
-    }
+    p(r) \propto \begin{cases}
+        r^m & \text{if } r_\mathrm{min} < r < r_\mathrm{max} \\
+        0 & \text{otherwise}
+    \end{cases}
 \end{equation*}
@@
 %
-\begin{eqnarray}
-    \fl f(\alpha, \delta) \propto
+\begin{multline}
+    f(\alpha, \delta) \propto
     \int_{0}^{\pi}
@@
         \right\}
-    \right] r^m \, dt_\oplus \, d\phi_c dr \, d\cos{\iota} \, d\psi.
-\end{eqnarray}
+    \right] \\
+    r^m \, dt_\oplus \, d\phi_c dr \, d\cos{\iota} \, d\psi.
+\end{multline}
 
@@
 %
-\begin{eqnarray}
-    \fl f(\alpha, \delta) \rightarrow
+\begin{multline}
+    f(\alpha, \delta) \rightarrow
     \int_{0}^{\pi}
@@
             \right|
-    \right]
-    r^m dr \, dt_\oplus \, d\cos{\iota} \, d\psi.
-\end{eqnarray}
+    \right] \\
+    r^m dr dt_\oplus \, d\cos{\iota} \, d\psi.
+\end{multline}
 
@@
 %
-If we apply the same change of variables to \Eref{eq:importance-sampling-difficult-integral}, we get
+If we apply the same change of variables to Equation~(\ref{eq:importance-sampling-difficult-integral}), we get
 %
@@
 If $f(x)$ and $g(x)$ have sufficiently similar behavior, then their quotient $f/g$ will be better behaved than $f$ itself, and more amenable to any given quadrature technique.
-When the transformed integral \Eref{eq:importance-sampling} is evaluated using Monte Carlo integration, this technique is referred to as importance sampling.
+When the transformed integral Equation~(\ref{eq:importance-sampling}) is evaluated using Monte Carlo integration, this technique is referred to as importance sampling.
 
 We use the same kind of change of variables with fixed\nobreakdashes-order Gaussian quadrature.
-We first find the approximate \ac{ML} distance $r_0$ using \Eref{eq:ml-distance}.
+We first find the approximate \ac{ML} distance $r_0$ using Equation~(\ref{eq:ml-distance}).
 If $r_0 < r_\mathrm{min}$ or $r_0 > r_\mathrm{max}$, then the ML peak lies outside of the limits of integration and the original integrand behaves like a low-order power law.
-In this case, we use 10\nobreakdashes-point Gaussian quadrature to evaluate \Eref{eq:distance-integral}.
+In this case, we use 10\nobreakdashes-point Gaussian quadrature to evaluate Equation~(\ref{eq:distance-integral}).
 
 If, on the other hand, $r_\mathrm{min} \leq r_0 \leq r_\mathrm{max}$, then the integrand contains a sharp peak that we can smooth out with importance sampling.
-We first re-scale the distance integral \Eref{eq:distance-integral} so that the peak value of the integrand is $\sim 1$:
+We first re-scale the distance integral Equation~(\ref{eq:distance-integral}) so that the peak value of the integrand is $\sim 1$:
 %
@@
 
-Last, there are a few special cases where \Eref{eq:distance-integral} can be evaluated exactly.
+Last, there are a few special cases where Equation~(\ref{eq:distance-integral}) can be evaluated exactly.
 If $p^2=0$ (which implies $b = 0$), then the integral may be expressed in terms of logarithms.
@@
 
-\ack \marginpar{This URL is not specific enough.}Source code for \ac{BAYESTAR} is available on the \ac{LIGO} \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/lalsuite.html}.
+\begin{acknowledgements}
+\marginpar{This URL is not specific enough.}Source code for \ac{BAYESTAR} is available on the \ac{LIGO} \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/lalsuite.html}.
 
@@
 This paper has \ac{LIGO} Document Number \ac{LIGO}\nobreakdashes-PXXXXXXX\nobreakdashes-vX.
+\end{acknowledgements}
 
-\section*{References}
-\bibliographystyle{iopart-num}
 \bibliography{apj-jour,bib/telescope,ms}

### 2014-08-14 ms.tex
Rework notation

@@
 \begin{equation}\label{eq:signal-model}
-    x_i (\omega) = h_i (\omega; \bm\theta) + n_i (\omega),
+    y_i (\omega) = x_i (\omega; \bm\theta) + n_i (\omega),
 \end{equation}
 %
-where $h_i (\omega; \bm\theta)$ is the \ac{GW} signal given a parameter vector $\bm\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[\left|n_i(\omega)\right|^2\right] + E\left[\left|n_i(-\omega)\right|^2\right] = 2 E\left[\left|n_i(\omega)\right|^2\right]$.
-We shall denote the combined observation from a network of detectors as $\mathbf x (\omega) \equiv \{x_i (\omega)\}_i$.
+where $x_i (\omega; \bm\theta)$ is the \ac{GW} signal given a parameter vector $\bm\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[\left|n_i(\omega)\right|^2\right] + E\left[\left|n_i(-\omega)\right|^2\right] = 2 E\left[\left|n_i(\omega)\right|^2\right]$.
+We shall denote the combined observation from a network of detectors as $\mathbf y (\omega) \equiv \{y_i (\omega)\}_i$.
 
-The likelihood, or the probability of obtaining the observation $\mathbf x$ conditioned on the value of $\bm\theta$, is a product of Gaussian distributions:
+The likelihood, or the probability of obtaining the observation $\mathbf y$ conditioned on the value of $\bm\theta$, is a product of Gaussian distributions:
 %
 \begin{multline}\label{eq:gaussian-likelihood}
-    \mathcal{L}(\mathbf x; \bm\theta) = \prod_i p(x_i | \bm\theta)
+    \mathcal{L}(\mathbf y; \bm\theta) = \prod_i p(y_i | \bm\theta)
         \\ \propto \exp \left[
-        - \frac{1}{2} \sum_i \int_0^\infty \frac{\left|x_i (\omega)
-            - h_i(\omega; \bm\theta) \right|^2}{S_i(\omega)} \, d\omega
+        - \frac{1}{2} \sum_i \int_0^\infty \frac{\left|y_i (\omega)
+            - x_i(\omega; \bm\theta) \right|^2}{S_i(\omega)} \, d\omega
     \right].
@@
 
-Assuming circular orbits and no spin precession, we can write the \ac{GW} signal in each detector as a linear combination of two basis waveforms $h_0$ and $h_{\pi/2}$.
+Assuming circular orbits and no spin precession, we can write the \ac{GW} signal in each detector as a linear combination of two basis waveforms $h = h_0$ and $h_{\pi/2}$.
 For nonprecessing systems, $h_0$ and $h_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the sine and cosine functions, being nearly orthogonal and out of phase by ${\pi/2}$ at all frequencies.
@@
 \begin{multline}\label{eq:full-signal-model}
-    h_i(\omega; \bm\theta) = e^{-i \omega t_\oplus}
+    x_i(\omega; \bm\theta) = e^{-i \omega t_\oplus}
     \frac{r_{1,i}}{r}
@@
     \left[
-    \frac{1}{2} \left(1 + \cos^2 \iota\right) \Re \left\{z\right\} - i
-    \left(\cos\iota\right) \Im z
+    \frac{1}{2} \left(1 + \cos^2 \iota\right) \Re \left\{\zeta\right\} - i
+    \left(\cos\iota\right) \Im \left\{\zeta\right\}
     \right]
@@
 \begin{equation}
-    z = e^{-2 i \psi} \left(
+    \zeta = e^{-2 i \psi} \left(
     F_{+,i}(\alpha, \delta, t_\oplus) +
@@
 \begin{equation}\label{eq:signal-model}
-    h_i (\omega; \bm\theta_i, \bm\theta_\mathrm{in}) = h_i (\omega; \rho_i, \gamma_i, \tau_i, \bm\theta_\mathrm{in}) = \frac{\rho_i}{\sigma_i} e^{i (\gamma_i - \omega \tau_i)} h(\omega; \bm\theta_\mathrm{in}).
+    x_i (\omega; \bm\theta_i, \bm\theta_\mathrm{in}) = x_i (\omega; \rho_i, \gamma_i, \tau_i, \bm\theta_\mathrm{in}) = \frac{\rho_i}{\sigma_i} e^{i (\gamma_i - \omega \tau_i)} h(\omega; \bm\theta_\mathrm{in}).
 \end{equation}
@@
 \begin{equation}
-y_i(\tau_i;\bm\theta_\mathrm{in}) = \frac{1}{\sigma_i (\bm\theta_\mathrm{in})} \int_0^\infty \frac{h^*(\omega; \bm\theta_\mathrm{in}) x_i(\omega) e^{i \omega \tau_i}}{S_i(\omega)} \,d\omega.
+z_i(\tau_i;\bm\theta_\mathrm{in}) = \frac{1}{\sigma_i (\bm\theta_\mathrm{in})} \int_0^\infty \frac{h^*(\omega; \bm\theta_\mathrm{in}) y_i(\omega) e^{i \omega \tau_i}}{S_i(\omega)} \,d\omega.
 \end{equation}
 %
-The \ac{ML} point estimates of the signal parameters, $\mathrm{MLE}(\mathbf{x}) = \{\{ \hat{\bm\theta}_i \}_i, \hat{\bm\theta}_\mathrm{in}\} = \{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$, are given by
+The \ac{ML} point estimates of the signal parameters, $\mathrm{MLE}(\mathbf{y}) = \{\{ \hat{\bm\theta}_i \}_i, \hat{\bm\theta}_\mathrm{in}\} = \{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$, are given by
 %
@@
         &=& \argmax_{\bm\theta_\mathrm{in}, \{\hat\tau_i\}_i}
-        \sum_i \left| y_i\left(\tau_i;
+        \sum_i \left| z_i\left(\tau_i;
         \bm\theta_\mathrm{in}\right) \right|^2, \\
-    \hat\rho_i &=& \left| y_i\left(\hat\tau_i;
+    \hat\rho_i &=& \left| z_i\left(\hat\tau_i;
         \hat{\bm\theta}_\mathrm{in}\right) \right|, \\
-    \hat\gamma_i &=& \arg y_i\left(\hat\tau_i;
+    \hat\gamma_i &=& \arg z_i\left(\hat\tau_i;
         \hat{\bm\theta}_\mathrm{in}\right).
@@
 
-In the Bayesian framework, the parameters are inferred from the data by forming the posterior distribution, $p(\bm\theta|\mathbf x)$, which describes the probability of the parameters given the observations.
-Bayes' rule relates the likelihood $p(\mathbf x|\bm\theta)$ to the posterior $p(\bm\theta|\mathbf x)$,
+In the Bayesian framework, the parameters are inferred from the data by forming the posterior distribution, $p(\bm\theta|\mathbf y)$, which describes the probability of the parameters given the observations.
+Bayes' rule relates the likelihood $p(\mathbf y|\bm\theta)$ to the posterior $p(\bm\theta|\mathbf y)$,
 %
 \begin{equation}\label{bayes}
-p(\bm\theta|\mathbf x) = \frac{p(\mathbf x|\bm\theta) p(\bm\theta)}{p(\mathbf x)},
+p(\bm\theta|\mathbf y) = \frac{p(\mathbf y|\bm\theta) p(\bm\theta)}{p(\mathbf y)},
 \end{equation}
 %
-introducing the prior distribution $p(\bm\theta)$ which encapsulates previous information about the paramters (for example, arising from earlier observations or from known physical bounds on the parameters), and the evidence $p(\mathbf x)$ which can be thought of as a normalization factor or as describing the parsimoniousness of the model.
+introducing the prior distribution $p(\bm\theta)$ which encapsulates previous information about the paramters (for example, arising from earlier observations or from known physical bounds on the parameters), and the evidence $p(\mathbf y)$ which can be thought of as a normalization factor or as describing the parsimoniousness of the model.
 
@@
 \begin{equation}\label{eq:marginal-posterior}
-    p(\bm\beta|\mathbf x) = \int \frac{p(\mathbf x|\bm\beta,\bm\lambda) p(\bm\beta,\bm\lambda)}{p(\mathbf x)} \,d\bm\lambda
+    p(\bm\beta|\mathbf y) = \int \frac{p(\mathbf y|\bm\beta,\bm\lambda) p(\bm\beta,\bm\lambda)}{p(\mathbf y)} \,d\bm\lambda
 \end{equation}
@@
 
-For the purpose of rapid sky localization, we assume that we do not have access to the \ac{GW} data $\mathbf{x}$ itself, and that our only contact with it is through the ML parameter estimates $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$.
+For the purpose of rapid sky localization, we assume that we do not have access to the \ac{GW} data $\mathbf{y}$ itself, and that our only contact with it is through the ML parameter estimates $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$.
 Although this is a significant departure from conventional GW parameter estimation techniques, we can still apply the full Bayesian machinery of Equation~(\ref{eq:marginal-posterior}) to compute a posterior distribution for the sky location.
@@
 
-To treat the ML estimates as measurements in and of themselves, we need to work out their likelihood, marginalized over the GW signal $\mathbf{x}$,
+To treat the ML estimates as measurements in and of themselves, we need to work out their likelihood, marginalized over the GW signal $\mathbf{y}$,
 %
@@
     \middle| \bm\theta\right)
-    \propto \int\limits_{\mathclap{\mathbf{x} | \mathrm{MLE}(\mathbf{x}) =
+    \propto \int\limits_{\mathclap{\mathbf{y} | \mathrm{MLE}(\mathbf{y}) =
         \{\{\hat{\bm\theta}_i\}_i,
         \hat{\bm\theta}_\mathrm{in}\}}}
-    p(\mathbf{x} | \bm\theta) \, p(\bm\theta)
-    \, d\mathbf{x}
+    p(\mathbf{y} | \bm\theta) \, p(\bm\theta)
+    \, d\mathbf{y}
 \end{equation}
 %
-The integral above could be written in terms of the matched filter signals $\left\{y_i\right\}_i$, such that $y_i\left(\hat\tau_i, \hat{\bm\theta}_\mathrm{in}\right) = \hat\rho_i e^{i \hat\gamma_i}$ and $\forall t$, $\left|y_i(t, \bm\theta_\mathrm{in})\right| \leq \hat\rho_i$.
+The integral above could be written in terms of the matched filter signals $\left\{z_i\right\}_i$, such that $z_i\left(\hat\tau_i, \hat{\bm\theta}_\mathrm{in}\right) = \hat\rho_i e^{i \hat\gamma_i}$ and $\forall t$, $\left|z_i(t, \bm\theta_\mathrm{in})\right| \leq \hat\rho_i$.
 The combination of the nontrivial integration limits and the Fourier transform in the matched filter seem like a problem for direct evaluation.
@@
 \begin{multline}\label{eq:gaussian-likelihood-spa}
-    \mathcal{L}\left(x_i; \rho_i, \gamma_i, \tau_i,
+    \mathcal{L}\left(y_i; \rho_i, \gamma_i, \tau_i,
         \bm\theta_\mathrm{in}\right) \\
     \propto \exp \left[
-        - \frac{1}{2} \int_0^\infty \frac{\left|x_i (\omega)
-            - h_i\left(x_i; \rho_i, \gamma_i, \tau_i,
+        - \frac{1}{2} \int_0^\infty \frac{\left|y_i (\omega)
+            - x_i\left(y_i; \rho_i, \gamma_i, \tau_i,
                 \bm\theta_\mathrm{in}\right)
@@
 %
-with $h_i(\omega; \rho_i, \gamma_i, \tau_i, \bm\theta_\mathrm{in})$ given by Equation~(\ref{eq:signal-model}).
+with $x_i(\omega; \rho_i, \gamma_i, \tau_i, \bm\theta_\mathrm{in})$ given by Equation~(\ref{eq:signal-model}).
 
-The Fisher information matrix for a measurement $x$ described by the unknown parameter vector $\bm{\theta}$ is
+The Fisher information matrix for a measurement $y$ described by the unknown parameter vector $\bm{\theta}$ is
 %
@@
         -\frac{\partial^2 \log
-            \mathcal{L}(x_i ; \bm\theta)}
+            \mathcal{L}(y_i ; \bm\theta)}
             {\partial \theta_j \theta_k}
@@
     \mathcal{I}_{jk} = \int_0^\infty \Re \left[
-        \left(\frac{\partial h_i}{\partial \theta_j}\right)^*
-        \left(\frac{\partial h_i}{\partial \theta_k}\right)
+        \left(\frac{\partial x_i}{\partial \theta_j}\right)^*
+        \left(\frac{\partial x_i}{\partial \theta_k}\right)
     \right] \frac{1}{S_i(\omega)} \, d\omega.
@@
 %
-This form is more useful because it involves only first derivatives, and of the signal $h (\omega)$ rather than the entire observation $x (\omega)$.
+This form is more useful because it involves only first derivatives, and of the signal $x_i (\omega)$ rather than the entire observation $y (\omega)$.
 In terms of the $k$th \ac{SNR}-weighted moment of angular frequency,
@@
 It must also have the same limiting behavior: it should be periodic in the phase error $\tilde{\gamma}_i$ and go to zero as $\tilde{\tau}_i \rightarrow \pm \infty$, $\hat{\rho}_i \rightarrow 0$, or $\hat{\rho}_i \rightarrow \infty$.
-Additionally, when $\tilde{\tau}_i = 0$, the distribution of ${\hat{\rho}_i}^2$ should reduce to a noncentral $\chi^2$ distribution with two degrees of freedom, centered about ${\rho_i}^2$, because the complex matched filter time series $y_i(t)$ is Gaussian.
+Additionally, when $\tilde{\tau}_i = 0$, the distribution of ${\hat{\rho}_i}^2$ should reduce to a noncentral $\chi^2$ distribution with two degrees of freedom, centered about ${\rho_i}^2$, because the complex matched filter time series $z_i(t)$ is Gaussian.
 
@@
 %
+\begin{widetext}
 \begin{eqnarray*}
@@
     &:=&
-    p\left(x_i(\omega) = h_i(\omega; \hat{\bm\theta}_i)
+    p\left(y_i(\omega) = x_i(\omega; \hat{\bm\theta}_i)
         \middle| \bm\theta \right)
@@
 \end{eqnarray*}
+\end{widetext}
 %

### 2014-09-15 ms.tex
Redo abstract

@@
 \begin{abstract}
-\marginpar{Wrong emphasis? I like Larry's idea of having this paper discuss several different tiers of sky localization with increasing accuracy and latency.}%
-Timely \acl{EM} follow\nobreakdashes-up of \acl{CBC} events detected by \acl{aLIGO} requires rapidly inferring the sky location from the \ac{GW} observations.
-Calculation of the posterior distribution of the sky location and intrinsic source parameters given the \ac{GW} strain takes hours with state\nobreakdashes-of\nobreakdashes-the\nobreakdashes-art \ac{MCMC} parameter estimation codes.
-By taking as the measurement not the \ac{GW} strain itself but the amplitude and \acl{TOA} of the putative signal at each detector, and by fixing the intrinsic source parameters, we have constructed a non-\ac{MCMC}, fully deterministic Bayesian parameter estimation algorithm that takes just seconds to produce sky maps of posterior probability.
+Advanced LIGO and Virgo should begin detecting gravitational waves from binary neutron star and neutron star\nobreakdashes--black hole mergers within the next few years.
+These sources are also predicted to power a variety of electromagnetic transients.
+Because these signatures can be fast and faint, locating and observing them hinges on rapidly inferring the sky location from the gravitational wave observations.
+MCMC methods for gravitational\nobreakdashes-wave parameter estimation can take hours or more.
+We describe a new, non-MCMC, fully deterministic Bayesian parameter estimation algorithm that takes just seconds to produce probability sky maps that are nearly as accurate.
 \end{abstract}

### 2014-09-15 ms.tex
Single column-wide figure

@@
 %
-where $x_i (\omega; \bm\theta)$ is the \ac{GW} signal given a parameter vector $\bm\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's \ac{WSS} Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[\left|n_i(\omega)\right|^2\right] + E\left[\left|n_i(-\omega)\right|^2\right] = 2 E\left[\left|n_i(\omega)\right|^2\right]$.
+where $x_i (\omega; \bm\theta)$ is the \ac{GW} signal given a parameter vector $\bm\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[\left|n_i(\omega)\right|^2\right] + E\left[\left|n_i(-\omega)\right|^2\right] = 2 E\left[\left|n_i(\omega)\right|^2\right]$.
 We shall denote the combined observation from a network of detectors as $\mathbf y (\omega) \equiv \{y_i (\omega)\}_i$.
 
-The likelihood, or the probability of obtaining the observation $\mathbf y$ conditioned on the value of $\bm\theta$, is a product of Gaussian distributions:
+The likelihood, or the probability of the observation $\mathbf y$ conditioned on the value of $\bm\theta$, is a product of Gaussian distributions:
 %
@@
 
-\begin{figure}
+\begin{figure*}
     \begin{center}
         \includegraphics{crlb_tau}
-
         \includegraphics{crlb_gamma}
@@
     \caption{\label{fig:crlb-tau}\ac{CRLB} on root\nobreakdashes-mean square timing uncertainty, using the likelihood for the full GW data (Equation~\ref{eq:gaussian-likelihood-spa}; dashed diagonal line) or the autocorrelation likelihood (Equation~\ref{eq:autocor-likelihood}; solid lines) with a selection of arrival time priors.}
-\end{figure}
+\end{figure*}
 

### 2014-09-16 ms.tex
More changes to abstract

@@
 \begin{abstract}
-Advanced LIGO and Virgo should begin detecting gravitational waves from binary neutron star and neutron star\nobreakdashes--black hole mergers within the next few years.
-These sources are also predicted to power a variety of electromagnetic transients.
-Because these signatures can be fast and faint, locating and observing them hinges on rapidly inferring the sky location from the gravitational wave observations.
+Within the next few years, Advanced LIGO and Virgo should detect gravitational waves from binary neutron star and neutron star\nobreakdashes--black hole mergers.
+These sources are also predicted to power a broad array of electromagnetic transients.
+Because the X\nobreakdashes-ray and optical signatures can be faint and fade rapidly, observing them hinges on rapidly inferring the sky location from the gravitational wave observations.
 MCMC methods for gravitational\nobreakdashes-wave parameter estimation can take hours or more.
-We describe a new, non-MCMC, fully deterministic Bayesian parameter estimation algorithm that takes just seconds to produce probability sky maps that are nearly as accurate.
+We introduce BAYESTAR, a rapid, Bayesian, non-MCMC sky localization algorithm that takes just seconds to produce probability sky maps that are comparable in accuracy to the full analysis.
+Prompt localizations from BAYESTAR will make it possible to search electromagnetic counterparts of compact binary mergers.
 \end{abstract}

### 2014-09-17 ms.tex
Remove unnecessary phrase
Rewrite intro

@@
 
-Ground-based \ac{GW} interferometers are entering the advanced detector era, positioning themselves to make profound discoveries about the universe.
-The \ac{aLIGO} and Vigo detectors will be part of a worldwide network that also includes KAGRA and hopefully LIGO-India.
-Coalescing neutron star binaries are among the most likely sources, with 0.4\nobreakdashes--400 detections per year~\cite{rates}.
-In addition to being an efficient source of \acp{GW}, the \ac{NS} may be tidally shredded before merging, providing fuel for an electromagnetic counterpart.
-The short-lived accretion flow may power a collimated relativistic jet, resulting in a short \ac{GRB} and an X-ray afterglow if it is aligned with the line of sight.
-An optical afterglow may follow minutes to hours later as the jet plows into the interstellar medium.
+The Advanced LIGO and Virgo ground-based \ac{GW} detectors will soon begin taking data \cite{LIGOObservingScenarios} in their `Advanced' configuration with significantly greater sensitivity than their initial runs through 2006.
+They will form the first parts of a sensitive global network to be later augmented by the Japanese KAGRA facility \cite{LCGT,KAGRAInterferometerDesign} and LIGO\nobreakdashes--India \cite{LIGOIndia}.
+Their most likely sources of \acp{GW} include 0.4\nobreakdashes--400 detections binary neutron star mergers per year~\cite{rates}.
+
+These systems are both efficient sources of \acp{GW} and potential sources of detectable electromagnetic transients \citep{MostPromisingEMCounterpart}.
+If one of the binary companions is a \ac{NS}, then it may be tidally shredded before merging.
+If the system's orbital axis is aligned with the line of sight, this may produce a short, hard \acp{GRB}
+\cite{1986ApJ...308L..43P,1989Natur.340..126E,1992ApJ...395L..83N,2011ApJ...732L...6R}.
+An X\nobreakdashes-ray and optical afterglow may follow minutes to hours later as the jet plows into the interstellar medium.
 As surrounding neutron-rich ejecta decay radioactively, an optical `kilonova' may be visible after $\sim$1 day \cite{metzger:2010}.
-Finally, as the ejecta plow into the interstellar medium, a faint radio afterglow could occur on a timescale of weeks to years \cite{Nakar:2011cw}.
+Finally, as the ejecta itself plow into the interstellar medium, a faint radio afterglow could rise and fall on timescales of weeks to years \cite{Nakar:2011cw}.
 
-There is therefore a strong case for performing electromagnetic followups of these sources of \acp{GW}.
-Indeed, the final science run of the intitial LIGO and Virgo instruments saw the first joint search for \acp{GW} from compact binaries and their electromagnetic counterparts \cite{abadie2012first}.
-As we prepare for the next generation of these searches there is a need for determining the location of the source as precisely as possible in as little time as possible.
+The on\nobreakdashes-axis optical afterglows and the recently hypothesized kilonova precursors \cite{KilonovaPrecursor} may be among the most straightforward electromagnetic components to pinpoint with existing and planned optical transient facilities such as the Zwicky Transient Facility~\citep{ZTF} and BlackGEM\footnote{\url{https://www.astro.ru.nl/wiki/research/blackgemarray}}.
+However, both of these signatures fade rapidly, on a timescale of hours.
+Because \ac{GW} detectors have inherently coarse sky position reconstruction ability \cite{FairhurstTriangulation,WenLocalizationAdvancedLIGO,FairhurstLocalizationAdvancedLIGO,2011PhRvD..84j4020V,RodriguezBasicParameterEstimation,NissankeLocalization,NissankeKasliwalEMCounterparts,KasliwalTwoDetectors,Grover:2013,SiderySkyLocalizationComparison,FirstTwoYears}, detecting these transients may depend upon rapidly calculating and disseminating \ac{GW} probability sky maps.
 
-In this paper we present a rapid and accurate method of sky localization that makes use of Bayesian methods.
-It differs from existing techniques in several important ways.
-In the first place, we fix the intrinsic parameters to their \ac{ML} estimates, as provided by the detection pipeline.
-This reduces the dimensionality of the parameter space we need to sample.
-In addition, the technique takes as its input the \ac{ML} estimates of the extrinsic parameters, instead of the $h(t)$ time series.
-This makes the likelihood itself much faster to evaluate.
-Finally, instead of of using \ac{MCMC} or some similar method for statistical sampling, we make use of a deterministic quadrature scheme.
-We call this technique \ac{BAYESTAR}%
+The final science run of the initial LIGO and Virgo instruments saw the first joint search for \ac{GW} and electromagnetic emission from compact binaries.
+This involved several advances in the \ac{GW} data analysis, including the first real\nobreakdashes-time detection and rapid sky localization pipeline \cite{abadie2012first} and the first practical \ac{MCMC} parameter estimation analysis of \ac{GW} signals \cite{S6PE}.
+Despite these technical achievements, there was an undesirable tradeoff between the speed of the rapid localization and the full parameter estimation: the former could analyze a detection candidate in minutes, whereas the latter took half a day; the latter decreased the area on the sky by a factor of 1/20 over the former but took 1000 times as long to run \cite{SiderySkyLocalizationComparison}.
+
+For the purpose of electromagnetic observations, we would like to have a nearly final sky localization within a timescale of minutes to hours.
+To that end, we have devised a rapid and accurate Bayesian sky localization method that takes mere seconds but achieves approximately the same accuracy as the full \ac{MCMC} analysis.
+We call this algorithm \ac{BAYESTAR}%
 %
@@
 .
-It is unique in that it bridges the detection and parameter estimation of \ac{GW} signals, two tasks that have until now involved very different numerical methods and time scales.
-We expect that \ac{BAYESTAR} will take on a key role in observing \ac{CBC} events in both \ac{GW} and optical channels during the Advanced \ac{LIGO} era.
+It differs from existing techniques in several important ways.
+In the first place, we treat the matched-filter detection pipeline as a measurement system in and of itself, treating the point\nobreakdashes-parameter estimates that it provides as the experimental input data rather than the full \ac{GW} time series that is used by the \ac{MCMC} analysis.
+This drastically reduces the dimensionality of both the data and the signal model.
+It also permits us to avoid directly computing the expensive post-Newtonian model waveforms, making the likelihood itself much faster to evaluate.
+Finally, instead of of using \ac{MCMC} or some similar method for statistical sampling, we make use of a deterministic quadrature scheme.
+This algorithm is unique in that it bridges the detection and parameter estimation of \ac{GW} signals, two tasks that have until now involved very different numerical methods and time scales.
+We expect that \ac{BAYESTAR} will take on a key role in observing \ac{CBC} events in both \ac{GW} and optical channels during the Advanced LIGO era.
 
@@
 This list of parameters involves some simplifying assumptions.
-Eccentricity is omitted: although it does play a major role in the evolution and waveforms of \ac{NSBH} and \ac{BBH} sources formed by dynamical capture~\cite{PhysRevD.87.043004}, \ac{BNS} systems formed by binary stellar evolution should circularize due to tidal interaction~\cite{0004-637X-572-1-407} and later \ac{GW} emission~\cite{PhysRev.136.B1224} long before the inspiral enters \ac{LIGO}'s frequency range of $\sim$10\nobreakdashes--1000~kHz.
+Eccentricity is omitted: although it does play a major role in the evolution and waveforms of \ac{NSBH} and \ac{BBH} sources formed by dynamical capture~\cite{PhysRevD.87.043004}, \ac{BNS} systems formed by binary stellar evolution should circularize due to tidal interaction~\cite{0004-637X-572-1-407} and later \ac{GW} emission~\cite{PhysRev.136.B1224} long before the inspiral enters LIGO's frequency range of $\sim$10\nobreakdashes--1000~kHz.
 Tidal deformabilities of the \acp{NS} are omitted because the signal imprinted by the companions' material properties is so small that it will only be detectable by an Einstein Telescope\nobreakdashes-class \ac{GW} observatory~\cite{PhysRevD.81.123016}.
@@
 
-The choice of prior is open to one's astrophysical preconceptions, but during \ac{S6} when \ac{LIGO}'s Bayesian \ac{CBC} parameter estimation pipelines were pioneered, the prior was taken to be isotropic in sky location and binary orientation, and uniform in volume, arrival time, and the component masses~\cite{2013arXiv1304.1775T}.
+The choice of prior is open to one's astrophysical preconceptions, but during \ac{S6} when LIGO's Bayesian \ac{CBC} parameter estimation pipelines were pioneered, the prior was taken to be isotropic in sky location and binary orientation, and uniform in volume, arrival time, and the component masses~\cite{2013arXiv1304.1775T}.
 
@@
 With the most sophisticated \ac{CBC} parameter estimation codes, it still takes days to process a single event.
-This delay is undesirable for planning targeted \ac{EM} follow\nobreakdashes-up searches of \ac{LIGO} events.
+This delay is undesirable for planning targeted \ac{EM} follow\nobreakdashes-up searches of LIGO events.
 
@@
 \begin{acknowledgements}
-\marginpar{This URL is not specific enough.}Source code for \ac{BAYESTAR} is available on the \ac{LIGO} \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/lalsuite.html}.
+\marginpar{This URL is not specific enough.}Source code for \ac{BAYESTAR} is available on the LIGO \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/lalsuite.html}.
 
@@
 
-\ac{LIGO} was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the \ac{NSF} and operates under cooperative agreement PHY\nobreakdashes-0107417.
+LIGO was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the \ac{NSF} and operates under cooperative agreement PHY\nobreakdashes-0107417.
 Some results were produced on the NEMO computing cluster operated by the Center for Gravitation and Cosmology at University of Wisconsin\nobreakdashes--Milwaukee under \ac{NSF} Grants PHY\nobreakdashes-0923409 and PHY\nobreakdashes-0600953.
 This research is supported by the \ac{NSF} through a Graduate Research Fellowship to L.S.
-This paper has \ac{LIGO} Document Number \ac{LIGO}\nobreakdashes-PXXXXXXX\nobreakdashes-vX.
+This paper has LIGO Document Number LIGO\nobreakdashes-PXXXXXXX\nobreakdashes-vX.
 \end{acknowledgements}

### 2014-09-17 ms.tex
marginpar to footnote
Rename section
Remove marginpar

@@
 
-\section{Theory of \ac{GW} detection and parameter estimation}
+\section{\ac{GW} detection and parameter estimation}
 
@@
 There are various ways to characterize the significance of a detection candidate, but in Gaussian noise the combined SNR $\hat\rho_\mathrm{c}$ suffices,
-%
-\marginpar{What is the least that I can get away with saying about false alarm rate estimation and likelihood ratios?}%
 %
@@
 %
-\marginpar{The Fish(er) factor. Go home, Leo.}%
+\footnote{The Fish(er) factor.}%
 %
@@
 \begin{acknowledgements}
-\marginpar{This URL is not specific enough.}Source code for \ac{BAYESTAR} is available on the LIGO \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/lalsuite.html}.
+Source code for \ac{BAYESTAR} is available as part of LALInference, the open source LIGO/Virgo parameter estimation toolchain.
+See \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/lalsuite.html} for further details.
 

### 2015-04-15 ms.tex
Use acronyms in environment
Add stuff from thesis

--- 2014-09-16 ms.tex
@@
 \documentclass[amsmath,amssymb,aps,prx,reprint,nopreprintnumbers,nofootinbib]{revtex4-1}
+\usepackage[nolist]{acronym}
 \usepackage{multirow}
@@
 \usepackage{protosem}
-\input{ligo-acronyms/acronyms}
 
@@
 
+\DeclareMathOperator{\Tr}{Tr}
+\DeclareMathOperator{\erf}{erf}
 \DeclareMathOperator{\cov}{cov}
@@
 
+\newcommand\transpose{\ensuremath{^{^\mathsf{T}}}}
+
 \begin{document}
@@
 \begin{abstract}
-Within the next few years, Advanced LIGO and Virgo should detect gravitational waves from binary neutron star and neutron star\nobreakdashes--black hole mergers.
+Within the next few years, Advanced LIGO and Virgo should detect \acp{GW} from binary neutron star and neutron star\nobreakdashes--black hole mergers.
 These sources are also predicted to power a broad array of electromagnetic transients.
@@
 
-Ground-based \ac{GW} interferometers are entering the advanced detector era, positioning themselves to make profound discoveries about the universe.
-The \ac{aLIGO} and Vigo detectors will be part of a worldwide network that also includes KAGRA and hopefully LIGO-India.
-Coalescing neutron star binaries are among the most likely sources, with 0.4\nobreakdashes--400 detections per year~\cite{rates}.
-In addition to being an efficient source of \acp{GW}, the \ac{NS} may be tidally shredded before merging, providing fuel for an electromagnetic counterpart.
-The short-lived accretion flow may power a collimated relativistic jet, resulting in a short \ac{GRB} and an X-ray afterglow if it is aligned with the line of sight.
-An optical afterglow may follow minutes to hours later as the jet plows into the interstellar medium.
-As surrounding neutron-rich ejecta decay radioactively, an optical `kilonova' may be visible after $\sim$1 day \cite{metzger:2010}.
-Finally, as the ejecta plow into the interstellar medium, a faint radio afterglow could occur on a timescale of weeks to years \cite{Nakar:2011cw}.
-
-There is therefore a strong case for performing electromagnetic followups of these sources of \acp{GW}.
-Indeed, the final science run of the intitial LIGO and Virgo instruments saw the first joint search for \acp{GW} from compact binaries and their electromagnetic counterparts \cite{abadie2012first}.
-As we prepare for the next generation of these searches there is a need for determining the location of the source as precisely as possible in as little time as possible.
-
-In this paper we present a rapid and accurate method of sky localization that makes use of Bayesian methods.
-It differs from existing techniques in several important ways.
-In the first place, we fix the intrinsic parameters to their \ac{ML} estimates, as provided by the detection pipeline.
-This reduces the dimensionality of the parameter space we need to sample.
-In addition, the technique takes as its input the \ac{ML} estimates of the extrinsic parameters, instead of the $h(t)$ time series.
-This makes the likelihood itself much faster to evaluate.
-Finally, instead of of using \ac{MCMC} or some similar method for statistical sampling, we make use of a deterministic quadrature scheme.
-We call this technique \ac{BAYESTAR}%
-%
-\footnote{A pun on the Cylon battleships in the American television series Battlestar Galactica.
+The \acf{LIGO} and Virgo will soon begin taking data \cite{LIGOObservingScenarios} in their `Advanced' configurations, ultimately attaining an order of magnitude more sensitivity than in their initial configurations through 2010.
+They will form the first parts of a sensitive global \ac{GW} detector network, to be later augmented by the Japanese KAGRA facility \cite{LCGT,KAGRAInterferometerDesign} and \ac{LIGO}\nobreakdashes--India \cite{LIGOIndia}.
+
+The most readily detectable sources of \acp{GW} include binary neutron star mergers, with 0.4\nobreakdashes--400 events per year within the reach of Advanced LIGO at final design sensitivity~\cite{rates}.
+These binary systems are not only efficient sources of \acp{GW}, but also potential sources of detectable \ac{EM} transients from the aftermath of the tidal disruption of the \acp{NS}.
+\citet{MostPromisingEMCounterpart} argue that the most promising \ac{EM} counterparts are the hypothesized optical/infrared ``kilonovae'' powered by the radioactivate decay of $r$\nobreakdashes-process elements synthesized within the neutron\nobreakdashes-rich ejecta.
+These are expected to be faint, red, and peak rapidly, reaching only $R \sim 24$~mag within a week, though rising several magnitudes brighter in the infrared \citep{KilonovaHighOpacities}.
+
+Several mechanisms could make the kilonovae brighter, bluer, and hence more readily detectable \citep{KilonovaPrecursor,KilonovaRedOrBlue}, but peak even earlier, within hours.
+If, as is widely believed \citep{1986ApJ...308L..43P,1989Natur.340..126E,1992ApJ...395L..83N,2011ApJ...732L...6R}, \ac{BNS} mergers are indeed progenitors of short \acp{GRB}, then a small (due to jet collimation) fraction of Advanced \ac{LIGO} events could also be accompanied by a bright optical afterglow, but this signature, likewise, would peak within hours or earlier.
+
+Adding to the challenge of detecting a faint, short-lived optical transient, \ac{GW} detector networks have inherently coarse position reconstruction accurate to tens to hundreds of deg$^2$ \citep{FairhurstTriangulation,WenLocalizationAdvancedLIGO,FairhurstLocalizationAdvancedLIGO,2011PhRvD..84j4020V,RodriguezBasicParameterEstimation,NissankeLocalization,NissankeKasliwalEMCounterparts,KasliwalTwoDetectors,Grover:2013,SiderySkyLocalizationComparison,FirstTwoYears}.
+Wide\nobreakdashes-field optical transient surveys such as the \acf{ZTF}, BlackGEM, or the \acf{LSST}, operated in ``target of opportunity'' mode, may be able to tile these large areas rapidly enough to find the one needle in the haystack that is connected to the \ac{GW} event, but clearly time will be of the utmost.
+
+The final science run of the initial LIGO and Virgo instruments saw the first joint search for \ac{GW} and electromagnetic emission from compact binaries.
+This involved several advances in the \ac{GW} data analysis \citep{CBCLowLatency}, including the first real\nobreakdashes-time matched\nobreakdashes-filter detection pipeline (\acs{MBTA}; \aclu{MBTA}), a semi\nobreakdashes-coherent, ad hoc rapid triangulation code (Timing++), and the first version of a rigorous Bayesian \ac{MCMC} parameter estimation code (LALINFERENCE; \citealt{S6PE})---all in service of the first search for X\nobreakdashes-ray \citep{SwiftFollowup} and optical \citep{OpticalImageAnalysis} counterparts of \ac{GW} triggers, by a consortium of facilities.
+Despite the technical achievements in the \ac{GW} data analysis, there was an undesirable tradeoff between the speed of the rapid localization and the full parameter estimation: the former could analyze a detection candidate in minutes, whereas the latter took half a day; the latter decreased the area on the sky by a factor of 1/20 over the former but took 1000 times as long to run \cite{SiderySkyLocalizationComparison}.
+
+The success of \ac{EM} follow\nobreakdashes-up of \ac{LIGO} events will depend critically on disseminating high quality sky localizations within a timescale of minutes to hours.
+To that end, we have devised a rapid and accurate Bayesian sky localization method that takes mere seconds to achieve approximately the same accuracy as the full \ac{MCMC} analysis.
+Our key insights are the following:
+%
+\begin{enumerate}
+    \item Nearly all of the information in the \ac{GW} time series that is informative for sky localization is encapsulated within the matched\nobreakdashes-filter estimates of the times, amplitudes, and phases on arrival at the detectors.
+To infer the position and distance of a \ac{GW} event, we only have to consider three numbers per detector rather than a densely sampled strain time series per detector.
+    \item The matched\nobreakdashes-filter pipeline can be treated as a measurement system in and of itself.
+Just like the strain time series from the detectors, the resultant times, amplitudes, and phases have a predictable and quantifiable measurement uncertainty that can be translated into a likelihood function suitable for Bayesian inference.
+    \item Errors in sky localization are semi\nobreakdashes-independent from errors in masses.
+If we care only about position reconstruction and not about jointly estimating masses as well, then we can reduce the dimensionality of the parameter estimation problem significantly.
+Moreover, this frees us of the need to directly compute the expensive post-Newtonian model waveforms, making the likelihood itself much faster to evaluate.
+    \item Thanks to a simple likelihood function and a well\nobreakdashes-characterized parameter space, we may dispense with costly and parallelization\nobreakdashes-resistant \ac{MCMC} integration, and instead perform the Bayesian marginalization with classic, deterministic, very low order Gaussian quadrature.
+\end{enumerate}
+%
+We call the resulting algorithm \ac{BAYESTAR}\footnote{A pun on the Cylon battleships in the American television series Battlestar Galactica.
 The defining characteristic of the Cylons is that they repeatedly defeat humanity by using their superhuman information\nobreakdashes-gathering ability to coordinate overwhelming forces.
-The name also suggests that, like the Cylons, \ac{GW} detectors may some day rise against us humans.}%
-%
-\footnote{We do not like to mention the final `L' in the acronym, because then it would be pronounced BAYESTARL, which sounds stupid.}%
-.
+The name also suggests that, like the Cylons, \ac{GW} detectors may some day rise against us humans.}\footnote{We do not like to mention the final `L' in the acronym, because then it would be pronounced BAYESTARL, which sounds stupid.}.
+It is as fast as Timing++, but nearly as accurate as the rigorous full parameter estimation.
 It is unique in that it bridges the detection and parameter estimation of \ac{GW} signals, two tasks that have until now involved very different numerical methods and time scales.
-We expect that \ac{BAYESTAR} will take on a key role in observing \ac{CBC} events in both \ac{GW} and optical channels during the Advanced \ac{LIGO} era.
-
-\section{Theory of \ac{GW} detection and parameter estimation}
-
-In the frequency domain, the strain observed by a single \ac{GW} interferometer is
+Beginning with the first Advanced \acs{LIGO} observing run, \ac{BAYESTAR} will provide localizations within minutes of the detection of any \ac{BNS} merger candidate, playing a a key role in enabling rapid follow-up observations.
+
+This paper is organized as follows.
+In Section~\ref{sec:detection}, we describe the \ac{GW} signal model and sketch the standard detection algorithm, the matched filter bank.
+In Section~\ref{sec:parameter-estimation}, we describe Bayesian inference formalism and the prevailing method for inferring the parameters of detected candidates, \ac{MCMC} sampling.
+In Section~\ref{sec:bayestar-likelihood}, we propose the \ac{BAYESTAR} likelihood as a model for the uncertainty in the matched\nobreakdashes-filter parameter estimates, and discuss its relationship to and consistency with the likelihood for the full \ac{GW} data.
+In Section~\ref{sec:prior}, we describe the input to \ac{BAYESTAR} supplied by the detection pipeline, and the prior distribution on parameters.
+In Section~\ref{sec:marginal-posterior}, we explain the integration scheme by which the posterior probability distribution is calculated for a given sky location.
+In Section~\ref{sec:adaptive-sampling}, we show a scheme whereby the sky posterior is sampled on an adaptive HEALPix grid.
+In Section~\ref{sec:run-time}, we report the running time of the algorithm on the hardware available on the \ac{LIGO} Data Grid.
+Finally, in Section~\ref{sec:case-study}, we quantify the sky localization performance on a comprehensive set of simulated \ac{GW} events.
+
+% Establish and describe the first rapid and tractable solution of the GW position reconstruction problem.
+% Use and abuse of Fisher matrix: use it as a clue.
+% Generality: replace GW antenna pattern and GW polarization model with any other kind of receiver.
+% Demonstration of a practical Bayesian analysis that does not use an MCMC method.
+
+\section{Signal model and detection}
+\label{sec:detection}
+
+In the \ac{TD}, the strain observed by a single \ac{GW} interferometer is
+%
+\begin{equation}
+    y_i(t) = x_i (t; \bm\theta) + n_i (t).
+\end{equation}
+%
+In the \ac{FD},
 %
 \begin{equation}\label{eq:signal-model}
-    y_i (\omega) = x_i (\omega; \bm\theta) + n_i (\omega),
-\end{equation}
-%
-where $x_i (\omega; \bm\theta)$ is the \ac{GW} signal given a parameter vector $\bm\theta$ that describes the \ac{GW} source, and $n_i (\omega)$ is that detector's Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[\left|n_i(\omega)\right|^2\right] + E\left[\left|n_i(-\omega)\right|^2\right] = 2 E\left[\left|n_i(\omega)\right|^2\right]$.
-We shall denote the combined observation from a network of detectors as $\mathbf y (\omega) \equiv \{y_i (\omega)\}_i$.
-
-The likelihood, or the probability of the observation $\mathbf y$ conditioned on the value of $\bm\theta$, is a product of Gaussian distributions:
+    Y_i (\omega) = \int_{-\infty}^\infty y(t) e^{-i \omega t} dt = X_i (\omega; \bm\theta) + N_i (\omega),
+\end{equation}
+%
+where $X_i (\omega; \bm\theta)$ is the \ac{GW} signal given a parameter vector $\bm\theta$ that describes the \ac{GW} source, and $N_i (\omega)$ is that detector's Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[\left|N_i(\omega)\right|^2\right] + E\left[\left|N_i(-\omega)\right|^2\right] = 2 E\left[\left|N_i(\omega)\right|^2\right]$.
+We shall denote the combined observation from a network of detectors as $\mathbf Y (\omega) \equiv \{Y_i (\omega)\}_i$.
+
+Under the assumptions that the detector noise is Gaussian and that the noise from different detectors are uncorrelated, the likelihood of the observation, $\mathbf y$, conditioned upon the parameters $\bm\theta$, is a product of Gaussian distributions:
 %
 \begin{multline}\label{eq:gaussian-likelihood}
-    \mathcal{L}(\mathbf y; \bm\theta) = \prod_i p(y_i | \bm\theta)
-        \\ \propto \exp \left[
-        - \frac{1}{2} \sum_i \int_0^\infty \frac{\left|y_i (\omega)
-            - x_i(\omega; \bm\theta) \right|^2}{S_i(\omega)} \, d\omega
+    \mathcal{L}(\mathbf Y; \bm\theta) = \prod_i p(Y_i | \bm\theta)
+        \\
+        \propto \exp \left[
+        - \frac{1}{2} \sum_i \int_0^\infty \frac{\left|Y_i (\omega)
+            - X_i(\omega; \bm\theta) \right|^2}{S_i(\omega)} \, d\omega
     \right].
@@
 
-A \ac{CBC} source is specified by a vector of extrinsic parameters describing its position and orientation and intrinsic parameters describing the physical properties of the binary components:
-%
-\begin{equation}
+A \ac{CBC} source is specified by a vector of extrinsic parameters describing its position and orientation, and intrinsic parameters describing the physical properties of the binary components\footnote{This list of parameters involves some simplifying assumptions.
+Eccentricity is omitted: although it does play a major role in the evolution and waveforms of \ac{NSBH} and \ac{BBH} sources formed by dynamical capture~\citep{PhysRevD.87.043004}, \ac{BNS} systems formed by binary stellar evolution should almost always circularize due to tidal interaction~\citep{0004-637X-572-1-407} and later \ac{GW} emission~\citep{PhysRev.136.B1224} long before the inspiral enters LIGO's frequency range of $\sim$10\nobreakdashes--1000~kHz.
+Tidal deformabilities of the \acp{NS} are omitted because the signal imprinted by the companions' material properties is so small that it will only be detectable by an Einstein Telescope\nobreakdashes-class \ac{GW} observatory~\citep{PhysRevD.81.123016}.
+Furthermore, in \ac{GW} detection efforts, especially those focused on \ac{BNS} systems, the component spins $\mathbf{S}_1$ and $\mathbf{S}_2$ are often assumed to be nonprecessing and aligned with the system's total angular momentum and condensed to a single scalar parameter $\chi$, or even neglected entirely: $\mathbf{S}_1 = \mathbf{S}_2 = 0$.}:
+%
+\begin{equation}\label{eq:params}
     \bm\theta = \left[
@@
 \end{equation}
-%
-This list of parameters involves some simplifying assumptions.
-Eccentricity is omitted: although it does play a major role in the evolution and waveforms of \ac{NSBH} and \ac{BBH} sources formed by dynamical capture~\cite{PhysRevD.87.043004}, \ac{BNS} systems formed by binary stellar evolution should circularize due to tidal interaction~\cite{0004-637X-572-1-407} and later \ac{GW} emission~\cite{PhysRev.136.B1224} long before the inspiral enters \ac{LIGO}'s frequency range of $\sim$10\nobreakdashes--1000~kHz.
-Tidal deformabilities of the \acp{NS} are omitted because the signal imprinted by the companions' material properties is so small that it will only be detectable by an Einstein Telescope\nobreakdashes-class \ac{GW} observatory~\cite{PhysRevD.81.123016}.
-Furthermore, in \ac{GW} detection efforts, especially those focused on \ac{BNS} systems, the component spins $\mathbf{S}_1$ and $\mathbf{S}_2$ are often assumed to be nonprecessing and aligned with the system's total angular momentum and condensed to a single scalar parameter $\chi$, or even neglected entirely, $\mathbf{S}_1 = \mathbf{S}_2 = 0$.
-
-Assuming circular orbits and no spin precession, we can write the \ac{GW} signal in each detector as a linear combination of two basis waveforms $h = h_0$ and $h_{\pi/2}$.
-For nonprecessing systems, $h_0$ and $h_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the sine and cosine functions, being nearly orthogonal and out of phase by ${\pi/2}$ at all frequencies.
-If $h_0$ and $h_{\pi/2}$ are Fourier transforms of real functions, then $h_0(\omega) = h_0^*(-\omega)$ and $h_{\pi/2}(\omega) = h_{\pi/2}^*(-\omega)$, and we can write (assuming an arbitrary phase convention)
-%
-\begin{equation}
-    h_{\pi/2}(\omega) = h_0(\omega) \cdot
+
+Assuming a nonprecessing circular orbit, we can write the \ac{GW} signal received by any detector as a linear combination of two basis waveforms, $H_0$ and $H_{\pi/2}$ \citep{PhysRevD.83.084002}.
+$H_0$ and $H_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the sine and cosine functions, being nearly orthogonal and out of phase by ${\pi/2}$ at all frequencies.
+If $H_0$ and $H_{\pi/2}$ are Fourier transforms of real functions, then $H_0(\omega) = H_0^*(-\omega)$ and $H_{\pi/2}(\omega) = H_{\pi/2}^*(-\omega)$, and we can write (assuming an arbitrary phase convention)
+%
+\begin{equation}
+    H_{\pi/2}(\omega) = H_0(\omega) \cdot
     \begin{cases}
@@
 %
-Then, we can write the signal model in a way that isolates all dependence on the extrinsic parameters $\bm\theta_\mathrm{ex}$ into the coefficients and all dependence on the intrinsic parameters $\bm\theta_\mathrm{in}$ into the basis waveform, by taking the Fourier transform of Equation~(2.8) of \cite{PhysRevD.83.084002}:
+For brevity, we define $H \equiv H_{0}$ and write all subsequent equations in terms of the $H$ basis vector alone.
+Then, we can write the signal model in a way that isolates all dependence on the extrinsic parameters, $\bm\theta_\mathrm{ex}$, into a few coefficients, and all dependence on the intrinsic parameters, $\bm\theta_\mathrm{in}$, into the basis waveform, by taking the Fourier transform of Equation~(2.8) of \cite{PhysRevD.83.084002}:
 %
 \begin{multline}\label{eq:full-signal-model}
-    x_i(\omega; \bm\theta) = e^{-i \omega t_\oplus}
+    X_i(\omega; \bm\theta) = e^{-i \omega (t_\oplus - \mathbf{d}_i \cdot \mathbf{n})}
     \frac{r_{1,i}}{r}
-    e^{2 i \phi_c} \\
+    e^{2 i \phi_c}
+    \\
     \left[
@@
     \right]
-    h(\omega; \bm\theta_\mathrm{in})
+    H(\omega; \bm\theta_\mathrm{in})
 \end{multline}
@@
 %
-The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, defined such that $0 \leq {F_{+,i}}^2 + {F_{\times,i}} \leq 1$.
-They depend on the orientation of detector $i$ as well as the sky location and sidereal time of the event, and are derived in~\cite{PhysRevD.63.042003}.
-The quantity $r_{1,i}$ is a fiducial distance at which detector $i$ would register SNR=1 for an optimally oriented binary (face\nobreakdashes-on, and in a direction perpendicular to the interferometer's arms),
-%
-\marginpar{I seem to have dropped a factor of 2 here...}%
+The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, defined such that $0 \leq {F_{+,i}}^2 + {F_{\times,i}}^2 \leq 1$.
+They depend on the orientation of detector $i$ as well as the sky location and sidereal time of the event and are presented in~\cite{ExcessPower}.
+In a coordinate system with the $x$ and $y$ axes aligned with the arms of a detector, the antenna pattern is given in spherical polar coordinates as
+%
+\begin{align}
+    F_+ &= -\frac{1}{2}(1 + \cos^2 \theta) \cos{2\phi}, \\
+    F_\times &= -\cos\theta \sin{2\phi}.
+\end{align}
+%
+The unit vector $\mathbf{d}_i$ represents the position of detector $i$ in units of light travel time.\footnote{When considering transient \ac{GW} sources such as those that we are concerned with in this thesis, the origin of the coordinate system is usually taken to be the geocenter.
+For long\nobreakdashes-duration signals such as those from statically deformed neutron stars, the solar system barycenter is a more natural choice.} The vector $\mathbf{n}$ is the direction of the source.
+The negative sign in the dot product $-\mathbf{d}_i \cdot \mathbf{n}$ is present because the direction of travel of the \ac{GW} signal is opposite to that of its sky location.
+The quantity $r_{1,i}$ is a fiducial distance at which detector $i$ would register SNR=1 for an optimally oriented binary (face\nobreakdashes-on, and in a direction perpendicular to the interferometer's arms):
 %
 \begin{equation}\label{eq:horizon}
-r_{1,i} = 1 / \sigma_i, \qquad {\sigma_i}^2 = \int_0^\infty \frac{\left|h(\omega; \mathbf \theta)\right|^2}{S_i(\omega)} \,d\omega.
-\end{equation}
-
-More succinctly, we can write the signal received by detector $i$ in terms of observable extrinsic parameters $\bm\theta_i = (\rho_i, \gamma_i, \tau_i)$, the amplitude $\rho_i$, phase $\gamma_i$, and time delay $\tau_i$ on arrival at detector $i$,
-%
-\begin{equation}\label{eq:signal-model}
-    x_i (\omega; \bm\theta_i, \bm\theta_\mathrm{in}) = x_i (\omega; \rho_i, \gamma_i, \tau_i, \bm\theta_\mathrm{in}) = \frac{\rho_i}{\sigma_i} e^{i (\gamma_i - \omega \tau_i)} h(\omega; \bm\theta_\mathrm{in}).
-\end{equation}
+r_{1,i} = 1 / \sigma_i, \qquad {\sigma_i}^2 = \int_0^\infty \frac{\left|H(\omega; \boldsymbol \theta_\mathrm{in})\right|^2}{S_i(\omega)} \,d\omega.
+\end{equation}
+
+More succinctly, we can write the signal received by detector $i$ in terms of observable extrinsic parameters $\bm\theta_i = (\rho_i, \gamma_i, \tau_i)$, the amplitude $\rho_i$, phase $\gamma_i$, and time delay $\tau_i$ on arrival at detector $i$:
+%
+\begin{multline}\label{eq:signal-model}
+    X_i (\omega; \bm\theta_i, \bm\theta_\mathrm{in})
+    \\
+    = X_i (\omega; \rho_i, \gamma_i, \tau_i, \bm\theta_\mathrm{in}) = \frac{\rho_i}{\sigma_i} e^{i (\gamma_i - \omega \tau_i)} H(\omega; \bm\theta_\mathrm{in}).
+\end{multline}
 
 The prevailing technique for detection of \acp{GW} from \acp{CBC} is to realize a \ac{MLE} from the likelihood in Equation~(\ref{eq:gaussian-likelihood}) and the signal model in Equation~(\ref{eq:signal-model}).
-Concretely, this results in a bank of matched filters, or the cross-correlation between a template waveform and the incoming data stream,
-%
-\begin{equation}
-z_i(\tau_i;\bm\theta_\mathrm{in}) = \frac{1}{\sigma_i (\bm\theta_\mathrm{in})} \int_0^\infty \frac{h^*(\omega; \bm\theta_\mathrm{in}) y_i(\omega) e^{i \omega \tau_i}}{S_i(\omega)} \,d\omega.
+Concretely, this results in a bank of matched filters, or cross-correlations between the incoming data stream and a collection of template waveforms,
+%
+\begin{equation}
+z_i(\tau_i;\bm\theta_\mathrm{in}) = \frac{1}{\sigma_i (\bm\theta_\mathrm{in})} \int_0^\infty \frac{H^*(\omega; \bm\theta_\mathrm{in}) Y_i(\omega) e^{i \omega \tau_i}}{S_i(\omega)} \,d\omega.
 \end{equation}
@@
 \begin{eqnarray}
+    \label{eq:optimal-tau}
     \hat{\bm\theta}_\mathrm{in}, \{\hat\tau_i\}_i
@@
         \bm\theta_\mathrm{in}\right) \right|^2, \\
+    \label{eq:optimal-rho}
     \hat\rho_i &=& \left| z_i\left(\hat\tau_i;
         \hat{\bm\theta}_\mathrm{in}\right) \right|, \\
+    \label{eq:optimal-gamma}
     \hat\gamma_i &=& \arg z_i\left(\hat\tau_i;
@@
 A detection candidate consists of $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$.
-There are various ways to characterize the significance of a detection candidate, but in Gaussian noise the combined SNR $\hat\rho_\mathrm{c}$ suffices,
-%
-\marginpar{What is the least that I can get away with saying about false alarm rate estimation and likelihood ratios?}%
-%
-\begin{equation}
-    \hat\rho_\mathrm{c} = \sqrt{\sum_i {\hat\rho_i}^2}.
-\end{equation}
-
-\subsection{Bayesian probability and parameter estimation}
-
-In the Bayesian framework, the parameters are inferred from the data by forming the posterior distribution, $p(\bm\theta|\mathbf y)$, which describes the probability of the parameters given the observations.
-Bayes' rule relates the likelihood $p(\mathbf y|\bm\theta)$ to the posterior $p(\bm\theta|\mathbf y)$,
-%
-\begin{equation}\label{bayes}
-p(\bm\theta|\mathbf y) = \frac{p(\mathbf y|\bm\theta) p(\bm\theta)}{p(\mathbf y)},
-\end{equation}
-%
-introducing the prior distribution $p(\bm\theta)$ which encapsulates previous information about the paramters (for example, arising from earlier observations or from known physical bounds on the parameters), and the evidence $p(\mathbf y)$ which can be thought of as a normalization factor or as describing the parsimoniousness of the model.
-
-The choice of prior is open to one's astrophysical preconceptions, but during \ac{S6} when \ac{LIGO}'s Bayesian \ac{CBC} parameter estimation pipelines were pioneered, the prior was taken to be isotropic in sky location and binary orientation, and uniform in volume, arrival time, and the component masses~\cite{2013arXiv1304.1775T}.
-
-In Bayesian inference, although it is often easy to write down the likelihood or even the full posterior in closed form, usually one is interested in only a subset $\bm\beta$ of all of the model's parameters, the others $\bm\lambda$ being nuisance parameters.
-In this case, we integrate away the nuisance parameters, forming the marginal posterior
-%
-\begin{equation}\label{eq:marginal-posterior}
-    p(\bm\beta|\mathbf y) = \int \frac{p(\mathbf y|\bm\beta,\bm\lambda) p(\bm\beta,\bm\lambda)}{p(\mathbf y)} \,d\bm\lambda
-\end{equation}
-%
-with $\bm\theta = (\bm\beta, \bm\lambda)$. For instance, for the purpose of locating a \ac{GW} source on the sky, all parameters but $(\alpha, \delta)$ are nuisance parameters.
-
-Bayesian parameter estimation has many advantages, including broad generality and the ability to make probabilistically meaningful statements even with very low \ac{SNR} measurements.
-However, in problems of even modest complexity, the marginalization step involves many-dimensional, ill-behaved integrals.
-The powerful \ac{MCMC} integration technique has become almost synonymous with Bayesian inference.
-Though powerful, \ac{MCMC} is inherently non-deterministic and resistant to parallelization, as well as (at least historically) slow.
-With the most sophisticated \ac{CBC} parameter estimation codes, it still takes days to process a single event.
-This delay is undesirable for planning targeted \ac{EM} follow\nobreakdashes-up searches of \ac{LIGO} events.
-
-In what follows, we propose a complementary rapid parameter estimation scheme that can produce reliable positions estimates within minutes of a detection.
-We can even use our scheme to speed up the full \ac{MCMC} analysis and make the refined parameter estimates available more quickly.
-The key difference is that we start not from the \ac{GW} signal itself, but from the point parameter estimates from the detection.
-By harnessing the detection pipeline in this manner, we arrive at a simpler Bayesian problem that is amenable to straightforward, deterministic, numerical quadrature.
-
-\section{Measurement uncertainty}
-
-For the purpose of rapid sky localization, we assume that we do not have access to the \ac{GW} data $\mathbf{y}$ itself, and that our only contact with it is through the ML parameter estimates $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$.
-Although this is a significant departure from conventional GW parameter estimation techniques, we can still apply the full Bayesian machinery of Equation~(\ref{eq:marginal-posterior}) to compute a posterior distribution for the sky location.
-
-\marginpar{This paragraph might make more sense in a conclusion.}%
-%
-There are many practical advantages of doing so.
-For one, there are difficulties in synchronously gathering together the calibrated GW strain data, auxiliary instrument channels, and data quality vetoes from all of the sites.
-The data consumed by the real\nobreakdashes-time detection pipeline are not necessarily final.
-Longer\nobreakdashes-running follow\nobreakdashes-up analyses can benefit from offline calibration, whereas the rapid sky localization need not re\nobreakdashes-analyze the online data.
-Moreover, the dimensionality of the problem is greatly reduced, and the problem becomes computationally easier.
-
-\subsection{Fisher matrix: single detector}
-
-To treat the ML estimates as measurements in and of themselves, we need to work out their likelihood, marginalized over the GW signal $\mathbf{y}$,
-%
-\begin{equation}\label{eq:detection-candidate-likelihood}
-    p\left(\{\hat{\bm\theta}_i\}_i,
-        \hat{\bm\theta}_\mathrm{in}
-    \middle| \bm\theta\right)
-    \propto \int\limits_{\mathclap{\mathbf{y} | \mathrm{MLE}(\mathbf{y}) =
-        \{\{\hat{\bm\theta}_i\}_i,
-        \hat{\bm\theta}_\mathrm{in}\}}}
-    p(\mathbf{y} | \bm\theta) \, p(\bm\theta)
-    \, d\mathbf{y}
-\end{equation}
-%
-The integral above could be written in terms of the matched filter signals $\left\{z_i\right\}_i$, such that $z_i\left(\hat\tau_i, \hat{\bm\theta}_\mathrm{in}\right) = \hat\rho_i e^{i \hat\gamma_i}$ and $\forall t$, $\left|z_i(t, \bm\theta_\mathrm{in})\right| \leq \hat\rho_i$.
-The combination of the nontrivial integration limits and the Fourier transform in the matched filter seem like a problem for direct evaluation.
-
-However, even without directly evaluating the distribution of the \ac{ML} estimator, we know many of its properties.
+There are various ways to characterize the significance of a detection candidate.
+In Gaussian noise, the maximum likelihood for the network is obtained by maximizing the network \ac{SNR}, $\rho_\mathrm{net}$,
+%
+\begin{equation}
+    \hat\rho_\mathrm{net} = \max_{\bm\theta} \sum_i {|z_i({\bm\theta})|}^2 = \sqrt{\sum_i {\hat\rho_i}^2};
+\end{equation}
+%
+this, therefore, is the simplest useful candidate ranking statistic.
+
+\section{Uncertainty and the Fisher matrix}
+\label{sec:fisher}
+
+We can predict the uncertainty in the \ac{ML} estimates without working out its full distribution.
 The \ac{CRLB} gives its covariance in the asymptotic limit of high \ac{SNR}.
-The \ac{CRLB} has been widely applied in \ac{GW} data analysis \cite{fairhurst:2009}.
+The \ac{CRLB} has been widely applied in \ac{GW} data analysis to estimate parameter estimation uncertainty (for example, \cite{1996PhRvD..53.3033B,FairhurstTriangulation,2009PhRvD..79h4032A,WenLocalizationAdvancedLIGO,LIGOObservingScenarios,FairhurstLIGOIndia})\footnote{The Fisher matrix is also used in construction of \ac{CBC} matched filter banks.
+The common procedure is to place templates uniformly according to the determinant of the signal space metric, which is the Fisher matrix.
+This is equivalent to uniformly sampling the Jeffreys prior.
+In practice, this is done either by constructing a hexagonal lattice \citep{PhysRevD.76.102004} or sampling stochastically \citep{2009PhRvD..80j4014H,2009PhRvD..80b4009V,SBank,2010PhRvD..81b4004M,2014PhRvD..89b4003P}}.
 We will momentarily consider the likelihood for a single detector:
@@
 \begin{multline}\label{eq:gaussian-likelihood-spa}
-    \mathcal{L}\left(y_i; \rho_i, \gamma_i, \tau_i,
-        \bm\theta_\mathrm{in}\right) \\
+    \mathcal{L}\left(Y_i; \rho_i, \gamma_i, \tau_i,
+        \bm\theta_\mathrm{in}\right)
+    \\
     \propto \exp \left[
-        - \frac{1}{2} \int_0^\infty \frac{\left|y_i (\omega)
-            - x_i\left(y_i; \rho_i, \gamma_i, \tau_i,
+        - \frac{1}{2} \int_0^\infty \frac{\left|Y_i (\omega)
+            - X_i\left(\omega; \rho_i, \gamma_i, \tau_i,
                 \bm\theta_\mathrm{in}\right)
@@
 %
-with $x_i(\omega; \rho_i, \gamma_i, \tau_i, \bm\theta_\mathrm{in})$ given by Equation~(\ref{eq:signal-model}).
-
-The Fisher information matrix for a measurement $y$ described by the unknown parameter vector $\bm{\theta}$ is
+with $X_i(\omega; \rho_i, \gamma_i, \tau_i, \bm\theta_\mathrm{in})$ given by Equation~(\ref{eq:signal-model}).
+
+The Fisher information matrix for a measurement $y$ described by the unknown parameter vector $\bm{\theta}$ is the conditional expectation value
 %
@@
     \mathcal{I}_{jk} = \mathrm{E} \, \left[
-        -\frac{\partial^2 \log
-            \mathcal{L}(y_i ; \bm\theta)}
-            {\partial \theta_j \theta_k}
+        \left(\frac{\partial \log
+            \mathcal{L}(Y_i ; \bm\theta)}
+            {\partial \theta_j}\right)
+        \left(\frac{\partial \log
+            \mathcal{L}(Y_i ; \bm\theta)}
+            {\partial \theta_k}\right)
     \middle| \bm\theta
@@
 %
-In other words, it is the expectation value, condition on the true parameter values, of the Jacobian of the log likelihood.
-This equation involves double derivatives, but substantial simplification is possible when---as in this case---the likelihood is Gaussian:
+Note that if $\log\mathcal{L}$ is twice differentiable in terms of $\bm\theta$, then the Fisher matrix can also be written in terms of second derivatives as
+%
+\begin{equation}\label{eq:general-fisher-matrix-second-derivatives}
+    \mathcal{I}_{jk} = \mathrm{E} \, \left[
+        -\frac{\partial^2 \log
+            \mathcal{L}(Y_i ; \bm\theta)}
+            {\partial \theta_j \partial \theta_k}
+    \middle| \bm\theta
+    \right].
+\end{equation}
+%
+In this form, we can recognize the Fisher matrix as the expectation value, conditioned upon the true parameter values, of the Hessian matrix of the log likelihood.
+It describes how strongly the likelihood depends, on average, on the parameters.
+If $\hat{\bm\theta}$ is an unbiased estimator of $\bm\theta$, $\tilde{\bm\theta} = \hat{\bm\theta} - \bm\theta$ is the measurement error, and $\Sigma = \mathrm{E} \, [\tilde{\bm\theta}\tilde{\bm\theta}\transpose]$ is the covariance of the measurement error, then the \ac{CRLB} says that $\Sigma \geq \mathcal{I}^{-1}$, in the sense that $\left(\Sigma - \mathcal{I}^{-1}\right)$ is positive semi\nobreakdashes-definite.
+
+When (as in our assumptions) the likelihood is Gaussian, Equation~(\ref{eq:general-fisher-matrix}) simplifies to
 %
@@
     \mathcal{I}_{jk} = \int_0^\infty \Re \left[
-        \left(\frac{\partial x_i}{\partial \theta_j}\right)^*
-        \left(\frac{\partial x_i}{\partial \theta_k}\right)
+        \left(\frac{\partial X_i}{\partial \theta_j}\right)^*
+        \left(\frac{\partial X_i}{\partial \theta_k}\right)
     \right] \frac{1}{S_i(\omega)} \, d\omega.
@@
 %
-This form is more useful because it involves only first derivatives, and of the signal $x_i (\omega)$ rather than the entire observation $y (\omega)$.
-In terms of the $k$th \ac{SNR}-weighted moment of angular frequency,
+This form is useful because it involves manipulating the signal $X_i (\omega)$ rather than the entire observation $Y (\omega)$. In terms of the $k$th \ac{SNR}-weighted moment of angular frequency,
 %
@@
             \mathcal{I}_{\bm\theta_i,\bm\theta_\mathrm{in}} \\
-            {\mathcal{I}_{\bm\theta_i,\bm\theta_\mathrm{in}}}^\intercal &
+            {\mathcal{I}_{\bm\theta_i,\bm\theta_\mathrm{in}}}\transpose &
             {\rho_i}^2 \mathcal{I}_{\bm\theta_\mathrm{in},\bm\theta_\mathrm{in}}
@@
 %
-(See also \cite{Grover:2013}.) To lowest post\nobreakdashes-Newtonian order, the intrinsic parameters affect the phase evolution but not the amplitude evolution, so for BNS systems we can assume $\mathcal{I}_{\rho_i,\bm\theta_\mathrm{in}} \approx 0$.
-%
-\marginpar{Show this?}%
-%
-When we form the Fisher matrix for the whole network, if the detectors all have the same noise \acp{PSD}, $S_1(\omega) = S_2(\omega) = \cdots = S_n(\omega) \equiv S(\omega)$, then we can change variables from phases and times on arrival $\left\{\gamma_i, \tau_i\right\}_i$ to appropriate differences between them, in order to cancel out all off-diagonal terms between the intrinsic and extrinsic parameters.
-Because all existing ground-based GW detectors have similar bandpasses, we restrict all of our further attention to just the extrinsic part of the Fisher matrix, $\mathcal{I}_{\bm\theta_i,\bm\theta_i}$.
-
-\subsection{\ac{CRLB} and physical interpretation}
-
-If $\hat{\bm\theta}$ is an unbiased estimator of $\bm\theta$, $\tilde{\bm\theta} = \hat{\bm\theta} - \bm\theta$ is the measurement error, and $\Sigma = \mathrm{E} \, [\tilde{\bm\theta}\tilde{\bm\theta}^\intercal]$ is the covariance of the measurement error, then the \ac{CRLB} says that $\Sigma \geq \mathcal{I}^{-1}$, in the sense that $\left(\Sigma - \mathcal{I}^{-1}\right)$ is positive semi\nobreakdashes-definite.
-Specifically, for our likelihood, the \ac{CRLB} implies that
-%
-\begin{equation}\label{eq:covariance-matrix}
-    \cov{
-        \left(
-        \begin{array}{c}
-            \tilde{\rho}_i \\
-            \tilde{\gamma}_i \\
-            \tilde{\tau}_i
-        \end{array}
-        \right)
-    } \geq \mathcal{I}^{-1} = \left(
-        \begin{array}{ccc}
-            1 & 0 & 0 \\
-            0 & {\rho_i}^2 {\overline{\omega^2}}_i/{\omega_{\mathrm{rms},i}}^2 & {\rho_i}^2 {\overline{\omega}}_i/{\omega_{\mathrm{rms},i}}^2 \\
-            0 & {\rho_i}^2 {\overline{\omega}}_i/{\omega_{\mathrm{rms},i}}^2 & {\rho_i}^2/{\omega_{\mathrm{rms},i}}^2
-        \end{array}
-    \right)
-\end{equation}
-%
-where ${\omega_{\mathrm{rms},i}}^2 = {\overline{\omega^2}}_i - {{\overline{\omega}}_i}^2$.
-Reading off the $\tau \tau$ element of the covariance matrix reproduces the timing accuracy in Equation~(24) of \cite{fairhurst:2009},
-%
-\begin{equation}\label{eq:timing-crlb}
-    \std \left(\hat{\tau}_i - \tau_i \right) \geq \sqrt{\left(\mathcal{I}^{-1}\right)_{\tau\tau}} = \frac{\rho_i}{\omega_{\mathrm{rms},i}}.
-\end{equation}
-%
-Furthermore, the Fisher matrix in Equation~(\ref{eq:fisher-matrix}) is block diagonal, which implies that estimation errors in the signal amplitude $\rho$ are uncorrelated with the phase $\gamma$ and time $\tau$.
-A sequence of two changes of variables lends some physical interpretation to the nature of the coupled estimation errors in $\gamma$ and $\tau$.
-
-First, we put the phase and time on the same footing by measuring the time in units of $1 / \sqrt{\overline{\omega^2}}$ with a change of variables from $\tau$ to $\gamma_\tau = \sqrt{\overline{\omega^2}} \tau$:
-%
-\begin{equation}
-    \mathcal{I}' = \bordermatrix{
-        ~ & \rho_i & \gamma_i & \gamma_{\tau,i} \cr
-        \rho_i & 1 & 0 & 0 \cr
-        \gamma_i & 0 & {\rho_i}^2 & -{\rho_i}^2\frac{{\overline{\omega}}_i}{\sqrt{{\overline{\omega^2}}_i}} \cr
-        \gamma_{\tau,i} & 0 & -{\rho_i}^2\frac{{\overline{\omega}}_i}{\sqrt{{\overline{\omega^2}}_i}} & {\rho_i}^2
-    }.
-\end{equation}
-%
-The second change of variables, from $\gamma$ and $\gamma_\tau$ to $\gamma_\pm = \frac{1}{\sqrt{2}}(\gamma \pm \gamma_\tau)$, diagonalizes the Fisher matrix:
-%
-\begin{equation}\label{eq:fisher-matrix-extrinsic-one-detector}
-    \mathcal{I}'' = \bordermatrix{
-        ~ & \rho_i & \gamma_{+,i} & \gamma_{-,i} \cr
-        \rho_i & 1 & 0 & 0 \cr
-        \gamma_{+,i} & 0 & \left(1 - \frac{\overline{\omega}_i}{\sqrt{\overline{\omega^2}_i}}\right){\rho_i}^2 & 0 \cr
-        \gamma_{-,i} & 0 & 0 & \left(1 + \frac{\overline{\omega}_i}{\sqrt{\overline{\omega^2}_i}}\right){\rho_i}^2
-    }.
-\end{equation}
-%
-Thus, in the appopriate time units, the \textit{sum and difference} of the phase and time of the signal are measured independently.
-
-\section{The \ac{BAYESTAR} likelihood}
-
-Although we may not be able to compute the detection candidate likelihood, Equation~(\ref{eq:detection-candidate-likelihood}), directly, we can make an educated guess that has many properties in common with it.
-Any valid approximate likelihood must have the same Fisher matrix Equation~(\ref{eq:fisher-matrix-extrinsic-one-detector}).
+(This is equivalent to an expression given in \citealt{Grover:2013}.) The information matrix elements that relate to the intrinsic parameters can also be expressed as linear combinations of the angular frequency moments.
+However, as we will see in the next section, we need not compute these matrix elements if we are only interested in sky localization accuracy.
+
+\section{Bayesian probability and parameter estimation}
+\label{sec:parameter-estimation}
+
+In the Bayesian framework, parameters are inferred from the data by forming the posterior distribution, $p(\bm\theta|\mathbf y)$, which describes the probability of the parameters given the observations.
+Bayes' rule relates the likelihood $p(\mathbf y|\bm\theta)$ to the posterior $p(\bm\theta|\mathbf y)$,
+%
+\begin{equation}\label{bayes}
+p(\bm\theta|\mathbf y) = \frac{p(\mathbf y|\bm\theta) p(\bm\theta)}{p(\mathbf y)},
+\end{equation}
+%
+introducing the prior distribution $p(\bm\theta)$ which encapsulates previous information about the paramters (for example, arising from earlier observations or from known physical bounds on the parameters), and the evidence $p(\mathbf y)$ which can be thought of as a normalization factor or as describing the parsimoniousness of the model.
+
+The choice of prior is determined by one's astrophysical assumptions.
+During \ac{S6} when LIGO's Bayesian \ac{CBC} parameter estimation pipelines were first deployed, the prior was taken to be isotropic in sky location and binary orientation, and uniform in volume, arrival time, and the component masses~\cite{S6PE}.
+
+In Bayesian inference, although it is often easy to write down the likelihood or even the full posterior in closed form, usually one is interested in only a subset $\bm\beta$ of all of the model's parameters, the others $\bm\lambda$ being nuisance parameters.
+In this case, we integrate away the nuisance parameters, forming the marginal posterior
+%
+\begin{equation}\label{eq:marginal-posterior}
+    p(\bm\beta|\mathbf y) = \int \frac{p(\mathbf y|\bm\beta,\bm\lambda) p(\bm\beta,\bm\lambda)}{p(\mathbf y)} \,d\bm\lambda
+\end{equation}
+%
+with $\bm\theta = (\bm\beta, \bm\lambda)$. For instance, for the purpose of locating a \ac{GW} source on the sky, all parameters but $(\alpha, \delta)$ are nuisance parameters.
+
+% Bayesian parameter estimation has many advantages, including broad generality and the ability to make probabilistically meaningful statements even with very low \ac{SNR} measurements.
+However, in problems of even modest complexity, the marginalization step involves many-dimensional, ill-behaved integrals.
+The powerful \ac{MCMC} integration technique has become almost synonymous with Bayesian inference.
+Though powerful, \ac{MCMC} is inherently non-deterministic and resistant to parallelization, as well as (at least historically) slow.
+With the most sophisticated \ac{CBC} parameter estimation codes, it still takes days to process a single event.
+This delay is undesirable for planning targeted \ac{EM} follow\nobreakdashes-up searches of LIGO events.
+
+\section{The \acs{BAYESTAR} likelihood}
+\label{sec:bayestar-likelihood}
+
+For the purpose of rapid sky localization, we assume that we do not have access to the \ac{GW} data $\mathbf{Y}$ itself, and that our only contact with it is through the \ac{ML} parameter estimates $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$.
+Although this is a significant departure from conventional \ac{GW} parameter estimation techniques, we can still apply the full Bayesian machinery of Equation~(\ref{eq:marginal-posterior}) to compute a posterior distribution for the sky location.
+
+The relevant likelihood is now the probability of the \ac{ML} estimates, conditioned upon the true parameter values, and marginalized over all possible \ac{GW} observations:
+%
+\begin{equation}\label{eq:detection-candidate-likelihood}
+    p\left(\{\hat{\bm\theta}_i\}_i,
+        \hat{\bm\theta}_\mathrm{in}
+    \middle| \bm\theta\right)
+    \propto \int\limits_{\mathclap{\mathbf{Y} | \mathrm{MLE}(\mathbf{Y}) =
+        \{\{\hat{\bm\theta}_i\}_i,
+        \hat{\bm\theta}_\mathrm{in}\}}}
+    p(\mathbf{Y} | \bm\theta) \, p(\bm\theta)
+    \, d\mathbf{Y}.
+\end{equation}
+%
+Although we may not be able to evaluate this equation directly, with some educated guesses we can create a likelihood that has many properties in common with it.
+Any valid approximate likelihood must have the same Fisher matrix as shown in Equation~(\ref{eq:fisher-matrix-extrinsic-one-detector}).
 It must also have the same limiting behavior: it should be periodic in the phase error $\tilde{\gamma}_i$ and go to zero as $\tilde{\tau}_i \rightarrow \pm \infty$, $\hat{\rho}_i \rightarrow 0$, or $\hat{\rho}_i \rightarrow \infty$.
-Additionally, when $\tilde{\tau}_i = 0$, the distribution of ${\hat{\rho}_i}^2$ should reduce to a noncentral $\chi^2$ distribution with two degrees of freedom, centered about ${\rho_i}^2$, because the complex matched filter time series $z_i(t)$ is Gaussian.
-
-These conditions could be satisfied by realizing a multivariate Gaussian distribution with covariance matrix $\Sigma = \mathcal{I}^\intercal$, and then replacing individual terms of the form $-\tilde{\theta}^2/2$ with $\cos{\tilde{\theta}}$.
-
-\marginpar{Actually, this does not give exactly the same Fisher matrix. Explain.}%
-%
-Another way is to plug the signal model from Equation~(\ref{eq:signal-model}) \emph{evaluated at the ML parameter estimates} into the single-detector likelihood in Equation~(\ref{eq:gaussian-likelihood-spa}):
+Additionally, when $\tilde{\tau}_i = 0$, the distribution of ${\hat{\rho}_i}^2$ should reduce to a noncentral $\chi^2$ distribution with two degrees of freedom, centered about ${\rho_i}^2$, because the complex matched filter time series $z_i(t)$ is Gaussian (under the ideal assumption the \ac{GW} strain time series is Gaussian).
+
+These conditions could be satisfied by realizing a multivariate Gaussian distribution with covariance matrix $\Sigma = \mathcal{I}^\intercal$, and then replacing individual quadratic terms in the exponent of the form $-\tilde{\theta}^2/2$ with $\cos{\tilde{\theta}}$.
+
+A more natural way is to plug the signal model from Equation~(\ref{eq:signal-model}) \emph{evaluated at the \ac{ML} parameter estimates} into the single-detector likelihood in Equation~(\ref{eq:gaussian-likelihood-spa}):
 %
 \begin{widetext}
-\begin{eqnarray*}
+\begin{equation}
     p\left(\hat{\bm\theta}_i \middle| \bm\theta \right)
-    &:=&
-    p\left(y_i(\omega) = x_i(\omega; \hat{\bm\theta}_i)
+    :=
+    p\left(Y_i(\omega) = X_i(\omega; \hat{\bm\theta})
         \middle| \bm\theta \right)
-    \\
-    &\propto& \exp \left[
-        - \frac{1}{2} \int_0^\infty \frac{\left|
-            \frac{\hat{\rho}_i}{\sigma_i(\hat{\bm\theta}_\mathrm{in})} e^{i (\hat\gamma_i - \omega \hat\tau_i)} h(\omega; \hat{\bm\theta}_\mathrm{in})
-            - \frac{\rho_i}{\sigma_i(\bm\theta_\mathrm{in})} e^{i (\gamma_i - \omega \tau_i)} h(\omega; \bm\theta_\mathrm{in})
-        \right|^2}{S_i(\omega)} \, d\omega
-    \right].
-\end{eqnarray*}
+    %\\
+    \propto \exp \Bigg[
+        - \frac{1}{2} \int_0^\infty \Bigg|
+            \frac{\hat{\rho}_i}{\sigma_i(\hat{\bm\theta}_\mathrm{in})} e^{i (\hat\gamma_i - \omega \hat\tau_i)} \frac{H(\omega; \hat{\bm\theta}_\mathrm{in})}{S_i(\omega)}
+    %\\
+            - \frac{\rho_i}{\sigma_i(\bm\theta_\mathrm{in})} e^{i (\gamma_i - \omega \tau_i)} \frac{H(\omega; \bm\theta_\mathrm{in})}{S_i(\omega)}
+        \Bigg|^2 \, d\omega
+    \Bigg].
+\end{equation}
 \end{widetext}
 %
-If we assume that $\hat{\bm\theta}_\mathrm{in} = \bm\theta_\mathrm{in}$, then this reduces to
-%
+If we further assume that \emph{the intrinsic parameters are equal to their \ac{ML} estimates}, $\bm\theta_\mathrm{in} = \hat{\bm\theta}_\mathrm{in}$, then this reduces to what we call the \emph{autocorrelation likelihood},
+%
+\begin{widetext}
 \begin{equation}\label{eq:autocor-likelihood}
@@
         \right\}
-    \right]
-\end{equation}
+    \right],
+\end{equation}
+\end{widetext}
 %
@@
 \begin{equation}\label{eq:autocorrelation-function}
-    a_i(t; \bm\theta_\mathrm{in}) := \frac{1}{{{\sigma_i}^2(\bm\theta_\mathrm{in})}} \int_0^\infty \frac{\left| h(\omega; \hat{\bm\theta}_\mathrm{in})\right|^2}{S_i(\omega)} e^{i \omega t} \,d\omega.
-\end{equation}
-
-To assemble the joint likelihood for the whole network, we just form the product from the individual detectors:
+    a_i(t; \bm\theta_\mathrm{in}) := \frac{1}{{{\sigma_i}^2(\bm\theta_\mathrm{in})}} \int_0^\infty \frac{\left| H(\omega; \hat{\bm\theta}_\mathrm{in})\right|^2}{S_i(\omega)} e^{i \omega t} \,d\omega.
+\end{equation}
+
+To assemble the joint likelihood for the whole network, we form the product of the autocorrelation likelihoods from the individual detectors:
 %
@@
     p\left(\left\{\hat\rho_i, \hat\gamma_i, \hat\tau_i\right\}_i
-        \middle| \left\{\rho_i, \gamma_i, \tau_i\right\}_i \right) \\ \propto
+        \middle| \left\{\rho_i, \gamma_i, \tau_i\right\}_i \right) \\
+    \propto
     \exp \left[ - \frac{1}{2} \sum_i {\hat\rho_i}^2 - \frac{1}{2} \sum_i {\rho_i}^2
@@
 \begin{equation*}
-    a^{(n)}(t) = i^n \overline{\omega^n};
+    a^{(n)}(t) = i^n \overline{\omega^n},
 \end{equation*}
 %
-for instance,
+with $\overline{\omega^n}$ defined in Equation~(\ref{eq:angular-frequency-moments}). For example, the first few derivatives are
 %
@@
     \qquad
-    a'(0) = i \overline{\omega},
+    \dot{a}(0) = i \overline{\omega},
     \qquad
-    a''(0) = -\overline{\omega^2}.
+    \ddot{a}(0) = -\overline{\omega^2}.
 \end{equation*}
 
-We can compute the Fisher matrix elements for the autocorrelation likelihood given by Equation~(\ref{eq:autocor-likelihood}), with detector subscript suppressed:
-%
-\begin{eqnarray}
+Using Equation~(\ref{eq:general-fisher-matrix-second-derivatives}), we can compute the Fisher matrix elements for the autocorrelation likelihood given by Equation~(\ref{eq:autocor-likelihood}), with detector subscript suppressed:
+%
+\begin{align}
     \nonumber
-    \mathcal{I}_{\rho\rho} &=& 1 \\
+    \mathcal{I}_{\rho\rho} &= 1, \\
     \nonumber
-    \mathcal{I}_{\rho\gamma} &=& 0 \\
+    \mathcal{I}_{\rho\gamma} &= 0, \\
     \nonumber
-    \mathcal{I}_{\rho\tau} &=& 0 \\
+    \mathcal{I}_{\rho\tau} &= 0, \\
     \label{eq:fisher-autocor-gamma-gamma}
-    \mathcal{I}_{\gamma\gamma} &=& \rho^2
-        \int_{-T}^T \left|a(t)\right|^2 w(t; \rho) dt \\
+    \mathcal{I}_{\gamma\gamma} &= \rho^2
+        \int_{-T}^T \left|a(t)\right|^2 w(t; \rho) dt, \\
     \label{eq:fisher-autocor-tau-tau}
-    \mathcal{I}_{\tau\tau} &=& -\rho^2
-        \int_{-T}^T \Re\left[a^*(t) a''(t)\right] w(t; \rho) dt \\
+    \mathcal{I}_{\tau\tau} &= -\rho^2
+        \int_{-T}^T \Re\left[a^*(t) \ddot{a}(t)\right] w(t; \rho) dt, \\
     \label{eq:fisher-autocor-gamma-tau}
-    \mathcal{I}_{\gamma\tau} &=& -\rho^2
-        \int_{-T}^T \Im\left[a^*(t) a'(t)\right] w(t; \rho) dt \\
-    \nonumber\textrm{where} && \\
-    w(t; \rho) &=& \frac{
+    \mathcal{I}_{\gamma\tau} &= -\rho^2
+        \int_{-T}^T \Im\left[a^*(t) \dot{a}(t)\right] w(t; \rho) dt,
+\intertext{where}
+    w(t; \rho) &= \frac{
         \displaystyle
@@
     }.
-\end{eqnarray}
+\end{align}
 %
@@
 Note that the minus signs are correct but a little confusing: despite them, $\mathcal{I}_{\gamma\gamma}, \mathcal{I}_{\tau\tau} \geq 0$ and $\mathcal{I}_{\gamma\tau} \leq 0$.
-The time integration limits $[-T, T]$ constitute a flat prior on arrival time, the time coincidence window between detectors.
-
-If we can show that the weighting function $w(t; \rho)$ approaches a Dirac delta function as $\rho \rightarrow \infty$, then the Fisher matrix for the autocorrelation likelihood approaches the Fisher matrix for the full GW data, Equation~(\ref{eq:fisher-matrix}), as $\rho \rightarrow \infty$.
+The time integration limits $[-T, T]$ correspond to a flat prior on arrival time, or a time coincidence window between detectors.
+
+We can show that the weighting function $w(t; \rho)$ approaches a Dirac delta function as $\rho \rightarrow \infty$, so that the Fisher matrix for the autocorrelation likelihood approaches the Fisher matrix for the full \ac{GW} data, Equation~(\ref{eq:fisher-matrix}), as $\rho \rightarrow \infty$.
 The Bessel functions asymptotically approach:
@@
 %
-\begin{equation*}
-    |a(t)|^2 = 1 + \frac{1}{2} \left(\frac{\partial^2}{\partial t^2}|a(t)|^2 \Bigg|_{t=0}\right) t^2 + \mathcal{O}(t^4)
-    = 1 - {\omega_\mathrm{rms}}^2 t^2 + \mathcal{O}(t^4).
-\end{equation*}
+\begin{align}
+    \nonumber
+    |a(t)|^2 &= 1 + \frac{1}{2} \left(\frac{\partial^2}{\partial t^2}|a(t)|^2 \Bigg|_{t=0}\right) t^2 + \mathcal{O}(t^4) \\
+    \nonumber
+    &= 1 - {\omega_\mathrm{rms}}^2 t^2 + \mathcal{O}(t^4).
+\end{align}
 %
@@
 %
-\begin{eqnarray*}
-    \mathcal{I}_{\gamma\gamma} &=& \rho^2 \cdot \textproto{D}_{\gamma\gamma}(\rho) \\
-    \mathcal{I}_{\tau\tau} &=& \rho^2 \overline{\omega^2} \cdot \textproto{D}_{\tau\tau}(\rho) \\
-    \mathcal{I}_{\gamma\tau} &=& -\rho^2 \overline{\omega} \cdot \textproto{D}_{\gamma\tau}(\rho)
-\end{eqnarray*}
-%
-Now, the $\textproto{D}_{ij}$ %
-%
-\marginpar{The Fish(er) factor. Go home, Leo.}%
-%
-contain the integrals from Equations~(\ref{eq:fisher-autocor-gamma-gamma}, \ref{eq:fisher-autocor-tau-tau}, \ref{eq:fisher-autocor-gamma-tau}) and encode the departure of the autocorrelation likelihood from the likelihood of the full data at low SNR.
-All of the $\textproto{D}_{ij}(\rho)$ are sigmoid-type functions that asymptotically approach 1 as $\rho \rightarrow \infty$.
-The transition SNR $\rho_\mathrm{crit}$ is largely the same for all three nontrivial matrix elements, and is determined by the time coincidence window $T$ and the signal bandwidth $\omega_\mathrm{rms}$.
-%
-\marginpar{The transition SNR is probably proportional to some power of $(\omega_\mathrm{rms} T)$.}
-
-In the limit of large SNR, our interpretation is that the point estimates $(\hat\rho, \hat\gamma, \hat\tau)$ contain all of the information about the underlying extrinsic parameters.
-
-On the other hand, in the low SNR limit, the diminishing value of $\textproto{D}_{ij}(\rho)$ reflects the fact that some information is lost when the full data $\mathbf{x}$ is discarded.
-Concretely, as the prior interval $T$ becomes large compared to 1/$\rho\omega_\mathrm{rms}$, the ML estimator becomes more and more prone to pick up a spurious noise fluctuation far from the true signal.
-Clearly, when the coincidence window $T$ is kept small as possible, more information is retained in the ML point estimates.
-Put another way, if $T$ is small, then the transition SNR $\rho_\mathrm{crit}$ is also small and fainter signals become useful for parameter estimation.
+\begin{align*}
+    \mathcal{I}_{\gamma\gamma} &= \rho^2 \cdot \textproto{D}_{\gamma\gamma}(\rho), \\
+    \mathcal{I}_{\tau\tau} &= \rho^2 \overline{\omega^2} \cdot \textproto{D}_{\tau\tau}(\rho), \\
+    \mathcal{I}_{\gamma\tau} &= -\rho^2 \overline{\omega} \cdot \textproto{D}_{\gamma\tau}(\rho).
+\end{align*}
+%
+Now, the $\textproto{D}_{ij}$%
+%
+\footnote{The Fish(er) factor.}
+%
+contain the integrals from Equations~(\ref{eq:fisher-autocor-gamma-gamma}, \ref{eq:fisher-autocor-tau-tau}, \ref{eq:fisher-autocor-gamma-tau}) and encode the departure of the autocorrelation likelihood from the likelihood of the full data at low \ac{SNR}.
+All of the $\textproto{D}_{ij}(\rho)$ are sigmoid-type functions that asymptotically approach 1 as $\rho \rightarrow \infty$ (see Figures~\ref{fig:crlb-tau} and~\ref{fig:fishfactor}).
+The transition \ac{SNR} $\rho_\mathrm{crit}$ is largely the same for all three nontrivial matrix elements, and is determined by the time coincidence window $T$ and the signal bandwidth $\omega_\mathrm{rms}$.
+
+In the limit of large \ac{SNR}, our interpretation is that the point estimates $(\hat\rho, \hat\gamma, \hat\tau)$ contain all of the information about the underlying extrinsic parameters.
+
+On the other hand, in the low \ac{SNR} limit, the diminishing value of $\textproto{D}_{ij}(\rho)$ reflects the fact that some information is lost when the full data $\mathbf{x}$ is discarded.
+Concretely, as the prior interval $T$ becomes large compared to 1/$\rho\omega_\mathrm{rms}$, the \ac{ML} estimator becomes more and more prone to picking up spurious noise fluctuations far from the true signal.
+Clearly, when the coincidence window $T$ is kept small as possible, more information is retained in the \ac{ML} point estimates.
+Put another way, if $T$ is small, then the transition \ac{SNR} $\rho_\mathrm{crit}$ is also small and fainter signals become useful for parameter estimation.
+In this way, the \ac{BAYESTAR} likelihood exhibits the \emph{threshold effect} that is well\nobreakdashes-known in communication and radar applications \citep{barankin1949locally,mcaulay1969barankin,mcaulay1971barankin}.
 
 \begin{figure*}
-    \begin{center}
-        \includegraphics{crlb_tau}
-        \includegraphics{crlb_gamma}
-    \end{center}
-    \caption{\label{fig:crlb-tau}\ac{CRLB} on root\nobreakdashes-mean square timing uncertainty, using the likelihood for the full GW data (Equation~\ref{eq:gaussian-likelihood-spa}; dashed diagonal line) or the autocorrelation likelihood (Equation~\ref{eq:autocor-likelihood}; solid lines) with a selection of arrival time priors.}
+    \centering
+        \includegraphics[width=\columnwidth]{crlb_tau}
+        \includegraphics[width=\columnwidth]{crlb_gamma}
+    \caption[\acl{CRLB} on time and phase accuracy]{\label{fig:crlb-tau}\ac{CRLB} on \ac{RMS} timing uncertainty and phase error, using the likelihood for the full \ac{GW} data (Equation~\ref{eq:gaussian-likelihood-spa}; dashed diagonal line) or the autocorrelation likelihood (Equation~\ref{eq:autocor-likelihood}; solid lines) with a selection of arrival time priors.}
 \end{figure*}
@@
 \begin{figure}
-    \begin{center}
+    \centering
     \includegraphics{fishfactor}
-    \end{center}
-    \caption{\label{fig:fishfactor}Ratio between Fisher matrix elements (solid: $\mathcal{I}_{\gamma\gamma}$, dashed: $\mathcal{I}_{\gamma\tau}$, dotted: $\mathcal{I}_{\tau\tau}$) for the the autocorrelation likelihood and the full GW data.
+    \caption[Ratio of Fisher matrix elements between autocorrelation likelihood and full \acs{GW} data]{\label{fig:fishfactor}Ratio between Fisher matrix elements (solid: $\textproto{D}_{\gamma\gamma}$, dashed: $\textproto{D}_{\gamma\tau}$, dotted: $\textproto{D}_{\tau\tau}$) for the autocorrelation likelihood and the full \ac{GW} data.
 Colors correspond to different arrival time priors as in Figure~\ref{fig:crlb-tau}.}
@@
 
-\section{The \ac{BAYESTAR} algorithm}
-
-Here we describe the numerical implementation of \ac{BAYESTAR}.
-
-\subsection{Prior and problem setup}
-
-The detection pipeline supplies a candidate $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$, and discretely sampled noise \acp{PSD} $S_i(\omega_j)$ for all detectors.
-\marginpar{Discuss single-detector threshold here?}
-
-We compute the GW signal for a source with intrinsic parameters equal to the detection pipeline's estimate, $h(\omega; \hat{\bm\theta}_\mathrm{in})$.
-Then we find the SNR=1 horizon distance $r_{1,i}$ for each detector by numerically integrating Equation~(\ref{eq:horizon}).
+In the following sections, we describe the numerical implementation of \ac{BAYESTAR}.
+
+\section{Prior and problem setup}
+\label{sec:prior}
+
+The detection pipeline supplies a candidate, $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$, and discretely sampled noise \acp{PSD}, $S_i(\omega_j)$, for all detectors.
+We compute the \ac{GW} signal for a source with intrinsic parameters equal to the detection pipeline's estimate, $H(\omega; \hat{\bm\theta}_\mathrm{in})$.
+Then we find the \ac{SNR}=1 horizon distance $r_{1,i}$ for each detector by numerically integrating Equation~(\ref{eq:horizon}).
 
@@
 The arrival time prior is connected to the origin of the detector coordinate system.
-Given the Earth-fixed coordinates of the detectors $\mathbf{n}_i$ and the arrival times $\tau_i$, we compute their averages weighted by the timing uncertainty formula:
+Given the Earth\nobreakdashes-fixed coordinates of the detectors $\mathbf{n}_i$ and the arrival times $\tau_i$, we compute their averages weighted by the timing uncertainty formula:
 %
@@
 %
-Then, we subtract off these means:
+Then, we subtract these means:
 %
@@
 %
-In these coordinates relative to the weighted detector array barycenter, the arrival time prior is uniform in $-T \leq t \leq T$, with $T = \max\limits_i |\mathbf{n}_i| / c + 5\,\textrm{ms}$.
+In these coordinates, now relative to the weighted detector array barycenter, the arrival time prior is uniform in $-T \leq t \leq T$, with $T = \max\limits_i |\mathbf{n}_i| / c + 5~\textrm{ms}$.
 
@@
         r^m & \text{if } r_\mathrm{min} < r < r_\mathrm{max} \\
-        0 & \text{otherwise}
+        0 & \text{otherwise},
     \end{cases}
@@
 We compute the autocorrelation function for each detector from $t = 0$ to $t = T$ at intervals of $\Delta t = 1/f_s$, where $f_s$ is the smallest power of two that is greater than or equal to the Nyquist rate.
-Because BNS signals typically terminate at about 1500\,Hz, a typical value for $\Delta t$ is $(4096\,\textrm{Hz})^{-1}$.
-We use a pruned \ac{FFT} because for BNS systems, the GW signal remains in LIGO's sensitive band for $\sim$100\nobreakdashes--1000\,s, whereas $T \sim 10$\,ms.\footnote{See \url{http://www.fftw.org/pruned.html} for a discussion of methods for computing the first $K$ samples of an \ac{FFT} of length $N$.}
-
-\subsection{Marginal posterior}
+Because BNS signals typically terminate at about 1500~Hz, a typical value for $\Delta t$ is $(4096\,\textrm{Hz})^{-1}$.
+We use a pruned \ac{FFT} because for BNS systems, the \ac{GW} signal remains in LIGO's sensitive band for $\sim$100\nobreakdashes--1000~s, whereas $T \sim 10$~ms.\footnote{See \url{http://www.fftw.org/pruned.html} for a discussion of methods for computing the first $K$ samples of an \ac{FFT} of length $N$.}
+
+\section{Marginal posterior}
+\label{sec:marginal-posterior}
+
+\begin{figure*}
+    \includegraphics[width=\textwidth]{illustration}
+    \caption{\label{fig:adaptive-illustration}Illustration of the \ac{BAYESTAR} adaptive mesh refinement scheme.}
+\end{figure*}
 
@@
     \int_{r_\mathrm{min}}^{r_\mathrm{max}}
-    \int_{0}^{2\pi} \\
-    %
+    \int_{0}^{2\pi}
+    \\
     \exp \left[ - \frac{1}{2} \sum_i {\rho_i}^2
@@
     \right] \\
-    r^m \, dt_\oplus \, d\phi_c dr \, d\cos{\iota} \, d\psi.
+    r^m d\phi_c \, dr \, dt_\oplus \, d\cos{\iota} \, d\psi.
 \end{multline}
 
-To marginalize over the coalescence phase, we can insert an arbitrary nuisance phase $\eta$ by making the replacement $\tilde{\gamma}_i \rightarrow \tilde{\gamma}_i + \eta$.
-Then integrating over $\eta$ from $0$ to $2\pi$ and suppressing normalization factors, we get
+To marginalize over the coalescence phase, we can write $\tilde{\gamma}_i = \tilde{\gamma}_i^\prime + 2\phi_c$. Then integrating over $\phi_c$ and suppressing normalization factors, we get
 %
@@
     \int_{-T}^{T}
-    \int_{r_\mathrm{min}}^{r_\mathrm{max}} \\
-    %
+    \int_{r_\mathrm{min}}^{r_\mathrm{max}}
+    \\
     \exp \left[ - \frac{1}{2} \sum_i {\rho_i}^2 \right] I_0 \left[
@@
     \right] \\
-    r^m dr dt_\oplus \, d\cos{\iota} \, d\psi.
+    r^m dr \, dt_\oplus \, d\cos{\iota} \, d\psi.
 \end{multline}
-
-\subsubsection{Integral over angles and time}
-
-The integrand is periodic in $\psi$, so simple Newton-Cotes quadrature over $\psi$ exhibits extremely rapid convergence.
+%
+In the above equation, we need not distinguish between $\tilde{\gamma}_i$ and $\tilde{\gamma}_i^\prime$ because the likelihood is now invariant under arbitrary phase shifts of all of the detectors' signals.
+
+\subsection{Integral over angles and time}
+
+The integrand is periodic in $\psi$, so simple Newton\nobreakdashes--Cotes quadrature over $\psi$ exhibits extremely rapid convergence.
 We therefore sample the posterior on a regular grid of 10 points from 0 to $\pi$.
 
-The integral over $\cos\iota$ converges with just as rapidly with Gauss-Legendre quadrature, so we use a 10\nobreakdashes-point Gauss-Legendre rule for integration over $\cos\iota$.
+The integral over $\cos\iota$ converges with just as rapidly with Gauss\nobreakdashes--Legendre quadrature, so we use a 10\nobreakdashes-point Gauss\nobreakdashes--Legendre rule for integration over $\cos\iota$.
 
 We sample $t_\oplus$ regularly from $-T$ to $T$ at intervals of $\Delta t$.
-This is typically $\sim 2 (10\,\mathrm{ms})(4096\,\mathrm{Hz}) \approx 80$\,samples.
-We use Catmull\nobreakdashes-Rom cubic splines to interpolate the real and imaginary parts of the autocorrelation functions between samples.
-
-\subsubsection{Integral over distance}
+This is typically $\sim 2 (10\,\mathrm{ms})(4096\,\mathrm{Hz}) \approx 80$~samples.
+We use Catmull\nobreakdashes--Rom cubic splines to interpolate the real and imaginary parts of the autocorrelation functions between samples.
+
+\subsection{Integral over distance}
 
@@
 \begin{equation}\label{eq:importance-sampling-difficult-integral}
-    \mathscr{F}(x_1, x_2) = \int_{x_1}^{x_2} f(x') dx'
+    \mathscr{F}(x_1, x_2) = \int_{x_1}^{x_2} f(x') dx'.
 \end{equation}
@@
 
-We use the same kind of change of variables with fixed\nobreakdashes-order Gaussian quadrature.
-We first find the approximate \ac{ML} distance $r_0$ using Equation~(\ref{eq:ml-distance}).
+We use the same kind of change of variables with fixed\nobreakdashes-order Gaussian quadrature. The distance integral can be written as
+%
+\begin{align}\label{eq:distance-integral}
+    \mathscr{F} &=
+        \int_{\mathrlap{r_\mathrm{min}}}^{\mathrlap{r_\mathrm{max}}}
+        \exp\left[-\frac{p^2}{r^2}\right]
+        I_0\left[\frac{b}{r}\right] r^m dr \nonumber \\
+    &=
+        \int_{\mathrlap{r_\mathrm{min}}}^{\mathrlap{r_\mathrm{max}}}
+        \exp\left[-\frac{p^2}{r^2}+\frac{b}{r}\right]
+        \overline{I}_0\left[\frac{b}{r}\right] r^m dr
+\end{align}
+%
+where
+%
+\begin{align}
+    p^2 &= \frac{1}{2} r^2 \sum_i {\rho_i}^2 \label{eq:likelihood-p-factor} \\
+    b &= r \left| \sum_i \hat\rho_i \rho_i e^{i \tilde{\gamma}_i}
+        a_i^*(\tilde{\tau}_i) \right| \label{eq:likelihood-b-factor} \\
+    \overline{I}_0(x) &= \exp(-|x|) I_0(x).
+\end{align}
+%
+The coefficients $p^2$ and $b$ are nonnegative and independent of distance.
+The symbol $\overline{I}_0$ denotes an exponentially scaled Bessel function.
+In the limit of large argument, $I_0(|x|) \sim \exp(|x|) / \sqrt{2 \pi |x|}$ \citep{Olver:2010:NHMF,NIST:DLMF}\footnote{\url{http://dlmf.nist.gov/10.40.E1}}.
+The scaled Bessel is useful for evaluation on a computer because it has a relatively small range $(0, 1]$ and varies slowly in proportion to $x^{1/2}$.
+If we neglect both the Bessel function and the $r^m$ prior, then the approximate likelihood $\exp(-p^2/r^2 + b/r)$ is maximized when
+%
+\begin{equation}\label{eq:ml-distance}
+    r = r_0 \equiv 2p^2/b.
+\end{equation}
+%
+with $p$ and $b$ defined above in Equations~(\ref{eq:likelihood-p-factor},~\ref{eq:likelihood-b-factor}).
 If $r_0 < r_\mathrm{min}$ or $r_0 > r_\mathrm{max}$, then the ML peak lies outside of the limits of integration and the original integrand behaves like a low-order power law.
@@
     g(r) = \exp\left[-\left(\frac{p}{r} - \frac{b}{2p}\right)^2\right]
-        \frac{1}{r^2}
+        \frac{1}{r^2},
 \end{equation}
@@
     \mathscr{G}(r) = \frac{\sqrt{\pi}}{p}
-        Q\left[\sqrt{2}\left(\frac{p}{r} - \frac{b}{2p}\right)\right]
+        Q\left[\sqrt{2}\left(\frac{p}{r} - \frac{b}{2p}\right)\right],
 \end{equation}
@@
         {r_0}^m
-        \int_{\mathscr{G}(r_\mathrm{min})}^{\mathscr{G}(r_\mathrm{max})}
+        \int\limits_{\mathclap{\mathscr{G}(r_\mathrm{min})}}^{\mathclap{\mathscr{G}(r_\mathrm{max})}}
         \overline{I}_0\left[\frac{b}{r}\right]
         \left[\frac{r}{r_0}\right]^m r^2
-        \Bigg|_{r = \mathscr{G}^{-1}(G)}
+        \Bigg|_{\mathrlap{r = \mathscr{G}^{-1}(G)}}
         dG.
@@
 %
-This integral is evaluated numerically with a 10\nobreakdashes-point Gauss-Legendre rule.
+This integral is evaluated numerically with a 10\nobreakdashes-point Gauss\nobreakdashes--Legendre rule.
 
@@
 
-\subsection{Adaptive mesh refinement}
+\section{Adaptive HEALPix sampling}
+\label{sec:adaptive-sampling}
 
@@
 There is a hierarchy of \ac{HEALPix} resolutions.
-A HEALPix resolution may be designated by its order $N$.
+A \ac{HEALPix} resolution may be designated by its order $N$.
 The $N=0$th order or base tiling has 12 pixels.
 At every successive order, each tile is subdivided into four new tiles.
-A resolution may also be referred to by the number of subdivisions along each side of the base tiles, $N_\mathrm{side} = N^2$.
+A resolution may also be referred to by the number of subdivisions along each side of the base tiles, $N_\mathrm{side} = 2^N$.
 There are $N^\mathrm{pix} = 12 {N_\mathrm{side}}^2$ pixels at any given resolution.
@@
 We begin by evaluating the posterior probability density at the center of each of the $N_{\mathrm{pix},0} = 3072$ pixels of an $N_{\mathrm{side},0}=16$ \ac{HEALPix} grid.
-At this resolution, each pixel has an area of 13.4\,deg$^2$.
+At this resolution, each pixel has an area of 13.4~deg$^2$.
 We then rank the pixels by contained probability (assuming constant probability density within a pixel), and subdivide the most probable $N_{\mathrm{pix},0}/4$ pixels into $N_{\mathrm{pix},0}$ new pixels.
@@
 We traverse the tree and flatten it into the highest resolution represented.
-The highest possible resolution is $N_\mathrm{side}=2^{11}$, with an area of $\approx 10^{-3}$\,deg$^2$ per pixel.\footnote{Although the resulting sky map contains $N_\mathrm{pix} \approx 5\times10^6$ pixels, at most $\approx 2\times10^4$ pixels have distinct values.
+The highest possible resolution is $N_\mathrm{side}=2^{11}$, with an area of $\approx 10^{-3}$~deg$^2$ per pixel.\footnote{Although the resulting sky map contains $N_\mathrm{pix} \approx 5\times10^6$ pixels, at most $\approx 2\times10^4$ pixels have distinct values.
 For the purpose of delivery to observers, therefore, the output is always \texttt{gzip}\nobreakdashes-compressed with a ratio of $\approx 250:1$.}
@@
 
-\section{Results}
+\section{Run time}
+\label{sec:run-time}
+
+Since \ac{BAYESTAR} is designed as one of the final steps in the real\nobreakdashes-time \ac{BNS} search, it is important to characterize how long it takes to calculate a sky map.
+We compiled \ac{BAYESTAR} with the Intel~C~Compiler~(\texttt{icc}) at the highest architecture\nobreakdashes-specific optimization setting (\texttt{-xhost}).
+We timed it under Scientific~Linux~6 on a Supermicro~1027GR\nobreakdashes-TRFT system with dual Intel Xeon E5-2670 CPUs clocked at 2.60GHz, providing a total of 16 cores and capable of executing 32 threads simultaneously (with hyperthreading).
+In Figure~\ref{fig:runtimes}, we show how long it took to calculate a localization with \ac{BAYESTAR} as the number of OpenMP threads was varied from 1 to 32.
+This is a violin plot, a smoothed vertical histogram.
+The red surface shows run times for a two\nobreakdashes-detector network (HL) modeled on the first scheduled Advanced \ac{LIGO} observing run in 2015, and the blue surface shows run times for a three\nobreakdashes-detector network (HLV) based on the second planned observing run in 2016.
+These are the two observing scenarios that will be discussed at length in the following chapter.
+
+Several features are apparent.
+First, at any number of threads, the two configurations have similar run times, although the 2016 events contain a subpopulation of outliers that take about 2.5 times as long as the 2015 events.
+These are probably due to taking one of the more expensive code branches in the distance integral.
+Second, the run times decrease proportionally to the number of threads.
+A slight flattening at 32 threads may be due to increasing use of hyperthreading.
+Based on experiences running \ac{BAYESTAR} on the 32\nobreakdashes-core (64 threads with hyperthreading) cluster login machine, we expect the almost ideal parallel speedup to continue on machines with more threads.
+With just one thread, the \ac{BAYESTAR} analysis takes 500\nobreakdashes--1500~s, already orders of magnitude faster than the full parameter estimation.
+With 32 threads, \ac{BAYESTAR} takes just 20\nobreakdashes--60~s.
+This latency is comparable to the other stages (data aggregation, trigger generation, alert distribution) in the real\nobreakdashes-time \ac{BNS} analysis.
+The 32\nobreakdashes-thread configuration is representative of how \ac{BAYESTAR} might be deployed in early Advanced \ac{LIGO}.\footnote{\ac{BAYESTAR} has been successfully ported to the Intel's Many Integrated Core~(MIC) architecture and has been tested in a 500 thread configuration on a system with dual Intel Xeon Phi 5110P coprocessors.}
 
 \begin{figure}
-    \begin{center}
-        \includegraphics{pp}
-    \end{center}
-    \caption{\label{fig:pp}$P-P$ plot.}
+    \centering
+    \includegraphics{runtimes}
+    \caption[\acs{BAYESTAR} run time]{\label{fig:runtimes}Violin plot of \ac{BAYESTAR} run times as the number of OpenMP threads is varied from 1 to 32.
+The 2015 scenario is shown in red and the 2016 scenario in blue.}
 \end{figure}
 
+\section{Case study}
+\label{sec:case-study}
+
+We have completed our description of the \ac{BAYESTAR} algorithm.
+In the following chapter, we will describe the first results with \ac{BAYESTAR}, a detailed case study of the sky localization capabilities of the earliest configurations of Advanced \ac{LIGO} and Virgo.
+Because it is typically three orders of magnitude faster than the conventional \ac{MCMC} analysis, for the first time we are able to generate detailed sky maps for thousands of simulated \ac{GW} sources.
+With such a large sample, we can for the first time unambiguously describe areas and morphologies that should be typical of networks of just two detectors.
+This case study will be the subject of the next chapter.
+
+\input{ligo-acronyms/acronyms}
+
 \begin{acknowledgements}
-\marginpar{This URL is not specific enough.}Source code for \ac{BAYESTAR} is available on the \ac{LIGO} \acl{DASWG} web site at \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/lalsuite.html}.
+Source code for \ac{BAYESTAR} is available as part of LALInference, the open source LIGO/Virgo parameter estimation toolchain.
+See \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/lalsuite.html} for further details.
 
@@
 
-\ac{LIGO} was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the \ac{NSF} and operates under cooperative agreement PHY\nobreakdashes-0107417.
+LIGO was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the \ac{NSF} and operates under cooperative agreement PHY\nobreakdashes-0107417.
 Some results were produced on the NEMO computing cluster operated by the Center for Gravitation and Cosmology at University of Wisconsin\nobreakdashes--Milwaukee under \ac{NSF} Grants PHY\nobreakdashes-0923409 and PHY\nobreakdashes-0600953.
 This research is supported by the \ac{NSF} through a Graduate Research Fellowship to L.S.
-This paper has \ac{LIGO} Document Number \ac{LIGO}\nobreakdashes-PXXXXXXX\nobreakdashes-vX.
+This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v1.
+
+We thank Britt Griswold for assistance with preparing Figure~\ref{fig:adaptive-illustration}.
 \end{acknowledgements}
 
+\appendix
+
+\section{Independence of intrinsic and extrinsic errors}
+\label{sec:independence-intrinsic-extrinsic}
+
+If all of the detectors have the same noise \acp{PSD} up to multiplicative factors, $c_1 S_1(\omega) = c_2 S_2(\omega) = \cdots = c_n S_n(\omega) \equiv S(\omega)$, then we can show that the errors in the intrinsic parameters (masses) are not correlated with sky position errors.
+This is because we can change variables from amplitudes, phases, and times to amplitude ratios, phase differences, and time differences.
+With $N$ detectors, we can form a single average amplitude, time, and phase, plus $N-1$ linearly independent differences.
+The averages are correlated with the intrinsic parameters, but neither are correlated with the differences.
+Since only the differences inform sky location, this gives us license to neglect uncertainty in masses when we are computing sky resolution.
+
+This is easiest to see if we make the temporary change of variables $\rho \rightarrow \varsigma = \log \rho$.
+This allows us to factor out the \ac{SNR} dependence from the single-detector Fisher matrix.
+The extrinsic part becomes
+%
+\begin{align}\label{eq:fisher-matrix}
+    \mathcal{I}_{\bm\theta_i,\bm\theta_i} &= \bordermatrix{
+        ~ & \varsigma_i & \gamma_i & \tau_i \cr
+        \varsigma_i & {\rho_i}^2 & 0 & 0 \cr
+        \gamma_i & 0 & {\rho_i}^2 & -{\rho_i}^2 {\overline{\omega}}_i \cr
+        \tau_i & 0 & -{\rho_i}^2 {\overline{\omega}}_i & {\rho_i}^2 {\overline{\omega^2}}_i
+    }
+    \\
+    \nonumber
+    &= {\rho_i}^2 \left( \begin{array}{ccc}
+        1 & 0 & 0 \\
+        0 & 1 & -{\overline{\omega}}_i \\
+        0 & -{\overline{\omega}}_i & {\overline{\omega^2}}_i
+    \end{array} \right).
+\end{align}
+%
+Due to our assumption that the detectors' \acp{PSD} are proportional to each other, the noise moments are the same for all detectors, ${\overline{\omega^k}}_i \equiv {\overline{\omega^k}}$.
+Then we can write the single-detector Fisher matrix as
+%
+\begin{equation}
+    \mathcal{I}_i = {\rho_i}^2 \left(
+        \begin{array}{cc}
+            A & B \\
+            B\transpose & C
+        \end{array}
+    \right),
+\end{equation}
+%
+with the top-left block $A$ comprising the extrinsic parameters and the bottom-right block $C$ the intrinsic parameters.
+
+Information is additive, so the Fisher matrix for the whole detector network is
+%
+\begin{equation}
+    \mathcal{I}_\mathrm{net} = \left(
+    \begin{array}{ccccc}
+    {\rho_1}^2 A & 0 & \cdots & 0 & {\rho_1}^2 B \\
+    0 & {\rho_2}^2 A & & \vdots & {\rho_1}^2 B \\
+    \vdots & & \ddots & 0 & \vdots \\
+    0 & 0 & \cdots & {\rho_N}^2 A & {\rho_N}^2 B \\
+    {\rho_1}^2 B\transpose & {\rho_2}^2 B\transpose & \cdots & {\rho_N}^2 B\transpose & {\rho_\mathrm{net}}^2 C
+    \end{array}
+    \right).
+\end{equation}
+%
+Now we introduce the change of variables that sacrifices the $N$th detector's extrinsic parameters for the network averages,
+%
+\begin{align}
+    &\begin{array}{c@{\quad\rightarrow\quad}l}
+    \varsigma_N & \overline{\varsigma} = \left(\sum_i {\rho_i}^2 \varsigma_i\right) / {\rho_\mathrm{net}}^2, \\
+    \gamma_N & \overline{\gamma} = \left(\sum_i {\rho_i}^2 \gamma_i\right) / {\rho_\mathrm{net}}^2, \\
+    \tau_N & \overline{\tau} = \left(\sum_i {\rho_i}^2 \tau_i\right) / {\rho_\mathrm{net}}^2, \\
+    \end{array}
+%
+\intertext{and replaces the first $N-1$ detectors' extrinsic parameters with differences,}
+%
+    &\left.
+    \begin{array}{c@{\quad\rightarrow\quad}l}
+    \varsigma_i & \delta\varsigma_i = \varsigma_i - \overline{\varsigma} \\
+    \gamma_i & \delta\gamma_i = \gamma_i - \overline{\gamma} \\
+    \tau_i & \delta\tau_i = \tau_i - \overline{\tau} \\
+    \end{array}
+    \right\} \text{ for } i = 1, \dots, N - 1.
+\end{align}
+%
+The Jacobian matrix that describes this change of variables is
+%
+\begin{equation}
+    J = \left(
+    \begin{array}{cccccc}
+    1 & 0 & \cdots & 0 & 1 & 0 \\
+    0 & 1 & & 0 & 1 & 0 \\
+    \vdots & & \ddots & & \vdots & \vdots \\
+    0 & 0 & \cdots & 1 & 1 & 0 \\
+    \frac{-{\rho_1}^2}{{\rho_N}^2} & \frac{-{\rho_2}^2}{{\rho_N}^2} & \cdots & \frac{-{\rho_{N-1}}^2}{{\rho_N}^2} & 1 & 0 \\
+    0 & 0 & \cdots & 0 & 0 & 1 \\
+    \end{array}
+    \right).
+\end{equation}
+%
+The transformed network Fisher matrix is block diagonal,
+%
+\begin{widetext}
+\begin{equation}
+    \mathcal{I}_\mathrm{net} \rightarrow J\transpose \mathcal{I}_\mathrm{net} J = \left(
+    \begin{array}{cccccc}
+    {\rho_1}^2(1+\frac{1}{{\rho_1}^4}) A & \frac{{\rho_1}^2 {\rho_2}^2}{{\rho_N}^2} A & \cdots & \frac{{\rho_1}^2 {\rho_{N-1}}^2}{{\rho_N}^2} A & 0 & 0 \\
+    \frac{{\rho_1}^2 {\rho_2}^2}{{\rho_N}^2} A & {\rho_2}^2(1+\frac{1}{{\rho_1}^4}) A & & \frac{{\rho_2}^2 {\rho_{N-1}}^2}{{\rho_N}^2} A & 0 & 0 \\
+    \vdots & & \ddots & \vdots & \vdots & \vdots \\
+    \frac{{\rho_1}^2 {\rho_{N-1}}^2}{{\rho_N}^2} A & \frac{{\rho_2}^2 {\rho_{N-1}}^2}{{\rho_N}^2} A & \cdots & {\rho_{N-1}}^2(1+\frac{1}{{\rho_1}^4}) A & 0 & 0 \\
+    0 & 0 & \cdots & 0 & {\rho_\mathrm{net}}^2 A & {\rho_\mathrm{net}}^2 B \\
+    0 & 0 & \cdots & 0 & {\rho_\mathrm{net}}^2 B\transpose & {\rho_\mathrm{net}}^2 C \\
+    \end{array}
+    \right).
+\end{equation}
+\end{widetext}
+%
+The top\nobreakdashes-left block contains $N-1$ relative amplitudes, phases, and times on arrival, all potentially correlated with each other.
+The bottom\nobreakdashes-right block contains the average amplitudes, phases, and times, as well as the masses.
+The averages and the masses are correlated with each other, but are not correlated with the differences.
+Because only the differences are informative for sky localization, we drop the intrinsic parameters from the rest of the Fisher matrix calculations in this chapter.
+
+\section{Interpretation of phase and time errors}
+
+We take a brief digression to discuss the physical interpretation of the time and amplitude errors. For our likelihood, the \ac{CRLB} implies that
+%
+\begin{equation}\label{eq:covariance-matrix}
+    \cov{
+        \left(
+        \begin{array}{c}
+            \tilde{\rho}_i \\
+            \tilde{\gamma}_i \\
+            \tilde{\tau}_i
+        \end{array}
+        \right)
+    } \geq \mathcal{I}^{-1} = \left(
+        \begin{array}{ccc}
+            1 & 0 & 0 \\
+            0 & {\rho_i}^2 {\overline{\omega^2}}_i/{\omega_{\mathrm{rms},i}}^2 & {\rho_i}^2 {\overline{\omega}}_i/{\omega_{\mathrm{rms},i}}^2 \\
+            0 & {\rho_i}^2 {\overline{\omega}}_i/{\omega_{\mathrm{rms},i}}^2 & {\rho_i}^2/{\omega_{\mathrm{rms},i}}^2
+        \end{array}
+    \right),
+\end{equation}
+%
+where ${\omega_{\mathrm{rms},i}}^2 = {\overline{\omega^2}}_i - {{\overline{\omega}}_i}^2$.
+Reading off the $\tau \tau$ element of the covariance matrix reproduces the timing accuracy in Equation~(24) of \cite{FairhurstTriangulation},
+%
+\begin{equation}\label{eq:timing-crlb}
+    \std \left(\hat{\tau}_i - \tau_i \right) \geq \sqrt{\left(\mathcal{I}^{-1}\right)_{\tau\tau}} = \frac{\rho_i}{\omega_{\mathrm{rms},i}}.
+\end{equation}
+
+The Fisher matrix in Equation~(\ref{eq:fisher-matrix}) is block diagonal, which implies that estimation errors in the signal amplitude $\rho$ are uncorrelated with the phase $\gamma$ and time $\tau$.
+A sequence of two changes of variables lends some physical interpretation to the nature of the coupled estimation errors in $\gamma$~and~$\tau$.
+
+First, we put the phase and time on the same footing by measuring the time in units of $1 / \sqrt{\overline{\omega^2}}$ with a change of variables from $\tau$ to $\gamma_\tau = \sqrt{\overline{\omega^2}} \tau$:
+%
+\begin{equation}
+    \mathcal{I}' = \bordermatrix{
+        ~ & \rho_i & \gamma_i & \gamma_{\tau,i} \cr
+        \rho_i & 1 & 0 & 0 \cr
+        \gamma_i & 0 & {\rho_i}^2 & -{\rho_i}^2\frac{{\overline{\omega}}_i}{\sqrt{{\overline{\omega^2}}_i}} \cr
+        \gamma_{\tau,i} & 0 & -{\rho_i}^2\frac{{\overline{\omega}}_i}{\sqrt{{\overline{\omega^2}}_i}} & {\rho_i}^2
+    }.
+\end{equation}
+%
+The second change of variables, from $\gamma$ and $\gamma_\tau$ to $\gamma_\pm = \frac{1}{\sqrt{2}}(\gamma \pm \gamma_\tau)$, diagonalizes the Fisher matrix:
+%
+\begin{equation}\label{eq:fisher-matrix-extrinsic-one-detector}
+    \mathcal{I}'' = \bordermatrix{
+        ~ & \rho_i & \gamma_{+,i} & \gamma_{-,i} \cr
+        \rho_i & 1 & 0 & 0 \cr
+        \gamma_{+,i} & 0 & \left(1 - \frac{\overline{\omega}_i}{\sqrt{\overline{\omega^2}_i}}\right){\rho_i}^2 & 0 \cr
+        \gamma_{-,i} & 0 & 0 & \left(1 + \frac{\overline{\omega}_i}{\sqrt{\overline{\omega^2}_i}}\right){\rho_i}^2
+    }.
+\end{equation}
+%
+Thus, in the appopriate time units, the \textit{sum and difference} of the phase and time of the signal are measured independently.
+
+\section{Position resolution}
+
+Finally, we will calculate the position resolution of a network of \ac{GW} detectors.
+We could launch directly into computing derivatives of the full signal model from Equation~(\ref{eq:full-signal-model}) with respect to all of the parameters, but this would result in a very complicated expression.
+Fortunately, we can take two shortcuts.
+First, since we showed in Section~\ref{sec:independence-intrinsic-extrinsic} that the intrinsic parameters are correlated only with an overall nuisance average arrival time, amplitude, and phase, we need not consider the derivatives with respect to mass at all.
+Second, we can reuse the extrinsic part of the single detector Fisher matrix from Equation~(\ref{eq:fisher-matrix}) by computing the much simpler Jacobian matrix to transform from the time, amplitude, and phase on arrival, to the parameters of interest.
+
+We begin by transforming the single\nobreakdashes-detector Fisher matrix from a polar to a rectangular representation of the complex amplitude given in Equations~(\ref{eq:optimal-rho}, \ref{eq:optimal-tau}), $\rho_i, \gamma_i \rightarrow \Re[z_i] = \rho_i \cos \gamma_i, \Im[z_i] = \rho_i \sin \gamma_i$:
+%
+\begin{equation}\label{eq:fisher-matrix-extrinsic-one-detector-cartesian}
+    \mathcal{I}_i = \bordermatrix{
+        ~ & \Re[z_i] & \Im[z_i] & \tau_i \cr
+        %
+        \Re[z_i] &
+        1 &
+        0 &
+        {\overline{\omega}}_i b_i \cr
+        %
+        \Im[z_i] &
+        0 &
+        1 &
+        -{\overline{\omega}}_i b_i \cr
+        %
+        \tau_i &
+        {\overline{\omega}}_i b_i &
+        -{\overline{\omega}}_i b_i &
+        {\rho_i}^2 {\overline{\omega^2}}_i
+    }.
+\end{equation}
+
+Consider a source in a ``standard'' orientation with the direction of propagation along the $+z$ axis, such that the \ac{GW} polarization tensor may be written in Cartesian coordinates as
+%
+\begin{equation}
+    H = \frac{1}{r} e^{2 i \phi_c} \left(
+    \begin{array}{ccc}
+        \frac{1}{2}(1 + \cos^2 \iota) & i \cos\iota & 0 \\
+        i \cos\iota & -\frac{1}{2}(1 + \cos^2 \iota) & 0 \\
+        0 & 0 & 0
+    \end{array}
+    \right).
+\end{equation}
+%
+Now introduce a rotation matrix $R$ that actively transforms this source to the Earth-relative polar coordinates $\theta, \phi$, and gives the source a polarization angle $\psi$ (adopting temporarily the notation $\mathrm{c}_\theta = \cos\theta,\,\mathrm{s}_\theta = \sin\theta$):
+%
+\begin{widetext}
+\begin{equation}
+    R = R_z(\phi) R_y(\theta) R_z(\psi) R_y(\pi)
+    = \left(
+        \begin{array}{ccc}
+            \mathrm{c}_\phi & -\mathrm{s}_\phi & 0 \\
+            \mathrm{s}_\phi & -\mathrm{c}_\phi & 0 \\
+            0 & 0 & 1
+        \end{array}
+    \right)
+    \left(
+        \begin{array}{ccc}
+            \mathrm{c}_\theta & 0 & \mathrm{s}_\theta \\
+            0 & 1 & 0 \\
+            -\mathrm{s}_\theta & 0 & \mathrm{c}_\theta
+        \end{array}
+    \right)
+    \left(
+        \begin{array}{ccc}
+            \mathrm{c}_\psi & -\mathrm{s}_\psi & 0 \\
+            \mathrm{s}_\psi & -\mathrm{c}_\psi & 0 \\
+            0 & 0 & 1
+        \end{array}
+    \right)
+    \left(
+        \begin{array}{ccc}
+            -1 & 0 & 0 \\
+            0 & 1 & 0 \\
+            0 & 0 & -1
+        \end{array}
+    \right).
+\end{equation}
+\end{widetext}
+%
+(The rightmost rotation reverses the propagation direction so that the wave is traveling \emph{from} the sky position $\theta, \phi$.) With the (symmetric) detector response tensor $D_i$, we can write the received amplitude and arrival time as
+%
+\begin{align}
+    z_i &= r_{1,i} \Tr \left[ D_i R \, H \, R\transpose \right], \\
+    \tau_i &= t_\oplus + {\mathbf{d}_i}\transpose R \, \mathbf{k}.
+\end{align}
+%
+Equivalently, we can absorb the rotation $R$ and the horizon distance $r_{1,i}$ into the polarization tensor, detector response tensors, and positions,
+%
+\begin{align}
+    H &\rightarrow H^\prime = R_z(\psi) \, R_y(\pi) \, H \, R_y(\pi)\transpose  \, R_z(\psi)\transpose, \\
+    D_i &\rightarrow D_i^\prime = r_{1,i} \, R_y(\theta)\transpose \, R_z(\phi)\transpose \, D_i \, R_z(\phi) \, R_y(\theta), \\
+    \mathbf{d}_i &\rightarrow \mathbf{d}_i^\prime = R_y(\theta)\transpose \, R_z(\phi)\transpose \, \mathbf{d}_i, \\
+    \mathbf{k} &\rightarrow \mathbf{k}^\prime = (0, 0, -1).
+\end{align}
+%
+Now the model becomes
+%
+\begin{align}
+    \label{eq:H-hplus-hcross}
+    H^\prime &= \left(
+        \begin{array}{ccc}
+            h_+ & h_\times & 0 \\
+            h_\times & -h_+ & 0 \\
+            0 & 0 & 0
+        \end{array}
+    \right), \\
+    \label{eq:z-hplus-hcross}
+    z_i &= \Tr \left[ D_i^\prime H^\prime \right] = h_+ (D_{00}^\prime - D_{11}^\prime) + 2 h_\times D_{01}^\prime, \\
+    \tau_i &= t_\oplus + (\mathbf{d}_i^\prime) \cdot \mathbf{k}, \\
+    %
+    \intertext{where}
+    %
+    \label{eq:hplus}
+    h_+ &= \frac{1}{r} e^{2 i \phi_c} \left[\frac{1}{2} (1+\cos^2\iota) \cos{2\psi} + i \cos\iota \sin{2\psi}\right], \\
+    \label{eq:hcross}
+    h_\times &= \frac{1}{r} e^{2 i \phi_c} \left[\frac{1}{2} (1+\cos^2\iota) \sin{2\psi} - i \cos\iota \cos{2\psi}\right].
+\end{align}
+%
+We insert an infinitesimal rotation $\delta R$ to perturb the source's orientation from the true value:
+%
+\begin{align}
+    z_i &= \Tr \left[ D_i^\prime (\delta R) H^\prime (\delta R)\transpose \right], \\
+    \tau_i &= t_\oplus + (\mathbf{d}_i^\prime)\transpose (\delta R) \mathbf{k}^\prime.
+\end{align}
+%
+We only need a first order expression for $\delta R$, because we will be taking products of first derivatives of it\footnote{Caution: the angles $\delta\theta$ and $\delta\phi$ represent displacements in two orthogonal directions, but are \emph{not} necessarily simply related to $\theta$ and $\phi$.}:
+%
+\begin{equation}
+    \delta R = \left(
+        \begin{array}{ccc}
+            1 & 0 & \delta\theta \\
+            0 & 1 & \delta\phi \\
+            -\delta\theta & -\delta\phi & 1
+        \end{array}
+    \right).
+\end{equation}
+%
+We construct a Jacobian matrix $J_i$ to transform from the single\nobreakdashes-detector observables $(\Re[z_i], \Im[z_i], \tau_i)$ to the position perturbations, polarization components, and geocentered arrival time \\ $(\delta\theta, \delta\phi, \Re[h_+], \Im[h_+], \Re[h_\times], \Im[h_\times], t_\oplus)$:
+%
+\begin{widetext}
+\begin{equation}\label{eq:jacobian-detector-tensor}
+    {J_i}\transpose = \bordermatrix{
+        ~ & \Re[z_i] & \Im[z_i] & \tau_i \cr
+        %
+        \delta\theta &
+        -2\Re[h_+]D_{02}^\prime -2\Re[h_\times]D_{12}^\prime &
+        -2\Im[h_+]D_{02}^\prime -2\Im[h_\times]D_{12}^\prime &
+        -d_0^\prime \cr
+        %
+        \delta\phi &
+        -2\Re[h_\times]D_{02}^\prime +2\Re[h_+]D_{12}^\prime &
+        -2\Im[h_\times]D_{02}^\prime +2\Im[h_+]D_{12}^\prime &
+        -d_1^\prime \cr
+        %
+        \Re[h_+] & D_{00}^\prime - D_{11}^\prime & 0 & 0 \cr
+        %
+        \Im[h_+] & 0 & D_{00}^\prime - D_{11}^\prime & 0 \cr
+        %
+        \Re[h_\times] & 2D_{01}^\prime & 0 & 0 \cr
+        %
+        \Im[h_\times] & 0 & 2D_{01}^\prime & 0 \cr
+        %
+        t_\oplus & 0 & 0 & 1
+    }.
+\end{equation}
+\end{widetext}
+%
+We transform and sum the information from each detector,
+%
+\begin{equation}\label{eq:summed-information}
+    \mathcal{I}_\mathrm{net} = \sum_i {J_i}\transpose \mathcal{I}_i {J_i}.
+\end{equation}
+
+\subsection{Marginalization over nuisance parameters}
+
+To extract an area from the Fisher matrix, we must first marginalize or discard the nuisance parameters.
+Note that marginalizing parameters of a multivariate Gaussian distribution amounts to simply dropping the relevant entries in the mean vector and covariance matrix.
+Since the information is the inverse of the covariance matrix, we need to invert the Fisher matrix, drop all but the first two rows and columns, and then invert again.
+
+This procedure has a shortcut called the Schur complement (see, for example, \citealt{numerical-recipes-inversion-partition}). Consider a partitioned square matrix $M$ and its inverse:
+%
+\begin{equation}
+    M = \left(
+        \begin{array}{cc}
+            A & B \\
+            C & D
+        \end{array}
+    \right), \qquad
+    M^{-1} = \left(
+        \begin{array}{cc}
+            \tilde{A} & \tilde{B} \\
+            \tilde{C} & \tilde{D}
+        \end{array}
+    \right).
+\end{equation}
+%
+If $A$ and $B$ are square matrices, then the upper-left block of the inverse can be written as
+%
+\begin{equation}
+    \tilde{A}^{-1} = A - B D^{-1} C.
+\end{equation}
+%
+If we partition the $\mathcal{I}_\mathrm{net}$ similarly, the $A$ block consists of the first two rows and columns and $D$ is the lower right block that describes all other parameters.
+Because the Fisher matrix is symmetric, the off\nobreakdashes-diagonal blocks satisfy $C = B\transpose$.
+Then the Schur complement
+%
+\begin{equation}\label{eq:marginal-fisher-matrix}
+    \mathcal{I}_\mathrm{marg} = A - B D^{-1} B\transpose
+\end{equation}
+%
+gives us the information matrix marginalized over all parameters but $\delta\theta$ and $\delta\phi$.
+
+\subsection{Spatial interpretation}
+
+How do we extract the dimensions of the localization from the Fisher matrix?
+If there are $N \leq 2$ detectors, then the Fisher matrix must be degenerate, because there are $3N$ measurements and 7 parameters:
+%
+\begin{equation*}
+    \left\{
+    \begin{array}{c}
+        \delta\theta \\
+        \delta\phi \\
+        \Re[h_+] \\
+        \Im[h_+] \\
+        \Re[h_\times] \\
+        \Im[h_\times] \\
+        t_\oplus
+    \end{array}
+    \right\}
+    = 7 \text{ parameters}
+    \qquad
+    \longleftrightarrow
+    \qquad
+    \left\{
+    \begin{array}{c}
+        \Re[z_i] \\
+        \Im[z_i] \\
+        \tau_i
+    \end{array}
+    \right\}
+    \times N = 3N \text{ observables.}
+\end{equation*}
+%
+Therefore, for $N=2$ detectors, the marginalized Fisher matrix $\mathcal{I}_\mathrm{marg}$ is singular.
+Its only nonzero eigenvalue $\lambda$ describes the width of an annulus on the sky.
+The width of the annulus that contains probability $p$ is given by
+%
+\begin{equation}\label{eq:l-pth-quantile}
+    L_p = 2 \sqrt{2} \erf^{-1}(p) / \sqrt{\lambda}.
+\end{equation}
+%
+The prefactor $2 \sqrt{2} / \erf^{-1}(p)$ is the central interval of a normal distribution that contains a probability $p$, and is $\approx 3.3$ for $p = 0.9$.
+\emph{Caution:} for two\nobreakdashes-detector networks, priors play an important role in practical parameter estimation and areas can be much smaller than one would predict from the Fisher matrix (see Chapter~\ref{chap:first2years} for more discussion).
+
+For $N \geq 3$ detectors, the parameters are over\nobreakdashes-constrained by the data and the Fisher matrix describes the dimensions of an ellipse.
+Within a circle of radius $r$ centered on the origin, the enclosed probability $p$ is
+%
+\begin{equation}
+    p = \int_0^{2\pi} \int_0^r \frac{1}{2\pi} e^{-s^2/2} s \, ds \, d\phi = 1 - e^{-r^2/2}.
+\end{equation}
+%
+Therefore the radius $r$ of the circle that contains a probability $p$ is
+%
+\begin{equation}
+    r = \sqrt{-2\ln (1 - p)}.
+\end{equation}
+%
+Suppose that the eigenvalues of the Fisher matrix are $\lambda_1$ and $\lambda_2$.
+This describes a $1\sigma$ uncertainty ellipse that has major and minor radii ${\lambda_1}^{-1/2}$, ${\lambda_2}^{-1/2}$, and area $A_{1\sigma} = \pi / \sqrt{\lambda_1 \lambda_2} = \pi / \sqrt{\det\mathcal{I}}$.
+Then the area of an ellipse containing probability $p$ is
+%
+\begin{equation}\label{eq:a-pth-quantile}
+    A_p = -2\pi\ln(1-p) / \sqrt{\det \mathcal{I}},
+\end{equation}
+%
+or, more memorably for the $90$th percentile, $A_{0.9} = 2\pi\ln(10) / \sqrt{\det{\mathcal{I}}}$.
+
+\subsection{Outline of calculation}
+\label{sec:fisher-matrix-area-outline}
+
+Using the above derivation, we arrive at a prediction for the sky resolution of a \ac{GW} detector network.
+We took some shortcuts that allowed us to avoid directly evaluating the complicated derivatives of the signal itself with respect to sky location.
+As a result, the expressions involved in each step are simple enough to be manually entered into a computer program.
+However, because the procedure involves several steps, we outline it once again below.
+
+\begin{enumerate}
+    \item Compute, for each detector, the horizon distance $r_{1,i}$, the angular frequency moments ${\overline{\omega}}_i$ and ${\overline{\omega^2}}_i$, and $(h_+, h_\times)$ from Equations~(\ref{eq:hplus},~\ref{eq:hcross}).
+(These can be reused for multiple source positions as long as the masses and the detector noise \acp{PSD} are the same.)
+    \item For a given $\phi, \theta, \psi$, compute the complex received amplitude $z_i$ from Equations~(\ref{eq:H-hplus-hcross},~\ref{eq:z-hplus-hcross}), the extrinsic Fisher matrix from Equation~(\ref{eq:fisher-matrix}), and the Jacobian from Equation~(\ref{eq:jacobian-detector-tensor}).
+    \item Sum the information from all detectors using Equation~\ref{eq:summed-information}.
+    \item Compute the marginalized Fisher matrix from the Schur complement using Equation~(\ref{eq:marginal-fisher-matrix}).
+    \item If there are two detectors, find the width $L_p$ of the ring describing the $p$th quantile using Equation~(\ref{eq:l-pth-quantile}).
+If there are three or more detectors, find the area $A_p$ of the $p$th quantile using Equation~(\ref{eq:a-pth-quantile}).
+    \item (Optionally, convert from (ste)radians to (square) degrees.)
+\end{enumerate}
+
+See code listing in Appendix~\ref{sec:fisher-matrix-code-listing}.
+
 \bibliography{apj-jour,bib/telescope,ms}

### 2015-04-16 importance-sampling.py
Add importance sampling figure

@@ -1,0 +1,100 @@
+#!/usr/bin/env python
+import matplotlib
+matplotlib.rcParams['text.usetex'] = False
+import seaborn
+from matplotlib import pyplot as plt
+from matplotlib.gridspec import GridSpec
+import numpy as np
+import scipy.stats
+import scipy.special
+seaborn.set_style('white')
+
+gcolor, fcolor = seaborn.color_palette('deep', 2)
+
+
+def Q(x):
+    return scipy.stats.norm.sf(x)
+
+nsamp = 500
+
+r0 = 0.33
+p = 1.07
+nrmin = 130
+nrmax=-240
+b = 2 * np.square(p) / r0
+
+r = np.linspace(0, 1, nsamp)
+diff = p/r - 0.5*b/p
+diff2 = np.square(diff)
+
+f = np.exp(-diff2) * np.square(r/r0)
+g = np.exp(-diff2) / np.square(r / r0)
+G = np.sqrt(np.pi) / p * Q(np.sqrt(2) * diff)
+fg = scipy.special.i0e(b / r) * np.square(r / r0) * np.square(r)
+
+# nrmin = 110
+# nrmax = -200
+rmin = r[nrmin]
+rmax = r[nrmax]
+Gmin = G[nrmin]
+Gmax = G[nrmax]
+
+common_line_style = {'color': 'black', 'linewidth': 1}
+connecting_common_line_style = {'color': 'black', 'linewidth': 1}
+
+fig = plt.figure(figsize=(5, 5))
+gs = GridSpec(2, 2)
+ax0 = fig.add_subplot(gs[0, 0])
+ax1 = fig.add_subplot(gs[1, 0], sharex=ax0)
+ax2 = fig.add_subplot(gs[1, 1], sharey=ax1)
+ax3 = fig.add_subplot(gs[0, 1], sharex=ax2, sharey=ax0, frameon=False)
+
+ax0.fill_between(r[:nrmin], f[:nrmin] / np.nanmax(f), alpha=0.5, color=fcolor)
+ax0.fill_between(r[nrmax:], f[nrmax:] / np.nanmax(f), alpha=0.5, color=fcolor)
+ax0.fill_between(r[nrmin:nrmax], f[nrmin:nrmax] / np.nanmax(f), color=fcolor)
+ax0.plot(r, g / np.nanmax(g), color=gcolor, alpha=0.5)
[…]

### 2015-04-16 ms.tex
Try to weave appendices into the paper better

@@
 \usepackage{protosem}
+
+\usepackage{listings}
+\lstset{ %
+    basicstyle=\footnotesize\ttfamily,
+    showspaces=false,
+    showstringspaces=false,
+    showtabs=false,
+    keepspaces,
+    frame=tb,
+    commentstyle=\color{red},
+    keywordstyle=\color{blue}
+}
 
@@
 
-\section{Uncertainty and the Fisher matrix}
+\subsection{Uncertainty and the Fisher matrix}
 \label{sec:fisher}
 
-We can predict the uncertainty in the \ac{ML} estimates without working out its full distribution.
-The \ac{CRLB} gives its covariance in the asymptotic limit of high \ac{SNR}.
+We can predict the uncertainty in the detection pipeline's \ac{ML} estimates using the \ac{CRLB}.
 The \ac{CRLB} has been widely applied in \ac{GW} data analysis to estimate parameter estimation uncertainty (for example, \cite{1996PhRvD..53.3033B,FairhurstTriangulation,2009PhRvD..79h4032A,WenLocalizationAdvancedLIGO,LIGOObservingScenarios,FairhurstLIGOIndia})\footnote{The Fisher matrix is also used in construction of \ac{CBC} matched filter banks.
@@
 This is equivalent to uniformly sampling the Jeffreys prior.
-In practice, this is done either by constructing a hexagonal lattice \citep{PhysRevD.76.102004} or sampling stochastically \citep{2009PhRvD..80j4014H,2009PhRvD..80b4009V,SBank,2010PhRvD..81b4004M,2014PhRvD..89b4003P}}.
+In practice, this is done either by constructing a hexagonal lattice \citep{PhysRevD.76.102004} or sampling stochastically \citep{2009PhRvD..80j4014H,2009PhRvD..80b4009V,SBank,2010PhRvD..81b4004M,2014PhRvD..89b4003P}.}.
+
 We will momentarily consider the likelihood for a single detector:
@@
 \end{equation}
-%
+
+The Fisher matrix describes how strongly the likelihood depends, on average, on the parameters.
+Furthermore, it provides an estimate of the mean-square error in the parameters.
+If $\hat{\bm\theta}$ is an unbiased estimator of $\bm\theta$, $\tilde{\bm\theta} = \hat{\bm\theta} - \bm\theta$ is the measurement error, and $\Sigma = \mathrm{E} \, [\tilde{\bm\theta}\tilde{\bm\theta}\transpose]$ is the covariance of the measurement error, then the \ac{CRLB} says that $\Sigma \geq \mathcal{I}^{-1}$, in the sense that $\left(\Sigma - \mathcal{I}^{-1}\right)$ is positive semi\nobreakdashes-definite.
+
 Note that if $\log\mathcal{L}$ is twice differentiable in terms of $\bm\theta$, then the Fisher matrix can also be written in terms of second derivatives as
@@
 \end{equation}
-%
-In this form, we can recognize the Fisher matrix as the expectation value, conditioned upon the true parameter values, of the Hessian matrix of the log likelihood.
-It describes how strongly the likelihood depends, on average, on the parameters.
-If $\hat{\bm\theta}$ is an unbiased estimator of $\bm\theta$, $\tilde{\bm\theta} = \hat{\bm\theta} - \bm\theta$ is the measurement error, and $\Sigma = \mathrm{E} \, [\tilde{\bm\theta}\tilde{\bm\theta}\transpose]$ is the covariance of the measurement error, then the \ac{CRLB} says that $\Sigma \geq \mathcal{I}^{-1}$, in the sense that $\left(\Sigma - \mathcal{I}^{-1}\right)$ is positive semi\nobreakdashes-definite.
 
@@
 \begin{equation}
+    \label{eq:single-detector-block-fisher-matrix}
     \mathcal{I}_i = \left(
@@
         \end{array}
-    \right)
-\end{equation}
-%
-where
+    \right).
+\end{equation}
+%
+The top-left block describes only the extrinsic parameters, and is given by
 %
@@
 %
-(This is equivalent to an expression given in \citealt{Grover:2013}.) The information matrix elements that relate to the intrinsic parameters can also be expressed as linear combinations of the angular frequency moments.
-However, as we will see in the next section, we need not compute these matrix elements if we are only interested in sky localization accuracy.
+(This is equivalent to an expression given in \cite{Grover:2013}.) The bottom row and right column of Equation~(\ref{eq:single-detector-block-fisher-matrix}) describe the intrinsic parameters and how they are coupled to the extrinsic parameters.
+We show in Appendix~\ref{sec:independence-intrinsic-extrinsic} that we need not consider the intrinsic parameters at all if we are concerned only with sky localization.
+
+The Fisher matrix has been much maligned within the \ac{GW} literature because the \ac{CRLB} is a good description of the parameter covariance in the limit of high \ac{SNR}, and can be abused by applying it in low-\ac{SNR} conditions, or close to the boundaries of the parameter space \cite{UseAbuseFisherMatrix}.
+However, we will put the \ac{CRLB} to good use to improve our qualitative understanding of the likelihood.
+Furthermore, the Fisher matrix in its own right---independent of its suitability to describe the parameter covariance---is a well-defined property of any likelihood function, and we will exploit it as such in Section~\ref{sec:bayestar-likelihood}.
+
+For our likelihood, the \ac{CRLB} implies that
+%
+\begin{equation}\label{eq:covariance-matrix}
+    \cov{
+        \left(
+        \begin{array}{c}
+            \tilde{\rho}_i \\
+            \tilde{\gamma}_i \\
+            \tilde{\tau}_i
+        \end{array}
+        \right)
+    } \geq \mathcal{I}^{-1} = \left(
+        \begin{array}{ccc}
+            1 & 0 & 0 \\
+            0 & {\rho_i}^2 {\overline{\omega^2}}_i/{\omega_{\mathrm{rms},i}}^2 & {\rho_i}^2 {\overline{\omega}}_i/{\omega_{\mathrm{rms},i}}^2 \\
+            0 & {\rho_i}^2 {\overline{\omega}}_i/{\omega_{\mathrm{rms},i}}^2 & {\rho_i}^2/{\omega_{\mathrm{rms},i}}^2
+        \end{array}
+    \right),
+\end{equation}
+%
+where ${\omega_{\mathrm{rms},i}}^2 = {\overline{\omega^2}}_i - {{\overline{\omega}}_i}^2$.
+This structure implies that errors in \ac{SNR} are uncorrelated with errors in time and phase, and that there is a particular sum and difference of the times and phases that are measured independently (see Appendix~\ref{sec:interpretation-of-errors}).
+
+Reading off the $\tau \tau$ element of the covariance matrix reproduces the timing accuracy in Equation~(24) of \cite{FairhurstTriangulation},
+%
+\begin{equation}\label{eq:timing-crlb}
+    \std \left(\hat{\tau}_i - \tau_i \right) \geq \sqrt{\left(\mathcal{I}^{-1}\right)_{\tau\tau}} = \frac{\rho_i}{\omega_{\mathrm{rms},i}}.
+\end{equation}
+%
+\citet{FairhurstTriangulation} goes on to frame the characteristic position reconstruction accuracy of a \ac{GW} detector network in terms of time delay triangulation, with the above formula describing the time of arrival uncertainty for each detector.
+In Appendix~\ref{sec:position-resolution}, we show how to extend this formalism to include the phases and amplitudes on arrival as well.
 
@@
 
-\begin{figure*}
-    \includegraphics[width=\textwidth]{illustration}
-    \caption{\label{fig:adaptive-illustration}Illustration of the \ac{BAYESTAR} adaptive mesh refinement scheme.}
-\end{figure*}
-
 The marginal posterior as a function of sky location is
@@
 
+\begin{figure}
+    \includegraphics[width=\columnwidth]{importance-sampling}
+    \caption{\label{fig:importance-sampling}Illustration of importance sampling scheme.
+The top-left panel shows the original integrand $f(r)$, the proposal integrand $g(r)$, and the limits of integration $r = r_\mathrm{min}$ and $r = r_\mathrm{max}$.
+The area of the dark shaded region is the integral of interest.
+The bottom-left panel shows the integral $\mathscr{G}(r) = \int_0^r g(r) dr$.
+The new variable of integration is $G = \mathscr{G}(r)$, from $G_\mathrm{min} = \mathscr{G}(r_\mathrm{min})$ to $G_\mathrm{max} = \mathscr{G}(r_\mathrm{max})$.
+The bottom-right panel shows the transformed integrand, $f(\mathscr{G}^{-1}(G))/g(\mathscr{G}^{-1}(G))$.
+The dark shaded region has the same area as in the top-left figure.}
+\end{figure}
+
 Last, there are a few special cases where Equation~(\ref{eq:distance-integral}) can be evaluated exactly.
@@
 
-Our sampling of the sky is relies completely on the \acl{HEALPix} (\acs{HEALPix}), a special data structure designed for all\nobreakdashes-sky maps.
+\begin{figure*}
+    \includegraphics[width=\textwidth]{illustration}
+    \caption{\label{fig:adaptive-illustration}Illustration of the \ac{BAYESTAR} adaptive HEALPix sampling scheme.}
+\end{figure*}
+
+Our sampling of the sky relies completely on the \acl{HEALPix} (\acs{HEALPix}), a special data structure designed for all\nobreakdashes-sky maps.
 \ac{HEALPix} divides the sky into equal\nobreakdashes-area pixels.
@@
 The red surface shows run times for a two\nobreakdashes-detector network (HL) modeled on the first scheduled Advanced \ac{LIGO} observing run in 2015, and the blue surface shows run times for a three\nobreakdashes-detector network (HLV) based on the second planned observing run in 2016.
-These are the two observing scenarios that will be discussed at length in the following chapter.
+These are the two observing scenarios that are discussed in \cite{FirstTwoYears}.
 
@@
 \section{Interpretation of phase and time errors}
-
-We take a brief digression to discuss the physical interpretation of the time and amplitude errors. For our likelihood, the \ac{CRLB} implies that
-%
-\begin{equation}\label{eq:covariance-matrix}
-    \cov{
-        \left(
-        \begin{array}{c}
-            \tilde{\rho}_i \\
-            \tilde{\gamma}_i \\
-            \tilde{\tau}_i
-        \end{array}
-        \right)
-    } \geq \mathcal{I}^{-1} = \left(
-        \begin{array}{ccc}
-            1 & 0 & 0 \\
-            0 & {\rho_i}^2 {\overline{\omega^2}}_i/{\omega_{\mathrm{rms},i}}^2 & {\rho_i}^2 {\overline{\omega}}_i/{\omega_{\mathrm{rms},i}}^2 \\
-            0 & {\rho_i}^2 {\overline{\omega}}_i/{\omega_{\mathrm{rms},i}}^2 & {\rho_i}^2/{\omega_{\mathrm{rms},i}}^2
-        \end{array}
-    \right),
-\end{equation}
-%
-where ${\omega_{\mathrm{rms},i}}^2 = {\overline{\omega^2}}_i - {{\overline{\omega}}_i}^2$.
-Reading off the $\tau \tau$ element of the covariance matrix reproduces the timing accuracy in Equation~(24) of \cite{FairhurstTriangulation},
-%
-\begin{equation}\label{eq:timing-crlb}
-    \std \left(\hat{\tau}_i - \tau_i \right) \geq \sqrt{\left(\mathcal{I}^{-1}\right)_{\tau\tau}} = \frac{\rho_i}{\omega_{\mathrm{rms},i}}.
-\end{equation}
+\label{sec:interpretation-of-errors}
 
@@
 \section{Position resolution}
+\label{sec:position-resolution}
 
@@
 %
+\begin{widetext}
 \begin{equation*}
@@
 \end{equation*}
+\end{widetext}
 %
@@
 The prefactor $2 \sqrt{2} / \erf^{-1}(p)$ is the central interval of a normal distribution that contains a probability $p$, and is $\approx 3.3$ for $p = 0.9$.
-\emph{Caution:} for two\nobreakdashes-detector networks, priors play an important role in practical parameter estimation and areas can be much smaller than one would predict from the Fisher matrix (see Chapter~\ref{chap:first2years} for more discussion).
+\emph{Caution:} for two\nobreakdashes-detector networks, priors play an important role in practical parameter estimation and areas can be much smaller than one would predict from the Fisher matrix.
 
@@
 
-See code listing in Appendix~\ref{sec:fisher-matrix-code-listing}.
+See code listing in Appendix~A.6 of \cite{leo-singer-thesis}.
 

### 2015-04-24 ms.tex
Clean-up of first few sections

@@
 Although we may not be able to evaluate this equation directly, with some educated guesses we can create a likelihood that has many properties in common with it.
-Any valid approximate likelihood must have the same Fisher matrix as shown in Equation~(\ref{eq:fisher-matrix-extrinsic-one-detector}).
+Any valid approximate likelihood must have the same Fisher matrix as shown in Equation~(\ref{eq:single-detector-block-fisher-matrix}).
 It must also have the same limiting behavior: it should be periodic in the phase error $\tilde{\gamma}_i$ and go to zero as $\tilde{\tau}_i \rightarrow \pm \infty$, $\hat{\rho}_i \rightarrow 0$, or $\hat{\rho}_i \rightarrow \infty$.
@@
 \end{equation}
-
+%
 To assemble the joint likelihood for the whole network, we form the product of the autocorrelation likelihoods from the individual detectors:
@@
 
+In the following section, we discuss some key properties of the autocorreltion likelihood.
+
 \subsection{Properties}
 
-First, observe that at the true parameter values, $\hat{\bm\theta}_i = \bm\theta_i$, the logarithms of Equation~(\ref{eq:autocor-likelihood}) and Equation~(\ref{eq:gaussian-likelihood-spa}) have the same Jacobian.
+First, the autocorrelation likelihood has the elegant feature that if we were to replace the autocorrelation function with the \ac{SNR} time series for the best-matching template, $z(\tau; \hat{\mathbf{\theta}}_\mathrm{in})$, we would recover the likelihood for the full \ac{GW} time series, evaluated at the \ac{ML} estimate of the intrinsic parameters.
+The numerical scheme that we will develop is thus equally applicable for rapid, coincidence-based localization, or as a fast extrinsic marginalization step for the full parameter estimation.
+
+Second, observe that at the true parameter values, $\hat{\bm\theta}_i = \bm\theta_i$, the logarithms of Equation~(\ref{eq:autocor-likelihood}) and Equation~(\ref{eq:gaussian-likelihood-spa}) have the same Jacobian.
 This is because the derivatives of the autocorrelation function are
@@
 
-In the following sections, we describe the numerical implementation of \ac{BAYESTAR}.
+In the following sections, we describe our prior and our numerical scheme to integrate over nuisance parameters, which amounts to the \ac{BAYESTAR} algorithm.
 
@@
 The resulting map is a tree structure that consists of a mix of several resolutions.
+This process is illustrated in Figure~\ref{fig:adaptive-illustration}.
+
 We traverse the tree and flatten it into the highest resolution represented.
@@
 
-\section{Run time}
+\section{Case study}
+\label{sec:case-study}
+
+We have completed our description of the \ac{BAYESTAR} algorithm.
+In the following chapter, we will describe the first results with \ac{BAYESTAR}, a detailed case study of the sky localization capabilities of the earliest configurations of Advanced \ac{LIGO} and Virgo.
+Because it is typically three orders of magnitude faster than the conventional \ac{MCMC} analysis, for the first time we are able to generate detailed sky maps for thousands of simulated \ac{GW} sources.
+With such a large sample, we can for the first time unambiguously describe areas and morphologies that should be typical of networks of just two detectors.
+This case study will be the subject of the next chapter.
+
+\subsection{Run time}
 \label{sec:run-time}
@@
 \end{figure}
-
-\section{Case study}
-\label{sec:case-study}
-
-We have completed our description of the \ac{BAYESTAR} algorithm.
-In the following chapter, we will describe the first results with \ac{BAYESTAR}, a detailed case study of the sky localization capabilities of the earliest configurations of Advanced \ac{LIGO} and Virgo.
-Because it is typically three orders of magnitude faster than the conventional \ac{MCMC} analysis, for the first time we are able to generate detailed sky maps for thousands of simulated \ac{GW} sources.
-With such a large sample, we can for the first time unambiguously describe areas and morphologies that should be typical of networks of just two detectors.
-This case study will be the subject of the next chapter.
 

### 2015-07-06 ms.tex
Some edits

@@
 
-The \acf{LIGO} and Virgo will soon begin taking data \cite{LIGOObservingScenarios} in their `Advanced' configurations, ultimately attaining an order of magnitude more sensitivity than in their initial configurations through 2010.
+The \acf{LIGO} and Virgo will soon begin taking data \cite{LIGOObservingScenarios} in their `Advanced' configurations, ultimately increasing their reach in volume within the local Universe by three orders of magnitude as compared to their initial configurations through 2010.
 They will form the first parts of a sensitive global \ac{GW} detector network, to be later augmented by the Japanese KAGRA facility \cite{LCGT,KAGRAInterferometerDesign} and \ac{LIGO}\nobreakdashes--India \cite{LIGOIndia}.
@@
 Several mechanisms could make the kilonovae brighter, bluer, and hence more readily detectable \citep{KilonovaPrecursor,KilonovaRedOrBlue}, but peak even earlier, within hours.
-If, as is widely believed \citep{1986ApJ...308L..43P,1989Natur.340..126E,1992ApJ...395L..83N,2011ApJ...732L...6R}, \ac{BNS} mergers are indeed progenitors of short \acp{GRB}, then a small (due to jet collimation) fraction of Advanced \ac{LIGO} events could also be accompanied by a bright optical afterglow, but this signature, likewise, would peak within hours or earlier.
-
-Adding to the challenge of detecting a faint, short-lived optical transient, \ac{GW} detector networks have inherently coarse position reconstruction accurate to tens to hundreds of deg$^2$ \citep{FairhurstTriangulation,WenLocalizationAdvancedLIGO,FairhurstLocalizationAdvancedLIGO,2011PhRvD..84j4020V,RodriguezBasicParameterEstimation,NissankeLocalization,NissankeKasliwalEMCounterparts,KasliwalTwoDetectors,Grover:2013,SiderySkyLocalizationComparison,FirstTwoYears}.
-Wide\nobreakdashes-field optical transient surveys such as the \acf{ZTF}, BlackGEM, or the \acf{LSST}, operated in ``target of opportunity'' mode, may be able to tile these large areas rapidly enough to find the one needle in the haystack that is connected to the \ac{GW} event, but clearly time will be of the utmost.
+If, as is widely believed \citep{1986ApJ...308L..43P,1989Natur.340..126E,1992ApJ...395L..83N,2011ApJ...732L...6R}, \ac{BNS} mergers are indeed progenitors of short \acp{GRB}, then a small (due to jet collimation) fraction of Advanced \ac{LIGO} events could also be accompanied by a bright optical afterglow, but this signature, likewise, would peak within hours or faster.
+
+Adding to the challenge of detecting a faint, short-lived optical transient, there is an extreme mismatch between the sky localization accuracy of \ac{GW} detector networks, $\sim 10--500$~deg$^2$ \citep{FairhurstTriangulation,WenLocalizationAdvancedLIGO,FairhurstLocalizationAdvancedLIGO,2011PhRvD..84j4020V,RodriguezBasicParameterEstimation,NissankeLocalization,NissankeKasliwalEMCounterparts,KasliwalTwoDetectors,Grover:2013,SiderySkyLocalizationComparison,FirstTwoYears}, and the \acp{FOV} of 1\nobreakdashes--8\,m\nobreakdashes-class optical telescopes.
+Wide\nobreakdashes-field optical transient facilites such as BlackGEM (0.6\,m/2.7\,deg$^2$), the \acl{ZTF} \cite{ZTF} (1.2\,m/47\,deg$^2$), DECam (4\,m/3\,deg$^2$), or the \acf{LSST} (8.4\,m/9.6\,deg$^2$), operated in ``target of opportunity'' mode, may be able to tile these large areas rapidly enough to find the one needle in the haystack that is connected to the \ac{GW} event.
+However, prompt and accurate \ac{GW} position reconstructions will be of the utmost importance for guiding the selection of fields to observe.
 
@@
 This involved several advances in the \ac{GW} data analysis \citep{CBCLowLatency}, including the first real\nobreakdashes-time matched\nobreakdashes-filter detection pipeline (\acs{MBTA}; \aclu{MBTA}), a semi\nobreakdashes-coherent, ad hoc rapid triangulation code (Timing++), and the first version of a rigorous Bayesian \ac{MCMC} parameter estimation code (LALINFERENCE; \citealt{S6PE})---all in service of the first search for X\nobreakdashes-ray \citep{SwiftFollowup} and optical \citep{OpticalImageAnalysis} counterparts of \ac{GW} triggers, by a consortium of facilities.
-Despite the technical achievements in the \ac{GW} data analysis, there was an undesirable tradeoff between the speed of the rapid localization and the full parameter estimation: the former could analyze a detection candidate in minutes, whereas the latter took half a day; the latter decreased the area on the sky by a factor of 1/20 over the former but took 1000 times as long to run \cite{SiderySkyLocalizationComparison}.
+Despite the technical achievements in the \ac{GW} data analysis, there was an undesirable tradeoff between the speed as well as accuracy of the rapid localization and the full parameter estimation: the former could analyze a detection candidate in minutes, whereas the latter took half a day; the latter decreased the area on the sky by a factor of 1/20 over the former but took 1000 times as long to run \cite{SiderySkyLocalizationComparison}.
 
@@
 Just like the strain time series from the detectors, the resultant times, amplitudes, and phases have a predictable and quantifiable measurement uncertainty that can be translated into a likelihood function suitable for Bayesian inference.
-    \item Errors in sky localization are semi\nobreakdashes-independent from errors in masses.
-If we care only about position reconstruction and not about jointly estimating masses as well, then we can reduce the dimensionality of the parameter estimation problem significantly.
+    \item The Fisher information matrix will provide clues as to suitable forms of this likelihood function.
+Recent \ac{GW} parameter estimation literature has largely rejected the Fisher matrix, but this is mostly on the grounds of the abuse of the related \ac{CRLB} outside its realm of validity (i.e., at low to moderate \ac{SNR}) \cite{BayesianBounds,UseAbuseFisherMatrix,FisherMatrixAsymptoticExpansions,InadequaciesFisherMatrix}.
+However, we recognize that the block structure of the Fisher matrix provides important insights, and is a useful quantity for checking the validity of the aforementioned likelihood function, quite independent of the \ac{CRLB}.
+    \item The Fisher matrix teaches us that errors in sky localization are semi\nobreakdashes-independent from errors in masses.
+This implies that if we care only about position reconstruction and not about jointly estimating masses as well, then we can reduce the dimensionality of the parameter estimation problem significantly.
 Moreover, this frees us of the need to directly compute the expensive post-Newtonian model waveforms, making the likelihood itself much faster to evaluate.
     \item Thanks to a simple likelihood function and a well\nobreakdashes-characterized parameter space, we may dispense with costly and parallelization\nobreakdashes-resistant \ac{MCMC} integration, and instead perform the Bayesian marginalization with classic, deterministic, very low order Gaussian quadrature.
+    \item The Bayesian inference scheme thus designed to operate on the matched\nobreakdashes-filter detection pipeline outputs could be trivially generalized to operate on the full \ac{GW} time series within the same computational constraints.
+This would yield a fast and coherent localization algorithm that is mathematically equivalent to the full \ac{MCMC} parameter estimation, restricted to extrinsic parameters (sky location, binary orientation, and distance).
 \end{enumerate}
 %
-We call the resulting algorithm \ac{BAYESTAR}\footnote{A pun on the Cylon battleships in the American television series Battlestar Galactica.
-The defining characteristic of the Cylons is that they repeatedly defeat humanity by using their superhuman information\nobreakdashes-gathering ability to coordinate overwhelming forces.
-The name also suggests that, like the Cylons, \ac{GW} detectors may some day rise against us humans.}\footnote{We do not like to mention the final `L' in the acronym, because then it would be pronounced BAYESTARL, which sounds stupid.}.
+We call this algorithm \ac{BAYESTAR}\footnote{A pun on the Cylon battleships in the American television series Battlestar Galactica.
+The defining characteristic of the Cylons is that they repeatedly defeat humanity by using their superhuman information\nobreakdashes-gathering ability to coordinate overwhelming forces.}\footnote{We do not like to mention the final `L' in the acronym, because then it would be pronounced BAYESTARL, which sounds stupid.}.
 It is as fast as Timing++, but nearly as accurate as the rigorous full parameter estimation.
 It is unique in that it bridges the detection and parameter estimation of \ac{GW} signals, two tasks that have until now involved very different numerical methods and time scales.
-Beginning with the first Advanced \acs{LIGO} observing run, \ac{BAYESTAR} will provide localizations within minutes of the detection of any \ac{BNS} merger candidate, playing a a key role in enabling rapid follow-up observations.
+Beginning with the first Advanced \acs{LIGO} observing run, \ac{BAYESTAR} will provide localizations within minutes of the detection of any \ac{BNS} merger candidate, playing a key role in enabling rapid follow-up observations.
 
@@
 We can predict the uncertainty in the detection pipeline's \ac{ML} estimates using the \ac{CRLB}.
-The \ac{CRLB} has been widely applied in \ac{GW} data analysis to estimate parameter estimation uncertainty (for example, \cite{1996PhRvD..53.3033B,FairhurstTriangulation,2009PhRvD..79h4032A,WenLocalizationAdvancedLIGO,LIGOObservingScenarios,FairhurstLIGOIndia})\footnote{The Fisher matrix is also used in construction of \ac{CBC} matched filter banks.
+The \ac{CRLB} has been widely applied in \ac{GW} data analysis to estimate parameter estimation uncertainty \cite{1996PhRvD..53.3033B,FairhurstTriangulation,2009PhRvD..79h4032A,WenLocalizationAdvancedLIGO,LIGOObservingScenarios,FairhurstLIGOIndia}\footnote{\label{footnote:template-banks}The Fisher matrix is also used in construction of \ac{CBC} matched filter banks.
 The common procedure is to place templates uniformly according to the determinant of the signal space metric, which is the Fisher matrix.
@@
 In practice, this is done either by constructing a hexagonal lattice \citep{PhysRevD.76.102004} or sampling stochastically \citep{2009PhRvD..80j4014H,2009PhRvD..80b4009V,SBank,2010PhRvD..81b4004M,2014PhRvD..89b4003P}.}.
+As we noted, there are significant caveats to the \ac{CRLB} at low or moderate \ac{SNR} \cite{BayesianBounds,UseAbuseFisherMatrix,FisherMatrixAsymptoticExpansions,InadequaciesFisherMatrix}.
+However, here we will be concerned more with gaining intuition from the block structure of the Fisher matrix than its numerical value.
+Furthermore, the Fisher matrix in its own right---independent of its suitability to describe the parameter covariance---is a well\nobreakdashes-defined property of any likelihood function, and we will exploit it as such in Section~\ref{sec:bayestar-likelihood}.
 
@@
 We show in Appendix~\ref{sec:independence-intrinsic-extrinsic} that we need not consider the intrinsic parameters at all if we are concerned only with sky localization.
-
-The Fisher matrix has been much maligned within the \ac{GW} literature because the \ac{CRLB} is a good description of the parameter covariance in the limit of high \ac{SNR}, and can be abused by applying it in low-\ac{SNR} conditions, or close to the boundaries of the parameter space \cite{UseAbuseFisherMatrix}.
-However, we will put the \ac{CRLB} to good use to improve our qualitative understanding of the likelihood.
-Furthermore, the Fisher matrix in its own right---independent of its suitability to describe the parameter covariance---is a well-defined property of any likelihood function, and we will exploit it as such in Section~\ref{sec:bayestar-likelihood}.
 
@@
         \int_{-T}^T \Im\left[a^*(t) \dot{a}(t)\right] w(t; \rho) dt,
-\intertext{where}
+\intertext{where}\label{eq:fisher-weight-factor}
     w(t; \rho) &= \frac{
@@
 %
-For large $\rho$, the exponential dominates and we can write:
+For large $\rho$, the exponents of $e^{\rho^2}$ dominate Equation~(\ref{eq:fisher-weight-factor}) and we can write:
 %
@@
 
-We have no explicit prior on the intrinsic parameters; in our analysis they are fixed at their \ac{ML} estimates, $\hat{\bm\theta}_\mathrm{in}$.
+We have no explicit prior on the intrinsic parameters; in our analysis they are fixed at their \ac{ML} estimates, $\hat{\bm\theta}_\mathrm{in}$.\footnote{As noted in footnote \ref{footnote:template-banks}, the detection template bank is typically designed to uniformly sample the Jeffreys prior on the intrinsic parameters.
+Due to the equivalence of marginalization and maximization with respect to a parameter under a Gaussian distribution, fixing the intrinsic parameters at their \ac{ML} estimates is roughly equivalent to selecting the Jeffreys prior.}
 
@@
 Because BNS signals typically terminate at about 1500~Hz, a typical value for $\Delta t$ is $(4096\,\textrm{Hz})^{-1}$.
-We use a pruned \ac{FFT} because for BNS systems, the \ac{GW} signal remains in LIGO's sensitive band for $\sim$100\nobreakdashes--1000~s, whereas $T \sim 10$~ms.\footnote{See \url{http://www.fftw.org/pruned.html} for a discussion of methods for computing the first $K$ samples of an \ac{FFT} of length $N$.}
+We use a pruned \ac{FFT} because for BNS systems, the \ac{GW} signal remains in LIGO's sensitive band for $\sim$100\nobreakdashes--1000~s, whereas $T \sim 10$~ms.\footnote{See \url{http://www.fftw.org/pruned.html} for a discussion of methods for computing the pruned \ac{FFT}, the first $K$ samples of an \ac{FFT} of length $N$.}
 
@@
 
-The integral over $\cos\iota$ converges with just as rapidly with Gauss\nobreakdashes--Legendre quadrature, so we use a 10\nobreakdashes-point Gauss\nobreakdashes--Legendre rule for integration over $\cos\iota$.
+The integral over $\cos\iota$ converges just as rapidly with Gauss\nobreakdashes--Legendre quadrature, so we use a 10\nobreakdashes-point Gauss\nobreakdashes--Legendre rule for integration over $\cos\iota$.
+
+\begin{figure*}
+    \begin{minipage}[t]{0.5\textwidth}
+        \centering
+        (a) Polarization angle \\
+        \includegraphics{polarization_angle_integral_convergence}
+    \end{minipage}%
+    \begin{minipage}[t]{0.5\textwidth}
+        \centering
+        (b) Inclination angle \\
+        \includegraphics{inclination_integral_convergence}
+    \end{minipage}
+    \caption{\label{fig:convergence}Relative error in \ac{BAYESTAR} integration scheme as a function of }
+\end{figure*}
 
@@
 %
-\begin{equation}
+\begin{multline}
     \mathscr{F} = \exp\left(\frac{b^2}{4p^2}\right)
         {r_0}^m
-        \int_{r_\mathrm{min}}^{r_\mathrm{max}}
+        \int_{r_\mathrm{min}}^{r_\mathrm{max}} \\
         \exp\left[-\left(\frac{p}{r} - \frac{b}{2p}\right)^2\right]
+        \overline{I}_0\left[\frac{b}{r}\right]
         \left[\frac{r}{r_0}\right]^m
         dr.
-\end{equation}
+\end{multline}
 %
@@
 %
-\begin{align}\label{eq:fisher-matrix}
+\begin{align}
     \mathcal{I}_{\bm\theta_i,\bm\theta_i} &= \bordermatrix{

### 2015-07-06 ms.tex
Redo distance section

@@
 
-The integrand is periodic in $\psi$, so simple Newton\nobreakdashes--Cotes quadrature over $\psi$ exhibits extremely rapid convergence.
+The integrand is periodic in $\psi$, so simple Newton\nobreakdashes--Cotes quadrature over $\psi$ exhibits extremely rapid convergence (see Figure~\ref{fig:angle-convergence}).
 We therefore sample the posterior on a regular grid of 10 points from 0 to $\pi$.
 
-The integral over $\cos\iota$ converges just as rapidly with Gauss\nobreakdashes--Legendre quadrature, so we use a 10\nobreakdashes-point Gauss\nobreakdashes--Legendre rule for integration over $\cos\iota$.
+The integral over $\cos\iota$ converges just as rapidly with Gauss\nobreakdashes--Legendre quadrature (see Figure~\ref{fig:angle-convergence}), so we use a 10\nobreakdashes-point Gauss\nobreakdashes--Legendre rule for integration over $\cos\iota$.
 
@@
     \end{minipage}
-    \caption{\label{fig:convergence}Relative error in \ac{BAYESTAR} integration scheme as a function of }
+    \caption{\label{fig:angle-convergence}Relative error in \ac{BAYESTAR} integration scheme as a function of the number of Gaussian quadrature nodes.
+The two panels describe (a) the integral over the polarization angle $\psi$ and (b) the integral over inclination angle $\iota$.}
 \end{figure*}
@@
 
-The innermost integral over $r$ is a little bit more challenging.
-The distance integrand has both a sharp maximum\nobreakdashes-likelihood peak and an extended power\nobreakdashes-law prior\nobreakdashes-dominated tail.
-At any values of the other parameters, either the peak or the tail might dominate.
-If the peak dominates, it may not be successfully resolved by any general\nobreakdashes-purpose fixed\nobreakdashes-order quadrature scheme.
-Any general adaptive integrator may take a long time to find the peak.
-
-Let's say that we want to evaluate an integral
-%
-\begin{equation}\label{eq:importance-sampling-difficult-integral}
-    \mathscr{F}(x_1, x_2) = \int_{x_1}^{x_2} f(x') dx'.
-\end{equation}
-%
-Suppose that this integral is resistant to direct application of standard quadrature techniques, but we have another function $g(x)$ that we can integrate analytically or at least numerically with relative ease,
-%
-\begin{equation}
-    \mathscr{G}(x) = G(x) - G(x_0) = \int_{x_0}^x g(x') dx'.
-\end{equation}
-%
-There is an \emph{ultimate}, though tautological, change of variables that makes it trivial to evaluate $\mathscr{G}(x)$,
-%
-\begin{equation}
-    \mathscr{G}(x) = \int_0^{\mathrlap{\mathscr{G}(x)}} dG.
-\end{equation}
-%
-If we apply the same change of variables to Equation~(\ref{eq:importance-sampling-difficult-integral}), we get
-%
-\begin{equation}\label{eq:importance-sampling}
-    \mathscr{F}(x_1, x_2) = \int_{\mathscr{G}(x_1)}^{\mathscr{G}(x_2)} \frac{f(\mathscr{G}^{-1}(G))}{g(\mathscr{G}^{-1}(G))} dG.
-\end{equation}
-%
-If $f(x)$ and $g(x)$ have sufficiently similar behavior, then their quotient $f/g$ will be better behaved than $f$ itself, and more amenable to any given quadrature technique.
-When the transformed integral Equation~(\ref{eq:importance-sampling}) is evaluated using Monte Carlo integration, this technique is referred to as importance sampling.
-
-We use the same kind of change of variables with fixed\nobreakdashes-order Gaussian quadrature. The distance integral can be written as
+The innermost integral over distance $r$ may be written as
 %
@@
 %
+or, completing the square,
+%
+\begin{align}
+    \mathscr{F} &= \exp\left[\frac{p^2}{{r_0}^2}\right]
+        \int_{\mathrlap{r_\mathrm{min}}}^{\mathrlap{r_\mathrm{max}}}
+        \exp\left[-\left(\frac{p}{r} - \frac{p}{r_0}\right)^2\right]
+        \overline{I}_0\left[\frac{2p^2}{r r_0}\right] r^m dr \\
+    &= \exp\left[\frac{p^2}{{r_0}^2}\right] \mathscr{G},
+\end{align}
+%
 where
@@
         a_i^*(\tilde{\tau}_i) \right| \label{eq:likelihood-b-factor} \\
+    r_0 &= 2 p^2 / b \\
     \overline{I}_0(x) &= \exp(-|x|) I_0(x).
@@
 %
-The coefficients $p^2$ and $b$ are nonnegative and independent of distance.
+The coefficients $p^2$ and $b$ are nonnegative and independent of distance. $p$ has a maximum value of
+%
+\begin{equation}
+    p_\mathrm{max} = \sqrt{\frac{1}{2} \sum_i \left(\frac{r_{1,i}}{r}\right)^2}.
+\end{equation}
+%
 The symbol $\overline{I}_0$ denotes an exponentially scaled Bessel function.
@@
 The scaled Bessel is useful for evaluation on a computer because it has a relatively small range $(0, 1]$ and varies slowly in proportion to $x^{1/2}$.
-If we neglect both the Bessel function and the $r^m$ prior, then the approximate likelihood $\exp(-p^2/r^2 + b/r)$ is maximized when
-%
-\begin{equation}\label{eq:ml-distance}
-    r = r_0 \equiv 2p^2/b.
-\end{equation}
-%
-with $p$ and $b$ defined above in Equations~(\ref{eq:likelihood-p-factor},~\ref{eq:likelihood-b-factor}).
-If $r_0 < r_\mathrm{min}$ or $r_0 > r_\mathrm{max}$, then the ML peak lies outside of the limits of integration and the original integrand behaves like a low-order power law.
-In this case, we use 10\nobreakdashes-point Gaussian quadrature to evaluate Equation~(\ref{eq:distance-integral}).
-
-If, on the other hand, $r_\mathrm{min} \leq r_0 \leq r_\mathrm{max}$, then the integrand contains a sharp peak that we can smooth out with importance sampling.
-We first re-scale the distance integral Equation~(\ref{eq:distance-integral}) so that the peak value of the integrand is $\sim 1$:
-%
-\begin{multline}
-    \mathscr{F} = \exp\left(\frac{b^2}{4p^2}\right)
-        {r_0}^m
-        \int_{r_\mathrm{min}}^{r_\mathrm{max}} \\
-        \exp\left[-\left(\frac{p}{r} - \frac{b}{2p}\right)^2\right]
-        \overline{I}_0\left[\frac{b}{r}\right]
-        \left[\frac{r}{r_0}\right]^m
-        dr.
-\end{multline}
-%
-The importance sampling integrand $g(r)$ is
-%
-\begin{equation}
-    g(r) = \exp\left[-\left(\frac{p}{r} - \frac{b}{2p}\right)^2\right]
-        \frac{1}{r^2},
-\end{equation}
-%
-with definite integral
-%
-\begin{equation}
-    \mathscr{G}(r) = \frac{\sqrt{\pi}}{p}
-        Q\left[\sqrt{2}\left(\frac{p}{r} - \frac{b}{2p}\right)\right],
-\end{equation}
-%
-where $Q$ is the cumulative distribution function for the upper tail of the standard normal distribution,
-%
-\begin{equation}
-    Q(x) = \int_x^\infty \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{{x'}^2}{2}\right] dx'.
-\end{equation}
-%
-Its inverse is
-%
-\begin{equation}
-    \mathscr{G}^{-1}(G) = \left[\frac{1}{r_0} + \frac{1}{\sqrt{2}p}
-    Q^{-1}\left(\frac{p G}{\sqrt{\pi}}\right)\right]^{-1}.
-\end{equation}
-%
-The transformed integral is
-%
-\begin{equation}
-    \mathscr{F} = \exp\left(\frac{b^2}{4p^2}\right)
-        {r_0}^m
-        \int\limits_{\mathclap{\mathscr{G}(r_\mathrm{min})}}^{\mathclap{\mathscr{G}(r_\mathrm{max})}}
-        \overline{I}_0\left[\frac{b}{r}\right]
-        \left[\frac{r}{r_0}\right]^m r^2
-        \Bigg|_{\mathrlap{r = \mathscr{G}^{-1}(G)}}
-        dG.
-\end{equation}
-%
-This integral is evaluated numerically with a 10\nobreakdashes-point Gauss\nobreakdashes--Legendre rule.
 
 \begin{figure}
-    \includegraphics[width=\columnwidth]{importance-sampling}
-    \caption{\label{fig:importance-sampling}Illustration of importance sampling scheme.
-The top-left panel shows the original integrand $f(r)$, the proposal integrand $g(r)$, and the limits of integration $r = r_\mathrm{min}$ and $r = r_\mathrm{max}$.
-The area of the dark shaded region is the integral of interest.
-The bottom-left panel shows the integral $\mathscr{G}(r) = \int_0^r g(r) dr$.
-The new variable of integration is $G = \mathscr{G}(r)$, from $G_\mathrm{min} = \mathscr{G}(r_\mathrm{min})$ to $G_\mathrm{max} = \mathscr{G}(r_\mathrm{max})$.
-The bottom-right panel shows the transformed integrand, $f(\mathscr{G}^{-1}(G))/g(\mathscr{G}^{-1}(G))$.
-The dark shaded region has the same area as in the top-left figure.}
+    \includegraphics{radial_integral_interpolant}
+    \caption{\label{fig:radial-integral-interpolant}Partition of the parameter space of the distance integral into three regions for (bi)cubic interpolation.}
 \end{figure}
 
-Last, there are a few special cases where Equation~(\ref{eq:distance-integral}) can be evaluated exactly.
-If $p^2=0$ (which implies $b = 0$), then the integral may be expressed in terms of logarithms.
-If $b = 0$ but $p^2 \neq 0$, then the integral may be expressed in terms of the exponential integral function,
-%
-\begin{equation}
-    E_n(x) = \int_1^\infty \frac{\exp (-xt)}{t^n} dt.
-\end{equation}
-
-The distance integral, therefore, takes either one or ten function evaluations.
+\subsubsection{Parameter grid}
+
+This integral is not particularly amenable to low-order Gaussian quadrature.
+However, luckily $\mathscr{G}$ is a very well-behaved function of $p$ and $r_0$, so we evaluate it using a lookup table and bicubic interpolation.
+The lookup table is produced in logarithmic coordinates
+%
+\begin{equation}
+    x = \log p, \qquad y = \log r_0.
+\end{equation}
+%
+As shown in Figure~\ref{fig:radial-integral-interpolant}, the function basically consists of a plateau region in the upper-left half of the plane delimited by the lines $y = x$ and $x = \log p_0$, with
+%
+\begin{equation}
+    p_0 = \frac{1}{2}
+    \begin{cases}
+        r_\mathrm{max} & \text{if } m \geq 0 \\
+        r_\mathrm{min} & \text{if } m < 0.
+    \end{cases}
+\end{equation}
+%
+We tabulate $\mathscr{G}$ on a $400 \times 400$ regular grid spanning the range
+%
+\begin{align}
+    x_0 &= \log \min(p_0, p_\mathrm{max}) \\
+    x_\mathrm{min} &= x_0 - (1 + \sqrt{2}) \alpha \\
+    x_\mathrm{max} &= \log p_\mathrm{max} \\
+    y_\mathrm{min} &= 2 x_0 - \sqrt{2} \alpha - x_\mathrm{max} \\
+    y_\mathrm{max} &= x_0 + \alpha.
+\end{align}
+
+\subsubsection{Lookup table construction}
+
+The lookup table for $\mathscr{G}$ is populated as follows.
+If we neglect both the Bessel function and the $r^m$ prior, then the approximate likelihood $\exp(-(p/r-p/r_0)^2)$ is maximized when $r = r_0$.
+The likelihood takes on a factor $\eta$ (say, $\eta=0.01$) of its maximum value when
+%
+\begin{equation}
+    r = r_\pm = \left(\frac{1}{r_0} \mp \frac{\sqrt{-\log\eta}}{p}\right)^{-1}.
+\end{equation}
+
+We have now identified up to five breakpoints that partition the distance integrand into up to four intervals with quantitatively distinct behavior.
+These intervals are depicted in Figure~\ref{fig:radial_integrand} with distance increasing from left to right.
+There is a left\nobreakdashes-hand or small distance tail in which the integrand is small and monotonically increasing, a left\nobreakdashes- and right\nobreakdashes-hand side of the maximum likelihood peak, and a right\nobreakdashes-hand tail in which the integrand is small and monotonically decreasing.
+These breakpoints are:
+%
+\begin{equation}
+    r_\mathrm{break} = \{ r \in
+    \left\{
+    \begin{array}{c}
+    r_\mathrm{min} \\
+    r_- \\
+    r_0 \\
+    r_+ \\
+    r_\mathrm{max}
+    \end{array}
+    \right\} :
+    r_\mathrm{min} \leq r
+    \leq r_\mathrm{max}\}.
+\end{equation}
+
+We use these breakpoints as initial subdivisions in an adaptive Gaussian quadrature algorithm\footnote{for instance, \ac{GSL}'s \texttt{gsl\_integrate\_qagp} function, \url{http://www.gnu.org/software/gsl/manual/html_node/QAGP-adaptive-integration-with-known-singular-points.html}}.
+This function estimates the integral over each subdivision and each interval's contribution to the total error, then subdivides the interval whose error contribution is largest.
+Subdivisions continue until a fixed total fractional error is reached.
+In this way, most integrand evaluations are expended on the most important distance interval, whether that happens to be the tails (when the posterior is dominated by the prior) or the peak (when the posterior is dominated by the observations).
+
+\begin{figure}
+    \begin{center}
+        \includegraphics[width=\columnwidth]{radial_integrand}
+    \end{center}
+    \caption{\label{fig:radial_integrand}Illustration of initial subdivisions for distance integration scheme.
+Distance increases from left to right.
+In the color version, the left\nobreakdashes-hand tail, the left\nobreakdashes- and right\nobreakdashes-hand sides of the maximum likelihood peak, and the right\nobreakdashes-hand tail, are colored cyan, red, green, and blue, respectively.}
+\end{figure}
+
+\subsubsection{Interpolation}
+
+The interpolant is evaluated slightly differently depending on which of the three regions marked I, II, and II in Figure~\ref{fig:radial-integral-interpolant} contains the point of interest.
+In region I, we use bicubic interpolation in $x$ and $y$.
+In region II, we use univariate cubic interpolation in $x$, with the sample points taken from the horizontal boundary between regions I and II.
+In region III, we use univariate cubic interpolation in $u = (x - y) / 2$, with the sample points taken from the downward diagonal boundary between regions I and III.
+This entire lookup table scheme is accurate to an absolute error of about $10^{-5}$ in $\mathscr{G}$.
+Finally, the distance integral $\mathscr{F}$ is obtained by multiplying the interpolated value of $\mathscr{G}$ by $\exp\left(p^2/{r_0}^2\right)$.
 

### 2015-07-06 scripts/radial_integral_interpolant.py

@@ -1,0 +1,62 @@
+from __future__ import division
+import numpy as np
+from matplotlib import pyplot as plt
+import scipy.integrate
+import scipy.special
+
+r1 = 0
+r2 = 0.25
+pmax = np.sqrt(0.5 * 3)
+k = 2
+
+def integrand(r, p, b, k):
+    return np.exp(-np.square(p/r - 0.5*b/p)) * scipy.special.i0e(b / r) * r**k
+
+@np.vectorize
+def integral(p, b, k):
+    value, error = scipy.integrate.quad(integrand, r1, r2, args=(p, b, k))
+    return value
+
+SQRT2 = np.sqrt(2)
+alpha = 4.0
+p0 = 0.5 * (r2 if k >= 0 else r1)
+x0 = np.log(min(p0, pmax))
+xmin = x0 - (1 + SQRT2) * alpha
+xmax = np.log(pmax)
+ymin = 2 * x0 - SQRT2 * alpha - xmax
+ymax = x0 + alpha
+
+x = np.linspace(xmin - 4, xmax, 200)
+y = np.linspace(ymin - 4, ymax + 4, 200)
+xx, yy = np.meshgrid(x, y)
+
+p = np.exp(xx)
+r0 = np.exp(yy)
+b = 2*np.square(p) / r0
+z = integral(p, b, k)
+
+fig = plt.figure(figsize=(3, 3))
+ax = plt.subplot(111, aspect=1)
+ax.imshow(z, origin='lower left', extent=[x.min(), x.max(), y.min(), y.max()], cmap='cool')
+ax.contour(x, y, z, levels=np.logspace(-4, np.log10(z.max()), 6), colors='black')
+ax.axvspan(xmax, xmax + 4, color='lightgray', alpha=0.25)
+
+# ax.set_frame_on(False)
+ax.set_xticks([x0 - (1 + SQRT2) * alpha, x0, xmax])
+for tick in ax.get_xticks():
+    ax.axvline(tick, color='black', linewidth=0.5, alpha=0.25)
+ax.set_xticklabels([r'$x_\mathrm{min}$', r'$x_0$', r'$x_\mathrm{max}$'])
+ax.set_yticks([2 * x0 - SQRT2 * alpha - xmax, x0, x0 + alpha])
+for tick in ax.get_yticks():
+    ax.axhline(tick, color='black', linewidth=0.5, alpha=0.25)
[…]

### 2015-07-06 scripts/inclination_integral_convergence.py
Update rate of convergence plots

@@ -7,7 +7,7 @@ fig = plt.figure(figsize=(3.5, 2.5))
 ax = fig.add_subplot(111)
 
 np.random.seed(0)
-nruns = 10
+nruns = 20
 neval = np.unique(np.round(np.logspace(0, 2, 50)).astype(int))
 
 def integrand(u, a, b, c, d, e, f):
@@ -24,20 +24,21 @@ for irun in range(nruns):
     relerr = np.abs(approx - exact) / exact
     ax.plot(neval, relerr, '-k')
 
-def integrand(u, a, b, c, d, e, f):
-    v = 1 + np.square(u)
-    I0arg1 = np.sqrt(np.square(c * u) + np.square(d * v))
-    I0arg2 = np.sqrt(np.square(e * u) + np.square(f * v))
-    return np.exp(-(np.square(a * u) + np.square(b * v)) + I0arg1 + I0arg2) * scipy.special.i0e(I0arg1) * scipy.special.i0e(I0arg2)
-
-for irun in range(nruns):
-    args = tuple(np.random.randn(6).tolist())
-    exact, _ = scipy.integrate.quad(
-        integrand, 0, 1, args, epsabs=1e-14, epsrel=1e-14, limit=200)
-    approx = np.asarray([
-        scipy.integrate.fixed_quad(integrand, 0, 1, args, n)[0] for n in neval])
-    relerr = np.abs(approx - exact) / exact
-    ax.plot(neval, relerr, '--k')
+# TOA+SNR
+# def integrand(u, a, b, c, d, e, f):
+#     v = 1 + np.square(u)
+#     I0arg1 = np.sqrt(np.square(c * u) + np.square(d * v))
+#     I0arg2 = np.sqrt(np.square(e * u) + np.square(f * v))
+#     return np.exp(-(np.square(a * u) + np.square(b * v)) + I0arg1 + I0arg2) * scipy.special.i0e(I0arg1) * scipy.special.i0e(I0arg2)
+#
+# for irun in range(nruns):
+#     args = tuple(np.random.randn(6).tolist())
+#     exact, _ = scipy.integrate.quad(
+#         integrand, 0, 1, args, epsabs=1e-14, epsrel=1e-14, limit=200)
+#     approx = np.asarray([
+#         scipy.integrate.fixed_quad(integrand, 0, 1, args, n)[0] for n in neval])
+#     relerr = np.abs(approx - exact) / exact
[…]

### 2015-07-06 scripts/polarization_angle_integral_convergence.py

@@ -7,7 +7,7 @@ fig = plt.figure(figsize=(3.5, 2.5))
 ax = fig.add_subplot(111)
 
 np.random.seed(0)
-nruns = 10
+nruns = 20
 neval = np.unique(np.round(np.logspace(0, 2, 50)).astype(int))
 
 def integrand(x, a, b, c, d, e, f):
@@ -26,19 +26,20 @@ for irun in range(nruns):
     relerr = np.abs(approx - exact) / exact
     ax.plot(neval, relerr, '-k')
 
-def integrand(x, a, b, c, d, e, f, g, h, i, j):
-    I0arg1 = np.sqrt(np.square(c * np.cos(x) + d * np.sin(x)) + np.square(e * np.cos(x) + f * np.sin(x)))
-    I0arg2 = np.sqrt(np.square(g * np.cos(x) + h * np.sin(x)) + np.square(i * np.cos(x) + j * np.sin(x)))
-    return np.exp(a * np.cos(x) + b * np.sin(x) + I0arg1 + I0arg2) * scipy.special.i0e(I0arg1) * scipy.special.i0e(I0arg2)
-
-for irun in range(nruns):
-    args = tuple(np.random.randn(10).tolist())
-    exact, _ = scipy.integrate.quad(
-        integrand, -1, 1, args, epsabs=1e-14, epsrel=1e-14, limit=200)
-    approx = np.asarray([
-        scipy.integrate.fixed_quad(integrand, -1, 1, args, n)[0] for n in neval])
-    relerr = np.abs(approx - exact) / exact
-    ax.plot(neval, relerr, '--k')
+# TOA+SNR
+# def integrand(x, a, b, c, d, e, f, g, h, i, j):
+#     I0arg1 = np.sqrt(np.square(c * np.cos(x) + d * np.sin(x)) + np.square(e * np.cos(x) + f * np.sin(x)))
+#     I0arg2 = np.sqrt(np.square(g * np.cos(x) + h * np.sin(x)) + np.square(i * np.cos(x) + j * np.sin(x)))
+#     return np.exp(a * np.cos(x) + b * np.sin(x) + I0arg1 + I0arg2) * scipy.special.i0e(I0arg1) * scipy.special.i0e(I0arg2)
+#
+# for irun in range(nruns):
+#     args = tuple(np.random.randn(10).tolist())
+#     exact, _ = scipy.integrate.quad(
[…]

### 2015-07-06 ms.tex
Rearrange figures
Add autocorrelation likelihood plot

@@
 In the following section, we discuss some key properties of the autocorreltion likelihood.
+
+\begin{figure*}
+    \includegraphics[width=\textwidth]{autocorr-likelihood}
+    \caption{\label{fig:autocorr-likelihood}The autocorrelation likelihood for a $(1.4, 1.4)$\,$M_\odot$ binary as observed by four detector configurations: from top to bottom, the final sensitivity achieved by the \acs{LIGO} Hanford, \acs{LIGO} Livingston, and Virgo detectors in their ``initial'' configuration, and the final Advanced \acs{LIGO} design sensitivity.
+The left panels show the noise amplitude spectral densities.
+The middle panels show the absolute value of the autocorrelation function.
+The right panel shows the phase-marginalized autocorrelation likelihood for \ac{SNR}=1,~2,~4,~and~8.
+In the right panel, the time scale is normalized by \ac{SNR} so that one can see that as \ac{SNR} increases, a central parabola is approached (the logarithm of a Gaussian distribution with standard deviation given by the Fisher matrix).}
+\end{figure*}
 
@@
 
+\begin{figure*}
+    \centering
+        \includegraphics[width=\columnwidth]{crlb_tau}
+        \includegraphics[width=\columnwidth]{crlb_gamma}
+    \caption[\acl{CRLB} on time and phase accuracy]{\label{fig:crlb-tau}\ac{CRLB} on \ac{RMS} timing uncertainty and phase error, using the likelihood for the full \ac{GW} data (Equation~\ref{eq:gaussian-likelihood-spa}; dashed diagonal line) or the autocorrelation likelihood (Equation~\ref{eq:autocor-likelihood}; solid lines) with a selection of arrival time priors.}
+\end{figure*}
+
 We can show that the weighting function $w(t; \rho)$ approaches a Dirac delta function as $\rho \rightarrow \infty$, so that the Fisher matrix for the autocorrelation likelihood approaches the Fisher matrix for the full \ac{GW} data, Equation~(\ref{eq:fisher-matrix}), as $\rho \rightarrow \infty$.
@@
 
-In the limit of large \ac{SNR}, our interpretation is that the point estimates $(\hat\rho, \hat\gamma, \hat\tau)$ contain all of the information about the underlying extrinsic parameters.
-
-On the other hand, in the low \ac{SNR} limit, the diminishing value of $\textproto{D}_{ij}(\rho)$ reflects the fact that some information is lost when the full data $\mathbf{x}$ is discarded.
-Concretely, as the prior interval $T$ becomes large compared to 1/$\rho\omega_\mathrm{rms}$, the \ac{ML} estimator becomes more and more prone to picking up spurious noise fluctuations far from the true signal.
-Clearly, when the coincidence window $T$ is kept small as possible, more information is retained in the \ac{ML} point estimates.
-Put another way, if $T$ is small, then the transition \ac{SNR} $\rho_\mathrm{crit}$ is also small and fainter signals become useful for parameter estimation.
-In this way, the \ac{BAYESTAR} likelihood exhibits the \emph{threshold effect} that is well\nobreakdashes-known in communication and radar applications \citep{barankin1949locally,mcaulay1969barankin,mcaulay1971barankin}.
-
-\begin{figure*}
-    \centering
-        \includegraphics[width=\columnwidth]{crlb_tau}
-        \includegraphics[width=\columnwidth]{crlb_gamma}
-    \caption[\acl{CRLB} on time and phase accuracy]{\label{fig:crlb-tau}\ac{CRLB} on \ac{RMS} timing uncertainty and phase error, using the likelihood for the full \ac{GW} data (Equation~\ref{eq:gaussian-likelihood-spa}; dashed diagonal line) or the autocorrelation likelihood (Equation~\ref{eq:autocor-likelihood}; solid lines) with a selection of arrival time priors.}
-\end{figure*}
-
 \begin{figure}
@@
 \end{figure}
+
+In the limit of large \ac{SNR}, our interpretation is that the point estimates $(\hat\rho, \hat\gamma, \hat\tau)$ contain all of the information about the underlying extrinsic parameters.
+
+On the other hand, in the low \ac{SNR} limit, the diminishing value of $\textproto{D}_{ij}(\rho)$ reflects the fact that some information is lost when the full data $\mathbf{x}$ is discarded.
+Concretely, as the prior interval $T$ becomes large compared to 1/$\rho\omega_\mathrm{rms}$, the \ac{ML} estimator becomes more and more prone to picking up spurious noise fluctuations far from the true signal.
+Clearly, when the coincidence window $T$ is kept small as possible, more information is retained in the \ac{ML} point estimates.
+Put another way, if $T$ is small, then the transition \ac{SNR} $\rho_\mathrm{crit}$ is also small and fainter signals become useful for parameter estimation.
+In this way, the \ac{BAYESTAR} likelihood exhibits the \emph{threshold effect} that is well\nobreakdashes-known in communication and radar applications \citep{barankin1949locally,mcaulay1969barankin,mcaulay1971barankin}.
 
@@
 
+\subsubsection{Parameter grid}
+
 \begin{figure}
@@
 \end{figure}
-
-\subsubsection{Parameter grid}
 
@@
 We have explained how we evaluate the marginal posterior at a given sky location. Now we must specify where we choose to evaluate it.
-
-\begin{figure*}
-    \includegraphics[width=\textwidth]{illustration}
-    \caption{\label{fig:adaptive-illustration}Illustration of the \ac{BAYESTAR} adaptive HEALPix sampling scheme.}
-\end{figure*}
 
@@
 
+\begin{figure*}
+    \includegraphics[width=\textwidth]{illustration}
+    \caption{\label{fig:adaptive-illustration}Illustration of the \ac{BAYESTAR} adaptive HEALPix sampling scheme.}
+\end{figure*}
+
 Within each iteration, all of the marginalization integrals may be evaluated in parallel.
@@
 These are the two observing scenarios that are discussed in \cite{FirstTwoYears}.
+
+\begin{figure}
+    \centering
+    \includegraphics{runtimes}
+    \caption[\acs{BAYESTAR} run time]{\label{fig:runtimes}Violin plot of \ac{BAYESTAR} run times as the number of OpenMP threads is varied from 1 to 32.
+The 2015 scenario is shown in red and the 2016 scenario in blue.}
+\end{figure}
 
@@
 The 32\nobreakdashes-thread configuration is representative of how \ac{BAYESTAR} might be deployed in early Advanced \ac{LIGO}.\footnote{\ac{BAYESTAR} has been successfully ported to the Intel's Many Integrated Core~(MIC) architecture and has been tested in a 500 thread configuration on a system with dual Intel Xeon Phi 5110P coprocessors.}
-
-\begin{figure}
-    \centering
-    \includegraphics{runtimes}
-    \caption[\acs{BAYESTAR} run time]{\label{fig:runtimes}Violin plot of \ac{BAYESTAR} run times as the number of OpenMP threads is varied from 1 to 32.
-The 2015 scenario is shown in red and the 2016 scenario in blue.}
-\end{figure}
 

### 2015-07-06 scripts/autocorr-likelihood.py

@@ -1,0 +1,184 @@
+from __future__ import division
+import numpy as np
+from matplotlib import pyplot as plt
+from matplotlib import contour as mcontour
+from matplotlib import gridspec as mgridspec
+import scipy.special
+from lalinference.bayestar import filter
+from lalinference.bayestar import timing
+import lal
+import lalsimulation
+import scipy.integrate
+import scipy.signal
+import os
+import seaborn as sns
+sns.set_style('white')
+
+def abs2(a):
+    return np.square(np.real(a)) + np.square(np.imag(a))
+
+data = [
+    ['LIGO-T1100338-v13-H1-SPECTRA-957474700-MODE.txt', "H1\nS6", False],
+    ['LIGO-T1100338-v13-L1-SPECTRA-951280082-MODE.txt', "L1\nS6", False],
+    ['LIGO-T1100338-v13-V1-SPECTRA-935662133-MODE.txt', "V1\nVSR2-3", False],
+    ['LIGO-T0900288-v3-ZERO_DET_high_P.txt', "aLIGO\ndesign", True]
+]
+
+mass1 = mass2 = 1.4
+
+fig = plt.figure(figsize=(11, 8.5))
+gridspec = mgridspec.GridSpec(len(data), 3, wspace=0.1, hspace=0.1)
+
+for i, (filename, label, need_square) in enumerate(data):
+    path = os.path.join('psds', filename)
+    f, S = np.loadtxt(path).T
+    if need_square:
+        S *= S
+    S = timing.InterpolatedPSD(f, S)
+
+    ax = fig.add_subplot(gridspec[i, 0])
+    f = np.logspace(0, 4, 1000)
+    ax.loglog(f, np.sqrt(S(f)), color='k')
+    ax.set_ylabel(label, rotation='horizontal', ha='right', labelpad=20)
+    ax.set_xlim(10, 1e4)
+    ax.set_ylim(1e-24, 1e-21)
+    if i == len(data) - 1:
+        ax.set_xlabel('frequency (Hz)')
+    else:
+        plt.setp(ax.get_xticklabels(), visible=False)
+    if i == 0:
+        ax.set_title('Noise amp. spectral density')
+
+    duration = lalsimulation.SimInspiralTaylorF2ReducedSpinChirpTime(
+        10, mass1*lal.MSUN_SI, mass2*lal.MSUN_SI, 0,
+        lalsimulation.PNORDER_THREE_POINT_FIVE)
[…]

### 2015-07-07 ms.tex
Add adaptive mesh example

@@
 A resolution may also be referred to by the number of subdivisions along each side of the base tiles, $N_\mathrm{side} = 2^N$.
-There are $N^\mathrm{pix} = 12 {N_\mathrm{side}}^2$ pixels at any given resolution.
+There are $N_\mathrm{pix} = 12 {N_\mathrm{side}}^2$ pixels at any given resolution.
 The \ac{HEALPix} projection uniquely specifies the coordinates of the center of each pixel by providing a mapping from the resolution and pixel index $(N_\mathrm{side}, i_\mathrm{pix})$ to right ascension and declination $(\alpha, \delta)$.
 
-We begin by evaluating the posterior probability density at the center of each of the $N_{\mathrm{pix},0} = 3072$ pixels of an $N_{\mathrm{side},0}=16$ \ac{HEALPix} grid.
+The \ac{BAYESTAR} adaptive sampling process works as follows.
+W begin by evaluating the posterior probability density at the center of each of the $N_{\mathrm{pix},0} = 3072$ pixels of an $N_{\mathrm{side},0}=16$ \ac{HEALPix} grid.
 At this resolution, each pixel has an area of 13.4~deg$^2$.
-We then rank the pixels by contained probability (assuming constant probability density within a pixel), and subdivide the most probable $N_{\mathrm{pix},0}/4$ pixels into $N_{\mathrm{pix},0}$ new pixels.
-We then evaluate the posterior again at the pixels that we subdivided, sort again, and repeat seven times, so that we have evaluated the posterior probability density a total of $8 N_{\mathrm{pix},0}$ times.
-On \emph{most} iterations, we descend one level deeper in \ac{HEALPix} resolution.
-The resulting map is a tree structure that consists of a mix of several resolutions.
+We then rank the pixels by contained probability (assuming constant probability density within a pixel), and subdivide the most probable $N_{\mathrm{pix},0}/4$ pixels into $N_{\mathrm{pix},0}$ new daughter pixels.
+We then evaluate the posterior again at the centers of the new daughter pixels, sort again, and repeat seven times.
+After the last iteration, we have evaluated the posterior probability density a total of $8 N_{\mathrm{pix},0}$ times.
+On \emph{most} subdivision steps, we descend one level deeper in \ac{HEALPix} resolution.
 This process is illustrated in Figure~\ref{fig:adaptive-illustration}.
-
-We traverse the tree and flatten it into the highest resolution represented.
-The highest possible resolution is $N_\mathrm{side}=2^{11}$, with an area of $\approx 10^{-3}$~deg$^2$ per pixel.\footnote{Although the resulting sky map contains $N_\mathrm{pix} \approx 5\times10^6$ pixels, at most $\approx 2\times10^4$ pixels have distinct values.
-For the purpose of delivery to observers, therefore, the output is always \texttt{gzip}\nobreakdashes-compressed with a ratio of $\approx 250:1$.}
 
@@
 \end{figure*}
+
+The resulting map is a tree structure that describes a mesh of pixels with different resolutions.
+An example \ac{BAYESTAR} subdivision is shown in Figure~\ref{fig:adaptive-illustration}.
+To convert this mesh into a FITS image,
+we traverse the tree and flatten it into the highest resolution represented.
+The highest possible resolution is $N_\mathrm{side}=2^{11}$, with an area of $\approx 10^{-3}$~deg$^2$ per pixel.\footnote{Although the resulting sky map contains $N_\mathrm{pix} \approx 5\times10^6$ pixels, at most $\approx 2\times10^4$ pixels have distinct values.
+For the purpose of delivery to observers, therefore, the output is always \texttt{gzip}\nobreakdashes-compressed with a ratio of $\approx 250:1$.}
+
+\begin{figure}
+    \includegraphics[width=\columnwidth]{adaptive_mesh}
+    \caption{\label{fig:adaptive-mesh}}
+\end{figure}
 
@@
 This is a violin plot, a smoothed vertical histogram.
-The red surface shows run times for a two\nobreakdashes-detector network (HL) modeled on the first scheduled Advanced \ac{LIGO} observing run in 2015, and the blue surface shows run times for a three\nobreakdashes-detector network (HLV) based on the second planned observing run in 2016.
+The magenta regions show run times for a two\nobreakdashes-detector network (HL) modeled on the first scheduled Advanced \ac{LIGO} observing run in 2015, and the blue regions show run times for a three\nobreakdashes-detector network (HLV) based on the second planned observing run in 2016.
 These are the two observing scenarios that are discussed in \cite{FirstTwoYears}.

### 2015-07-07 scripts/adaptive_mesh.py

@@ -1,0 +1,74 @@
+# -*- coding: utf-8 -*-
+from __future__ import division
+import matplotlib
+matplotlib.rcParams['text.usetex'] = False
+from lalinference import fits
+from lalinference import cmap
+from lalinference import plot
+from wcsaxes import WCS
+import numpy as np
+from matplotlib import pyplot as plt
+import healpy as hp
+from matplotlib import collections
+from matplotlib import patches
+from matplotlib import ticker
+
+ra0 = 130
+dec0 = -10
+phi0 = np.deg2rad(ra0)
+theta0 = 0.5 * np.pi - np.deg2rad(dec0)
+# ra0 = 137.8
+# dec0 = -39.9
+xyz0 = hp.ang2vec(0.5*np.pi - np.deg2rad(dec0), np.deg2rad(ra0))
+
+prob, meta = fits.read_sky_map('../cbc/emfollowup/papers/first2years/data/2015/compare/18951/bayestar.fits.gz', nest=True)
+
+def get_vertices(m):
+    m = np.copy(m)
+    top_npix = len(m)
+    top_nside = hp.npix2nside(top_npix)
+    top_order = int(np.log2(top_nside))
+    for order in range(top_order + 1):
+        nside = 1 << order
+        npix = hp.nside2npix(nside)
+        stride = 1 << (2 * (top_order - order))
+        keep = (hp.pix2vec(nside, np.arange(npix), nest=True) * np.expand_dims(xyz0, 1)).sum(0) >= np.cos(np.deg2rad(30))
+        if order < top_order:
+            mm = m.reshape((-1, stride))
+            keep &= (mm[:, :-1] == mm[:, 1:]).all(axis=1)
+            m += hp.ud_grade(np.where(keep, np.nan, 0), nside_out=top_nside, order_in='NEST', order_out='NEST')
+        else:
+            keep &= ~np.isnan(m)
+        for ipix in np.flatnonzero(keep):
+            boundaries = hp.boundaries(nside, ipix, nest=True, step=1).T
+            theta, phi = hp.vec2ang(boundaries)
+            ra = np.rad2deg(phi)
+            dec = np.rad2deg(0.5 * np.pi - theta)
+            vertices = np.column_stack((ra, dec))
+            yield vertices
+vertices = list(get_vertices(prob))
[…]

### 2015-07-07 ms.tex
Add caption

@@
 In Section~\ref{sec:marginal-posterior}, we explain the integration scheme by which the posterior probability distribution is calculated for a given sky location.
-In Section~\ref{sec:adaptive-sampling}, we show a scheme whereby the sky posterior is sampled on an adaptive HEALPix grid.
+In Section~\ref{sec:adaptive-sampling}, we show a scheme whereby the sky posterior is sampled on an adaptive \acs{HEALPix} grid.
 In Section~\ref{sec:run-time}, we report the running time of the algorithm on the hardware available on the \ac{LIGO} Data Grid.
@@
 
-Our sampling of the sky relies completely on the \acl{HEALPix} (\acs{HEALPix}), a special data structure designed for all\nobreakdashes-sky maps.
+Our sampling of the sky relies completely on the \acl{HEALPix} (\acsu{HEALPix}), a special data structure designed for all\nobreakdashes-sky maps.
 \ac{HEALPix} divides the sky into equal\nobreakdashes-area pixels.
@@
     \includegraphics[width=\columnwidth]{adaptive_mesh}
-    \caption{\label{fig:adaptive-mesh}}
+    \caption{\label{fig:adaptive-mesh}An example multi\nobreakdashes-resultion \ac{HEALPix} mesh arising from the \ac{BAYESTAR} sampling scheme. This is event 18951 from \cite{FirstTwoYears}.}
 \end{figure}

### 2015-07-07 ms.tex
Refer to autocorrelation likelihood figure
Rework acknowledgements

@@
 %
+Some example autocorrelation functions and corresponding likelihoods are shown in Figure~\ref{fig:autocorr-likelihood}.
 To assemble the joint likelihood for the whole network, we form the product of the autocorrelation likelihoods from the individual detectors:
@@
 \begin{acknowledgements}
+This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v1.
+
+LIGO was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the \ac{NSF} and operates under cooperative agreement PHY\nobreakdashes-0107417.
+
+This research is supported by the \ac{NSF} through a Graduate Research Fellowship to L.S.
+
+L.S. thanks the Aspen Center for Physics and \ac{NSF} Grant \#1066293 for hospitality during the editing of this paper.
+
 Source code for \ac{BAYESTAR} is available as part of LALInference, the open source LIGO/Virgo parameter estimation toolchain.
@@
 
-LIGO was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the \ac{NSF} and operates under cooperative agreement PHY\nobreakdashes-0107417.
 Some results were produced on the NEMO computing cluster operated by the Center for Gravitation and Cosmology at University of Wisconsin\nobreakdashes--Milwaukee under \ac{NSF} Grants PHY\nobreakdashes-0923409 and PHY\nobreakdashes-0600953.
-This research is supported by the \ac{NSF} through a Graduate Research Fellowship to L.S.
-This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v1.
 

### 2015-07-07 plot_runtimes.py
Adjust axes limits for runtime plot

@@ -14,7 +14,7 @@ plt.setp(artists['bodies'], color='blue')
 ax.xaxis.set_ticks(range(6))
 ax.xaxis.set_major_formatter(ticker.FuncFormatter(lambda x, _: str(2**x)))
 ax.set_yscale('log')
-ax.set_ylim(10, 3000)
+ax.set_ylim(1, 1000)
 ax.set_xlabel('threads')
 ax.set_ylabel('run time (s)')
 plt.subplots_adjust(bottom=0.15)

### 2015-07-07 ms.tex
Describe two observing runs

@@
 Because the X\nobreakdashes-ray and optical signatures can be faint and fade rapidly, observing them hinges on rapidly inferring the sky location from the gravitational wave observations.
-MCMC methods for gravitational\nobreakdashes-wave parameter estimation can take hours or more.
-We introduce BAYESTAR, a rapid, Bayesian, non-MCMC sky localization algorithm that takes just seconds to produce probability sky maps that are comparable in accuracy to the full analysis.
+\ac{MCMC} methods for gravitational\nobreakdashes-wave parameter estimation can take hours or more.
+We introduce BAYESTAR, a rapid, Bayesian, non-\ac{MCMC} sky localization algorithm that takes just seconds to produce probability sky maps that are comparable in accuracy to the full analysis.
 Prompt localizations from BAYESTAR will make it possible to search electromagnetic counterparts of compact binary mergers.
@@
 
-The innermost integral over distance $r$ may be written as
+The innermost integral over distance\footnote{The distance integral is now performed differently from what we initially described in \cite{leo-singer-thesis}; the method described in the present work is about an order of magnitude faster.} $r$ may be written as
 %
@@
     \includegraphics[width=\columnwidth]{adaptive_mesh}
-    \caption{\label{fig:adaptive-mesh}An example multi\nobreakdashes-resultion \ac{HEALPix} mesh arising from the \ac{BAYESTAR} sampling scheme. This is event 18951 from \cite{FirstTwoYears}.}
+    \caption{\label{fig:adaptive-mesh}An example multi\nobreakdashes-resultion \ac{HEALPix} mesh arising from the \ac{BAYESTAR} sampling scheme (plotted in a cylindrical projection).
+This is event 18951 from \cite{FirstTwoYears}.}
 \end{figure}
@@
 We have completed our description of the \ac{BAYESTAR} algorithm.
-In the following chapter, we will describe the first results with \ac{BAYESTAR}, a detailed case study of the sky localization capabilities of the earliest configurations of Advanced \ac{LIGO} and Virgo.
-Because it is typically three orders of magnitude faster than the conventional \ac{MCMC} analysis, for the first time we are able to generate detailed sky maps for thousands of simulated \ac{GW} sources.
-With such a large sample, we can for the first time unambiguously describe areas and morphologies that should be typical of networks of just two detectors.
-This case study will be the subject of the next chapter.
+In \cite{FirstTwoYears}, the authors presented a comprehensive and astrophysically realistic sample of simulated \ac{BNS} mergers.
+We focused on the first two planned Advanced LIGO and Virgo observing runs as desribed in \cite{LIGOObservingScenarios}.
+That work presented a catalog of 500 sky localizations from \ac{BAYESTAR} and LALINFERENCE, and dealt with the quantitave position reconstruction accuracy as well as the qualitative sky morphologies.
+In the present work, we will use the same data set, but instead focus on demonstrating the correctness and performance of the \ac{BAYESTAR} algorithm.
+
+\subsection{Observing scenarios}
+
+To review the assumptions made in \cite{FirstTwoYears}, the two scenarios are:
+
+\paragraph*{2015} The first Advanced \acs{LIGO} observing run, or ``O1'', scheduled to start in September 2015 and continue for three months.
+There are only two detectors participating in this run: \acs{LIGO} Hanford (``H'') and \acs{LIGO} Livingston (``L'').
+Both detectors are expected to operate with a direction\nobreakdashes-averaged \ac{BNS} merger range of 40\nobreakdashes--80\,Mpc (though ongoing Advanced \acs{LIGO} commissioning suggests that the higher end of this range will be achieved).
+As a result of having only two detectors, most localizations are long, thin arcs a few degrees wide and tens to hundreds of degrees long.
+The median 90\% credible area is about 600\,deg$^2$.
+
+\paragraph*{2016} The second observing run, ``O1'', with the two between Advanced \acs{LIGO} detectors, upgraded to a \ac{BNS} range of 80\nobreakdashes--120\,Mpc, operated jointly with the newly commissioned Advanced Virgo detector (``V''), operating at a range of 20\nobreakdashes--60\,Mpc.
+The run is envisioned as lasting for six months in 2016\nobreakdashes--2017.
+The detectors are assumed to have random and independent 80\% duty cycles.
+Consequently, all three detectors (``HLV'') are in science mode about half of the time, with the remaining time divided roughly equally between each of the possible pairs (``HL'', ``HV'', or ``LV'') and one or fewer detectors (at least two \ac{GW} facilities are required for a detection).
+Virgo's range is assumed to be somewhat less than \acs{LIGO}'s because its commissioning timetable is about a year behind.
+Although the simulated signals are generally to weak in Virgo to trigger the matched\nobreakdashes-filter pipeline and contribute to \emph{detection}, even these sub\nobreakdashes-threshold signals aid in position reconstruction with LALINFERENCE by lifting degeneracies.
+As a result, the median 90\% credible area decreases to about 200\,deg$^2$.
+
+\subsection{Self-consistency of credible levels}
+
+One might naively expect that self\nobreakdashes-consistency would require the 90\% confidence area and searched area histograms to intersect at 90\% of detections, but this is not generally required because the posteriors of different events have widely different dimensions.
+However, it is true that 90\% of sources should be found within their respective 90\% confidence contours.
+This can be formalized into a graphical self\nobreakdashes-consistency test called a probability\nobreakdashes--probability ($P$\nobreakdashes--$P$) plot (see \citealt{SiderySkyLocalizationComparison} for applications in \ac{GW} parameter estimation).
+For each event, one follows the water\nobreakdashes-filling procedure to walk outward from the most probable pixel until one reaches the pixel that contains the true sky location, noting the total probability $p$ that has been traversed.
+
+A cumulative histogram of $p$ should be diagonal, within a binomial confidence band.
+It is already well established that LALINFERENCE\_NEST/MCMC/BAMBI satisfy the $P$\nobreakdashes--$P$ plot test when deployed with accurate templates and reasonable priors.
+We found at first that \ac{BAYESTAR}'s $P$\nobreakdashes--$P$ plots tended to sag below the diagonal, indicating that though the accuracy (i.e., searched area) was comparable to the stochastic samplers, the \emph{precision} was overstated with confidence intervals that were only about 70\% of the correct area.
+This was rectified by pre\nobreakdashes-scaling the \acp{SNR} from GSTLAL\_INSPIRAL by a factor of 0.83 prior to running \ac{BAYESTAR}.
+This correction factor suggests that, for example, an \ac{SNR} 10 trigger from GSTLAL\_INSPIRAL has the \emph{effective information content} of an \ac{SNR} 8.3 signal.
+The missing information may be due to losses from the discreteness of the template bank, from the \ac{SVD}, from mismatch between the matched\nobreakdashes-filter templates and the simulated signals, from the small but nonzero correlations between masses and intrinsic parameters, or from elsewhere within the detection pipeline.
+The correction is hard\nobreakdashes-coded into the rapid localization.
+With it, the $P$\nobreakdashes--$P$ plots are diagonalized without negatively affecting the searched area (see Figure~\ref{fig:pp}).
+
+\begin{figure}
+    \begin{center}
+        \includegraphics{pp}
+    \end{center}
+    \caption[\acs{BAYESTAR} $P$\nobreakdashes--$P$ plots]{\label{fig:pp}$P$\nobreakdashes--$P$ plots for \ac{BAYESTAR} localizations in the 2015 and 2016 configurations.
+The gray lozenge around the diagonal is a target 95\% confidence band derived from a binomial distribution.}
+\end{figure}
 
@@
 First, at any number of threads, the two configurations have similar run times, although the 2016 events contain a subpopulation of outliers that take about 2.5 times as long as the 2015 events.
-These are probably due to taking one of the more expensive code branches in the distance integral.
+These are probably due to taking one of the more expensive code branches in the distance integral interpolation.
 Second, the run times decrease proportionally to the number of threads.
@@
 Based on experiences running \ac{BAYESTAR} on the 32\nobreakdashes-core (64 threads with hyperthreading) cluster login machine, we expect the almost ideal parallel speedup to continue on machines with more threads.
-With just one thread, the \ac{BAYESTAR} analysis takes 500\nobreakdashes--1500~s, already orders of magnitude faster than the full parameter estimation.
+With just one thread, the \ac{BAYESTAR} analysis takes 100\nobreakdashes--400~s, already orders of magnitude faster than the full parameter estimation.
 With 32 threads, \ac{BAYESTAR} takes just 20\nobreakdashes--60~s.

### 2015-07-08 ms.tex
Add area histogram

@@
 The \ac{BAYESTAR} adaptive sampling process works as follows.
-W begin by evaluating the posterior probability density at the center of each of the $N_{\mathrm{pix},0} = 3072$ pixels of an $N_{\mathrm{side},0}=16$ \ac{HEALPix} grid.
+We begin by evaluating the posterior probability density at the center of each of the $N_{\mathrm{pix},0} = 3072$ pixels of an $N_{\mathrm{side},0}=16$ \ac{HEALPix} grid.
 At this resolution, each pixel has an area of 13.4~deg$^2$.
@@
 Within each iteration, all of the marginalization integrals may be evaluated in parallel.
-In our C language implementation, the inner loop over pixels is parallelized with OpenMP\footnote{http://openmp.org/}.
+In our C language implementation of \ac{BAYESTAR}, the inner loop over pixels is parallelized with OpenMP\footnote{http://openmp.org/}.
 
@@
 
-\subsection{Self-consistency of credible levels}
-
-One might naively expect that self\nobreakdashes-consistency would require the 90\% confidence area and searched area histograms to intersect at 90\% of detections, but this is not generally required because the posteriors of different events have widely different dimensions.
+All simulated sources have component masses distributed uniformly between 1.2 and 1.6\,$M_\odot$ and randomly oriented spins with dimensionless magnitudes $\chi$ between -0.05 and +0.05.
+Sky positions and binary orientations are random and isotropic.
+Distances are drawn uniformly from ${D_\mathrm{L}}^3$, reflecting a uniform source population (neglecting cosmological effects, which are small within the Advanced \acs{LIGO} \ac{BNS} range).
+
+The simulated waveforms were deposited in Gaussian noise that has been filtered to have the \acp{PSD} consistent with \cite{LIGOObservingScenarios}.
+They were detected using the real\nobreakdashes-time matched\nobreakdashes-filter pipeline, GSTLAL\_INSPIRAL \citep{Cannon:2011vi}.
+Candidates with estimated \acp{FAR} less than $10^{-2}$\,yr$^{-1}$ were considered to be detections.
+Because using Gaussian noise results in lower \acp{FAR} than would be calculated in realistically glitchy detector noise, we imposed an additional detection threshold on the network \ac{SNR}, $\hat{\rho} \geq 12$, which has been found to correspond to a comparable \ac{FAR} in the initial \ac{LIGO} runs.
+For further analysis of the effect of glitchy noise on detection and sky localization, see \cite{BerryLocalization}.
+
+\subsection{Areas}
+
+\begin{figure*}
+    \includegraphics{area-hist}
+    \caption{\label{fig:area-hist}Cumulative histograms of sky area, broken down by observing run and detector network.
+The plots in the left column show the 90\% credible area and the plots in the right column show the searched area.
+From top bottom, the rows refer to the following observing scenarios/network configurations: 2015/HL, 2016/HL, 2016/HV, 2016/LV, 2016/HLV.
+The shaded regions represent the 95\% confidence bounds.
+The magenta lines are represent \ac{BAYESTAR} and the blue lines LALINFERENCE.
+Where relevant, dotted lines show all events in the given network configuratoin and solid lines show only events for which the matched\nobreakdashes-filter pipeline triggered on all operating detectors.
+Note that statistically significant differences in areas between the \ac{BAYESTAR} and LALINFERENCE localizations occur only for events that were below the detection threshold in one or more detectors.}
+\end{figure*}
+
+\subsection{Self-consistency}
+
+\textbf{need to rewrite this paragraph, copied from first2years paper} One might naively expect that self\nobreakdashes-consistency would require the 90\% confidence area and searched area histograms to intersect at 90\% of detections, but this is not generally required because the posteriors of different events have widely different dimensions.
 However, it is true that 90\% of sources should be found within their respective 90\% confidence contours.

### 2015-07-08 scripts/area-hist.py

@@ -1,0 +1,85 @@
+from __future__ import division
+from __future__ import print_function
+import numpy as np
+import scipy.stats
+from matplotlib import pyplot as plt
+from matplotlib import ticker as mticker
+from matplotlib import gridspec
+import astropy.table
+import os.path
+import operator
+# from common import *
+
+keys = ('area90', 'searched-area')
+labels = (r'area of 90\% confidence region (deg$^2$)', r'searched area (deg$^2$)')
+# colors = np.asarray([[230, 159, 0], [86, 180, 233]]) / 255
+colors = [[0.58, 0.15, 0.47], [0.19, 0.54, 0.71]]
+
+fig = plt.figure(figsize=(6, 8.5))
+gs = gridspec.GridSpec(5, 2, wspace=0.15, hspace=0.2)
+
+for irun, run in enumerate(['2015', '2016']):
+    filename = '../cbc/emfollowup/papers/first2years/data/{run}/{run}_coinc.txt'.format(run=run)
+    data = astropy.table.Table.read(filename, format='ascii.cds')
+
+    # Select only events that have LALInference localizations
+    data = data[~data['pe-area90'].mask]
+
+    for inet, net in enumerate(['HL', 'HV', 'LV', 'HLV']):
+        group = data[data['network'] == net]
+        if len(group) <= 0:
+            continue
+        for ikey, (key, label) in enumerate(zip(keys, labels)):
+            # plt.figure(figsize=(3.5, 2.5))
+            # plt.subplots_adjust(left=0.175, right=0.85, top=0.95, bottom=0.175)
+            ax = fig.add_subplot(gs[irun + inet, ikey])
+            ax.set_xlim(10, 1000)
+            ax.set_xscale('log')
+            ax.set_ylim(0, 100)
+            ax.yaxis.set_major_formatter(mticker.FormatStrFormatter(r'%d\%%'))
+            for complete, linestyle in zip([False, True], [':', '-']):
+                if not complete and len(net) <= 2:
+                    continue
+                if complete:
+                    group2 = group[reduce(operator.and_, [~group['snr-' + ifo].mask for ifo in net])]
+                else:
+                    group2 = group
+
+                count_simulated_detections = float(len(group2))
+                counts = np.arange(0, len(group2) + 1)
+                pct = counts / count_simulated_detections * 100
+                pct = np.concatenate((pct, [pct[-1]]))
[…]

### 2015-07-08 ms.tex
Add P-P plots

@@
 
-Adding to the challenge of detecting a faint, short-lived optical transient, there is an extreme mismatch between the sky localization accuracy of \ac{GW} detector networks, $\sim 10--500$~deg$^2$ \citep{FairhurstTriangulation,WenLocalizationAdvancedLIGO,FairhurstLocalizationAdvancedLIGO,2011PhRvD..84j4020V,RodriguezBasicParameterEstimation,NissankeLocalization,NissankeKasliwalEMCounterparts,KasliwalTwoDetectors,Grover:2013,SiderySkyLocalizationComparison,FirstTwoYears}, and the \acp{FOV} of 1\nobreakdashes--8\,m\nobreakdashes-class optical telescopes.
-Wide\nobreakdashes-field optical transient facilites such as BlackGEM (0.6\,m/2.7\,deg$^2$), the \acl{ZTF} \cite{ZTF} (1.2\,m/47\,deg$^2$), DECam (4\,m/3\,deg$^2$), or the \acf{LSST} (8.4\,m/9.6\,deg$^2$), operated in ``target of opportunity'' mode, may be able to tile these large areas rapidly enough to find the one needle in the haystack that is connected to the \ac{GW} event.
+Adding to the challenge of detecting a faint, short-lived optical transient, there is an extreme mismatch between the sky localization accuracy of \ac{GW} detector networks, $\sim$10\nobreakdashes--500\,deg$^2$ \citep{FairhurstTriangulation,WenLocalizationAdvancedLIGO,FairhurstLocalizationAdvancedLIGO,2011PhRvD..84j4020V,RodriguezBasicParameterEstimation,NissankeLocalization,NissankeKasliwalEMCounterparts,KasliwalTwoDetectors,Grover:2013,SiderySkyLocalizationComparison,FirstTwoYears}, and the \acp{FOV} of 1\nobreakdashes--8\,m\nobreakdashes-class optical telescopes.
+Wide\nobreakdashes-field optical transient facilites such as BlackGEM~(0.6\,m/2.7\,deg$^2$), the \acl{ZTF}~(\acsu{ZTF}; 1.2\,m/47\,deg$^2$) \cite{ZTF}, the \acl{DECAM}~(\acsu{DECAM}; 4\,m/3\,deg$^2$), or the \acl{LSST}~(\acsu{LSST}; 8.4\,m/9.6\,deg$^2$), operated in ``target of opportunity'' mode, may be able to tile these large areas rapidly enough to find the one needle in the haystack that is connected to the \ac{GW} event.
 However, prompt and accurate \ac{GW} position reconstructions will be of the utmost importance for guiding the selection of fields to observe.
@@
 \subsection{Properties}
+\label{sec:properties}
 
@@
 
+\subsection{Detection and localization}
+
 The simulated waveforms were deposited in Gaussian noise that has been filtered to have the \acp{PSD} consistent with \cite{LIGOObservingScenarios}.
 They were detected using the real\nobreakdashes-time matched\nobreakdashes-filter pipeline, GSTLAL\_INSPIRAL \citep{Cannon:2011vi}.
-Candidates with estimated \acp{FAR} less than $10^{-2}$\,yr$^{-1}$ were considered to be detections.
-Because using Gaussian noise results in lower \acp{FAR} than would be calculated in realistically glitchy detector noise, we imposed an additional detection threshold on the network \ac{SNR}, $\hat{\rho} \geq 12$, which has been found to correspond to a comparable \ac{FAR} in the initial \ac{LIGO} runs.
-For further analysis of the effect of glitchy noise on detection and sky localization, see \cite{BerryLocalization}.
+Candidates with estimated \acp{FAR} less than $10^{-2}$\,yr$^{-1}$ were considered to be ``detections.'' Because using Gaussian noise results in lower \acp{FAR} than would be calculated in realistically glitchy detector noise, we imposed an additional detection threshold on the network \ac{SNR}, $\hat{\rho} \geq 12$, which has been found to correspond to a comparable \ac{FAR} in the initial \ac{LIGO} runs.\footnote{See \cite{BerryLocalization} for an analysis of the effect of glitchy noise on detection and parameter estimation.} Localizations for the ``detections'' were generated with \ac{BAYESTAR} as well as the functionally equivalent and interchangeable LALINFERENCE\_MCMC, LALINFERENCE\_NEST, and LALINFERENCE\_BAMBI samplers (collectively referred to as LALINFERENCE).
 
 \subsection{Areas}
+
+We measured sky localization areas for each event as follows.
+First, we ranked the \ac{HEALPix} pixels by descending posterior probability.
+Then, we computed the cumulative sum of the pixels in that order.
+Finally, we searched for the index of the pixel whose cumulative sum was equal to a given value: for example, 0.9 if we are interested in the 90\% credible area.
+That pixel index times the area per pixel is the area of the smallest region of the specified credible level.
+
+We can construct a second measure, called the searched area, as the smallest such constructed area that contains the true location of the source.
+A telescope with a \ac{FOV} that is small compared to the characteristic scale of the posterior would intercept the true location of the source after covering the searched area.
+This measure is mainly useful because it measures the accuracy of the sky localization independently of the precision: that is to say, it treats the sky map as merely a ranking statistic.
+
+Histograms of the 90\% credible area and the searched area are shown in Figure~\ref{fig:area-hist}, broken down by observing scenario (2015 or 2016) and detector network (HL, HV, LV, or HLV).
+Note that there are no stastically significant differences in areas between \ac{BAYESTAR} and LALINFERENCE, with the exception in the 2016/HLV configuration, for which some LALINFERENCE sky maps span about an order of magnitude less area than \ac{BAYESTAR}.
+If we consider only events for which all three detectors contained a signal that was loud enough to trigger the matched\nobreakdashes-filter pipeline, the difference becomes much smaller and insignificant within 95\% error bars.
+This is because if the signal is too weak to trigger the detection pipeline in one of the detectors, then \ac{BAYESTAR} receives no information about that detector.
+This issue does not occur in the two\nobreakdashes-detector configurations (HL, HV, or LV) because two or more triggers are required to report a detection candidate.
 
@@
 
+This is a significant issue for the 2016 configuration, because the most accurate localizations are possible when all three detectors are operating.
+However, there may be a simple remedy.
+As we noted in Section~\ref{sec:properties}, the \ac{BAYESTAR} likelihood can be modified to use, instead of the times, phases, and amplitudes on arrival, the full complex matched\nobreakdashes-filter time series from all detectors.
+The detection pipeline, GSTLAL\_INSPIRAL, would have to be modifed to save and transmit a small interval of the complex \ac{SNR} time series (perhaps a few tens of milliseconds) around the time of each detection candidate.
+In addition to supplying the missing information for sub\nobreakdashes-threshold signals, this would make \ac{BAYESTAR} mathematically equivalent to the LALINFERENCE analysis, but with the intrinsic parameters fixed to their maximum\nobreakdashes-likelihood values.
+This idea will be pursued in future work.
+
 \subsection{Self-consistency}
@@
 A cumulative histogram of $p$ should be diagonal, within a binomial confidence band.
-It is already well established that LALINFERENCE\_NEST/MCMC/BAMBI satisfy the $P$\nobreakdashes--$P$ plot test when deployed with accurate templates and reasonable priors.
+It is already well established that LALINFERENCE localizations satisfy the $P$\nobreakdashes--$P$ plot test when deployed with accurate templates and reasonable priors.
 We found at first that \ac{BAYESTAR}'s $P$\nobreakdashes--$P$ plots tended to sag below the diagonal, indicating that though the accuracy (i.e., searched area) was comparable to the stochastic samplers, the \emph{precision} was overstated with confidence intervals that were only about 70\% of the correct area.
@@
 
-\begin{figure}
+\begin{figure*}
     \begin{center}
-        \includegraphics{pp}
+        \includegraphics[width=0.45\textwidth]{2015-pp}
+        \includegraphics[width=0.45\textwidth]{2016-pp}
     \end{center}
-    \caption[\acs{BAYESTAR} $P$\nobreakdashes--$P$ plots]{\label{fig:pp}$P$\nobreakdashes--$P$ plots for \ac{BAYESTAR} localizations in the 2015 and 2016 configurations.
+    \caption[\acs{BAYESTAR} $P$\nobreakdashes--$P$ plots]{\label{fig:pp}$P$\nobreakdashes--$P$ plots for \ac{BAYESTAR} and LALINFERENCE localizations in the 2015 and 2016 configurations.
 The gray lozenge around the diagonal is a target 95\% confidence band derived from a binomial distribution.}
-\end{figure}
+\end{figure*}
 
@@
 The averages and the masses are correlated with each other, but are not correlated with the differences.
-Because only the differences are informative for sky localization, we drop the intrinsic parameters from the rest of the Fisher matrix calculations in this chapter.
+Because only the differences are informative for sky localization, we drop the intrinsic parameters from the rest of the Fisher matrix calculations in this appendix.
 

### 2015-07-08 ms.tex
Update parallelization and run time sections

@@
 \subsection{Integral over distance}
-
-The innermost integral over distance\footnote{The distance integral is now performed differently from what we initially described in \cite{leo-singer-thesis}; the method described in the present work is about an order of magnitude faster.} $r$ may be written as
+\label{sec:distance}
+
+The distance integral is now performed differently from what we initially described in \cite{leo-singer-thesis,BerryLocalization}; the method described in the present work is about an order of magnitude faster.
+The innermost integral over distance $r$ may be written as
 %
@@
 
-Within each iteration, all of the marginalization integrals may be evaluated in parallel.
-In our C language implementation of \ac{BAYESTAR}, the inner loop over pixels is parallelized with OpenMP\footnote{http://openmp.org/}.
+\section{Parallelization}
+
+\ac{MCMC} and similar stochastic schemes are typically very resistant to parallelization.
+However, \ac{BAYESTAR} is completely deterministic and easily parallelizable because each pixel can be evaluated independently from all of the others.
+\ac{BAYESTAR} consists of nine computationally intensive loops: the generation of the distance integral lookup table and the eight loops over pixels in the adaptive HEALPix sampling step.
+The iterations of each loop are distributed across multiple cores using OpenMP\footnote{http://openmp.org/} compiler directives.
+In Section~\ref{sec:run-time}, we will show that \ac{BAYESTAR}'s run time is almost perfectly proportional to the number of cores, demonstrating that the serial sections (the sorts between the adaptation steps) are a negligible contribution to the overall wall clock time.
 
@@
 That pixel index times the area per pixel is the area of the smallest region of the specified credible level.
+This area can be thought of measuring the precision of the sky localization: it is a measure of the scale of the posterior distribution.
 
@@
 A telescope with a \ac{FOV} that is small compared to the characteristic scale of the posterior would intercept the true location of the source after covering the searched area.
-This measure is mainly useful because it measures the accuracy of the sky localization independently of the precision: that is to say, it treats the sky map as merely a ranking statistic.
+This measure is mainly useful because it measures the accuracy of the sky localization independently of the precision.
+In other words, it treats the sky map as merely a ranking statistic.
 
@@
 
-\subsection{Self-consistency}
-
-\textbf{need to rewrite this paragraph, copied from first2years paper} One might naively expect that self\nobreakdashes-consistency would require the 90\% confidence area and searched area histograms to intersect at 90\% of detections, but this is not generally required because the posteriors of different events have widely different dimensions.
-However, it is true that 90\% of sources should be found within their respective 90\% confidence contours.
-This can be formalized into a graphical self\nobreakdashes-consistency test called a probability\nobreakdashes--probability ($P$\nobreakdashes--$P$) plot (see \citealt{SiderySkyLocalizationComparison} for applications in \ac{GW} parameter estimation).
-For each event, one follows the water\nobreakdashes-filling procedure to walk outward from the most probable pixel until one reaches the pixel that contains the true sky location, noting the total probability $p$ that has been traversed.
-
-A cumulative histogram of $p$ should be diagonal, within a binomial confidence band.
+\subsection{Self consistency}
+
+As we observed above, the area of a given credible region describes the precision of the sky localization, whereas the searched area describes the accuracy.
+However, self consistency requires that the two are related.
+For example, we should find that on average 90\% of events have their true locations contained within their respective 90\% credible regions.
+More generally, if we make a cumulative histogram of the credible levels corresponding to the searched areas of all of the events, then we should obtain a diagonal line (with small deviations due to finite sample size).
+This test, popularized for \ac{GW} data analysis by \cite{SiderySkyLocalizationComparison}, is a necessary but not sufficient condition for the validity of any Bayesian parameter estimation scheme.
+
 It is already well established that LALINFERENCE localizations satisfy the $P$\nobreakdashes--$P$ plot test when deployed with accurate templates and reasonable priors.
-We found at first that \ac{BAYESTAR}'s $P$\nobreakdashes--$P$ plots tended to sag below the diagonal, indicating that though the accuracy (i.e., searched area) was comparable to the stochastic samplers, the \emph{precision} was overstated with confidence intervals that were only about 70\% of the correct area.
+We found at first that \ac{BAYESTAR}'s $P$\nobreakdashes--$P$ plots tended to sag below the diagonal, indicating that though the accuracy (i.e., searched area) was comparable to LALINFERENCE, the \emph{precision} was overstated, with confidence intervals that were only about 70\% of the correct area.
 This was rectified by pre\nobreakdashes-scaling the \acp{SNR} from GSTLAL\_INSPIRAL by a factor of 0.83 prior to running \ac{BAYESTAR}.
@@
 Based on experiences running \ac{BAYESTAR} on the 32\nobreakdashes-core (64 threads with hyperthreading) cluster login machine, we expect the almost ideal parallel speedup to continue on machines with more threads.
-With just one thread, the \ac{BAYESTAR} analysis takes 100\nobreakdashes--400~s, already orders of magnitude faster than the full parameter estimation.
-With 32 threads, \ac{BAYESTAR} takes just 20\nobreakdashes--60~s.
-This latency is comparable to the other stages (data aggregation, trigger generation, alert distribution) in the real\nobreakdashes-time \ac{BNS} analysis.
-The 32\nobreakdashes-thread configuration is representative of how \ac{BAYESTAR} might be deployed in early Advanced \ac{LIGO}.\footnote{\ac{BAYESTAR} has been successfully ported to the Intel's Many Integrated Core~(MIC) architecture and has been tested in a 500 thread configuration on a system with dual Intel Xeon Phi 5110P coprocessors.}
+
+With just one thread, the \ac{BAYESTAR} analysis takes 87\nobreakdashes--384~s, already orders of magnitude faster than the full parameter estimation.
+With 32 threads, \ac{BAYESTAR} takes just 4\nobreakdashes--15~s.
+This latency is comparable to the other stages (data aggregation, trigger generation, alert distribution) in the real\nobreakdashes-time \ac{BNS} analysis; therefore any significant further speedup would require significant changes to all of the analyses preceeding \ac{BAYESTAR}.
+The 32\nobreakdashes-thread configuration is representative of how \ac{BAYESTAR} might be deployed in early Advanced \ac{LIGO}.\footnote{\ac{BAYESTAR} has been successfully ported to the Intel's Many Integrated Core~(MIC) architecture and has been tested in a 500 thread configuration on a system with dual Intel Xeon Phi 5110P coprocessors.} For comparison, sky localization with LALINFERENCE takes about 100 hours \cite{BerryLocalization}.
+
+Note that this benchmark shows \ac{BAYESTAR} to be an order of magnitude faster than what was reported in \cite{leo-singer-thesis,BerryLocalization} due to the changes in the distance integration scheme that we noted in Section~\ref{sec:distance}.
 

### 2015-07-08 ms.tex
Add conclusion

@@
 With 32 threads, \ac{BAYESTAR} takes just 4\nobreakdashes--15~s.
-This latency is comparable to the other stages (data aggregation, trigger generation, alert distribution) in the real\nobreakdashes-time \ac{BNS} analysis; therefore any significant further speedup would require significant changes to all of the analyses preceeding \ac{BAYESTAR}.
+In practice, \ac{BAYESTAR}'s data handling (reading the detectors' \acp{PSD}, communicating with the \ac{GW} candidate database, writing FITS files) take an additional 15~s or so, though this overhead could be reduced by parallelizing many of these steps.
+The overall latency is comparable to the other stages (data aggregation, trigger generation, alert distribution) in the real\nobreakdashes-time \ac{BNS} analysis; therefore any significant further speedup would require significant changes through Advanced \acs{LIGO} computing and infrastructure.
 The 32\nobreakdashes-thread configuration is representative of how \ac{BAYESTAR} might be deployed in early Advanced \ac{LIGO}.\footnote{\ac{BAYESTAR} has been successfully ported to the Intel's Many Integrated Core~(MIC) architecture and has been tested in a 500 thread configuration on a system with dual Intel Xeon Phi 5110P coprocessors.} For comparison, sky localization with LALINFERENCE takes about 100 hours \cite{BerryLocalization}.
@@
 Note that this benchmark shows \ac{BAYESTAR} to be an order of magnitude faster than what was reported in \cite{leo-singer-thesis,BerryLocalization} due to the changes in the distance integration scheme that we noted in Section~\ref{sec:distance}.
+
+\section{Conclusion}
+
+We have presented a novel, fast, accurate, Bayesian algorithm for inferring the sky locations of compact binary merger sources that may soon be detected by advanced ground\nobreakdashes-based \ac{GW} detectors.
+For \ac{BNS} systems with small spins, we have shown that \ac{BAYESTAR} produces sky maps that are as accurate as the full \ac{MCMC} parameter estimation code, but can do so within $\sim$10 seconds after a detection.
+Still faster response times should be possible in the future (if warranted) by deploying \ac{BAYESTAR} on machines with more cores, or by distributing \ac{BAYESTAR} across multiple computers.
+
+Following a \ac{BNS} merger, the signal will be detected by the matched\nobreakdashes-filter pipeline within tens of seconds; an alert containing the time and estimated significance of the event can be distributed almost immediately (although a human validation stage that may be present at the beginning of the first observing run may introduce some additional latency).
+The localization from \ac{BAYESTAR} will be available tens of seconds to a minute later.
+Finally, the refined localiation and the detailed estimates of masses and spins from LALINFERENCE will be distributed hours to days later.
+
+Relevant timescales for possible \ac{EM} counterparts to \ac{GW} signals include seconds (the prompt \ac{GRB} signature), hundreds of seconds (extended emission and X\nobreakdashes-ray plateaus that are observed for some short \acp{GRB}), minutes to hours (X\nobreakdashes-ray and optical afterglow), hours to days (the kilonova or the blue flashes associated with unbound ejecta or disk winds), and days to years (the radio afterglow).
+For the first time, we are able to provide accurate localizations before the peak of any of these \ac{EM} signatures (except for the short \ac{GRB} or any pre\nobreakdashes-merger signal).
+Even for components like the kilonova that should peak within hours to days, the availability of the localizations within seconds might provide a window of several hours to obtain tiled images of the area before the \ac{EM} emission begins.
+These could be used as reference images, crucial at optical wavelengths for establishing the rapid rise and quickly distinguishing from slower background transients.
 

### 2015-07-08 ms.tex
Add some acknowledgements

@@
 
+We thank John Veitch and Will Farr for chairing a review of the analysis and code. We thank Britt Griswold for assistance with preparing Figure~\ref{fig:adaptive-illustration}.
+\end{acknowledgements}
+
 LIGO was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the \ac{NSF} and operates under cooperative agreement PHY\nobreakdashes-0107417.
 
-This research is supported by the \ac{NSF} through a Graduate Research Fellowship to L.S.
-
-L.S. thanks the Aspen Center for Physics and \ac{NSF} Grant \#1066293 for hospitality during the editing of this paper.
-
-Source code for \ac{BAYESTAR} is available as part of LALInference, the open source LIGO/Virgo parameter estimation toolchain.
-See \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/lalsuite.html} for further details.
-
-Some of the results in this paper have been derived using \ac{HEALPix} \cite{healpix}.
+This research was supported by the \ac{NSF} through a Graduate Research Fellowship to L.S.
+L.S.
+thanks the Aspen Center for Physics and \ac{NSF} Grant \#1066293 for hospitality during the editing of this paper.
+
+Source code for \ac{BAYESTAR} is available as part of LALInference\footnote{\url{https://ligo-vcs.phys.uwm.edu/cgit/lalsuite/tree/lalinference}}, the open source LIGO/Virgo parameter estimation toolchain, which is in turn part of LALSuite\footnote{\url{http://www.lsc-group.phys.uwm.edu/daswg/projects/lalsuite.html}}.
+This research made use of Astropy\footnote{\url{http://www.astropy.org}} \citep{astropy}, a community-developed core Python package for astronomy.
+Some of the results in this paper have been derived using \ac{HEALPix}\footnote{\url{http://healpix.sourceforge.net}} \cite{healpix}.
 
 Some results were produced on the NEMO computing cluster operated by the Center for Gravitation and Cosmology at University of Wisconsin\nobreakdashes--Milwaukee under \ac{NSF} Grants PHY\nobreakdashes-0923409 and PHY\nobreakdashes-0600953.
-
-We thank Britt Griswold for assistance with preparing Figure~\ref{fig:adaptive-illustration}.
-\end{acknowledgements}
 

### 2015-07-08 ms.tex
Add more eccentricity references

@@
 A \ac{CBC} source is specified by a vector of extrinsic parameters describing its position and orientation, and intrinsic parameters describing the physical properties of the binary components\footnote{This list of parameters involves some simplifying assumptions.
-Eccentricity is omitted: although it does play a major role in the evolution and waveforms of \ac{NSBH} and \ac{BBH} sources formed by dynamical capture~\citep{PhysRevD.87.043004}, \ac{BNS} systems formed by binary stellar evolution should almost always circularize due to tidal interaction~\citep{0004-637X-572-1-407} and later \ac{GW} emission~\citep{PhysRev.136.B1224} long before the inspiral enters LIGO's frequency range of $\sim$10\nobreakdashes--1000~kHz.
+Eccentricity is omitted: although it may play a major role in the evolution and waveforms of rare close binaries formed by dynamical capture~\citep{PhysRevD.87.043004,2013PhRvD..87l7501H,2014PhRvD..90h4016H}, \ac{BNS} systems formed by binary stellar evolution should almost always circularize due to tidal interaction~\citep{0004-637X-572-1-407} and later \ac{GW} emission~\citep{PhysRev.136.B1224} long before the inspiral enters LIGO's frequency range of $\sim$10\nobreakdashes--1000~kHz.
 Tidal deformabilities of the \acp{NS} are omitted because the signal imprinted by the companions' material properties is so small that it will only be detectable by an Einstein Telescope\nobreakdashes-class \ac{GW} observatory~\citep{PhysRevD.81.123016}.

### 2015-07-09 ms.tex
Correct system for benchmark

@@
 Since \ac{BAYESTAR} is designed as one of the final steps in the real\nobreakdashes-time \ac{BNS} search, it is important to characterize how long it takes to calculate a sky map.
-We compiled \ac{BAYESTAR} with the Intel~C~Compiler~(\texttt{icc}) at the highest architecture\nobreakdashes-specific optimization setting (\texttt{-xhost}).
-We timed it under Scientific~Linux~6 on a Supermicro~1027GR\nobreakdashes-TRFT system with dual Intel Xeon E5-2670 CPUs clocked at 2.60GHz, providing a total of 16 cores and capable of executing 32 threads simultaneously (with hyperthreading).
+We compiled \ac{BAYESTAR} with the Intel~C~Compiler~(\texttt{icc}) at the highest architecture\nobreakdashes-specific optimization setting (\texttt{-ipo~-O3~-xhost}).
+We timed it under Scientific~Linux~6.1 on a Supermicro SuperServer 6028TP\nobreakdashes-HTTR system with dual 8\nobreakdashes-core Intel Xeon E5\nobreakdashes-2630~v3 CPUs clocked at 2.40\,GHz, capable of executing 32 threads simultaneously (with hyperthreading).
 In Figure~\ref{fig:runtimes}, we show how long it took to calculate a localization with \ac{BAYESTAR} as the number of OpenMP threads was varied from 1 to 32.
@@
 Second, the run times decrease proportionally to the number of threads.
-A slight flattening at 32 threads may be due to increasing use of hyperthreading.
-Based on experiences running \ac{BAYESTAR} on the 32\nobreakdashes-core (64 threads with hyperthreading) cluster login machine, we expect the almost ideal parallel speedup to continue on machines with more threads.
+Based on experiences running \ac{BAYESTAR} on the 32\nobreakdashes-core (64 threads with hyperthreading) cluster login machine, we expect the almost ideal parallel speedup to continue on machines with even more processors.
 

### 2015-07-13 ms.tex
Fix typo

@@
 Virgo's range is assumed to be somewhat less than \acs{LIGO}'s because its commissioning timetable is about a year behind.
-Although the simulated signals are generally to weak in Virgo to trigger the matched\nobreakdashes-filter pipeline and contribute to \emph{detection}, even these sub\nobreakdashes-threshold signals aid in position reconstruction with LALINFERENCE by lifting degeneracies.
+Although the simulated signals are generally too weak in Virgo to trigger the matched\nobreakdashes-filter pipeline and contribute to \emph{detection}, even these sub\nobreakdashes-threshold signals aid in position reconstruction with LALINFERENCE by lifting degeneracies.
 As a result, the median 90\% credible area decreases to about 200\,deg$^2$.

### 2015-07-14 ms.tex
Explicitly write extrinsic-only likelihood

@@
 
-First, the autocorrelation likelihood has the elegant feature that if we were to replace the autocorrelation function with the \ac{SNR} time series for the best-matching template, $z(\tau; \hat{\mathbf{\theta}}_\mathrm{in})$, we would recover the likelihood for the full \ac{GW} time series, evaluated at the \ac{ML} estimate of the intrinsic parameters.
-The numerical scheme that we will develop is thus equally applicable for rapid, coincidence-based localization, or as a fast extrinsic marginalization step for the full parameter estimation.
+First, the autocorrelation likelihood has the elegant feature that if we were to replace the autocorrelation function with the \ac{SNR} time series for the best-matching template, $z(\tau; \hat{\mathbf{\theta}}_\mathrm{in})$, we would recover the likelihood for the full \ac{GW} time series, evaluated at the \ac{ML} estimate of the intrinsic parameters, \emph{viz.}:
+%
+\begin{equation}\label{eq:extrinsic-only-likelihood}
+    % p\left(z_i(\hat{\tau}, \hat{\bm\theta}_\mathrm{in})
+        % \middle| \rho_i, \gamma_i, \tau_i \right) \propto \\
+    \exp \left[ - \frac{1}{2} \sum_i {\rho_i}^2
+        + \sum_i \rho_i \Re \left\{ e^{-i \gamma_i} z_i^*(\tau_i)
+        \right\}
+    \right].
+\end{equation}
+%
+(We have ommitted the term $\int |Y_i(\omega)|^2/S(\omega)d\omega$, which takes the place of the earlyier ${\hat{\rho}_i}^2$ term and is only important for normalization.) The numerical scheme that we will develop is thus equally applicable for rapid, coincidence-based localization, or as a fast extrinsic marginalization step for the full parameter estimation.
 

### 2015-07-14 ms.tex
Some of Chris' comments

@@
 
-When (as in our assumptions) the likelihood is Gaussian, Equation~(\ref{eq:general-fisher-matrix}) simplifies to
+When (as in our assumptions) the likelihood is Gaussian,\footnote{This assumes that the merger occurs at a frequency outside the sensitive ``bucket'' of the detector's noise \ac{PSD}.
+There are additional terms if the \ac{GW} spectrum drops to zero within the sensitive bandwidth of the detector, as can be the case for \ac{NSBH} mergers; see \cite{2014CQGra..31o5005M}.} Equation~(\ref{eq:general-fisher-matrix}) simplifies to
 %

### 2015-07-14 ms.tex
Add future work section

@@
 
-All simulated sources have component masses distributed uniformly between 1.2 and 1.6\,$M_\odot$ and randomly oriented spins with dimensionless magnitudes $\chi$ between -0.05 and +0.05.
+All simulated sources have component masses distributed uniformly between 1.2 and 1.6\,$M_\odot$ and randomly oriented spins with dimensionless magnitudes $\chi = c|\mathbf{S}|/Gm^2$ between -0.05 and +0.05.
 Sky positions and binary orientations are random and isotropic.
@@
 Note that this benchmark shows \ac{BAYESTAR} to be an order of magnitude faster than what was reported in \cite{leo-singer-thesis,BerryLocalization} due to the changes in the distance integration scheme that we noted in Section~\ref{sec:distance}.
+
+\section{Future work}
+
+One immediately pressing direction for future work is to address the issue of sub\nobreakdashes-threshold signals, as this will be a major issue when Advanced Virgo comes online in 2016\nobreakdashes--2017.
+Using the full \ac{SNR} time series in place of the autocorrelation function seems like a promising avenue; implementing this requires some infrastructure changes to both the matched filter pipeline and \ac{BAYESTAR}.
+
+A more open question is how to account for spin precession.
+The simulations in \cite{FirstTwoYears} and in this paper featured extremely modest spins of $\chi \leq 0.05$, consistent with the fastest known pulsars in binaries \citep{2003Natur.426..531B,DetectingBNSSystemsWithSpin}.
+The signals were detected using a template bank that lacked spins entirely.
+Farr et al.
+(in preparation) show that using nonspinning \ac{BNS} templates for parameter estimation has negligible impact on sky localization.
+However, if one or both companions are spinning as fast as a millisecond pulsar, $\chi \sim 0.4$ \cite{FastestSpinningMillisecondPulsar}, or even near breakup, $\chi \sim 0.7$, then the orbital plane may precess; in this case, spins can no longer be neglected for detection \cite{DetectingBNSSystemsWithSpin} and may also be important for parameter estimation.
+Since spinning \ac{BNS} searches are still an active area of development, \ac{BAYESTAR}'s sky localization accuracy in this regime should be reexamined in the future.
 

### 2015-07-31 ms.tex
Bump DCC version
Add detector references

@@
 
-The \acf{LIGO} and Virgo will soon begin taking data \cite{LIGOObservingScenarios} in their `Advanced' configurations, ultimately increasing their reach in volume within the local Universe by three orders of magnitude as compared to their initial configurations through 2010.
+The \acl{LIGO} (\acsu{LIGO}; \cite{AdvancedLIGOHarry,AdvancedLIGO}) and Virgo \cite{AdvancedVirgo} will soon begin taking data \cite{LIGOObservingScenarios} in their `Advanced' configurations, ultimately increasing their reach in volume within the local Universe by three orders of magnitude as compared to their initial configurations through 2010.
 They will form the first parts of a sensitive global \ac{GW} detector network, to be later augmented by the Japanese KAGRA facility \cite{LCGT,KAGRAInterferometerDesign} and \ac{LIGO}\nobreakdashes--India \cite{LIGOIndia}.
@@
 \begin{acknowledgements}
-This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v1.
+This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v2.
 

### 2015-07-31 ms.tex
Add MBTA reference

@@
 The final science run of the initial LIGO and Virgo instruments saw the first joint search for \ac{GW} and electromagnetic emission from compact binaries.
-This involved several advances in the \ac{GW} data analysis \citep{CBCLowLatency}, including the first real\nobreakdashes-time matched\nobreakdashes-filter detection pipeline (\acs{MBTA}; \aclu{MBTA}), a semi\nobreakdashes-coherent, ad hoc rapid triangulation code (Timing++), and the first version of a rigorous Bayesian \ac{MCMC} parameter estimation code (LALINFERENCE; \citealt{S6PE})---all in service of the first search for X\nobreakdashes-ray \citep{SwiftFollowup} and optical \citep{OpticalImageAnalysis} counterparts of \ac{GW} triggers, by a consortium of facilities.
+This involved several advances in the \ac{GW} data analysis \citep{CBCLowLatency}, including the first real\nobreakdashes-time matched\nobreakdashes-filter detection pipeline (\acs{MBTA}, \aclu{MBTA}; \cite{LIGOVirgoInspiralPipelineComparison}), a semi\nobreakdashes-coherent, ad hoc rapid triangulation code (Timing++), and the first version of a rigorous Bayesian \ac{MCMC} parameter estimation code (LALINFERENCE; \citealt{S6PE})---all in service of the first search for X\nobreakdashes-ray \citep{SwiftFollowup} and optical \citep{OpticalImageAnalysis} counterparts of \ac{GW} triggers, by a consortium of facilities.
 Despite the technical achievements in the \ac{GW} data analysis, there was an undesirable tradeoff between the speed as well as accuracy of the rapid localization and the full parameter estimation: the former could analyze a detection candidate in minutes, whereas the latter took half a day; the latter decreased the area on the sky by a factor of 1/20 over the former but took 1000 times as long to run \cite{SiderySkyLocalizationComparison}.

### 2015-07-31 ms.tex
Cite effective Fisher matrix paper

@@
     \item The Fisher information matrix will provide clues as to suitable forms of this likelihood function.
-Recent \ac{GW} parameter estimation literature has largely rejected the Fisher matrix, but this is mostly on the grounds of the abuse of the related \ac{CRLB} outside its realm of validity (i.e., at low to moderate \ac{SNR}) \cite{BayesianBounds,UseAbuseFisherMatrix,FisherMatrixAsymptoticExpansions,InadequaciesFisherMatrix}.
+Recent \ac{GW} parameter estimation literature has largely rejected the Fisher matrix,\footnote{Though not entirely; see \cite{EffectiveFisherMatrix}.} but this is mostly on the grounds of the abuse of the related \ac{CRLB} outside its realm of validity (i.e., at low to moderate \ac{SNR}) \cite{BayesianBounds,UseAbuseFisherMatrix,FisherMatrixAsymptoticExpansions,InadequaciesFisherMatrix}.
 However, we recognize that the block structure of the Fisher matrix provides important insights, and is a useful quantity for checking the validity of the aforementioned likelihood function, quite independent of the \ac{CRLB}.

### 2015-08-10 ms.tex
WIP: Add some more future work stuff

@@
 Using the full \ac{SNR} time series in place of the autocorrelation function seems like a promising avenue; implementing this requires some infrastructure changes to both the matched filter pipeline and \ac{BAYESTAR}.
+Along these lines, we refer the reader to \cite{PankowRapidPE} for a similar, non-\ac{MCMC} approach to rapid exploration of the full parameter space.
+
+Although the response time of \ac{BAYESTAR} has been driven by the anticipated timescales for kilonova and afterglow emission, a recurring question is whether there is any detectable \ac{EM} signal in the seconds before, during, and after the merger itself.
+An obvious example would be prompt gamma\nobreakdashes-ray emission, though more exotic ...
+
+Since the \ac{GW} inspiral signal is in principal detectable for up to hundreds of seconds before merger, one could imagine positioning rapidly slewing instruments to search for prompt emission ...
 

### 2015-08-10 ms.tex
Add distance interpolant rate of convergence test

@@
 The interpolant is evaluated slightly differently depending on which of the three regions marked I, II, and II in Figure~\ref{fig:radial-integral-interpolant} contains the point of interest.
-In region I, we use bicubic interpolation in $x$ and $y$.
-In region II, we use univariate cubic interpolation in $x$, with the sample points taken from the horizontal boundary between regions I and II.
-In region III, we use univariate cubic interpolation in $u = (x - y) / 2$, with the sample points taken from the downward diagonal boundary between regions I and III.
-This entire lookup table scheme is accurate to an absolute error of about $10^{-5}$ in $\mathscr{G}$.
+In region I, we use bicubic interpolation of $\log \mathcal{G}$ in $x$ and $y$.
+In region II, we use univariate cubic interpolation of $\log \mathcal{G}$ in $x$, with the sample points taken from the horizontal boundary between regions I and II.
+In region III, we use univariate cubic interpolation of $\log \mathcal{G}$ in $u = (x - y) / 2$, with the sample points taken from the downward diagonal boundary between regions I and III.
 Finally, the distance integral $\mathscr{F}$ is obtained by multiplying the interpolated value of $\mathscr{G}$ by $\exp\left(p^2/{r_0}^2\right)$.
+For a $400\times400$ grid, the entire lookup table scheme is accurate to a relative error of about $10^{-5}$ in $\mathscr{F}$ (see Figure~\ref{fig:distance_interpolant_convergence}).
+
+\begin{figure}
+    \begin{center}
+        \includegraphics[width=\columnwidth]{distance_interpolant_convergence}
+    \end{center}
+    \caption{\label{fig:distance_interpolant_convergence}Relative error in \ac{BAYESTAR} distance integral interpolation scheme as a function of the size of the grid.}
+\end{figure}
 

### 2015-08-10 scripts/distance_interpolant_convergence.py

@@ -1,0 +1,40 @@
+import numpy as np
+from matplotlib import pyplot as plt
+from lalinference.bayestar.sky_map import LogRadialIntegrator
+from astropy.utils.console import ProgressBar
+
+fig = plt.figure(figsize=(3.5, 2.5))
+ax = fig.add_subplot(111)
+
+r1 = 0.0
+r2 = 0.25
+k = 2
+pmax = 1.0
+
+np.random.seed(0)
+nruns = 10
+x = np.random.uniform(-5, np.log(pmax), nruns)
+y = np.random.uniform(-5, 5, nruns)
+p = np.exp(x)
+b = 2 * np.exp(2*x - y)
+
+npoints = np.unique(np.round(np.logspace(np.log10(2), 3, 50)).astype(int))
+data = np.empty((len(npoints), len(p)))
+
+for i, n in enumerate(ProgressBar(npoints)):
+    integrator = LogRadialIntegrator(r1, r2, k, pmax, n)
+    data[i] = [integrator(*_) for _ in zip(p, b)]
+
+relerr = np.abs(np.exp(data[:-1] - data[-1]) - 1)
+npoints = npoints[:-1]
+ax.plot(npoints, relerr, '-k')
+
+ax.set_xscale('log')
+ax.set_yscale('log')
+ax.set_ylim(1e-10, 1)
+ax.set_xlabel('number of grid points')
+ax.set_ylabel('relative error')
+
+fig.subplots_adjust(left=0.175, bottom=0.15, right=0.95, top=0.95)
+
+fig.savefig('distance_interpolant_convergence.pdf')

### 2015-08-14 ms.tex
Increment DCC version
Update runtimes

@@
 
-With just one thread, the \ac{BAYESTAR} analysis takes 87\nobreakdashes--384~s, already orders of magnitude faster than the full parameter estimation.
-With 32 threads, \ac{BAYESTAR} takes just 4\nobreakdashes--15~s.
+With just one thread, the \ac{BAYESTAR} analysis takes 76\nobreakdashes--356~s, already orders of magnitude faster than the full parameter estimation.
+With 32 threads, \ac{BAYESTAR} takes just 4\nobreakdashes--13~s.
 In practice, \ac{BAYESTAR}'s data handling (reading the detectors' \acp{PSD}, communicating with the \ac{GW} candidate database, writing FITS files) take an additional 15~s or so, though this overhead could be reduced by parallelizing many of these steps.
@@
 \begin{acknowledgements}
-This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v2.
+This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v3.
 

### 2015-08-14 ms.tex
In-preparation reference for Sathya and Chu

@@
 
-Although the response time of \ac{BAYESTAR} has been driven by the anticipated timescales for kilonova and afterglow emission, a recurring question is whether there is any detectable \ac{EM} signal in the seconds before, during, and after the merger itself.
-An obvious example would be prompt gamma\nobreakdashes-ray emission, though more exotic ...
-
-Since the \ac{GW} inspiral signal is in principal detectable for up to hundreds of seconds before merger, one could imagine positioning rapidly slewing instruments to search for prompt emission ...
-
 A more open question is how to account for spin precession.
@@
 Since spinning \ac{BNS} searches are still an active area of development, \ac{BAYESTAR}'s sky localization accuracy in this regime should be reexamined in the future.
+
+Although the response time of \ac{BAYESTAR} has been driven by the anticipated timescales for kilonova and afterglow emission, a recurring question is whether there is any detectable \ac{EM} signal in the seconds before, during, and after the merger itself.
+Since the \ac{GW} inspiral signal is in principal detectable for up to hundreds of seconds before merger, one could imagine positioning rapidly slewing instruments to search for any prompt emission.
+This concept was explored by the authors \citep{Cannon:2011vi}, and will be revisted in further detail in both Sathyaprakash et al.
+(in preparation) and Chu et al.
+(in preparation).
 

### 2015-08-14 ms.tex
Nonbreaking spaces

@@
 The signals were detected using a template bank that lacked spins entirely.
-Farr et al.
+Farr et~al.
 (in preparation) show that using nonspinning \ac{BNS} templates for parameter estimation has negligible impact on sky localization.
@@
 Since the \ac{GW} inspiral signal is in principal detectable for up to hundreds of seconds before merger, one could imagine positioning rapidly slewing instruments to search for any prompt emission.
-This concept was explored by the authors \citep{Cannon:2011vi}, and will be revisted in further detail in both Sathyaprakash et al.
-(in preparation) and Chu et al.
+This concept was explored by the authors \citep{Cannon:2011vi}, and will be revisted in further detail in both Sathyaprakash et~al.
+(in preparation) and Chu et~al.
 (in preparation).

### 2015-08-14 ms.tex
Add Holz citation

@@
 
+On the topic of very low\nobreakdashes-latency localization, we also recommend Chen~\&~Holz (in preparation), who propose a rapid localization scheme that is similar to ours, but even faster: their likelihood is strictly Gaussian, so one more marginalization integral (the integral over arrival time) can be performed analytically.
+
 \section{Conclusion}

### 2015-08-14 ms.tex
Define distance-independent amplitude

@@
 The distance integral is now performed differently from what we initially described in \cite{leo-singer-thesis,BerryLocalization}; the method described in the present work is about an order of magnitude faster.
-The innermost integral over distance $r$ may be written as
+We define $\rho_i = \omega_i / r$ in order to absorb all of the distance-independent terms in the amplitudes into $\omega_i$, and then define
+%
+\begin{align}
+    p^2 &= \frac{1}{2} \sum_i {\omega_i}^2 \label{eq:likelihood-p-factor} \\
+    b &= \left| \sum_i \hat\rho_i \omega_i e^{i \tilde{\gamma}_i}
+        a_i^*(\tilde{\tau}_i) \right|. \label{eq:likelihood-b-factor} \end{align}
+%
+The innermost integral over distance $r$ may then be written as
 %
@@
 \begin{align}
-    p^2 &= \frac{1}{2} r^2 \sum_i {\rho_i}^2 \label{eq:likelihood-p-factor} \\
-    b &= r \left| \sum_i \hat\rho_i \rho_i e^{i \tilde{\gamma}_i}
-        a_i^*(\tilde{\tau}_i) \right| \label{eq:likelihood-b-factor} \\
     r_0 &= 2 p^2 / b \\

### 2015-08-14 scripts/autocorr-likelihood.py
Save as png to fit in arXiv file size limits

@@ -181,4 +181,4 @@ for i, (filename, label, need_square) in enumerate(data):
     #     ax.yaxis.set_label_position('right')
     # else:
     #     plt.setp(ax.get_yticklabels(), visible=False)
-fig.savefig('autocorr-likelihood.pdf')
+fig.savefig('autocorr-likelihood.png')

### 2015-08-14 angle_integral.py
Remove unused script

@@ -1,30 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-from __future__ import division
-"""
-Create a figure that depicts the angular integration scheme.
-"""
-__author__ = "Leo Singer <leo.singer@ligo.org>"
-
-from matplotlib import pyplot as plt
-from mpl_toolkits.mplot3d import Axes3D
-import numpy as np
-import sys
-
-npoints = 15
-u = np.arange(- (npoints // 2), np.ceil(npoints / 2)) / (npoints // 2)
-phi = np.pi * u
-
-phi, u = np.meshgrid(phi, u)
-
-plt.figure(figsize=(5, 5))
-plt.scatter(phi, u)
-plt.ylim(-1.1, 1.1)
-plt.xlim(-1.1*np.pi, 1.1*np.pi)
-plt.xticks([-np.pi, 0, np.pi], [u'-π', '0', u'π'])
-plt.yticks([-1, 0, 1])
-
-if len(sys.argv) > 1:
-    plt.savefig(sys.argv[1])
-else:
-    plt.show()

### 2015-08-25 ms.tex
Add publish target
Missing word

@@
 \usepackage{mathrsfs}
-\usepackage{bib/aas_macros}
+\usepackage{aas_macros}
 \usepackage[colorlinks]{hyperref}
@@
 In the limit of large argument, $I_0(|x|) \sim \exp(|x|) / \sqrt{2 \pi |x|}$ \citep{Olver:2010:NHMF,NIST:DLMF}\footnote{\url{http://dlmf.nist.gov/10.40.E1}}.
-The scaled Bessel is useful for evaluation on a computer because it has a relatively small range $(0, 1]$ and varies slowly in proportion to $x^{1/2}$.
+The scaled Bessel function is useful for evaluation on a computer because it has a relatively small range $(0, 1]$ and varies slowly in proportion to $x^{1/2}$.
 

### 2015-09-23 ms.tex
Adjustments for post-O1 submission

@@
 
-The \acl{LIGO} (\acsu{LIGO}; \cite{AdvancedLIGOHarry,AdvancedLIGO}) and Virgo \cite{AdvancedVirgo} will soon begin taking data \cite{LIGOObservingScenarios} in their `Advanced' configurations, ultimately increasing their reach in volume within the local Universe by three orders of magnitude as compared to their initial configurations through 2010.
-They will form the first parts of a sensitive global \ac{GW} detector network, to be later augmented by the Japanese KAGRA facility \cite{LCGT,KAGRAInterferometerDesign} and \ac{LIGO}\nobreakdashes--India \cite{LIGOIndia}.
+The \acl{LIGO} (\acsu{LIGO}; \cite{AdvancedLIGOHarry,AdvancedLIGO}) and Virgo \cite{AdvancedVirgo} have just begun taking data \cite{LIGOObservingScenarios} in their `Advanced' configurations.
+They will ultimately increase their reach in volume within the local Universe by three orders of magnitude as compared to their initial configurations through 2010.
+They form the first parts of a sensitive global \ac{GW} detector network, to be later augmented by the Japanese KAGRA facility \cite{LCGT,KAGRAInterferometerDesign} and \ac{LIGO}\nobreakdashes--India \cite{LIGOIndia}.
 
@@
 It is unique in that it bridges the detection and parameter estimation of \ac{GW} signals, two tasks that have until now involved very different numerical methods and time scales.
-Beginning with the first Advanced \acs{LIGO} observing run, \ac{BAYESTAR} will provide localizations within minutes of the detection of any \ac{BNS} merger candidate, playing a key role in enabling rapid follow-up observations.
+Beginning with the first Advanced \acs{LIGO} observing run, \ac{BAYESTAR} is providing localizations within minutes of the detection of any \ac{BNS} merger candidate, playing a key role in enabling rapid follow-up observations.
 
@@
 Using the full \ac{SNR} time series in place of the autocorrelation function seems like a promising avenue; implementing this requires some infrastructure changes to both the matched filter pipeline and \ac{BAYESTAR}.
-Along these lines, we refer the reader to \cite{PankowRapidPE} for a similar, non-\ac{MCMC} approach to rapid exploration of the full parameter space.
-
-A more open question is how to account for spin precession.
+Along these lines, we also refer the reader to \cite{PankowRapidPE} for a similar, non-\ac{MCMC} approach to rapid exploration of the full parameter space.
+
+A more open\nobreakdashes-ended question is how to account for spin precession.
 The simulations in \cite{FirstTwoYears} and in this paper featured extremely modest spins of $\chi \leq 0.05$, consistent with the fastest known pulsars in binaries \citep{2003Natur.426..531B,DetectingBNSSystemsWithSpin}.
@@
 
-On the topic of very low\nobreakdashes-latency localization, we also recommend Chen~\&~Holz (in preparation), who propose a rapid localization scheme that is similar to ours, but even faster: their likelihood is strictly Gaussian, so one more marginalization integral (the integral over arrival time) can be performed analytically.
+On the topic of very low\nobreakdashes-latency localization, we also recommend Chen~\&~Holz (in preparation), who propose a rapid localization scheme that is similar to ours, but even faster because it makes some additional compromises: their likelihood is strictly Gaussian, so one more marginalization integral (the integral over arrival time) can be performed analytically.
 

### 2015-09-27 ms.tex
Add PACS codes
Less specificity in abstract

@@
-\documentclass[amsmath,amssymb,aps,prx,reprint,nopreprintnumbers,nofootinbib]{revtex4-1}
+\documentclass[amsmath,amssymb,aps,prx,reprint,nopreprintnumbers,nofootinbib,showpacs]{revtex4-1}
 \usepackage[nolist]{acronym}
@@
 \affiliation{\acs{LIGO} Laboratory, California Institute of Technology, Pasadena, CA 91125, USA}
+\pacs{04.80.Nn, 04.30.Tv, 02.50.Tt}
 
@@
 These sources are also predicted to power a broad array of electromagnetic transients.
-Because the X\nobreakdashes-ray and optical signatures can be faint and fade rapidly, observing them hinges on rapidly inferring the sky location from the gravitational wave observations.
+Because the electromagnetic signatures can be faint and fade rapidly, observing them hinges on rapidly inferring the sky location from the gravitational wave observations.
 \ac{MCMC} methods for gravitational\nobreakdashes-wave parameter estimation can take hours or more.

### 2015-11-19 ms.tex
Remove Larry's Caltech e-mail address
Increment DCC number
Revert "Remove Larry's Caltech e-mail address"
This reverts commit c33eb20af5eeffb6ff33fac783c06f2d5c0570b6.
Remove some comments that have been addressed
Remove qualifications about phase convention
This is, in fact, the phase convention used by FINDCHIRP and LALSimulation.

@@
 Finally, in Section~\ref{sec:case-study}, we quantify the sky localization performance on a comprehensive set of simulated \ac{GW} events.
-
-% Establish and describe the first rapid and tractable solution of the GW position reconstruction problem.
-% Use and abuse of Fisher matrix: use it as a clue.
-% Generality: replace GW antenna pattern and GW polarization model with any other kind of receiver.
-% Demonstration of a practical Bayesian analysis that does not use an MCMC method.
 
@@
 Assuming a nonprecessing circular orbit, we can write the \ac{GW} signal received by any detector as a linear combination of two basis waveforms, $H_0$ and $H_{\pi/2}$ \citep{PhysRevD.83.084002}.
-$H_0$ and $H_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the sine and cosine functions, being nearly orthogonal and out of phase by ${\pi/2}$ at all frequencies.
-If $H_0$ and $H_{\pi/2}$ are Fourier transforms of real functions, then $H_0(\omega) = H_0^*(-\omega)$ and $H_{\pi/2}(\omega) = H_{\pi/2}^*(-\omega)$, and we can write (assuming an arbitrary phase convention)
+$H_0$ and $H_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the cosine and sine functions, being orthogonal and out of phase by ${\pi/2}$ at all frequencies.
+If $H_0$ and $H_{\pi/2}$ are Fourier transforms of real functions, then $H_0(\omega) = H_0^*(-\omega)$ and $H_{\pi/2}(\omega) = H_{\pi/2}^*(-\omega)$, and we can write
 %
@@
 \begin{acknowledgements}
-This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v3.
+This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v4.
 

### 2015-11-19 ms.tex
Fix spelling error

@@
 %
-(We have ommitted the term $\int |Y_i(\omega)|^2/S(\omega)d\omega$, which takes the place of the earlyier ${\hat{\rho}_i}^2$ term and is only important for normalization.) The numerical scheme that we will develop is thus equally applicable for rapid, coincidence-based localization, or as a fast extrinsic marginalization step for the full parameter estimation.
+(We have ommitted the term $\int |Y_i(\omega)|^2/S(\omega)d\omega$, which takes the place of the earlier ${\hat{\rho}_i}^2$ term and is only important for normalization.) The numerical scheme that we will develop is thus equally applicable for rapid, coincidence-based localization, or as a fast extrinsic marginalization step for the full parameter estimation.
 

### 2015-11-19 ms.tex
Adjust positioning of transpose symbol
Define alpha

@@
             \mathcal{I}_{\bm\theta_i,\bm\theta_\mathrm{in}} \\
-            {\mathcal{I}_{\bm\theta_i,\bm\theta_\mathrm{in}}}\transpose &
+            {\mathcal{I}_{\bm\theta_i,\bm\theta_\mathrm{in}}^{^\mathsf{T}}} &
             {\rho_i}^2 \mathcal{I}_{\bm\theta_\mathrm{in},\bm\theta_\mathrm{in}}
@@
     y_\mathrm{min} &= 2 x_0 - \sqrt{2} \alpha - x_\mathrm{max} \\
-    y_\mathrm{max} &= x_0 + \alpha.
+    y_\mathrm{max} &= x_0 + \alpha
 \end{align}
+%
+where $\alpha = 4$ is a constant parameter that determines the extent of the grid.
 

### 2015-11-19 ms.tex
Fix typo

@@
 
-The interpolant is evaluated slightly differently depending on which of the three regions marked I, II, and II in Figure~\ref{fig:radial-integral-interpolant} contains the point of interest.
+The interpolant is evaluated slightly differently depending on which of the three regions marked I, II, and III in Figure~\ref{fig:radial-integral-interpolant} contains the point of interest.
 In region I, we use bicubic interpolation of $\log \mathcal{G}$ in $x$ and $y$.

### 2015-11-19 ms.tex
Cite HEALPix again

@@
 
-Our sampling of the sky relies completely on the \acl{HEALPix} (\acsu{HEALPix}), a special data structure designed for all\nobreakdashes-sky maps.
+Our sampling of the sky relies completely on the \acl{HEALPix} (\acsu{HEALPix}, \cite{healpix}), a special data structure designed for all\nobreakdashes-sky maps.
 \ac{HEALPix} divides the sky into equal\nobreakdashes-area pixels.

### 2015-11-19 ms.tex
Fix incorrect figure reference
Switch to absolute magnitude

@@
 They will ultimately increase their reach in volume within the local Universe by three orders of magnitude as compared to their initial configurations through 2010.
-They form the first parts of a sensitive global \ac{GW} detector network, to be later augmented by the Japanese KAGRA facility \cite{LCGT,KAGRAInterferometerDesign} and \ac{LIGO}\nobreakdashes--India \cite{LIGOIndia}.
+They form the first parts of a sensitive global \ac{GW} detector network, to be later augmented by the Japanese KAGRA facility \cite{LCGT,KAGRAInterferometerDesign} and \ac{LIGO}\nobreakdashes--India~\cite{LIGOIndia}.
 
@@
 \citet{MostPromisingEMCounterpart} argue that the most promising \ac{EM} counterparts are the hypothesized optical/infrared ``kilonovae'' powered by the radioactivate decay of $r$\nobreakdashes-process elements synthesized within the neutron\nobreakdashes-rich ejecta.
-These are expected to be faint, red, and peak rapidly, reaching only $R \sim 24$~mag within a week, though rising several magnitudes brighter in the infrared \citep{KilonovaHighOpacities}.
+These are expected to be faint, red, and peak rapidly, reaching an absolute magnitude of only $M_R \sim -13$ within a week, though rising several magnitudes brighter in the infrared~\citep{KilonovaHighOpacities}.
 
@@
 This involved several advances in the \ac{GW} data analysis \citep{CBCLowLatency}, including the first real\nobreakdashes-time matched\nobreakdashes-filter detection pipeline (\acs{MBTA}, \aclu{MBTA}; \cite{LIGOVirgoInspiralPipelineComparison}), a semi\nobreakdashes-coherent, ad hoc rapid triangulation code (Timing++), and the first version of a rigorous Bayesian \ac{MCMC} parameter estimation code (LALINFERENCE; \citealt{S6PE})---all in service of the first search for X\nobreakdashes-ray \citep{SwiftFollowup} and optical \citep{OpticalImageAnalysis} counterparts of \ac{GW} triggers, by a consortium of facilities.
-Despite the technical achievements in the \ac{GW} data analysis, there was an undesirable tradeoff between the speed as well as accuracy of the rapid localization and the full parameter estimation: the former could analyze a detection candidate in minutes, whereas the latter took half a day; the latter decreased the area on the sky by a factor of 1/20 over the former but took 1000 times as long to run \cite{SiderySkyLocalizationComparison}.
+Despite the technical achievements in the \ac{GW} data analysis, there was an undesirable tradeoff between the speed as well as accuracy of the rapid localization and the full parameter estimation: the former could analyze a detection candidate in minutes, whereas the latter took half a day; the latter decreased the area on the sky by a factor of 1/20 over the former but took 1000 times as long to run~\cite{SiderySkyLocalizationComparison}.
 
@@
 The resulting map is a tree structure that describes a mesh of pixels with different resolutions.
-An example \ac{BAYESTAR} subdivision is shown in Figure~\ref{fig:adaptive-illustration}.
-To convert this mesh into a FITS image,
-we traverse the tree and flatten it into the highest resolution represented.
+An example \ac{BAYESTAR} subdivision is shown in Figure~\ref{fig:adaptive-mesh}.
+To convert this mesh into a FITS image, we traverse the tree and flatten it into the highest resolution represented.
 The highest possible resolution is $N_\mathrm{side}=2^{11}$, with an area of $\approx 10^{-3}$~deg$^2$ per pixel.\footnote{Although the resulting sky map contains $N_\mathrm{pix} \approx 5\times10^6$ pixels, at most $\approx 2\times10^4$ pixels have distinct values.

### 2015-11-19 ms.tex
Prior on psi matches between equation and text
Advanced Virgo has not yet started

@@
 
-The \acl{LIGO} (\acsu{LIGO}; \cite{AdvancedLIGOHarry,AdvancedLIGO}) and Virgo \cite{AdvancedVirgo} have just begun taking data \cite{LIGOObservingScenarios} in their `Advanced' configurations.
-They will ultimately increase their reach in volume within the local Universe by three orders of magnitude as compared to their initial configurations through 2010.
-They form the first parts of a sensitive global \ac{GW} detector network, to be later augmented by the Japanese KAGRA facility \cite{LCGT,KAGRAInterferometerDesign} and \ac{LIGO}\nobreakdashes--India~\cite{LIGOIndia}.
+The \acl{LIGO} (\acsu{LIGO}; \cite{AdvancedLIGOHarry,AdvancedLIGO}) has just begun taking data \cite{LIGOObservingScenarios} in its `Advanced' configuration.
+The two \ac{LIGO} detectors will ultimately increase their reach in volume within the local Universe by three orders of magnitude as compared to their initial configurations through 2010.
+They form the first parts of a sensitive global \ac{GW} detector network, soon to be augmented by Advanced Virgo \cite{AdvancedVirgo} and later by the Japanese KAGRA facility \cite{LCGT,KAGRAInterferometerDesign} and \ac{LIGO}\nobreakdashes--India~\cite{LIGOIndia}.
 
@@
 
-Finally, the prior is uniform in $-1 \leq \cos\iota \leq 1$ and $0 \leq \psi < 2\pi$.
+Finally, the prior is uniform in $-1 \leq \cos\iota \leq 1$ and $0 \leq \psi < \pi$.
 

### 2015-12-07 ms.tex
Increment DCC version
Corrected proofs

@@
 
-\title[Rapid-response Bayesian sky localization]{WHOOMP! (There It Is) \\
-Rapid Bayesian position reconstruction for gravitational\nobreakdashes-wave transients}
+\title[Rapid-response Bayesian sky localization]{WHOOMP! (There it is):\\Rapid Bayesian position reconstruction for gravitational\nobreakdashes-wave transients}
 \author{Leo P. Singer}
@@
 \email{larryp@caltech.edu}
-\affiliation{\acs{LIGO} Laboratory, California Institute of Technology, Pasadena, CA 91125, USA}
+\affiliation{\acs{LIGO} Laboratory, California Institute of Technology, Pasadena, California 91125, USA}
+\received{28 September 2015}
 \pacs{04.80.Nn, 04.30.Tv, 02.50.Tt}
@@
 \begin{abstract}
-Within the next few years, Advanced LIGO and Virgo should detect \acp{GW} from binary neutron star and neutron star\nobreakdashes--black hole mergers.
-These sources are also predicted to power a broad array of electromagnetic transients.
-Because the electromagnetic signatures can be faint and fade rapidly, observing them hinges on rapidly inferring the sky location from the gravitational wave observations.
-\ac{MCMC} methods for gravitational\nobreakdashes-wave parameter estimation can take hours or more.
-We introduce BAYESTAR, a rapid, Bayesian, non-\ac{MCMC} sky localization algorithm that takes just seconds to produce probability sky maps that are comparable in accuracy to the full analysis.
-Prompt localizations from BAYESTAR will make it possible to search electromagnetic counterparts of compact binary mergers.
+Within the next few years, Advanced LIGO and Virgo should detect \aclp{GW} from \acl{BNS} and \acl{NSBH} mergers.
+These sources are also predicted to power a broad array of \acl{EM} transients.
+Because the \acl{EM} signatures can be faint and fade rapidly, observing them hinges on rapidly inferring the sky location from the \acl{GW} observations.
+\acl{MCMC} methods for gravitational\nobreakdashes-wave parameter estimation can take hours or more.
+We introduce BAYESTAR, a rapid, Bayesian, non-\acl{MCMC} sky localization algorithm that takes just seconds to produce probability sky maps that are comparable in accuracy to the full analysis.
+Prompt localizations from BAYESTAR will make it possible to search \acl{EM} counterparts of compact binary mergers.
 \end{abstract}
@@
 
-The \acl{LIGO} (\acsu{LIGO}; \cite{AdvancedLIGOHarry,AdvancedLIGO}) has just begun taking data \cite{LIGOObservingScenarios} in its `Advanced' configuration.
-The two \ac{LIGO} detectors will ultimately increase their reach in volume within the local Universe by three orders of magnitude as compared to their initial configurations through 2010.
-They form the first parts of a sensitive global \ac{GW} detector network, soon to be augmented by Advanced Virgo \cite{AdvancedVirgo} and later by the Japanese KAGRA facility \cite{LCGT,KAGRAInterferometerDesign} and \ac{LIGO}\nobreakdashes--India~\cite{LIGOIndia}.
-
-The most readily detectable sources of \acp{GW} include binary neutron star mergers, with 0.4\nobreakdashes--400 events per year within the reach of Advanced LIGO at final design sensitivity~\cite{rates}.
-These binary systems are not only efficient sources of \acp{GW}, but also potential sources of detectable \ac{EM} transients from the aftermath of the tidal disruption of the \acp{NS}.
-\citet{MostPromisingEMCounterpart} argue that the most promising \ac{EM} counterparts are the hypothesized optical/infrared ``kilonovae'' powered by the radioactivate decay of $r$\nobreakdashes-process elements synthesized within the neutron\nobreakdashes-rich ejecta.
-These are expected to be faint, red, and peak rapidly, reaching an absolute magnitude of only $M_R \sim -13$ within a week, though rising several magnitudes brighter in the infrared~\citep{KilonovaHighOpacities}.
+The Laser Interferometer Gravitational Wave Observatory (\acsu{LIGO})~\cite{AdvancedLIGOHarry,AdvancedLIGO}) has just begun taking data \cite{LIGOObservingScenarios} in its ``Advanced'' configuration.
+The two \ac{LIGO} detectors will ultimately increase their reach in volume within the local Universe by 3 orders of magnitude as compared to their initial configurations through 2010.
+They form the first parts of a sensitive global gravitational-wave (\acsu{GW}) detector network, soon to be augmented by Advanced Virgo~\cite{AdvancedVirgo} and later by the Japanese KAGRA facility~\cite{LCGT,KAGRAInterferometerDesign} and \ac{LIGO}\nobreakdashes--India~\cite{LIGOIndia}.
+
+The most readily detectable sources of \acp{GW} include binary neutron star mergers, with 0.4\nobreakdashes--400 events per year within the reach of Advanced LIGO at its final design sensitivity~\cite{rates}.
+These binary systems are not only efficient sources of \acp{GW} but also potential sources of detectable \ac{EM} transients from the aftermath of the tidal disruption of the \acp{NS}.
+\citet{MostPromisingEMCounterpart} argue that the most promising \ac{EM} counterparts are the hypothesized optical/infrared ``kilonovae'' powered by the radioactive decay of $r$\nobreakdashes-process elements synthesized within the neutron\nobreakdashes-rich ejecta.
+These are expected to be faint and red and to peak rapidly, reaching an absolute magnitude of only $M_R \sim -13$ within a week, though rising several magnitudes brighter in the infrared~\citep{KilonovaHighOpacities}.
 
@@
 Adding to the challenge of detecting a faint, short-lived optical transient, there is an extreme mismatch between the sky localization accuracy of \ac{GW} detector networks, $\sim$10\nobreakdashes--500\,deg$^2$ \citep{FairhurstTriangulation,WenLocalizationAdvancedLIGO,FairhurstLocalizationAdvancedLIGO,2011PhRvD..84j4020V,RodriguezBasicParameterEstimation,NissankeLocalization,NissankeKasliwalEMCounterparts,KasliwalTwoDetectors,Grover:2013,SiderySkyLocalizationComparison,FirstTwoYears}, and the \acp{FOV} of 1\nobreakdashes--8\,m\nobreakdashes-class optical telescopes.
-Wide\nobreakdashes-field optical transient facilites such as BlackGEM~(0.6\,m/2.7\,deg$^2$), the \acl{ZTF}~(\acsu{ZTF}; 1.2\,m/47\,deg$^2$) \cite{ZTF}, the \acl{DECAM}~(\acsu{DECAM}; 4\,m/3\,deg$^2$), or the \acl{LSST}~(\acsu{LSST}; 8.4\,m/9.6\,deg$^2$), operated in ``target of opportunity'' mode, may be able to tile these large areas rapidly enough to find the one needle in the haystack that is connected to the \ac{GW} event.
+Wide\nobreakdashes-field optical transient facilities such as BlackGEM~(0.6\,m/2.7\,deg$^2$), the \acl{ZTF}~(1.2\,m/47\,deg$^2$)~\cite{ZTF}, the \acl{DECAM}~(4\,m/3\,deg$^2$), or the \acl{LSST}~(8.4\,m/9.6\,deg$^2$), operated in ``target of opportunity'' mode, may be able to tile these large areas rapidly enough to find the one needle in the haystack that is connected to the \ac{GW} event.
 However, prompt and accurate \ac{GW} position reconstructions will be of the utmost importance for guiding the selection of fields to observe.
 
-The final science run of the initial LIGO and Virgo instruments saw the first joint search for \ac{GW} and electromagnetic emission from compact binaries.
-This involved several advances in the \ac{GW} data analysis \citep{CBCLowLatency}, including the first real\nobreakdashes-time matched\nobreakdashes-filter detection pipeline (\acs{MBTA}, \aclu{MBTA}; \cite{LIGOVirgoInspiralPipelineComparison}), a semi\nobreakdashes-coherent, ad hoc rapid triangulation code (Timing++), and the first version of a rigorous Bayesian \ac{MCMC} parameter estimation code (LALINFERENCE; \citealt{S6PE})---all in service of the first search for X\nobreakdashes-ray \citep{SwiftFollowup} and optical \citep{OpticalImageAnalysis} counterparts of \ac{GW} triggers, by a consortium of facilities.
+The final science run of the initial LIGO and Virgo instruments saw the first joint search for \ac{GW} and \ac{EM} emission from compact binaries.
+This involved several advances in the \ac{GW} data analysis \citep{CBCLowLatency}, including the first real\nobreakdashes-time matched\nobreakdashes-filter detection pipeline, the \acl{MBTA}~\cite{LIGOVirgoInspiralPipelineComparison}; a semi\nobreakdashes-coherent, \textit{ad hoc} rapid triangulation code, Timing++; and the first version of a rigorous Bayesian \ac{MCMC} parameter estimation code, LALINFERENCE~\cite{S6PE}---all in service of the first search for X\nobreakdashes-ray \citep{SwiftFollowup} and optical \citep{OpticalImageAnalysis} counterparts of \ac{GW} triggers, by a consortium of facilities.
 Despite the technical achievements in the \ac{GW} data analysis, there was an undesirable tradeoff between the speed as well as accuracy of the rapid localization and the full parameter estimation: the former could analyze a detection candidate in minutes, whereas the latter took half a day; the latter decreased the area on the sky by a factor of 1/20 over the former but took 1000 times as long to run~\cite{SiderySkyLocalizationComparison}.
 
-The success of \ac{EM} follow\nobreakdashes-up of \ac{LIGO} events will depend critically on disseminating high quality sky localizations within a timescale of minutes to hours.
+The success of \ac{EM} follow\nobreakdashes-up of \ac{LIGO} events will depend critically on disseminating high quality sky localizations within a time scale of minutes to hours.
 To that end, we have devised a rapid and accurate Bayesian sky localization method that takes mere seconds to achieve approximately the same accuracy as the full \ac{MCMC} analysis.
@@
     \item The Fisher information matrix will provide clues as to suitable forms of this likelihood function.
-Recent \ac{GW} parameter estimation literature has largely rejected the Fisher matrix,\footnote{Though not entirely; see \cite{EffectiveFisherMatrix}.} but this is mostly on the grounds of the abuse of the related \ac{CRLB} outside its realm of validity (i.e., at low to moderate \ac{SNR}) \cite{BayesianBounds,UseAbuseFisherMatrix,FisherMatrixAsymptoticExpansions,InadequaciesFisherMatrix}.
-However, we recognize that the block structure of the Fisher matrix provides important insights, and is a useful quantity for checking the validity of the aforementioned likelihood function, quite independent of the \ac{CRLB}.
+Recent \ac{GW} parameter estimation literature has largely rejected the Fisher matrix,\footnote{Though not entirely; see \cite{EffectiveFisherMatrix}.} but this is mostly on the grounds of the abuse of the related \ac{CRLB} outside its realm of validity of moderate to high \ac{SNR}~\cite{BayesianBounds,UseAbuseFisherMatrix,FisherMatrixAsymptoticExpansions,InadequaciesFisherMatrix}.
+However, we recognize that the block structure of the Fisher matrix provides important insights and is a useful quantity for checking the validity of the aforementioned likelihood function, quite independent of the \ac{CRLB}.
     \item The Fisher matrix teaches us that errors in sky localization are semi\nobreakdashes-independent from errors in masses.
@@
 Moreover, this frees us of the need to directly compute the expensive post-Newtonian model waveforms, making the likelihood itself much faster to evaluate.
-    \item Thanks to a simple likelihood function and a well\nobreakdashes-characterized parameter space, we may dispense with costly and parallelization\nobreakdashes-resistant \ac{MCMC} integration, and instead perform the Bayesian marginalization with classic, deterministic, very low order Gaussian quadrature.
+    \item Thanks to a simple likelihood function and a well\nobreakdashes-characterized parameter space, we may dispense with costly and parallelization\nobreakdashes-resistant \ac{MCMC} integration and instead perform the Bayesian marginalization with classic, deterministic, very low order Gaussian quadrature.
     \item The Bayesian inference scheme thus designed to operate on the matched\nobreakdashes-filter detection pipeline outputs could be trivially generalized to operate on the full \ac{GW} time series within the same computational constraints.
@@
 %
-We call this algorithm \ac{BAYESTAR}\footnote{A pun on the Cylon battleships in the American television series Battlestar Galactica.
-The defining characteristic of the Cylons is that they repeatedly defeat humanity by using their superhuman information\nobreakdashes-gathering ability to coordinate overwhelming forces.}\footnote{We do not like to mention the final `L' in the acronym, because then it would be pronounced BAYESTARL, which sounds stupid.}.
-It is as fast as Timing++, but nearly as accurate as the rigorous full parameter estimation.
+We call this algorithm \ac{BAYESTAR}.\footnote{A pun on the Cylon battleships in the American television series Battlestar Galactica.
+The defining characteristic of the Cylons is that they repeatedly defeat humanity by using their superhuman information\nobreakdashes-gathering ability to coordinate overwhelming forces.}\footnote{We do not like to mention the final `L' in the acronym, because then it would be pronounced BAYESTARL, which sounds stupid.} It is as fast as Timing++ but nearly as accurate as the rigorous full parameter estimation.
 It is unique in that it bridges the detection and parameter estimation of \ac{GW} signals, two tasks that have until now involved very different numerical methods and time scales.
@@
 This paper is organized as follows.
-In Section~\ref{sec:detection}, we describe the \ac{GW} signal model and sketch the standard detection algorithm, the matched filter bank.
-In Section~\ref{sec:parameter-estimation}, we describe Bayesian inference formalism and the prevailing method for inferring the parameters of detected candidates, \ac{MCMC} sampling.
-In Section~\ref{sec:bayestar-likelihood}, we propose the \ac{BAYESTAR} likelihood as a model for the uncertainty in the matched\nobreakdashes-filter parameter estimates, and discuss its relationship to and consistency with the likelihood for the full \ac{GW} data.
-In Section~\ref{sec:prior}, we describe the input to \ac{BAYESTAR} supplied by the detection pipeline, and the prior distribution on parameters.
-In Section~\ref{sec:marginal-posterior}, we explain the integration scheme by which the posterior probability distribution is calculated for a given sky location.
-In Section~\ref{sec:adaptive-sampling}, we show a scheme whereby the sky posterior is sampled on an adaptive \acs{HEALPix} grid.
-In Section~\ref{sec:run-time}, we report the running time of the algorithm on the hardware available on the \ac{LIGO} Data Grid.
-Finally, in Section~\ref{sec:case-study}, we quantify the sky localization performance on a comprehensive set of simulated \ac{GW} events.
+In Sec.~\ref{sec:detection}, we describe the \ac{GW} signal model and sketch the standard detection algorithm, the matched\nobreakdashes-filter bank.
+In Sec.~\ref{sec:parameter-estimation}, we describe Bayesian inference formalism and the prevailing method for inferring the parameters of detected candidates, \ac{MCMC} sampling.
+In Sec.~\ref{sec:bayestar-likelihood}, we propose the \ac{BAYESTAR} likelihood as a model for the uncertainty in the matched\nobreakdashes-filter parameter estimates, and discuss its relationship to and consistency with the likelihood for the full \ac{GW} data.
+In Sec.~\ref{sec:prior}, we describe the input to \ac{BAYESTAR} supplied by the detection pipeline, and the prior distribution on parameters.
+In Sec.~\ref{sec:marginal-posterior}, we explain the integration scheme by which the posterior probability distribution is calculated for a given sky location.
+In Sec.~\ref{sec:adaptive-sampling}, we show a scheme whereby the sky posterior is sampled on an adaptive \ac{HEALPix} grid.
+In Sec.~\ref{sec:run-time}, we report the running time of the algorithm on the hardware available on the \ac{LIGO} Data Grid.
+Finally, in Sec.~\ref{sec:case-study}, we quantify the sky localization performance on a comprehensive set of simulated \ac{GW} events.
 
@@
 
-In the \ac{TD}, the strain observed by a single \ac{GW} interferometer is
+In the \acl{TD}, the strain observed by a single \ac{GW} interferometer is
 %
@@
 %
-In the \ac{FD},
+In the \acl{FD},
 %
@@
 %
-where $X_i (\omega; \bm\theta)$ is the \ac{GW} signal given a parameter vector $\bm\theta$ that describes the \ac{GW} source, and $N_i (\omega)$ is that detector's Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[\left|N_i(\omega)\right|^2\right] + E\left[\left|N_i(-\omega)\right|^2\right] = 2 E\left[\left|N_i(\omega)\right|^2\right]$.
+where $X_i (\omega; \bm\theta)$ is the \ac{GW} signal given a parameter vector $\bm\theta$ that describes the \ac{GW} source and $N_i (\omega)$ is that detector's Gaussian noise with one\nobreakdashes-sided \ac{PSD} $S_i(\omega) = E\left[\left|N_i(\omega)\right|^2\right] + E\left[\left|N_i(-\omega)\right|^2\right] = 2 E\left[\left|N_i(\omega)\right|^2\right]$.
 We shall denote the combined observation from a network of detectors as $\mathbf Y (\omega) \equiv \{Y_i (\omega)\}_i$.
 
-Under the assumptions that the detector noise is Gaussian and that the noise from different detectors are uncorrelated, the likelihood of the observation, $\mathbf y$, conditioned upon the parameters $\bm\theta$, is a product of Gaussian distributions:
+Under the assumptions that the detector noise is Gaussian and that the noise from different detectors is uncorrelated, the likelihood of the observation, $\mathbf y$, conditioned upon the parameters $\bm\theta$, is a product of Gaussian distributions:
 %
@@
 
-A \ac{CBC} source is specified by a vector of extrinsic parameters describing its position and orientation, and intrinsic parameters describing the physical properties of the binary components\footnote{This list of parameters involves some simplifying assumptions.
-Eccentricity is omitted: although it may play a major role in the evolution and waveforms of rare close binaries formed by dynamical capture~\citep{PhysRevD.87.043004,2013PhRvD..87l7501H,2014PhRvD..90h4016H}, \ac{BNS} systems formed by binary stellar evolution should almost always circularize due to tidal interaction~\citep{0004-637X-572-1-407} and later \ac{GW} emission~\citep{PhysRev.136.B1224} long before the inspiral enters LIGO's frequency range of $\sim$10\nobreakdashes--1000~kHz.
+A \ac{CBC} source is specified by a vector of extrinsic parameters describing its position and orientation and intrinsic parameters describing the physical properties of the binary components\footnote{This list of parameters involves some simplifying assumptions.
+Eccentricity is omitted; although it may play a major role in the evolution and waveforms of rare close binaries formed by dynamical capture~\citep{PhysRevD.87.043004,2013PhRvD..87l7501H,2014PhRvD..90h4016H}, \ac{BNS} systems formed by binary stellar evolution should almost always circularize due to tidal interaction~\citep{0004-637X-572-1-407} and later \ac{GW} emission~\citep{PhysRev.136.B1224} long before the inspiral enters LIGO's frequency range of $\sim$10\nobreakdashes--1000~kHz.
 Tidal deformabilities of the \acp{NS} are omitted because the signal imprinted by the companions' material properties is so small that it will only be detectable by an Einstein Telescope\nobreakdashes-class \ac{GW} observatory~\citep{PhysRevD.81.123016}.
@@
 
-Assuming a nonprecessing circular orbit, we can write the \ac{GW} signal received by any detector as a linear combination of two basis waveforms, $H_0$ and $H_{\pi/2}$ \citep{PhysRevD.83.084002}.
+Assuming a non\nobreakdashes-precessing circular orbit, we can write the \ac{GW} signal received by any detector as a linear combination of two basis waveforms, $H_0$ and $H_{\pi/2}$ \citep{PhysRevD.83.084002}.
 $H_0$ and $H_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the cosine and sine functions, being orthogonal and out of phase by ${\pi/2}$ at all frequencies.
@@
 For brevity, we define $H \equiv H_{0}$ and write all subsequent equations in terms of the $H$ basis vector alone.
-Then, we can write the signal model in a way that isolates all dependence on the extrinsic parameters, $\bm\theta_\mathrm{ex}$, into a few coefficients, and all dependence on the intrinsic parameters, $\bm\theta_\mathrm{in}$, into the basis waveform, by taking the Fourier transform of Equation~(2.8) of \cite{PhysRevD.83.084002}:
+Then, we can write the signal model in a way that isolates all dependence on the extrinsic parameters, $\bm\theta_\mathrm{ex}$, into a few coefficients and all dependence on the intrinsic parameters, $\bm\theta_\mathrm{in}$, into the basis waveform, by taking the Fourier transform of Eq.~(2.8) of Ref.~\cite{PhysRevD.83.084002},
 %
@@
 The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, defined such that $0 \leq {F_{+,i}}^2 + {F_{\times,i}}^2 \leq 1$.
-They depend on the orientation of detector $i$ as well as the sky location and sidereal time of the event and are presented in~\cite{ExcessPower}.
+They depend on the orientation of detector $i$ as well as the sky location and sidereal time of the event and are presented in Ref.~\cite{ExcessPower}.
 In a coordinate system with the $x$ and $y$ axes aligned with the arms of a detector, the antenna pattern is given in spherical polar coordinates as
@@
 %
-The unit vector $\mathbf{d}_i$ represents the position of detector $i$ in units of light travel time.\footnote{When considering transient \ac{GW} sources such as those that we are concerned with in this thesis, the origin of the coordinate system is usually taken to be the geocenter.
+The unit vector $\mathbf{d}_i$ represents the position of detector $i$ in units of light travel time.\footnote{When considering transient \ac{GW} sources such as those that we are concerned with in this article, the origin of the coordinate system is usually taken to be the geocenter.
 For long\nobreakdashes-duration signals such as those from statically deformed neutron stars, the solar system barycenter is a more natural choice.} The vector $\mathbf{n}$ is the direction of the source.
 The negative sign in the dot product $-\mathbf{d}_i \cdot \mathbf{n}$ is present because the direction of travel of the \ac{GW} signal is opposite to that of its sky location.
-The quantity $r_{1,i}$ is a fiducial distance at which detector $i$ would register SNR=1 for an optimally oriented binary (face\nobreakdashes-on, and in a direction perpendicular to the interferometer's arms):
+The quantity $r_{1,i}$ is a fiducial distance at which detector $i$ would register \ac{SNR}=1 for an optimally oriented binary (face\nobreakdashes-on, and in a direction perpendicular to the interferometer's arms):
 %
@@
 
-The prevailing technique for detection of \acp{GW} from \acp{CBC} is to realize a \ac{MLE} from the likelihood in Equation~(\ref{eq:gaussian-likelihood}) and the signal model in Equation~(\ref{eq:signal-model}).
+The prevailing technique for detection of \acp{GW} from \acp{CBC} is to realize a \ac{ML} estimator (MLE) from the likelihood in Eq.~(\ref{eq:gaussian-likelihood}) and the signal model in Eq.~(\ref{eq:signal-model}).
 Concretely, this results in a bank of matched filters, or cross-correlations between the incoming data stream and a collection of template waveforms,
@@
 We can predict the uncertainty in the detection pipeline's \ac{ML} estimates using the \ac{CRLB}.
-The \ac{CRLB} has been widely applied in \ac{GW} data analysis to estimate parameter estimation uncertainty \cite{1996PhRvD..53.3033B,FairhurstTriangulation,2009PhRvD..79h4032A,WenLocalizationAdvancedLIGO,LIGOObservingScenarios,FairhurstLIGOIndia}\footnote{\label{footnote:template-banks}The Fisher matrix is also used in construction of \ac{CBC} matched filter banks.
+The \ac{CRLB} has been widely applied in \ac{GW} data analysis to estimate parameter estimation uncertainty \cite{1996PhRvD..53.3033B,FairhurstTriangulation,2009PhRvD..79h4032A,WenLocalizationAdvancedLIGO,LIGOObservingScenarios,FairhurstLIGOIndia}.\footnote{\label{footnote:template-banks}The Fisher matrix is also used in construction of \ac{CBC} matched-filter banks.
 The common procedure is to place templates uniformly according to the determinant of the signal space metric, which is the Fisher matrix.
 This is equivalent to uniformly sampling the Jeffreys prior.
-In practice, this is done either by constructing a hexagonal lattice \citep{PhysRevD.76.102004} or sampling stochastically \citep{2009PhRvD..80j4014H,2009PhRvD..80b4009V,SBank,2010PhRvD..81b4004M,2014PhRvD..89b4003P}.}.
-As we noted, there are significant caveats to the \ac{CRLB} at low or moderate \ac{SNR} \cite{BayesianBounds,UseAbuseFisherMatrix,FisherMatrixAsymptoticExpansions,InadequaciesFisherMatrix}.
+In practice, this is done either by constructing a hexagonal lattice~\citep{PhysRevD.76.102004} or sampling stochastically~\citep{2009PhRvD..80j4014H,2009PhRvD..80b4009V,SBank,2010PhRvD..81b4004M,2014PhRvD..89b4003P}.} As we noted, there are significant caveats to the \ac{CRLB} at low or moderate \ac{SNR} \cite{BayesianBounds,UseAbuseFisherMatrix,FisherMatrixAsymptoticExpansions,InadequaciesFisherMatrix}.
 However, here we will be concerned more with gaining intuition from the block structure of the Fisher matrix than its numerical value.
-Furthermore, the Fisher matrix in its own right---independent of its suitability to describe the parameter covariance---is a well\nobreakdashes-defined property of any likelihood function, and we will exploit it as such in Section~\ref{sec:bayestar-likelihood}.
+Furthermore, the Fisher matrix in its own right---independent of its suitability to describe the parameter covariance---is a well\nobreakdashes-defined property of any likelihood function, and we will exploit it as such in Sec.~\ref{sec:bayestar-likelihood}.
 
@@
 %
-with $X_i(\omega; \rho_i, \gamma_i, \tau_i, \bm\theta_\mathrm{in})$ given by Equation~(\ref{eq:signal-model}).
+with $X_i(\omega; \rho_i, \gamma_i, \tau_i, \bm\theta_\mathrm{in})$ given by Eq.~(\ref{eq:signal-model}).
 
@@
 Furthermore, it provides an estimate of the mean-square error in the parameters.
-If $\hat{\bm\theta}$ is an unbiased estimator of $\bm\theta$, $\tilde{\bm\theta} = \hat{\bm\theta} - \bm\theta$ is the measurement error, and $\Sigma = \mathrm{E} \, [\tilde{\bm\theta}\tilde{\bm\theta}\transpose]$ is the covariance of the measurement error, then the \ac{CRLB} says that $\Sigma \geq \mathcal{I}^{-1}$, in the sense that $\left(\Sigma - \mathcal{I}^{-1}\right)$ is positive semi\nobreakdashes-definite.
+If $\hat{\bm\theta}$ is an unbiased estimator of $\bm\theta$, $\tilde{\bm\theta} = \hat{\bm\theta} - \bm\theta$ is the measurement error, and $\Sigma = \mathrm{E} \, [\tilde{\bm\theta}\tilde{\bm\theta}\transpose]$ is the covariance of the measurement error, then the \ac{CRLB} says that $\Sigma \geq \mathcal{I}^{-1}$, in the sense that $\left(\Sigma - \mathcal{I}^{-1}\right)$ is positive semidefinite.
 
@@
 When (as in our assumptions) the likelihood is Gaussian,\footnote{This assumes that the merger occurs at a frequency outside the sensitive ``bucket'' of the detector's noise \ac{PSD}.
-There are additional terms if the \ac{GW} spectrum drops to zero within the sensitive bandwidth of the detector, as can be the case for \ac{NSBH} mergers; see \cite{2014CQGra..31o5005M}.} Equation~(\ref{eq:general-fisher-matrix}) simplifies to
+There are additional terms if the \ac{GW} spectrum drops to zero within the sensitive bandwidth of the detector, as can be the case for \acl{NSBH} mergers; see Ref.~\cite{2014CQGra..31o5005M}.} Eq.~(\ref{eq:general-fisher-matrix}) simplifies to
 %
@@
 %
-(This is equivalent to an expression given in \cite{Grover:2013}.) The bottom row and right column of Equation~(\ref{eq:single-detector-block-fisher-matrix}) describe the intrinsic parameters and how they are coupled to the extrinsic parameters.
+(This is equivalent to an expression given in Ref.~\cite{Grover:2013}.) The bottom row and right column of Eq.~(\ref{eq:single-detector-block-fisher-matrix}) describe the intrinsic parameters and how they are coupled to the extrinsic parameters.
 We show in Appendix~\ref{sec:independence-intrinsic-extrinsic} that we need not consider the intrinsic parameters at all if we are concerned only with sky localization.
@@
 where ${\omega_{\mathrm{rms},i}}^2 = {\overline{\omega^2}}_i - {{\overline{\omega}}_i}^2$.
-This structure implies that errors in \ac{SNR} are uncorrelated with errors in time and phase, and that there is a particular sum and difference of the times and phases that are measured independently (see Appendix~\ref{sec:interpretation-of-errors}).
-
-Reading off the $\tau \tau$ element of the covariance matrix reproduces the timing accuracy in Equation~(24) of \cite{FairhurstTriangulation},
+This structure implies that errors in the \ac{SNR} are uncorrelated with errors in time and phase and that there is a particular sum and difference of the times and phases that are measured independently (see Appendix~\ref{sec:interpretation-of-errors}).
+
+Reading off the $\tau \tau$ element of the covariance matrix reproduces the timing accuracy in Eq.~(24) of Ref.~\cite{FairhurstTriangulation},
 %
@@
 In the Bayesian framework, parameters are inferred from the data by forming the posterior distribution, $p(\bm\theta|\mathbf y)$, which describes the probability of the parameters given the observations.
-Bayes' rule relates the likelihood $p(\mathbf y|\bm\theta)$ to the posterior $p(\bm\theta|\mathbf y)$,
+Bayes's rule relates the likelihood $p(\mathbf y|\bm\theta)$ to the posterior $p(\bm\theta|\mathbf y)$,
 %
@@
 %
-introducing the prior distribution $p(\bm\theta)$ which encapsulates previous information about the paramters (for example, arising from earlier observations or from known physical bounds on the parameters), and the evidence $p(\mathbf y)$ which can be thought of as a normalization factor or as describing the parsimoniousness of the model.
+introducing the prior distribution $p(\bm\theta)$ which encapsulates previous information about the parameters (for example, arising from earlier observations or from known physical bounds on the parameters) and the evidence $p(\mathbf y)$ which can be thought of as a normalization factor or as describing the parsimoniousness of the model.
 
 The choice of prior is determined by one's astrophysical assumptions.
-During \ac{S6} when LIGO's Bayesian \ac{CBC} parameter estimation pipelines were first deployed, the prior was taken to be isotropic in sky location and binary orientation, and uniform in volume, arrival time, and the component masses~\cite{S6PE}.
+During \acl{S6} when LIGO's Bayesian \ac{CBC} parameter estimation pipelines were first deployed, the prior was taken to be isotropic in the sky location and binary orientation and uniform in volume, arrival time, and the component masses~\cite{S6PE}.
 
@@
 
-\section{The \acs{BAYESTAR} likelihood}
+\section{\acs{BAYESTAR} likelihood}
 \label{sec:bayestar-likelihood}
 
-For the purpose of rapid sky localization, we assume that we do not have access to the \ac{GW} data $\mathbf{Y}$ itself, and that our only contact with it is through the \ac{ML} parameter estimates $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$.
-Although this is a significant departure from conventional \ac{GW} parameter estimation techniques, we can still apply the full Bayesian machinery of Equation~(\ref{eq:marginal-posterior}) to compute a posterior distribution for the sky location.
+For the purpose of rapid sky localization, we assume that we do not have access to the \ac{GW} data $\mathbf{Y}$ itself and that our only contact with it is through the \ac{ML} parameter estimates $\{\left\{ \hat\rho_i, \hat\gamma_i, \hat\tau_i \right\}_i, \hat{\bm\theta}_\mathrm{in}\}$.
+Although this is a significant departure from conventional \ac{GW} parameter estimation techniques, we can still apply the full Bayesian machinery of Eq.~(\ref{eq:marginal-posterior}) to compute a posterior distribution for the sky location.
 
@@
 Although we may not be able to evaluate this equation directly, with some educated guesses we can create a likelihood that has many properties in common with it.
-Any valid approximate likelihood must have the same Fisher matrix as shown in Equation~(\ref{eq:single-detector-block-fisher-matrix}).
+Any valid approximate likelihood must have the same Fisher matrix as shown in Eq.~(\ref{eq:single-detector-block-fisher-matrix}).
 It must also have the same limiting behavior: it should be periodic in the phase error $\tilde{\gamma}_i$ and go to zero as $\tilde{\tau}_i \rightarrow \pm \infty$, $\hat{\rho}_i \rightarrow 0$, or $\hat{\rho}_i \rightarrow \infty$.
-Additionally, when $\tilde{\tau}_i = 0$, the distribution of ${\hat{\rho}_i}^2$ should reduce to a noncentral $\chi^2$ distribution with two degrees of freedom, centered about ${\rho_i}^2$, because the complex matched filter time series $z_i(t)$ is Gaussian (under the ideal assumption the \ac{GW} strain time series is Gaussian).
-
-These conditions could be satisfied by realizing a multivariate Gaussian distribution with covariance matrix $\Sigma = \mathcal{I}^\intercal$, and then replacing individual quadratic terms in the exponent of the form $-\tilde{\theta}^2/2$ with $\cos{\tilde{\theta}}$.
-
-A more natural way is to plug the signal model from Equation~(\ref{eq:signal-model}) \emph{evaluated at the \ac{ML} parameter estimates} into the single-detector likelihood in Equation~(\ref{eq:gaussian-likelihood-spa}):
+Additionally, when $\tilde{\tau}_i = 0$, the distribution of ${\hat{\rho}_i}^2$ should reduce to a noncentral $\chi^2$ distribution with 2 degrees of freedom, centered about ${\rho_i}^2$, because the complex matched-filter time series $z_i(t)$ is Gaussian (under the ideal assumption the \ac{GW} strain time series is Gaussian).
+
+These conditions could be satisfied by realizing a multivariate Gaussian distribution with covariance matrix $\Sigma = \mathcal{I}^\intercal$ and then replacing individual quadratic terms in the exponent of the form $-\tilde{\theta}^2/2$ with $\cos{\tilde{\theta}}$.
+
+A more natural way is to plug the signal model from Eq.~(\ref{eq:signal-model}) evaluated at the \ac{ML} parameter estimates into the single-detector likelihood in Eq.~(\ref{eq:gaussian-likelihood-spa}):
 %
@@
 %
-If we further assume that \emph{the intrinsic parameters are equal to their \ac{ML} estimates}, $\bm\theta_\mathrm{in} = \hat{\bm\theta}_\mathrm{in}$, then this reduces to what we call the \emph{autocorrelation likelihood},
+If we further assume that the intrinsic parameters are equal to their \ac{ML} estimates, $\bm\theta_\mathrm{in} = \hat{\bm\theta}_\mathrm{in}$, then this reduces to what we call the autocorrelation likelihood,
 %
@@
 %
-Some example autocorrelation functions and corresponding likelihoods are shown in Figure~\ref{fig:autocorr-likelihood}.
+Some example autocorrelation functions and corresponding likelihoods are shown in Fig.~\ref{fig:autocorr-likelihood}.
 To assemble the joint likelihood for the whole network, we form the product of the autocorrelation likelihoods from the individual detectors:
@@
 
-In the following section, we discuss some key properties of the autocorreltion likelihood.
+In the following section, we discuss some key properties of the autocorrelation likelihood.
 
@@
     \includegraphics[width=\textwidth]{autocorr-likelihood}
-    \caption{\label{fig:autocorr-likelihood}The autocorrelation likelihood for a $(1.4, 1.4)$\,$M_\odot$ binary as observed by four detector configurations: from top to bottom, the final sensitivity achieved by the \acs{LIGO} Hanford, \acs{LIGO} Livingston, and Virgo detectors in their ``initial'' configuration, and the final Advanced \acs{LIGO} design sensitivity.
+    \caption{\label{fig:autocorr-likelihood}The autocorrelation likelihood for a $(1.4, 1.4)$\,$M_\odot$ binary as observed by four detector configurations: from top to bottom, the final sensitivity achieved by the \acs{LIGO} Hanford, \acs{LIGO} Livingston, and Virgo detectors in their ``initial'' configuration and the final Advanced \acs{LIGO} design sensitivity.
 The left panels show the noise amplitude spectral densities.
@@
 The right panel shows the phase-marginalized autocorrelation likelihood for \ac{SNR}=1,~2,~4,~and~8.
-In the right panel, the time scale is normalized by \ac{SNR} so that one can see that as \ac{SNR} increases, a central parabola is approached (the logarithm of a Gaussian distribution with standard deviation given by the Fisher matrix).}
+In the right panel, the time scale is normalized by the \ac{SNR} so that one can see that as the \ac{SNR} increases, a central parabola is approached (the logarithm of a Gaussian distribution with standard deviation given by the Fisher matrix).}
 \end{figure*}
@@
 
-First, the autocorrelation likelihood has the elegant feature that if we were to replace the autocorrelation function with the \ac{SNR} time series for the best-matching template, $z(\tau; \hat{\mathbf{\theta}}_\mathrm{in})$, we would recover the likelihood for the full \ac{GW} time series, evaluated at the \ac{ML} estimate of the intrinsic parameters, \emph{viz.}:
+First, the autocorrelation likelihood has the elegant feature that if we were to replace the autocorrelation function with the \ac{SNR} time series for the best-matching template, $z(\tau; \hat{\mathbf{\theta}}_\mathrm{in})$, we would recover the likelihood for the full \ac{GW} time series, evaluated at the \ac{ML} estimate of the intrinsic parameters, viz.:
 %
@@
 %
-(We have ommitted the term $\int |Y_i(\omega)|^2/S(\omega)d\omega$, which takes the place of the earlier ${\hat{\rho}_i}^2$ term and is only important for normalization.) The numerical scheme that we will develop is thus equally applicable for rapid, coincidence-based localization, or as a fast extrinsic marginalization step for the full parameter estimation.
-
-Second, observe that at the true parameter values, $\hat{\bm\theta}_i = \bm\theta_i$, the logarithms of Equation~(\ref{eq:autocor-likelihood}) and Equation~(\ref{eq:gaussian-likelihood-spa}) have the same Jacobian.
+[We have omitted the term $\int |Y_i(\omega)|^2/S(\omega)d\omega$, which takes the place of the earlier ${\hat{\rho}_i}^2$ term and is only important for normalization.] The numerical scheme that we will develop is thus equally applicable for rapid, coincidence-based localization, or as a fast extrinsic marginalization step for the full parameter estimation.
+
+Second, observe that at the true parameter values, $\hat{\bm\theta}_i = \bm\theta_i$, the logarithms of Eqs.~(\ref{eq:autocor-likelihood}) and~(\ref{eq:gaussian-likelihood-spa}) have the same Jacobian.
 This is because the derivatives of the autocorrelation function are
@@
 %
-with $\overline{\omega^n}$ defined in Equation~(\ref{eq:angular-frequency-moments}). For example, the first few derivatives are
+with $\overline{\omega^n}$ defined in Eq.~(\ref{eq:angular-frequency-moments}). For example, the first few derivatives are
 %
@@
 
-Using Equation~(\ref{eq:general-fisher-matrix-second-derivatives}), we can compute the Fisher matrix elements for the autocorrelation likelihood given by Equation~(\ref{eq:autocor-likelihood}), with detector subscript suppressed:
+Using Eq.~(\ref{eq:general-fisher-matrix-second-derivatives}), we can compute the Fisher matrix elements for the autocorrelation likelihood given by Eq.~(\ref{eq:autocor-likelihood}), with the detector subscript suppressed,
 %
@@
 Matrix elements that are not listed have values that are implied by the symmetry of the Fisher matrix.
-Note that the minus signs are correct but a little confusing: despite them, $\mathcal{I}_{\gamma\gamma}, \mathcal{I}_{\tau\tau} \geq 0$ and $\mathcal{I}_{\gamma\tau} \leq 0$.
-The time integration limits $[-T, T]$ correspond to a flat prior on arrival time, or a time coincidence window between detectors.
+Note that the minus signs are correct but a little confusing; despite them, $\mathcal{I}_{\gamma\gamma}, \mathcal{I}_{\tau\tau} \geq 0$ and $\mathcal{I}_{\gamma\tau} \leq 0$.
+The time integration limits $[-T, T]$ correspond to a flat prior on arrival time or a time coincidence window between detectors.
 
@@
         \includegraphics[width=\columnwidth]{crlb_gamma}
-    \caption[\acl{CRLB} on time and phase accuracy]{\label{fig:crlb-tau}\ac{CRLB} on \ac{RMS} timing uncertainty and phase error, using the likelihood for the full \ac{GW} data (Equation~\ref{eq:gaussian-likelihood-spa}; dashed diagonal line) or the autocorrelation likelihood (Equation~\ref{eq:autocor-likelihood}; solid lines) with a selection of arrival time priors.}
+    \caption[\acl{CRLB} on time and phase accuracy]{\label{fig:crlb-tau}\ac{CRLB} on \acl{RMS} timing uncertainty and phase error, using the likelihood for the full \ac{GW} data [Eq.~(\ref{eq:gaussian-likelihood-spa}); dashed diagonal line] or the autocorrelation likelihood [Eq.~(\ref{eq:autocor-likelihood}); solid lines] with a selection of arrival time priors.}
 \end{figure*}
 
-We can show that the weighting function $w(t; \rho)$ approaches a Dirac delta function as $\rho \rightarrow \infty$, so that the Fisher matrix for the autocorrelation likelihood approaches the Fisher matrix for the full \ac{GW} data, Equation~(\ref{eq:fisher-matrix}), as $\rho \rightarrow \infty$.
-The Bessel functions asymptotically approach:
+We can show that the weighting function $w(t; \rho)$ approaches a Dirac delta function as $\rho \rightarrow \infty$, so that the Fisher matrix for the autocorrelation likelihood approaches the Fisher matrix for the full \ac{GW} data, Eq.~(\ref{eq:fisher-matrix}), as $\rho \rightarrow \infty$.
+The Bessel functions asymptotically approach
 %
@@
 %
-For large $\rho$, the exponents of $e^{\rho^2}$ dominate Equation~(\ref{eq:fisher-weight-factor}) and we can write:
+For large $\rho$, the exponents of $e^{\rho^2}$ dominate Eq.~(\ref{eq:fisher-weight-factor}), and we can write
 %
@@
 
-We can now write the Fisher matrix for the autocorrelation likelihood in a way that makes a comparison to the full signal model explicit. Define:
+We can now write the Fisher matrix for the autocorrelation likelihood in a way that makes a comparison to the full signal model explicit. Define
 %
@@
 %
-contain the integrals from Equations~(\ref{eq:fisher-autocor-gamma-gamma}, \ref{eq:fisher-autocor-tau-tau}, \ref{eq:fisher-autocor-gamma-tau}) and encode the departure of the autocorrelation likelihood from the likelihood of the full data at low \ac{SNR}.
-All of the $\textproto{D}_{ij}(\rho)$ are sigmoid-type functions that asymptotically approach 1 as $\rho \rightarrow \infty$ (see Figures~\ref{fig:crlb-tau} and~\ref{fig:fishfactor}).
+contain the integrals from Eqs.~(\ref{eq:fisher-autocor-gamma-gamma},~\ref{eq:fisher-autocor-tau-tau},~\ref{eq:fisher-autocor-gamma-tau}) and encode the departure of the autocorrelation likelihood from the likelihood of the full data at a low \ac{SNR}.
+All of the $\textproto{D}_{ij}(\rho)$ are sigmoid-type functions that asymptotically approach 1 as $\rho \rightarrow \infty$ (see Figs.~\ref{fig:crlb-tau} and~\ref{fig:fishfactor}).
 The transition \ac{SNR} $\rho_\mathrm{crit}$ is largely the same for all three nontrivial matrix elements, and is determined by the time coincidence window $T$ and the signal bandwidth $\omega_\mathrm{rms}$.
@@
     \caption[Ratio of Fisher matrix elements between autocorrelation likelihood and full \acs{GW} data]{\label{fig:fishfactor}Ratio between Fisher matrix elements (solid: $\textproto{D}_{\gamma\gamma}$, dashed: $\textproto{D}_{\gamma\tau}$, dotted: $\textproto{D}_{\tau\tau}$) for the autocorrelation likelihood and the full \ac{GW} data.
-Colors correspond to different arrival time priors as in Figure~\ref{fig:crlb-tau}.}
+Colors correspond to different arrival time priors as in Fig.~\ref{fig:crlb-tau}.}
 \end{figure}
@@
 
-On the other hand, in the low \ac{SNR} limit, the diminishing value of $\textproto{D}_{ij}(\rho)$ reflects the fact that some information is lost when the full data $\mathbf{x}$ is discarded.
+On the other hand, in the low \ac{SNR} limit, the diminishing value of $\textproto{D}_{ij}(\rho)$ reflects the fact that some information is lost when the full data $\mathbf{x}$ are discarded.
 Concretely, as the prior interval $T$ becomes large compared to 1/$\rho\omega_\mathrm{rms}$, the \ac{ML} estimator becomes more and more prone to picking up spurious noise fluctuations far from the true signal.
-Clearly, when the coincidence window $T$ is kept small as possible, more information is retained in the \ac{ML} point estimates.
+Clearly, when the coincidence window $T$ is kept as small as possible, more information is retained in the \ac{ML} point estimates.
 Put another way, if $T$ is small, then the transition \ac{SNR} $\rho_\mathrm{crit}$ is also small and fainter signals become useful for parameter estimation.
-In this way, the \ac{BAYESTAR} likelihood exhibits the \emph{threshold effect} that is well\nobreakdashes-known in communication and radar applications \citep{barankin1949locally,mcaulay1969barankin,mcaulay1971barankin}.
-
-In the following sections, we describe our prior and our numerical scheme to integrate over nuisance parameters, which amounts to the \ac{BAYESTAR} algorithm.
+In this way, the \ac{BAYESTAR} likelihood exhibits the \emph{threshold effect} that is well known in communication and radar applications \citep{barankin1949locally,mcaulay1969barankin,mcaulay1971barankin}.
+
+In the following sections, we describe our prior and our numerical schemes to integrate over nuisance parameters, which together amount to the \ac{BAYESTAR} algorithm.
 
@@
 We compute the \ac{GW} signal for a source with intrinsic parameters equal to the detection pipeline's estimate, $H(\omega; \hat{\bm\theta}_\mathrm{in})$.
-Then we find the \ac{SNR}=1 horizon distance $r_{1,i}$ for each detector by numerically integrating Equation~(\ref{eq:horizon}).
+Then, we find the \ac{SNR}=1 horizon distance $r_{1,i}$ for each detector by numerically integrating Eq.~(\ref{eq:horizon}).
 
@@
 %
-where $m=2$ for a prior that is uniform in volume, and $m=-1$ for a prior that is uniform in the logarithm of the distance.
-If a distance prior is not specified, the default is uniform in volume out to the maximum SNR=4 horizon distance:
+where $m=2$ for a prior that is uniform in volume and $m=-1$ for a prior that is uniform in the logarithm of the distance.
+If a distance prior is not specified, the default is uniform in volume out to the maximum \ac{SNR}=4 horizon distance:
 %
@@
 
-We compute the autocorrelation function for each detector from $t = 0$ to $t = T$ at intervals of $\Delta t = 1/f_s$, where $f_s$ is the smallest power of two that is greater than or equal to the Nyquist rate.
+We compute the autocorrelation function for each detector from $t = 0$ to $t = T$ at intervals of $\Delta t = 1/f_s$, where $f_s$ is the smallest power of 2 that is greater than or equal to the Nyquist rate.
 Because BNS signals typically terminate at about 1500~Hz, a typical value for $\Delta t$ is $(4096\,\textrm{Hz})^{-1}$.
@@
 
-The marginal posterior as a function of sky location is
+The marginal posterior as a function of the sky location is
 %
@@
 
-To marginalize over the coalescence phase, we can write $\tilde{\gamma}_i = \tilde{\gamma}_i^\prime + 2\phi_c$. Then integrating over $\phi_c$ and suppressing normalization factors, we get
+To marginalize over the coalescence phase, we can write $\tilde{\gamma}_i = \tilde{\gamma}_i^\prime + 2\phi_c$. Then, integrating over $\phi_c$ and suppressing normalization factors, we get
 %
@@
 
-The integrand is periodic in $\psi$, so simple Newton\nobreakdashes--Cotes quadrature over $\psi$ exhibits extremely rapid convergence (see Figure~\ref{fig:angle-convergence}).
-We therefore sample the posterior on a regular grid of 10 points from 0 to $\pi$.
-
-The integral over $\cos\iota$ converges just as rapidly with Gauss\nobreakdashes--Legendre quadrature (see Figure~\ref{fig:angle-convergence}), so we use a 10\nobreakdashes-point Gauss\nobreakdashes--Legendre rule for integration over $\cos\iota$.
+The integrand is periodic in $\psi$, so simple Newton\nobreakdashes--Cotes quadrature over $\psi$ exhibits extremely rapid convergence (see Fig.~\ref{fig:angle-convergence}).
+We therefore sample the posterior on a regular grid of ten points from 0 to $\pi$.
+
+The integral over $\cos\iota$ converges just as rapidly with Gauss\nobreakdashes--Legendre quadrature (see Fig.~\ref{fig:angle-convergence}), so we use a ten\nobreakdashes-point Gauss\nobreakdashes--Legendre rule for integration over $\cos\iota$.
 
@@
     \end{minipage}
-    \caption{\label{fig:angle-convergence}Relative error in \ac{BAYESTAR} integration scheme as a function of the number of Gaussian quadrature nodes.
-The two panels describe (a) the integral over the polarization angle $\psi$ and (b) the integral over inclination angle $\iota$.}
+    \caption{\label{fig:angle-convergence}Relative error in the \ac{BAYESTAR} integration scheme as a function of the number of Gaussian quadrature nodes.
+The two panels describe (a) the integral over the polarization angle $\psi$ and (b) the integral over the inclination angle $\iota$.}
 \end{figure*}
@@
 
-The distance integral is now performed differently from what we initially described in \cite{leo-singer-thesis,BerryLocalization}; the method described in the present work is about an order of magnitude faster.
-We define $\rho_i = \omega_i / r$ in order to absorb all of the distance-independent terms in the amplitudes into $\omega_i$, and then define
+The distance integral is now performed differently from what we initially described in Refs.~\cite{leo-singer-thesis,BerryLocalization}; the method described in the present work is about an order of magnitude faster.
+We define $\rho_i = \omega_i / r$ in order to absorb all of the distance-independent terms in the amplitudes into $\omega_i$ and then define
 %
@@
 %
-The coefficients $p^2$ and $b$ are nonnegative and independent of distance. $p$ has a maximum value of
+The coefficients $p^2$ and $b$ are non\nobreakdashes-negative and independent of distance. $p$ has a maximum value of
 %
@@
 %
-As shown in Figure~\ref{fig:radial-integral-interpolant}, the function basically consists of a plateau region in the upper-left half of the plane delimited by the lines $y = x$ and $x = \log p_0$, with
+As shown in Fig.~\ref{fig:radial-integral-interpolant}, the function basically consists of a plateau region in the upper-left half of the plane delimited by the lines $y = x$ and $x = \log p_0$, with
 %
@@
 We have now identified up to five breakpoints that partition the distance integrand into up to four intervals with quantitatively distinct behavior.
-These intervals are depicted in Figure~\ref{fig:radial_integrand} with distance increasing from left to right.
+These intervals are depicted in Fig.~\ref{fig:radial_integrand} with the distance increasing from left to right.
 There is a left\nobreakdashes-hand or small distance tail in which the integrand is small and monotonically increasing, a left\nobreakdashes- and right\nobreakdashes-hand side of the maximum likelihood peak, and a right\nobreakdashes-hand tail in which the integrand is small and monotonically decreasing.
-These breakpoints are:
+These breakpoints are
 %
@@
 
-We use these breakpoints as initial subdivisions in an adaptive Gaussian quadrature algorithm\footnote{for instance, \ac{GSL}'s \texttt{gsl\_integrate\_qagp} function, \url{http://www.gnu.org/software/gsl/manual/html_node/QAGP-adaptive-integration-with-known-singular-points.html}}.
-This function estimates the integral over each subdivision and each interval's contribution to the total error, then subdivides the interval whose error contribution is largest.
+We use these breakpoints as initial subdivisions in an adaptive Gaussian quadrature algorithm\footnote{For instance, \acl{GSL}'s \texttt{gsl\_integrate\_qagp} function, \url{http://www.gnu.org/software/gsl/manual/html_node/QAGP-adaptive-integration-with-known-singular-points.html}.}.
+This function estimates the integral over each subdivision and each interval's contribution to the total error, then subdivides the interval that contributes the most to the error.
 Subdivisions continue until a fixed total fractional error is reached.
@@
     \end{center}
-    \caption{\label{fig:radial_integrand}Illustration of initial subdivisions for distance integration scheme.
-Distance increases from left to right.
+    \caption{\label{fig:radial_integrand}Illustration of initial subdivisions for the distance integration scheme.
+The distance increases from left to right.
 In the color version, the left\nobreakdashes-hand tail, the left\nobreakdashes- and right\nobreakdashes-hand sides of the maximum likelihood peak, and the right\nobreakdashes-hand tail, are colored cyan, red, green, and blue, respectively.}
@@
 
-The interpolant is evaluated slightly differently depending on which of the three regions marked I, II, and III in Figure~\ref{fig:radial-integral-interpolant} contains the point of interest.
+The interpolant is evaluated slightly differently depending on which of the three regions marked I, II, and III in Fig.~\ref{fig:radial-integral-interpolant} contains the point of interest.
 In region I, we use bicubic interpolation of $\log \mathcal{G}$ in $x$ and $y$.
@@
 Finally, the distance integral $\mathscr{F}$ is obtained by multiplying the interpolated value of $\mathscr{G}$ by $\exp\left(p^2/{r_0}^2\right)$.
-For a $400\times400$ grid, the entire lookup table scheme is accurate to a relative error of about $10^{-5}$ in $\mathscr{F}$ (see Figure~\ref{fig:distance_interpolant_convergence}).
+For a $400\times400$ grid, the entire lookup table scheme is accurate to a relative error of about $10^{-5}$ in $\mathscr{F}$ (see Fig.~\ref{fig:distance_interpolant_convergence}).
 
@@
     \end{center}
-    \caption{\label{fig:distance_interpolant_convergence}Relative error in \ac{BAYESTAR} distance integral interpolation scheme as a function of the size of the grid.}
+    \caption{\label{fig:distance_interpolant_convergence}Relative error in the \ac{BAYESTAR} distance integral interpolation scheme as a function of the size of the grid.}
 \end{figure}
@@
 
-Our sampling of the sky relies completely on the \acl{HEALPix} (\acsu{HEALPix}, \cite{healpix}), a special data structure designed for all\nobreakdashes-sky maps.
+Our sampling of the sky relies completely on \ac{HEALPix}~\cite{healpix}, a special data structure designed for all\nobreakdashes-sky maps.
 \ac{HEALPix} divides the sky into equal\nobreakdashes-area pixels.
@@
 At this resolution, each pixel has an area of 13.4~deg$^2$.
-We then rank the pixels by contained probability (assuming constant probability density within a pixel), and subdivide the most probable $N_{\mathrm{pix},0}/4$ pixels into $N_{\mathrm{pix},0}$ new daughter pixels.
+We then rank the pixels by contained probability (assuming constant probability density within a pixel) and subdivide the most probable $N_{\mathrm{pix},0}/4$ pixels into $N_{\mathrm{pix},0}$ new daughter pixels.
 We then evaluate the posterior again at the centers of the new daughter pixels, sort again, and repeat seven times.
-After the last iteration, we have evaluated the posterior probability density a total of $8 N_{\mathrm{pix},0}$ times.
+By the end of the last iteration, we have evaluated the posterior probability density a total of $8 N_{\mathrm{pix},0}$ times.
 On \emph{most} subdivision steps, we descend one level deeper in \ac{HEALPix} resolution.
-This process is illustrated in Figure~\ref{fig:adaptive-illustration}.
+This process is illustrated in Fig.~\ref{fig:adaptive-illustration}.
 
@@
 The resulting map is a tree structure that describes a mesh of pixels with different resolutions.
-An example \ac{BAYESTAR} subdivision is shown in Figure~\ref{fig:adaptive-mesh}.
-To convert this mesh into a FITS image, we traverse the tree and flatten it into the highest resolution represented.
+An example \ac{BAYESTAR} subdivision is shown in Fig.~\ref{fig:adaptive-mesh}.
+To convert this mesh into a \ac{FITS}~\cite{FITS} image, we traverse the tree and flatten it into the highest resolution represented.
 The highest possible resolution is $N_\mathrm{side}=2^{11}$, with an area of $\approx 10^{-3}$~deg$^2$ per pixel.\footnote{Although the resulting sky map contains $N_\mathrm{pix} \approx 5\times10^6$ pixels, at most $\approx 2\times10^4$ pixels have distinct values.
@@
     \includegraphics[width=\columnwidth]{adaptive_mesh}
-    \caption{\label{fig:adaptive-mesh}An example multi\nobreakdashes-resultion \ac{HEALPix} mesh arising from the \ac{BAYESTAR} sampling scheme (plotted in a cylindrical projection).
-This is event 18951 from \cite{FirstTwoYears}.}
+    \caption{\label{fig:adaptive-mesh}An example multiresolution \ac{HEALPix} mesh arising from the \ac{BAYESTAR} sampling scheme (plotted in a cylindrical projection).
+This is event 18951 from Ref.~\cite{FirstTwoYears}.}
 \end{figure}
@@
 \ac{BAYESTAR} consists of nine computationally intensive loops: the generation of the distance integral lookup table and the eight loops over pixels in the adaptive HEALPix sampling step.
-The iterations of each loop are distributed across multiple cores using OpenMP\footnote{http://openmp.org/} compiler directives.
-In Section~\ref{sec:run-time}, we will show that \ac{BAYESTAR}'s run time is almost perfectly proportional to the number of cores, demonstrating that the serial sections (the sorts between the adaptation steps) are a negligible contribution to the overall wall clock time.
+The iterations of each loop are distributed across multiple cores using OpenMP\footnote{\url{http://openmp.org/}.} compiler directives.
+In Sec.~\ref{sec:run-time}, we will show that \ac{BAYESTAR}'s run time is almost perfectly proportional to the number of cores, demonstrating that the serial sections (the sorts between the adaptation steps) are a negligible contribution to the overall wall clock time.
 
@@
 We have completed our description of the \ac{BAYESTAR} algorithm.
-In \cite{FirstTwoYears}, the authors presented a comprehensive and astrophysically realistic sample of simulated \ac{BNS} mergers.
-We focused on the first two planned Advanced LIGO and Virgo observing runs as desribed in \cite{LIGOObservingScenarios}.
-That work presented a catalog of 500 sky localizations from \ac{BAYESTAR} and LALINFERENCE, and dealt with the quantitave position reconstruction accuracy as well as the qualitative sky morphologies.
-In the present work, we will use the same data set, but instead focus on demonstrating the correctness and performance of the \ac{BAYESTAR} algorithm.
+In Ref.~\cite{FirstTwoYears}, the authors presented a comprehensive and astrophysically realistic sample of simulated \ac{BNS} mergers.
+We focused on the first two planned Advanced LIGO and Virgo observing runs as desribed in Ref.~\cite{LIGOObservingScenarios}.
+That work presented a catalog of 500 sky localizations from \ac{BAYESTAR} and LALINFERENCE and dealt with the quantitative position reconstruction accuracy as well as the qualitative sky morphologies.
+In the present work, we will use the same data set but instead focus on demonstrating the correctness and performance of the \ac{BAYESTAR} algorithm.
 
@@
 
-To review the assumptions made in \cite{FirstTwoYears}, the two scenarios are:
-
-\paragraph*{2015} The first Advanced \acs{LIGO} observing run, or ``O1'', scheduled to start in September 2015 and continue for three months.
-There are only two detectors participating in this run: \acs{LIGO} Hanford (``H'') and \acs{LIGO} Livingston (``L'').
+To review the assumptions made in Ref.~\cite{FirstTwoYears}, the two scenarios are:
+
+\textit{2015}.---The first Advanced \acs{LIGO} observing run, or ``O1,'' scheduled to start in September 2015 and continue for three months.
+There are only two detectors participating in this run: \acs{LIGO} Hanford (H) and \acs{LIGO} Livingston (L).
 Both detectors are expected to operate with a direction\nobreakdashes-averaged \ac{BNS} merger range of 40\nobreakdashes--80\,Mpc (though ongoing Advanced \acs{LIGO} commissioning suggests that the higher end of this range will be achieved).
@@
 
-\paragraph*{2016} The second observing run, ``O1'', with the two between Advanced \acs{LIGO} detectors, upgraded to a \ac{BNS} range of 80\nobreakdashes--120\,Mpc, operated jointly with the newly commissioned Advanced Virgo detector (``V''), operating at a range of 20\nobreakdashes--60\,Mpc.
+\textit{2016}.---The second observing run, ``O2,'' with the two between Advanced \acs{LIGO} detectors, upgraded to a \ac{BNS} range of 80\nobreakdashes--120\,Mpc, operated jointly with the newly commissioned Advanced Virgo detector (V), operating at a range of 20\nobreakdashes--60\,Mpc.
 The run is envisioned as lasting for six months in 2016\nobreakdashes--2017.
 The detectors are assumed to have random and independent 80\% duty cycles.
-Consequently, all three detectors (``HLV'') are in science mode about half of the time, with the remaining time divided roughly equally between each of the possible pairs (``HL'', ``HV'', or ``LV'') and one or fewer detectors (at least two \ac{GW} facilities are required for a detection).
-Virgo's range is assumed to be somewhat less than \acs{LIGO}'s because its commissioning timetable is about a year behind.
-Although the simulated signals are generally too weak in Virgo to trigger the matched\nobreakdashes-filter pipeline and contribute to \emph{detection}, even these sub\nobreakdashes-threshold signals aid in position reconstruction with LALINFERENCE by lifting degeneracies.
+Consequently, all three detectors (HLV) are in science mode about half of the time, with the remaining time divided roughly equally between each of the possible pairs (HL, HV, or LV) and one or fewer detectors (at least two \ac{GW} facilities are required for a detection).
+Virgo's range is assumed to be somewhat less than \acs{LIGO}'s because its commissioning time table is about a year behind.
+Although the simulated signals are generally too weak in Virgo to trigger the matched\nobreakdashes-filter pipeline and contribute to \emph{detection}, even these subthreshold signals aid in position reconstruction with LALINFERENCE by lifting degeneracies.
 As a result, the median 90\% credible area decreases to about 200\,deg$^2$.
@@
 
-The simulated waveforms were deposited in Gaussian noise that has been filtered to have the \acp{PSD} consistent with \cite{LIGOObservingScenarios}.
+The simulated waveforms were deposited in Gaussian noise that has been filtered to have the \acp{PSD} consistent with Ref.~\cite{LIGOObservingScenarios}.
 They were detected using the real\nobreakdashes-time matched\nobreakdashes-filter pipeline, GSTLAL\_INSPIRAL \citep{Cannon:2011vi}.
-Candidates with estimated \acp{FAR} less than $10^{-2}$\,yr$^{-1}$ were considered to be ``detections.'' Because using Gaussian noise results in lower \acp{FAR} than would be calculated in realistically glitchy detector noise, we imposed an additional detection threshold on the network \ac{SNR}, $\hat{\rho} \geq 12$, which has been found to correspond to a comparable \ac{FAR} in the initial \ac{LIGO} runs.\footnote{See \cite{BerryLocalization} for an analysis of the effect of glitchy noise on detection and parameter estimation.} Localizations for the ``detections'' were generated with \ac{BAYESTAR} as well as the functionally equivalent and interchangeable LALINFERENCE\_MCMC, LALINFERENCE\_NEST, and LALINFERENCE\_BAMBI samplers (collectively referred to as LALINFERENCE).
+Candidates with estimated \acp{FAR} less than $10^{-2}$\,yr$^{-1}$ were considered to be ``detections.'' Because using Gaussian noise results in lower \acp{FAR} than would be calculated in realistically glitchy detector noise, we imposed an additional detection threshold on the network \ac{SNR}, $\hat{\rho} \geq 12$, which has been found to correspond to a comparable \ac{FAR} in the initial \ac{LIGO} runs.\footnote{See Ref.~\cite{BerryLocalization} for an analysis of the effect of glitchy noise on detection and parameter estimation.} Localizations for the detections were generated with \ac{BAYESTAR} as well as the functionally equivalent and interchangeable LALINFERENCE\_MCMC, LALINFERENCE\_NEST, and LALINFERENCE\_BAMBI samplers (collectively referred to as LALINFERENCE).
 
@@
 Then, we computed the cumulative sum of the pixels in that order.
-Finally, we searched for the index of the pixel whose cumulative sum was equal to a given value: for example, 0.9 if we are interested in the 90\% credible area.
+Finally, we searched for the index of the pixel of which the cumulative sum was equal to a given value: for example, 0.9 if we are interested in the 90\% credible area.
 That pixel index times the area per pixel is the area of the smallest region of the specified credible level.
@@
 
-Histograms of the 90\% credible area and the searched area are shown in Figure~\ref{fig:area-hist}, broken down by observing scenario (2015 or 2016) and detector network (HL, HV, LV, or HLV).
-Note that there are no stastically significant differences in areas between \ac{BAYESTAR} and LALINFERENCE, with the exception in the 2016/HLV configuration, for which some LALINFERENCE sky maps span about an order of magnitude less area than \ac{BAYESTAR}.
+Histograms of the 90\% credible area and the searched area are shown in Fig.~\ref{fig:area-hist}, broken down by observing scenario (2015 or 2016) and detector network (HL, HV, LV, or HLV).
+Note that there are no statistically significant differences in areas between \ac{BAYESTAR} and LALINFERENCE, with the exception in the 2016/HLV configuration, for which some LALINFERENCE sky maps span about an order of magnitude less area than \ac{BAYESTAR}.
 If we consider only events for which all three detectors contained a signal that was loud enough to trigger the matched\nobreakdashes-filter pipeline, the difference becomes much smaller and insignificant within 95\% error bars.
@@
 The plots in the left column show the 90\% credible area and the plots in the right column show the searched area.
-From top bottom, the rows refer to the following observing scenarios/network configurations: 2015/HL, 2016/HL, 2016/HV, 2016/LV, 2016/HLV.
+From top bottom, the rows refer to the following observing scenarios/network configurations: 2015/HL, 2016/HL, 2016/HV, 2016/LV, and 2016/HLV.
 The shaded regions represent the 95\% confidence bounds.
-The magenta lines are represent \ac{BAYESTAR} and the blue lines LALINFERENCE.
-Where relevant, dotted lines show all events in the given network configuratoin and solid lines show only events for which the matched\nobreakdashes-filter pipeline triggered on all operating detectors.
+The magenta lines represent \ac{BAYESTAR} and the blue lines LALINFERENCE.
+Where relevant, dotted lines show all events in the given network configuration and solid lines show only events for which the matched\nobreakdashes-filter pipeline triggered on all operating detectors.
 Note that statistically significant differences in areas between the \ac{BAYESTAR} and LALINFERENCE localizations occur only for events that were below the detection threshold in one or more detectors.}
@@
 However, there may be a simple remedy.
-As we noted in Section~\ref{sec:properties}, the \ac{BAYESTAR} likelihood can be modified to use, instead of the times, phases, and amplitudes on arrival, the full complex matched\nobreakdashes-filter time series from all detectors.
-The detection pipeline, GSTLAL\_INSPIRAL, would have to be modifed to save and transmit a small interval of the complex \ac{SNR} time series (perhaps a few tens of milliseconds) around the time of each detection candidate.
-In addition to supplying the missing information for sub\nobreakdashes-threshold signals, this would make \ac{BAYESTAR} mathematically equivalent to the LALINFERENCE analysis, but with the intrinsic parameters fixed to their maximum\nobreakdashes-likelihood values.
+As we noted in Sec.~\ref{sec:properties}, the \ac{BAYESTAR} likelihood can be modified to use, instead of the times, phases, and amplitudes on arrival, the full complex matched\nobreakdashes-filter time series from all detectors.
+The detection pipeline, GSTLAL\_INSPIRAL, would have to be modified to save and transmit a small interval of the complex \ac{SNR} time series (perhaps a few tens of milliseconds) around the time of each detection candidate.
+In addition to supplying the missing information for subthreshold signals, this would make \ac{BAYESTAR} mathematically equivalent to the LALINFERENCE analysis, but with the intrinsic parameters fixed to their maximum\nobreakdashes-likelihood values.
 This idea will be pursued in future work.
 
-\subsection{Self consistency}
+\subsection{Self-consistency}
 
 As we observed above, the area of a given credible region describes the precision of the sky localization, whereas the searched area describes the accuracy.
-However, self consistency requires that the two are related.
+However, self-consistency requires that the two are related.
 For example, we should find that on average 90\% of events have their true locations contained within their respective 90\% credible regions.
 More generally, if we make a cumulative histogram of the credible levels corresponding to the searched areas of all of the events, then we should obtain a diagonal line (with small deviations due to finite sample size).
-This test, popularized for \ac{GW} data analysis by \cite{SiderySkyLocalizationComparison}, is a necessary but not sufficient condition for the validity of any Bayesian parameter estimation scheme.
+This test, popularized for \ac{GW} data analysis by Ref.~\cite{SiderySkyLocalizationComparison}, is a necessary but not sufficient condition for the validity of any Bayesian parameter estimation scheme.
 
@@
 We found at first that \ac{BAYESTAR}'s $P$\nobreakdashes--$P$ plots tended to sag below the diagonal, indicating that though the accuracy (i.e., searched area) was comparable to LALINFERENCE, the \emph{precision} was overstated, with confidence intervals that were only about 70\% of the correct area.
-This was rectified by pre\nobreakdashes-scaling the \acp{SNR} from GSTLAL\_INSPIRAL by a factor of 0.83 prior to running \ac{BAYESTAR}.
-This correction factor suggests that, for example, an \ac{SNR} 10 trigger from GSTLAL\_INSPIRAL has the \emph{effective information content} of an \ac{SNR} 8.3 signal.
-The missing information may be due to losses from the discreteness of the template bank, from the \ac{SVD}, from mismatch between the matched\nobreakdashes-filter templates and the simulated signals, from the small but nonzero correlations between masses and intrinsic parameters, or from elsewhere within the detection pipeline.
-The correction is hard\nobreakdashes-coded into the rapid localization.
-With it, the $P$\nobreakdashes--$P$ plots are diagonalized without negatively affecting the searched area (see Figure~\ref{fig:pp}).
+This was rectified by prescaling the \acp{SNR} from GSTLAL\_INSPIRAL by a factor of 0.83 prior to running \ac{BAYESTAR}.
+This correction factor suggests that, for example, a \ac{SNR}=10 trigger from GSTLAL\_INSPIRAL has the \emph{effective information content} of a \ac{SNR}=8.3 signal.
+The missing information may be due to losses from the discreteness of the template bank, from the \acl{SVD}, from mismatch between the matched\nobreakdashes-filter templates and the simulated signals, from the small but nonzero correlations between masses and intrinsic parameters, or from elsewhere within the detection pipeline.
+The correction is hard coded into the rapid localization.
+With it, the $P$\nobreakdashes--$P$ plots are diagonalized without negatively affecting the searched area (see Fig.~\ref{fig:pp}).
 
@@
 We timed it under Scientific~Linux~6.1 on a Supermicro SuperServer 6028TP\nobreakdashes-HTTR system with dual 8\nobreakdashes-core Intel Xeon E5\nobreakdashes-2630~v3 CPUs clocked at 2.40\,GHz, capable of executing 32 threads simultaneously (with hyperthreading).
-In Figure~\ref{fig:runtimes}, we show how long it took to calculate a localization with \ac{BAYESTAR} as the number of OpenMP threads was varied from 1 to 32.
+In Fig.~\ref{fig:runtimes}, we show how long it took to calculate a localization with \ac{BAYESTAR} as the number of OpenMP threads was varied from 1 to 32.
 This is a violin plot, a smoothed vertical histogram.
 The magenta regions show run times for a two\nobreakdashes-detector network (HL) modeled on the first scheduled Advanced \ac{LIGO} observing run in 2015, and the blue regions show run times for a three\nobreakdashes-detector network (HLV) based on the second planned observing run in 2016.
-These are the two observing scenarios that are discussed in \cite{FirstTwoYears}.
+These are the two observing scenarios that are discussed in Ref.~\cite{FirstTwoYears}.
 
@@
 With 32 threads, \ac{BAYESTAR} takes just 4\nobreakdashes--13~s.
-In practice, \ac{BAYESTAR}'s data handling (reading the detectors' \acp{PSD}, communicating with the \ac{GW} candidate database, writing FITS files) take an additional 15~s or so, though this overhead could be reduced by parallelizing many of these steps.
-The overall latency is comparable to the other stages (data aggregation, trigger generation, alert distribution) in the real\nobreakdashes-time \ac{BNS} analysis; therefore any significant further speedup would require significant changes through Advanced \acs{LIGO} computing and infrastructure.
-The 32\nobreakdashes-thread configuration is representative of how \ac{BAYESTAR} might be deployed in early Advanced \ac{LIGO}.\footnote{\ac{BAYESTAR} has been successfully ported to the Intel's Many Integrated Core~(MIC) architecture and has been tested in a 500 thread configuration on a system with dual Intel Xeon Phi 5110P coprocessors.} For comparison, sky localization with LALINFERENCE takes about 100 hours \cite{BerryLocalization}.
-
-Note that this benchmark shows \ac{BAYESTAR} to be an order of magnitude faster than what was reported in \cite{leo-singer-thesis,BerryLocalization} due to the changes in the distance integration scheme that we noted in Section~\ref{sec:distance}.
+In practice, \ac{BAYESTAR}'s data handling (reading the detectors' \acp{PSD}, communicating with the \ac{GW} candidate database, writing \ac{FITS} files) takes an additional 15~s or so, though this overhead could be reduced by parallelizing many of these steps.
+The overall latency is comparable to the other stages (data aggregation, trigger generation, alert distribution) in the real\nobreakdashes-time \ac{BNS} analysis; therefore, any significant further speedup would require significant changes through Advanced \acs{LIGO} computing and infrastructure.
+The 32\nobreakdashes-thread configuration is representative of how \ac{BAYESTAR} might be deployed in early Advanced~\ac{LIGO}.\footnote{\ac{BAYESTAR} has been successfully ported to the Intel's Many Integrated Core architecture and has been tested in a 500 thread configuration on a system with dual Intel Xeon Phi 5110P coprocessors.} For comparison, sky localization with LALINFERENCE takes about 100\,h~\cite{BerryLocalization}.
+
+Note that this benchmark shows \ac{BAYESTAR} to be an order of magnitude faster than what was reported in Refs.~\cite{leo-singer-thesis,BerryLocalization} due to the changes in the distance integration scheme that we noted in Sec.~\ref{sec:distance}.
 
@@
 
-One immediately pressing direction for future work is to address the issue of sub\nobreakdashes-threshold signals, as this will be a major issue when Advanced Virgo comes online in 2016\nobreakdashes--2017.
-Using the full \ac{SNR} time series in place of the autocorrelation function seems like a promising avenue; implementing this requires some infrastructure changes to both the matched filter pipeline and \ac{BAYESTAR}.
-Along these lines, we also refer the reader to \cite{PankowRapidPE} for a similar, non-\ac{MCMC} approach to rapid exploration of the full parameter space.
+One immediately pressing direction for future work is to address the issue of subthreshold signals, as this will be a major issue when Advanced Virgo comes online in 2016\nobreakdashes--2017.
+Using the full \ac{SNR} time series in place of the autocorrelation function seems like a promising avenue; implementing this requires some infrastructure changes to both the matched-filter pipeline and \ac{BAYESTAR}.
+Along these lines, we also refer the reader to Ref.~\cite{PankowRapidPE} for a similar, non\nobreakdashes-\ac{MCMC} approach to the rapid exploration of the full parameter space.
 
 A more open\nobreakdashes-ended question is how to account for spin precession.
-The simulations in \cite{FirstTwoYears} and in this paper featured extremely modest spins of $\chi \leq 0.05$, consistent with the fastest known pulsars in binaries \citep{2003Natur.426..531B,DetectingBNSSystemsWithSpin}.
+The simulations in Ref.~\cite{FirstTwoYears} and in this paper featured extremely modest spins of $\chi \leq 0.05$, consistent with the fastest known pulsars in binaries \citep{2003Natur.426..531B,DetectingBNSSystemsWithSpin}.
 The signals were detected using a template bank that lacked spins entirely.
-Farr et~al.
-(in preparation) show that using nonspinning \ac{BNS} templates for parameter estimation has negligible impact on sky localization.
+Ref.~\cite{FirstTwoYearsSpin} shows that using nonspinning \ac{BNS} templates for parameter estimation has negligible impact on sky localization.
 However, if one or both companions are spinning as fast as a millisecond pulsar, $\chi \sim 0.4$ \cite{FastestSpinningMillisecondPulsar}, or even near breakup, $\chi \sim 0.7$, then the orbital plane may precess; in this case, spins can no longer be neglected for detection \cite{DetectingBNSSystemsWithSpin} and may also be important for parameter estimation.
@@
 
-Although the response time of \ac{BAYESTAR} has been driven by the anticipated timescales for kilonova and afterglow emission, a recurring question is whether there is any detectable \ac{EM} signal in the seconds before, during, and after the merger itself.
+Although the response time of \ac{BAYESTAR} has been driven by the anticipated time scales for kilonova and afterglow emission, a recurring question is whether there is any detectable \ac{EM} signal in the seconds before, during, and after the merger itself.
 Since the \ac{GW} inspiral signal is in principal detectable for up to hundreds of seconds before merger, one could imagine positioning rapidly slewing instruments to search for any prompt emission.
-This concept was explored by the authors \citep{Cannon:2011vi}, and will be revisted in further detail in both Sathyaprakash et~al.
-(in preparation) and Chu et~al.
-(in preparation).
-
-On the topic of very low\nobreakdashes-latency localization, we also recommend Chen~\&~Holz (in preparation), who propose a rapid localization scheme that is similar to ours, but even faster because it makes some additional compromises: their likelihood is strictly Gaussian, so one more marginalization integral (the integral over arrival time) can be performed analytically.
+This concept was explored by the authors \citep{Cannon:2011vi}.
+
+On the topic of very low\nobreakdashes-latency localization, we also recommend Chen~\&~Holz~\cite{ChenHolzRapidLocalization}, who propose a rapid localization scheme that is similar to ours, but even faster because it makes some additional compromises: their likelihood is strictly Gaussian, so one more marginalization integral (the integral over arrival time) can be performed analytically.
 
@@
 We have presented a novel, fast, accurate, Bayesian algorithm for inferring the sky locations of compact binary merger sources that may soon be detected by advanced ground\nobreakdashes-based \ac{GW} detectors.
-For \ac{BNS} systems with small spins, we have shown that \ac{BAYESTAR} produces sky maps that are as accurate as the full \ac{MCMC} parameter estimation code, but can do so within $\sim$10 seconds after a detection.
-Still faster response times should be possible in the future (if warranted) by deploying \ac{BAYESTAR} on machines with more cores, or by distributing \ac{BAYESTAR} across multiple computers.
+For \ac{BNS} systems with small spins, we have shown that \ac{BAYESTAR} produces sky maps that are as accurate as the full \ac{MCMC} parameter estimation code but can do so within $\sim$10\,s after a detection.
+Still faster response times should be possible in the future (if warranted) by deploying \ac{BAYESTAR} on machines with more cores or by distributing \ac{BAYESTAR} across multiple computers.
 
@@
 The localization from \ac{BAYESTAR} will be available tens of seconds to a minute later.
-Finally, the refined localiation and the detailed estimates of masses and spins from LALINFERENCE will be distributed hours to days later.
-
-Relevant timescales for possible \ac{EM} counterparts to \ac{GW} signals include seconds (the prompt \ac{GRB} signature), hundreds of seconds (extended emission and X\nobreakdashes-ray plateaus that are observed for some short \acp{GRB}), minutes to hours (X\nobreakdashes-ray and optical afterglow), hours to days (the kilonova or the blue flashes associated with unbound ejecta or disk winds), and days to years (the radio afterglow).
-For the first time, we are able to provide accurate localizations before the peak of any of these \ac{EM} signatures (except for the short \ac{GRB} or any pre\nobreakdashes-merger signal).
+Finally, the refined localization and the detailed estimates of masses and spins from LALINFERENCE will be distributed hours to days later.
+
+Relevant time scales for possible \ac{EM} counterparts to \ac{GW} signals include seconds (the prompt \ac{GRB} signature), hundreds of seconds (extended emission and X\nobreakdashes-ray plateaus that are observed for some short \acp{GRB}), minutes to hours (X\nobreakdashes-ray and optical afterglow), hours to days (the kilonova or the blue flashes associated with unbound ejecta or disk winds), and days to years (the radio afterglow).
+For the first time, we are able to provide accurate localizations before the peak of any of these \ac{EM} signatures (except for the short \ac{GRB} or any premerger signal).
 Even for components like the kilonova that should peak within hours to days, the availability of the localizations within seconds might provide a window of several hours to obtain tiled images of the area before the \ac{EM} emission begins.
@@
 \begin{acknowledgements}
-This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v4.
-
-We thank John Veitch and Will Farr for chairing a review of the analysis and code. We thank Britt Griswold for assistance with preparing Figure~\ref{fig:adaptive-illustration}.
+This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v6.
+
+We thank John Veitch and Will Farr for chairing a review of the analysis and code. We thank Britt Griswold for assistance with preparing Fig.~\ref{fig:adaptive-illustration}.
 \end{acknowledgements}
 
-LIGO was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the \ac{NSF} and operates under cooperative agreement PHY\nobreakdashes-0107417.
-
-This research was supported by the \ac{NSF} through a Graduate Research Fellowship to L.S.
-L.S.
+LIGO was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the \ac{NSF} and operates under Cooperative Agreement No.~PHY\nobreakdashes-0107417.
+
+This research was supported by the \ac{NSF} through a Graduate Research Fellowship to L.~S.
+L.~S.
 thanks the Aspen Center for Physics and \ac{NSF} Grant \#1066293 for hospitality during the editing of this paper.
 
-Source code for \ac{BAYESTAR} is available as part of LALInference\footnote{\url{https://ligo-vcs.phys.uwm.edu/cgit/lalsuite/tree/lalinference}}, the open source LIGO/Virgo parameter estimation toolchain, which is in turn part of LALSuite\footnote{\url{http://www.lsc-group.phys.uwm.edu/daswg/projects/lalsuite.html}}.
-This research made use of Astropy\footnote{\url{http://www.astropy.org}} \citep{astropy}, a community-developed core Python package for astronomy.
-Some of the results in this paper have been derived using \ac{HEALPix}\footnote{\url{http://healpix.sourceforge.net}} \cite{healpix}.
-
-Some results were produced on the NEMO computing cluster operated by the Center for Gravitation and Cosmology at University of Wisconsin\nobreakdashes--Milwaukee under \ac{NSF} Grants PHY\nobreakdashes-0923409 and PHY\nobreakdashes-0600953.
+Source code for \ac{BAYESTAR} is available as part of LALInference,\footnote{\url{https://ligo-vcs.phys.uwm.edu/cgit/lalsuite/tree/lalinference}.} the open source LIGO/Virgo parameter estimation toolchain, which is in turn part of LALSuite.\footnote{\url{http://www.lsc-group.phys.uwm.edu/daswg/projects/lalsuite.html}.} This research made use of Astropy\footnote{\url{http://www.astropy.org}.}~\citep{astropy}, a community-developed core Python package for astronomy.
+Some of the results in this paper have been derived using \ac{HEALPix}\footnote{\url{http://healpix.sourceforge.net}.}~\cite{healpix}.
+
+Some results were produced on the NEMO computing cluster operated by the Center for Gravitation and Cosmology at University of Wisconsin\nobreakdashes--Milwaukee under \ac{NSF} Grants No.~PHY\nobreakdashes-0923409 and No.~PHY\nobreakdashes-0600953.
 
@@
 The averages are correlated with the intrinsic parameters, but neither are correlated with the differences.
-Since only the differences inform sky location, this gives us license to neglect uncertainty in masses when we are computing sky resolution.
+Since only the differences inform the sky location, this gives us license to neglect uncertainty in masses when we are computing the sky resolution.
 
@@
 Due to our assumption that the detectors' \acp{PSD} are proportional to each other, the noise moments are the same for all detectors, ${\overline{\omega^k}}_i \equiv {\overline{\omega^k}}$.
-Then we can write the single-detector Fisher matrix as
+Then, we can write the single-detector Fisher matrix as
 %
@@
 The averages and the masses are correlated with each other, but are not correlated with the differences.
-Because only the differences are informative for sky localization, we drop the intrinsic parameters from the rest of the Fisher matrix calculations in this appendix.
+Because only the differences are informative for sky localization, we drop the intrinsic parameters from the rest of the Fisher matrix calculations in the Appendix.
 
@@
 
-The Fisher matrix in Equation~(\ref{eq:fisher-matrix}) is block diagonal, which implies that estimation errors in the signal amplitude $\rho$ are uncorrelated with the phase $\gamma$ and time $\tau$.
+The Fisher matrix in Eq.~(\ref{eq:fisher-matrix}) is block diagonal, which implies that estimation errors in the signal amplitude $\rho$ are uncorrelated with the phase $\gamma$ and time $\tau$.
 A sequence of two changes of variables lends some physical interpretation to the nature of the coupled estimation errors in $\gamma$~and~$\tau$.
@@
 %
-Thus, in the appopriate time units, the \textit{sum and difference} of the phase and time of the signal are measured independently.
+Thus, in the appropriate time units, the \textit{sum and difference} of the phase and time of the signal are measured independently.
 
@@
 Finally, we will calculate the position resolution of a network of \ac{GW} detectors.
-We could launch directly into computing derivatives of the full signal model from Equation~(\ref{eq:full-signal-model}) with respect to all of the parameters, but this would result in a very complicated expression.
+We could launch directly into computing derivatives of the full signal model from Eq.~(\ref{eq:full-signal-model}) with respect to all of the parameters, but this would result in a very complicated expression.
 Fortunately, we can take two shortcuts.
-First, since we showed in Section~\ref{sec:independence-intrinsic-extrinsic} that the intrinsic parameters are correlated only with an overall nuisance average arrival time, amplitude, and phase, we need not consider the derivatives with respect to mass at all.
-Second, we can reuse the extrinsic part of the single detector Fisher matrix from Equation~(\ref{eq:fisher-matrix}) by computing the much simpler Jacobian matrix to transform from the time, amplitude, and phase on arrival, to the parameters of interest.
-
-We begin by transforming the single\nobreakdashes-detector Fisher matrix from a polar to a rectangular representation of the complex amplitude given in Equations~(\ref{eq:optimal-rho}, \ref{eq:optimal-tau}), $\rho_i, \gamma_i \rightarrow \Re[z_i] = \rho_i \cos \gamma_i, \Im[z_i] = \rho_i \sin \gamma_i$:
+First, since we showed in Appendix~\ref{sec:independence-intrinsic-extrinsic} that the intrinsic parameters are correlated only with an overall nuisance average arrival time, amplitude, and phase, we need not consider the derivatives with respect to mass at all.
+Second, we can reuse the extrinsic part of the single-detector Fisher matrix from Eq.~(\ref{eq:fisher-matrix}) by computing the much simpler Jacobian matrix to transform from the time, amplitude, and phase on arrival to the parameters of interest.
+
+We begin by transforming the single\nobreakdashes-detector Fisher matrix from a polar to a rectangular representation of the complex amplitude given in Eqs.~(\ref{eq:optimal-rho}, \ref{eq:optimal-tau}), $\rho_i, \gamma_i \rightarrow \Re[z_i] = \rho_i \cos \gamma_i, \Im[z_i] = \rho_i \sin \gamma_i$:
 %
@@
 %
-We only need a first order expression for $\delta R$, because we will be taking products of first derivatives of it\footnote{Caution: the angles $\delta\theta$ and $\delta\phi$ represent displacements in two orthogonal directions, but are \emph{not} necessarily simply related to $\theta$ and $\phi$.}:
+We only need a first-order expression for $\delta R$, because we will be taking products of first derivatives of it\footnote{Caution: the angles $\delta\theta$ and $\delta\phi$ represent displacements in two orthogonal directions, but are \emph{not} necessarily simply related to $\theta$ and $\phi$.}:
 %
@@
 How do we extract the dimensions of the localization from the Fisher matrix?
-If there are $N \leq 2$ detectors, then the Fisher matrix must be degenerate, because there are $3N$ measurements and 7 parameters:
-%
-\begin{widetext}
-\begin{equation*}
+If there are $N \leq 2$ detectors, then the Fisher matrix must be degenerate, because there are $3N$ measurements and seven parameters:
+%
+\begin{align*}
     \left\{
@@
     \right\}
-    = 7 \text{ parameters}
-    \qquad
+    =& \; 7 \text{ parameters} \;
     \longleftrightarrow
-    \qquad
     \left\{
@@
     \end{array}
-    \right\}
-    \times N = 3N \text{ observables.}
-\end{equation*}
-\end{widetext}
+    \right\} \\
+    &\times N = 3N \text{ observables.}
+\end{align*}
 %
@@
 
-For $N \geq 3$ detectors, the parameters are over\nobreakdashes-constrained by the data and the Fisher matrix describes the dimensions of an ellipse.
+For $N \geq 3$ detectors, the parameters are overconstrained by the data, and the Fisher matrix describes the dimensions of an ellipse.
 Within a circle of radius $r$ centered on the origin, the enclosed probability $p$ is
@@
 This describes a $1\sigma$ uncertainty ellipse that has major and minor radii ${\lambda_1}^{-1/2}$, ${\lambda_2}^{-1/2}$, and area $A_{1\sigma} = \pi / \sqrt{\lambda_1 \lambda_2} = \pi / \sqrt{\det\mathcal{I}}$.
-Then the area of an ellipse containing probability $p$ is
+Then, the area of an ellipse containing probability $p$ is
 %
@@
 Using the above derivation, we arrive at a prediction for the sky resolution of a \ac{GW} detector network.
-We took some shortcuts that allowed us to avoid directly evaluating the complicated derivatives of the signal itself with respect to sky location.
+We took some shortcuts that allowed us to avoid directly evaluating the complicated derivatives of the signal itself with respect to the sky location.
 As a result, the expressions involved in each step are simple enough to be manually entered into a computer program.
@@
 \begin{enumerate}
-    \item Compute, for each detector, the horizon distance $r_{1,i}$, the angular frequency moments ${\overline{\omega}}_i$ and ${\overline{\omega^2}}_i$, and $(h_+, h_\times)$ from Equations~(\ref{eq:hplus},~\ref{eq:hcross}).
+    \item Compute, for each detector, the horizon distance $r_{1,i}$, the angular frequency moments ${\overline{\omega}}_i$ and ${\overline{\omega^2}}_i$, and $(h_+, h_\times)$ from Eqs.~(\ref{eq:hplus},~\ref{eq:hcross}).
 (These can be reused for multiple source positions as long as the masses and the detector noise \acp{PSD} are the same.)
-    \item For a given $\phi, \theta, \psi$, compute the complex received amplitude $z_i$ from Equations~(\ref{eq:H-hplus-hcross},~\ref{eq:z-hplus-hcross}), the extrinsic Fisher matrix from Equation~(\ref{eq:fisher-matrix}), and the Jacobian from Equation~(\ref{eq:jacobian-detector-tensor}).
-    \item Sum the information from all detectors using Equation~\ref{eq:summed-information}.
-    \item Compute the marginalized Fisher matrix from the Schur complement using Equation~(\ref{eq:marginal-fisher-matrix}).
-    \item If there are two detectors, find the width $L_p$ of the ring describing the $p$th quantile using Equation~(\ref{eq:l-pth-quantile}).
-If there are three or more detectors, find the area $A_p$ of the $p$th quantile using Equation~(\ref{eq:a-pth-quantile}).
-    \item (Optionally, convert from (ste)radians to (square) degrees.)
+    \item For a given $\phi, \theta, \psi$, compute the complex received amplitude $z_i$ from Eqs.~(\ref{eq:H-hplus-hcross},~\ref{eq:z-hplus-hcross}), the extrinsic Fisher matrix from Eq.~(\ref{eq:fisher-matrix}), and the Jacobian from Eq.~(\ref{eq:jacobian-detector-tensor}).
+    \item Sum the information from all detectors using Eq.~\ref{eq:summed-information}.
+    \item Compute the marginalized Fisher matrix from the Schur complement using Eq.~(\ref{eq:marginal-fisher-matrix}).
+    \item If there are two detectors, find the width $L_p$ of the ring describing the $p$th quantile using Eq.~(\ref{eq:l-pth-quantile}).
+If there are three or more detectors, find the area $A_p$ of the $p$th quantile using Eq.~(\ref{eq:a-pth-quantile}).
+    \item {[}Optionally, convert from (ste)radians to (square) degrees.{]}
 \end{enumerate}
 
-See code listing in Appendix~A.6 of \cite{leo-singer-thesis}.
+See the code listing in Appendix~A.6 of Ref.~\cite{leo-singer-thesis}.
 

### 2015-12-21 ms.tex
Remove fun part of title
PRD editors have no sense of humor.
Fix equation in proof
Increment DCC version

@@
 
-\title[Rapid-response Bayesian sky localization]{WHOOMP! (There it is):\\Rapid Bayesian position reconstruction for gravitational\nobreakdashes-wave transients}
+\title[Rapid-response Bayesian sky localization]{Rapid Bayesian position reconstruction for gravitational\nobreakdashes-wave transients}
 \author{Leo P. Singer}
@@
 \begin{equation}
-    p_\mathrm{max} = \sqrt{\frac{1}{2} \sum_i \left(\frac{r_{1,i}}{r}\right)^2}.
+    p_\mathrm{max} = \sqrt{\frac{1}{2} \sum_i \left(\frac{r_{1,i}}{r_\mathrm{max}}\right)^2}.
 \end{equation}
@@
 \begin{acknowledgements}
-This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v6.
+This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v7.
 

### 2015-12-21 ms.tex
Fix bad parentheses

@@
 
-The Laser Interferometer Gravitational Wave Observatory (\acsu{LIGO})~\cite{AdvancedLIGOHarry,AdvancedLIGO}) has just begun taking data \cite{LIGOObservingScenarios} in its ``Advanced'' configuration.
+The Laser Interferometer Gravitational Wave Observatory (\acsu{LIGO})~\cite{AdvancedLIGOHarry,AdvancedLIGO} has just begun taking data \cite{LIGOObservingScenarios} in its ``Advanced'' configuration.
 The two \ac{LIGO} detectors will ultimately increase their reach in volume within the local Universe by 3 orders of magnitude as compared to their initial configurations through 2010.

### 2015-12-23 ms.tex
Expand Ref. abbreviation at beginning of sentence
Break up some equations across multiple lines
Fix some typos found in proof

@@
 %
-\begin{widetext}
-\begin{equation}
+\begin{multline}
     p\left(\hat{\bm\theta}_i \middle| \bm\theta \right)
@@
         \middle| \bm\theta \right)
-    %\\
+    \\
     \propto \exp \Bigg[
@@
             \frac{\hat{\rho}_i}{\sigma_i(\hat{\bm\theta}_\mathrm{in})} e^{i (\hat\gamma_i - \omega \hat\tau_i)} \frac{H(\omega; \hat{\bm\theta}_\mathrm{in})}{S_i(\omega)}
-    %\\
+    \\
             - \frac{\rho_i}{\sigma_i(\bm\theta_\mathrm{in})} e^{i (\gamma_i - \omega \tau_i)} \frac{H(\omega; \bm\theta_\mathrm{in})}{S_i(\omega)}
@@
     \Bigg].
-\end{equation}
-\end{widetext}
+\end{multline}
 %
@@
 %
-\begin{widetext}
-\begin{equation}\label{eq:autocor-likelihood}
+\begin{multline}\label{eq:autocor-likelihood}
     p\left(\hat\rho_i, \hat\gamma_i, \hat\tau_i
-        \middle| \rho_i, \gamma_i, \tau_i \right) \propto
+        \middle| \rho_i, \gamma_i, \tau_i \right)
+    \\
+    \propto
     \exp \left[ - \frac{1}{2}{\hat\rho_i}^2 - \frac{1}{2}{\rho_i}^2
@@
     \right],
-\end{equation}
-\end{widetext}
+\end{multline}
 %
@@
 That pixel index times the area per pixel is the area of the smallest region of the specified credible level.
-This area can be thought of measuring the precision of the sky localization: it is a measure of the scale of the posterior distribution.
+This area can be thought of as measuring the precision of the sky localization: it is a measure of the scale of the posterior distribution.
 
@@
 The signals were detected using a template bank that lacked spins entirely.
-Ref.~\cite{FirstTwoYearsSpin} shows that using nonspinning \ac{BNS} templates for parameter estimation has negligible impact on sky localization.
+Reference~\cite{FirstTwoYearsSpin} shows that using nonspinning \ac{BNS} templates for parameter estimation has negligible impact on sky localization.
 However, if one or both companions are spinning as fast as a millisecond pulsar, $\chi \sim 0.4$ \cite{FastestSpinningMillisecondPulsar}, or even near breakup, $\chi \sim 0.7$, then the orbital plane may precess; in this case, spins can no longer be neglected for detection \cite{DetectingBNSSystemsWithSpin} and may also be important for parameter estimation.
@@
 
-This procedure has a shortcut called the Schur complement (see, for example, \citealt{numerical-recipes-inversion-partition}). Consider a partitioned square matrix $M$ and its inverse:
+This procedure has a shortcut called the Schur complement (see, for example, \citeauthor{numerical-recipes-inversion-partition}~\cite{numerical-recipes-inversion-partition}).
+Consider a partitioned square matrix $M$ and its inverse:
 %

### 2016-01-15 ms.tex
Add date published
Change to PRD style
Increment DCC version

@@
-\documentclass[amsmath,amssymb,aps,prx,reprint,nopreprintnumbers,nofootinbib,showpacs]{revtex4-1}
+\documentclass[amsmath,amssymb,aps,prd,reprint,nopreprintnumbers,nofootinbib,showpacs]{revtex4-1}
 \usepackage[nolist]{acronym}
@@
 \received{28 September 2015}
+\published{14 January 2016}
 \pacs{04.80.Nn, 04.30.Tv, 02.50.Tt}
@@
 \begin{acknowledgements}
-This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v7.
+This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v8.
 
